---
title:        "Univariate Analysis for Inflammation"
subtitle:     "Based on Hilmer, et al., (2022). *Blood-brain barrier disruption measured by albumin index correlates with inflammatory fluid biomarkers*"
author:       Azadeh Golduzian
date:         last-modified # today, now, last-modified
date-format:  long          # full, long, medium, short, iso,   https://quarto.org/docs/reference/dates.html
format:
  html:
    theme:                  litera
    highlight-style:        atom-one
    page-layout:            full      # article, full   # https://quarto.org/docs/output-formats/page-layout.html
    toc:                    true
    toc-depth:              4
    toc-location:           body      # left, body, right
    number-sections:        true      # true, false
    number-depth:           3
    code-fold:              show      # true (initially hidden), false, show (initially shown)
    code-tools:                       # menu top-right to show/hide all code
      toggle:               true
      caption:              "Code"    # none
      source:               false
    code-overflow:          scroll    # scroll, wrap
    code-block-bg:          true
    code-block-border-left: "#30B0E0"
    df-print:               paged     # default, kable, tibble, paged   # https://quarto.org/docs/computations/r.html
fig-width:                  6
fig-height:                 4
execute: # https://quarto.org/docs/computations/execution-options.html, https://quarto.org/docs/computations/r.html
  cache:    false   # false, true
  eval:     true    # true, false  Evaluate the code chunk (if false, just echos the code into the output).
  echo:     true    # true, false  Include the source code in output
---

<!---
# Erik's compile commands in R:
  setwd("D:/Dropbox/StatAcumen/consult/Authorship/2014_GaryRosenberg_UNM_NIH_VascularDementia_MRI/2024_ADRC/Analyses/202501_ADRC_Permeability")
  fn_qmd <- "Erhardt_Rosenberg_ADRC_Fluid_Permeability_Analysis_20250219.qmd"
  quarto::quarto_render(input = fn_qmd)
-->

```{r init, echo=FALSE}
#options(width = 80)
#options(warn  = -1)
options(str   = strOptions(list.len = 1e3))
options(knitr.kable.NA = '') # Display NAs as blanks
my_seed <- 34567
set.seed(my_seed)
```

--------------------------------------------------------------------------------

# Next actions

2/19/2025 Comparison presentation:

<https://docs.google.com/presentation/d/10neeggTjkKFs3SFFL9hJXMTKkTZ0W43RY4E2Myppx1o/edit#slide=id.g33159024d0d_0_357>



--------------------------------------------------------------------------------

# Data preliminaries

```{r chunk-003}
library(erikmisc)
library(tidyverse)
library(labelled)
library(corrplot)
library(broom)
# install.packages("dunn.test")
library(dunn.test)
library(ggpubr)
# devtools::install_github("vincentarelbundock/modelsummary")
# install.packages("modelsummary")
library(modelsummary)

dir.create("All_figures_for_paper", showWarnings = FALSE)
```


```{r}
load ("/Users/azadehgolduzian/Dropbox/2024_ADRC/Data/Data_out/current/ADRC4_ADRC3_MarkVCID2_VCI_ALL.RData")

load("dat_sub_Data_inflammation_analysis_paper_06052025.RData"
     )
load("dat_sub_and_Var_for_inflammation_analysis.RData")

```


# ╔══════════════════════════════════════════════════════════════════════════════╗
# ║           UNIVARIATE HEATMAP: AIC-SELECTED REGRESSION MODELS               ║
# ╚══════════════════════════════════════════════════════════════════════════════╝
#
# PURPOSE:
#   For each combination of (response variable × biomarker covariate), fit a
#   linear regression model and display the retained coefficients as a color-
#   coded heatmap. This is the primary screening tool for identifying which
#   inflammatory biomarkers associate with BBB permeability measures.
#
# RESPONSE VARIABLES (columns in the heatmap):
#   1. Albumin Index (log2)          — CSF measure of BBB permeability
#   2. 3T MRI Permeability (log2)    — Ktrans, imaging measure of BBB permeability
#   3. CSF GFAP (log2)               — Astrocyte activation marker (CSF)
#   4. Plasma GFAP (log2)            — Astrocyte activation marker (plasma)
#   [5. PC1 Composite Score]         — Optional, only if PC1 block was run above
#
# COVARIATES (rows in the heatmap):
#   All CSF, plasma, and MRI biomarkers in Var$csf, Var$plasma, Var$mri
#   EXCEPT the response variables themselves (excluded to prevent self-regression)
#
# MODEL STRUCTURE (for each response × covariate pair):
#   Full model:    Y ~ age + sex + X + age:X + sex:X
#   Selection:     AIC-based bidirectional stepwise (f_model_sel function)
#   The selected model keeps only the terms that improve AIC
#
# WHAT EACH HEATMAP COLUMN MEANS:
#   n          — sample size for that response × covariate pair
#   Age        — coefficient for age main effect
#   Gender     — coefficient for sex main effect
#   X          — coefficient for the biomarker (main association of interest)
#   Age:X      — age × biomarker interaction (association changes with age)
#   Gender:X   — sex × biomarker interaction (association differs by sex)
#
# HOW TO READ THE HEATMAP:
#   Color      — direction and magnitude of the coefficient
#                Red = positive association, Blue = negative association
#                White/gray = term not retained by AIC (coefficient = 0 or NA)
#   Stars      — p-value significance: * p<0.05, ** p<0.01, *** p<0.001
#   Colored but no star = AIC retained the term but p >= 0.05
#                        (important! colored = scientifically relevant,
#                         not just p-value driven)
#
# OUTPUT:
#   dat_plot          — long-format tibble with all coefficients and p-values
#   dat_plot_modified — dat_plot with cleaned labels for plotting
#   p_combined_AIC    — the final combined heatmap figure (CSF+MRI | Plasma)
#   Saved to: All_figures_for_paper/p_combined_AIC.png
#
# NOTE ON MULTIPLE COMPARISONS:
#   This is an exploratory analysis. AIC is the primary selection criterion,
#   not p-value thresholds. Colored tiles without stars directly address
#   concerns about p-value fishing — the model retained these terms on
#   information-theoretic grounds regardless of significance.
#   FDR correction code is available but commented out below.
# ─────────────────────────────────────────────────────────────────────────────
```{r}

dat_var_label_matching <-
  tibble::tribble(
    ~Var                                    , ~Group  , ~Type                          , ~Scale
  #, "ptid"                                 ,         ,                                ,
  #, "dx_current"                           ,         ,                                ,
  #, "sex"                                  ,         ,                                ,
  #, "demo_edu_yrs"                         ,         ,                                ,
  #, "age"                                  ,         ,                                ,
  ##### CSF                                 ,         ,                                ,
  , "fl_albumin_index_log2"                 , "CSF"   , "CSF Permeability"             , "log2"
  , "fl_igg_index_log2"                     , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf_igg_log2"                       , "CSF"   , "CSF Other"                    , "log2"
  , "fl_msd_csf_mmp_1_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_2_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_3_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_9_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_csf__ALL__mmp_10_log2"              , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_angio_csf_bfgf_log2"                , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_flt_1_log2"               , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_tie2_log2"                , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_a_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_c_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_d_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_cyto_csf_ifn_g_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_csf__ALL__il_1b_log2"               , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_2_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_4_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_6_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_8_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_10_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_12p70_log2"             , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_13_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_tnf_a_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_csf__ALL__nfl_log2"                 , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf__ALL__plgf_log2"                , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf__ALL__gfap_log2"                , "CSF"   , "CSF Other"                    , "log2"
  , "fl_abeta_4240_ratio_log2"              , "CSF"   , "CSF AD"                       , "log2"
  , "fl_csf__ALL__ptau_log2"                , "CSF"   , "CSF AD"                       , "log2"
  , "fl_myelin_basic_protein_log2"          , "CSF"   , "CSF Other"                    , "log2"
  ##### Plasma                              ,         ,                                ,
  , "fl_serum_igg_log2"                     , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__mmp_1_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_2_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_3_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_9_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_10_log2"           , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__bfgf_log2"             , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__flt_1_log2"            , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__tie2_log2"             , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_a_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_c_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_d_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_cyto_heparin_plasma_ifn_g_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_plasma__ALL__il_1b_log2"            , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_2_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_4_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_6_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_8_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_10_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_12p70_log2"  , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_13_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_tnf_a_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_plasma__ALL__nfl_log2"              , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__plgf_log2"             , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__gfap_log2"             , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__abeta_40_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__abeta_42_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__ptau_181_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__ptau_217_log2"         , "Plasma", "Plasma AD"                    , "log2"
  ##### MRI                                 ,         ,                                ,
  , "mri_percent_vol_hippo"                 , "MRI"   , "MRI"                          , "org"
  # , "mri_ucd_wmhv_percent"                  , "MRI"   , "MRI Other"                    , "org"  # 2/19/2025 too few observations
  , "mri_mean_cort_thick_mm"                , "MRI"   , "MRI"                          , "org"
  , "mri_perm_3t_tot"                       , "MRI"   , "MRI"                          , "org"
  , "mri_perm_3t_tot_log2"                  , "MRI"   , "MRI"                          , "log2"
  , "mri_dti_psmd"                          , "MRI"   , "MRI"                          , "org"
  , "mri_dti_freewater_mean"                , "MRI"   , "MRI"                          , "org"
  )

dat_var_label_matching <- dat_var_label_matching |> 
  mutate(Type_short = Type |> str_replace(pattern = "^CSF |^Plasma ", replacement = ""))

# Add this RIGHT AFTER dat_var_label_matching is defined
dat_var_label_matching <- dat_var_label_matching %>%
  dplyr::bind_rows(
    tibble::tibble(
      Var        = "PC1",
      Group      = "CSF",          # treat as CSF so it appears in panel A
      Type       = "CSF Other",
      Scale      = "pc1",
      Type_short = "Other"
    )
  )

print(dat_var_label_matching, n=Inf)

var_list_plot<-
  c(
    Var$csf
  , Var$plasma
  , Var$mri
  )

#normalize so color scale is meaningful

# data for demographics, before z-score scaling
dat_sheet_demo <-
  #dat_sheet
  dat_sub

dat_sheet <-
  #dat_sheet %>%
  dat_sub %>%
  dplyr::mutate(
    # to allow log transformation when wmhv==0
  #  percent_wmhv  = percent_wmhv + 0.00001
  #  fl_cyto_csf_il_4 = ifelse(fl_cyto_csf_il_4 == 0.001, NA, fl_cyto_csf_il_4)
  #  across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), e_log_shift, min_add = 1, base_log = 2)
    across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), scale)
  , across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), as.numeric)
  )

# ╔══════════════════════════════════════════════════════════════════════════════╗
# ║           OPTIONAL: ADD PC1 AS A 5TH RESPONSE VARIABLE                     ║
# ║                                                                              ║
# ║  This block creates a composite inflammatory score (PC1) from 5 CSF         ║
# ║  biomarkers and adds it to dat_sheet so it can be used as a response        ║
# ║  variable in the univariate heatmap alongside Albumin, Ktrans, CSF GFAP,   ║
# ║  and Plasma GFAP.                                                            ║
# ║                                                                              ║
# ║  IF YOU DO NOT WANT PC1 IN THE HEATMAP:                                     ║
# ║    1. Comment out this entire block (from ╔ to ╚)                           ║
# ║    2. Keep list_var_resp WITHOUT "PC1" (see below)                          ║
# ║    3. Keep ind_var_covar WITHOUT "PC1"                                      ║
# ║    4. Keep ind_row WITHOUT "PC1"                                            ║
# ║    5. Keep l_var_resp WITHOUT "PC1"                                         ║
# ║    6. Keep Label_resp levels WITHOUT "PC1 Composite Inflammatory Score"     ║
# ╚══════════════════════════════════════════════════════════════════════════════╝
# 
# vars_pca <- c(
#   "fl_albumin_index_log2",
#   "fl_csf__ALL__gfap_log2",
#   "fl_igg_index_log2",
#   "fl_cyto_csf_il_8_log2",
#   "fl_cyto_csf_il_6_log2"
# )
# 
# dat_pca <- dat_sub %>%
#   dplyr::select(ptid, all_of(vars_pca)) %>%
#   tidyr::drop_na()
# 
# cat("N for PCA:", nrow(dat_pca), "\n")
# 
# pca_fit <- prcomp(
#   dat_pca %>% dplyr::select(all_of(vars_pca)),
#   center = TRUE,
#   scale. = TRUE
# )
# 
# dat_pca$PC1_raw <- pca_fit$x[, 1]
# 
# cor_pc1_albumin <- cor(
#   dat_pca$PC1_raw,
#   dat_pca$fl_albumin_index_log2,
#   method = "spearman",
#   use    = "complete.obs"
# )
# cat("Spearman cor(PC1, Albumin):", round(cor_pc1_albumin, 3), "\n")
# 
# dat_pca$PC1 <- dat_pca$PC1_raw * ifelse(cor_pc1_albumin < 0, -1, 1)
# 
# # Merge PC1 into dat_sheet
# dat_sheet <- dat_sheet %>%
#   dplyr::left_join(
#     dat_pca %>% dplyr::select(ptid, PC1),
#     by = "ptid"
#   )
# 
# labelled::var_label(dat_sheet[["PC1"]]) <- "PC1 Composite Inflammatory Score"
# cat("PC1 added for", sum(!is.na(dat_sheet$PC1)), "participants\n")

# ── END OF OPTIONAL PC1 BLOCK ─────────────────────────────────────────────────





  # %>%
  #dplyr::mutate_if(is.numeric, scale) %>%

# label dataset
for (n_label in as.character(unlist(Var))) {
  ## i_label = 1
  labelled::var_label(dat_sheet     [[ n_label ]]) <- labelled::var_label(dat_sheet_demo[[ n_label ]])
  #labelled::var_label(dat_sheet_demo[[ dat_var_label_matching$Var[i_label] ]]) <- dat_var_label_matching$Label[i_label]
}

#get regression coefficients for reduced model

source("./model_sel_new.R")

lm_criterion <- c("AIC", "BIC")[1]

## model selection
f_model_sel <-
  function(
    dat_this      = dat_lm
  , lm_criterion  = c("AIC", "BIC")[1]
  ) {

  if (lm_criterion == "AIC") { AIC_k = 2                     }
  if (lm_criterion == "BIC") { AIC_k = log(nrow( dat_this )) }

  results_list     <- list()
  results_list$mod <- f_formula_init()
  results_list$fit <- list()
  results_list$AIC <- list()

  for (i_mod in seq_len(length(results_list$mod))) {
    ## i_mod = 1

    results_list$fit[[i_mod]] <-
      lm(
        formula = results_list$mod[[i_mod]]
      , data    = dat_this
      )

    results_list$AIC[[i_mod]] <-
      results_list$fit[[i_mod]] %>%
      AIC(k = AIC_k)
  }

  ## some models have a term with 0 df, skip this model
  # first remove models with no error estimate
  results_list$AIC[ results_list$AIC == -Inf ] <- Inf
  # then remove models with 0 df
  sw_df_zero <- TRUE
  while(sw_df_zero) {
    # select lowest AIC, index
    ind_min_AIC <-
      results_list$AIC %>%
      unlist() %>%
      which.min()

    sw_df_zero <-
      any(
        results_list$fit[[ind_min_AIC]] %>%
        car::Anova(type = 3, singular.ok = TRUE) %>%
        pull(Df)
          ==
        0
      )

    if (sw_df_zero) {
      print("Skipping 0-df model")
      #results_list$mod[[ind_min_AIC]] <- NULL
      #results_list$fit[[ind_min_AIC]] <- NULL
      #results_list$AIC[[ind_min_AIC]] <- NULL
      results_list$AIC[[ind_min_AIC]] <- Inf
    } # if
  } # while

  # selected model
  results <-
    results_list$fit[[ind_min_AIC]]

  return(results)
} # f_model_sel


list_var_resp <-
  c("fl_albumin_index_log2", 
    "mri_perm_3t_tot_log2",
    "fl_csf__ALL__gfap_log2",
    "fl_plasma__ALL__gfap_log2"
     # "PC1"  
    )
ind_var_covar <-
  #c(totransform.log2, tokeep) %notin%
  var_list_plot %notin%
    c("ptid","age","sex","demo_edu_yrs","dx_current","fl_albumin_index_log2",
      "mri_perm_3t_tot_log2",
      "mri_perm_3t_tot",
      "fl_csf__ALL__gfap_log2",
      "fl_plasma__ALL__gfap_log2"
      # "PC1"  

    )
list_var_covar <-
  var_list_plot[ind_var_covar]

results <- list()

for (i_resp in seq_along(list_var_resp)) {
  for (i_covar in seq_along(list_var_covar)) {
    ## i_resp = 2
    ## i_covar = 6
    ## i_resp = 1
    ## i_covar = 1
    ## i_resp = 1
    ## i_covar = 23

    var_resp  <- list_var_resp[i_resp]
    var_covar <- list_var_covar[i_covar]

    print(paste0("BEGIN  Resp: ", i_resp, " ", var_resp, ",  Covar: ", i_covar, " ", var_covar, "  ===================================================="))

    if (i_covar == 1) {
      results[[var_resp]] <- list()
    }
    results[[var_resp]][[var_covar]] <- list()

    # subset of data for linear model
    dat_lm <-
      dat_sheet %>%
      dplyr::select(
        all_of(var_resp)
      , all_of(var_covar)
      , age
      , sex
      #, demo_edu_yrs  # 3/9/2022 Educ added
      ) %>%
      dplyr::mutate(
        Y = !!sym(var_resp)
      , X = !!sym(var_covar)
      ) %>%
      tidyr::drop_na() %>%
      tidyr::as_tibble()
    n_sample <- nrow(dat_lm)

    if (n_sample == 0) {
      next
    }


    form_lm <-
      as.formula(
        paste0(
        #  var_resp
          "Y"
        , " ~ "
        , "age"
        , " + "
        , "sex"
        #, " + "
        #, "demo_edu_yrs"    # 3/9/2022 Educ added
        , " + "
        #, var_covar
        , "X"
        , " + "
        #, paste0("age", ":", var_covar)
        , paste0("age", ":", "X")
        , " + "
        #, paste0("sex", ":", var_covar)
        , paste0("sex", ":", "X")
        #, " + "
        #, paste0("demo_edu_yrs", ":", "X")    # 3/9/2022 Educ added
        )
      )

    results[[var_resp]][[var_covar]][["form_lm"]] <- form_lm
    results[[var_resp]][[var_covar]][["dat_lm" ]] <- dat_lm

    results[[var_resp]][[var_covar]][["lm_fit_full"]] <-
      lm(
        formula = form_lm
      , data    = dat_lm
      )

    ## results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]] <-
    ##   stats::step(
    ##     object = results[[var_resp]][[var_covar]][["lm_fit_full"]]
    ##   , direction = "both"
    ##   , test = "F"
    ##   , trace = 0
    ##   , k = log(nrow(dat_lm))
    ##   )

    results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]] <-
      f_model_sel(
        dat_this      = dat_lm
      , lm_criterion  = lm_criterion
      )


    results[[var_resp]][[var_covar]][["lm_fit_full"]]$df.residual

    ### 7/8/2022 report all results
    ## skip if model selection removed our x variable
    if (!("X" %in% names(coef(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]])))) {
      print("Model selection removed X variable")
      results[[var_resp]][[var_covar]]$sw_sig_X <- FALSE
      #next
    } else {
      results[[var_resp]][[var_covar]]$sw_sig_X <- TRUE
    }

    print(summary(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]))

    # plot residuals
tryCatch(
  plot(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]], which = c(1)),
  error = function(e) message("Skipping residual plot (n too small): ", e$message)
)
    # Scatterplot for fitted reduced model

    covar_names_red <- names(coef(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]))

    dat_pred <-
      tidyr::expand_grid(
        X = dat_lm[["X"]] %>% range()
      , age  = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
      , sex     = dat_lm[["sex"]] %>% unique()
      #, demo_edu_yrs = c(0, dat_lm[["demo_edu_yrs"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
      )

    ## if (
    ##   ("age"  %in% covar_names_red) &
    ##   ("sexMale" %in% covar_names_red) #&
    ##   #("demo_edu_yrs" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     , age  = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     , sex     = dat_lm[["sex"]] %>% unique()
    ##     , demo_edu_yrs = c(0, dat_lm[["demo_edu_yrs"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     )
    ## } else if (
    ##   ("age" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     , age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     #, sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## } else if (
    ##   ("sexMale" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     #, age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     , sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## } else {
    ##   # only X is significant
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     #, age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     #, sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## }

    dat_pred$Y <-
      predict(
        results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]
      , newdata = dat_pred
      )

    dat_pred <-
      dat_pred %>%
      dplyr::mutate(
        age  = age  %>% factor(labels = c("10%", "Mean", "90%"))
      #, demo_edu_yrs = demo_edu_yrs %>% factor(labels = c("10%", "Mean", "90%"))
      )

    ## if ("age" %in% covar_names_red) {
    ##   dat_pred <-
    ##     dat_pred %>%
    ##     dplyr::mutate(
    ##       age = age %>% factor(labels = c("10%", "Mean", "90%"))
    ##     )
    ## }
    ## if ("demo_edu_yrs" %in% covar_names_red) {
    ##   dat_pred <-
    ##     dat_pred %>%
    ##     dplyr::mutate(
    ##       demo_edu_yrs = demo_edu_yrs %>% factor(labels = c("10%", "Mean", "90%"))
    ##     )
    ## }

    library(ggplot2)
    if ("sexFemale" %in% covar_names_red) {
      p <- ggplot(dat_lm, aes(x = X, y = Y, shape = sex, colour = sex))
    } else {
      p <- ggplot(dat_lm, aes(x = X, y = Y))
    }
    p <- p + theme_bw()
    p <- p + geom_point(size = 2, alpha = 1/2)
    if ("age" %in% covar_names_red) {
      p <- p + geom_line(data = dat_pred, aes(linetype = age), size = 0.75)
    } else {
      p <- p + geom_line(data = dat_pred, size = 0.75)
    }
    # 3/9/2022 Educ added  XXX

    p <- p + scale_linetype_manual(values = c(2, 1, 3))
    p <- p + labs(  title = NULL
                  #, subtitle = "Progress and Starting Current"
                  , x = labelled::var_label(dat_sheet[[var_covar]])
                  , y = labelled::var_label(dat_sheet[[var_resp ]])
                  #, caption = paste0(  "Model fit averages the numeric variables of Age, A1c, and PAM-10\n"
                  #                   , "  over Program, Diagnosis, Gender, and General health.\n"
                  #                   , "Dotted vertical lines indicate cutoffs between depression categories:\n"
                  #                   , "  None-minimal, Mild, Moderate, Moderately Severe, and Severe."
                  #                   )
                  , colour    = "Gender"
                  , shape     = "Gender"  # "Imputed"
                  , linetype  = "Age"  #"Diagnosis"
                  #, fill      = "Diagnosis"
                  )
    #p <- p + theme(plot.caption = element_text(hjust = 0))
    #p <- p + theme(legend.position = "bottom")
    #p <- p + guides(colour = guide_legend(nrow = 2), shape = guide_legend(nrow = 2))
    n_sample_annotate <- paste0("N=",n_sample)
    #p <- p + annotate("text", -Inf, -Inf, label=n_sample_annotate, hjust=-0.3, vjust=-1, colour = "gray25")
    p <- p + annotate("text", x=-Inf, y= Inf, label=n_sample_annotate, hjust=-0.3, vjust=2, colour = "gray25")
    print(p)

    results[[var_resp]][[var_covar]][["plot_lm_fit_red_BIC"]] <-
      p

    print(paste0("END    Resp: ", i_resp, " ", var_resp, ",  Covar: ", i_covar, " ", var_covar, "  ===================================================="))
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")


  } # i_covar
} # i_resp

dat_lm
#create matrices
ind_row <-
  #c(totransform.log2, tokeep) %notin%
  #  c("ptid","age","sex","demo_edu_yrs","dx_current","fl_albumin_index_log2","mri_perm_3t_tot")
  as.character(unlist(Var)) %notin%
    c("ptid","age","sex"
      ,"demo_edu_yrs",
      "dx_current",
      "fl_albumin_index_log2"
      ,"mri_perm_3t_tot_log2",
      "fl_plasma__ALL__gfap_log2",
      "fl_csf__ALL__gfap_log2"
       # "PC1"  
      )

n_row <-
  #c(totransform.log2, tokeep)[ind_row]
  as.character(unlist(Var))[ind_row]

l_row <-
  length(n_row)

n_col <-
  #c("Age", "Gender", "Education", "X", "Age:X", "Gender:X", "Education:X")
  c("Age", "Gender",              "X", "Age:X", "Gender:X"               )
l_col <-
  length(n_col)

# fill matrices with the signficant estimates, leave other cells blank

l_var_resp<-
  c("fl_albumin_index_log2",
    "mri_perm_3t_tot_log2",
    "fl_csf__ALL__gfap_log2",
    "fl_plasma__ALL__gfap_log2"
    # "PC1"  

    )

results_fit <- list()
i_results <- 0

for (i_resp in seq_along(l_var_resp)) {
  # i_resp = 1
  for (i_covar in seq_along(n_row)) {
    ## i_covar = 1

    var_row <- n_row[i_covar]
    var_col <- l_var_resp[i_resp]

    exists <- results[[var_col]][[var_row]]

    if(length(results[[var_col]][[var_row]]) == 0) {
      next
    }

    i_results <- i_results + 1

    results_fit[[ i_results ]] <-
      results[[var_col]][[var_row]][["lm_fit_red_BIC"]] |>
      summary() |>
      broom::tidy() |>
      dplyr::mutate(
        term  =
          term |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("sexFemale")
          , replacement = "Gender"
          ) |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("age")
          , replacement = "Age"
          )
      , Covar =
          term |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("X")
          , replacement = var_row
          )
      , resp  = var_col
      , covar = var_row
      , n     = results[[var_col]][[var_row]][[ "dat_lm" ]] |> nrow()
      ) |>
      dplyr::relocate(
        resp
      , covar
      )

  } # i_covar
} # i_resp

dat_plot <-
  results_fit |>
  dplyr::bind_rows() |>
  dplyr::rename(
  #  Var   = covar
    Coef  = estimate
  , pvals = p.value
  ) %>%
  dplyr::mutate(
    pvals_stars =
      pvals %>%
      e_pval_stars(
        cutpoints = c(0, 0.001, 0.01, 0.05, 1)
      , symbols = c("***", "**", "*", " ")
      )
  ) %>%
  left_join(
    dat_var_label_matching
  , by = dplyr::join_by(covar == Var)
  ) %>%
  left_join(
    e_data_var_labels_table_from_data(dat_sheet)
  , by = dplyr::join_by(covar == Var)
  ) %>%
  dplyr::rename(
    Label_covar = Label
  ) %>%
  left_join(
    e_data_var_labels_table_from_data(dat_sheet)
  , by = dplyr::join_by(resp == Var)
  ) %>%
  dplyr::rename(
    Label_resp = Label
  ) %>%
  dplyr::mutate(
    Label_covar = Label_covar |> factor()
  , Label_resp  = Label_resp  |> factor()
  # , term        = term        |> factor(levels = c("(Intercept)", n_col))
  , Type_short  = Type_short  |> factor(levels = c("AD", "Angio", "Cytokines", "MMP", "Other", "MRI"))
  )
# dat_plot <- dat_plot %>%
#   dplyr::mutate(
#     Label_resp = dplyr::if_else(
#       resp == "PC1" & is.na(as.character(Label_resp)),
#       "PC1 Composite Inflammatory Score",
#       as.character(Label_resp)
#     ),
#     Label_resp = factor(Label_resp)
#   )
# 
# # Verify PC1 rows exist with correct label
# dat_plot %>%
#   dplyr::filter(resp == "PC1") %>%
#   dplyr::distinct(resp, Label_resp) %>%
#   print()

# ── CORRECTED FDR block ──────────────────────────────────────────────────────

# Adjust ALL biomarker-related terms (X, Age:X, Gender:X)
# Done SEPARATELY per response variable (they are independent families)

dat_plot <- dat_plot %>%
  dplyr::group_by(resp) %>%                          # separate correction per outcome
  dplyr::mutate(
    pvals_adj = dplyr::case_when(
      term %in% c("X", "Age:X", "Gender:X") ~ p.adjust(pvals, method = "fdr"),
      TRUE ~ pvals                                   # Age, Gender, Intercept: unadjusted
    ),
    pvals_stars = pvals_adj %>%
      e_pval_stars(
        cutpoints = c(0, 0.001, 0.01, 0.05, 1),
        symbols   = c("***", "**", "*", " ")
      )
  ) %>%
  dplyr::ungroup()

# ────────────────────────────────────────────────────────────────────────────

# ────────────────────────────────────────────────────────────────────────────



# ────────────────────────────────────────────────────────────────────────────

#dat_plot <-
#  dplyr::full_join(
#    dplyr::bind_rows(
#      #coef_mat %>%
#      results_fit %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "Albumin Index"
#      )
#    , coef_mat_mri %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "MRI WM Permeability"
#      )
#    ) %>%
#    tidyr::pivot_longer(
#      cols = -c(Var, RespVar)
#    , names_to = "Covar"
#    , values_to = "Coef"
#    )
#  , dplyr::bind_rows(
#      p_mat %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "Albumin Index"
#      )
#    , p_mat_mri %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "MRI WM Permeability"
#      )
#    ) %>%
#    tidyr::pivot_longer(
#      cols = -c(Var, RespVar)
#    , names_to = "Covar"
#    , values_to = "pvals"
#    )
#  , by = c("Var", "RespVar", "Covar")
#  ) %>%
#  dplyr::mutate(
#  #  Var = Var %>% stringr::str_replace_all(pattern = stringr::fixed("_"), replacement = " ")
#  #, Var = Var %>% stringr::str_replace_all(c("fl" = "","csf" = "","t 1" = "flt-1"))
#    pvals_stars =
#      pvals %>%
#      e_pval_stars(
#        cutpoints = c(0, 0.001, 0.01, 0.05, 1)
#      , symbols = c("***", "**", "*", " ")
#      )
#  ) %>%
#  left_join(
#    dat_var_label_matching
#  ) %>%
#  left_join(
#    e_data_var_labels_table_from_data(dat_sheet)
#  ) %>%
#  dplyr::mutate(
#    Label = Label |> factor()
#  , Covar = Covar %>% factor(levels = n_col)
#  )


dat_plot <- dat_plot %>%
  dplyr::mutate(
    Label_covar = Label_covar %>%
      # Plasma Cytokines - shorten the long MSD V-PLEX names
      stringr::str_replace_all("Heparin Plasma IFN-γ MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IFN-γ") %>%
      stringr::str_replace_all("Heparin Plasma IL-10 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-10") %>%
      stringr::str_replace_all("Heparin Plasma IL-12p70 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-12p70") %>%
      stringr::str_replace_all("Heparin Plasma IL-13 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-13") %>%
      stringr::str_replace_all("Heparin Plasma IL-2 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-2") %>%
      stringr::str_replace_all("Heparin Plasma IL-4 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-4") %>%
      stringr::str_replace_all("Heparin Plasma IL-6 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-6") %>%
      stringr::str_replace_all("Heparin Plasma IL-8 MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma IL-8") %>%
      stringr::str_replace_all("Heparin Plasma TNF-a MSD V-PLEX Proinflammatory Panel 1 Human Kit", "Plasma TNF-α") %>%
      
      # CSF Cytokines
      stringr::str_replace_all("CSF Neuro IFN-g \\(ALL\\)", "CSF IFN-γ") %>%
      stringr::str_replace_all("CSF Neuro IL-10 \\(ALL\\)", "CSF IL-10") %>%
      stringr::str_replace_all("CSF IL-12p70", "CSF IL-12p70") %>%
      stringr::str_replace_all("CSF IL-13", "CSF IL-13") %>%
      stringr::str_replace_all("CSF IL-1b", "CSF IL-1β") %>%
      stringr::str_replace_all("CSF IL-2", "CSF IL-2") %>%
      stringr::str_replace_all("CSF IL-4", "CSF IL-4") %>%
      stringr::str_replace_all("CSF IL-6", "CSF IL-6") %>%
      stringr::str_replace_all("CSF IL-8", "CSF IL-8") %>%
      stringr::str_replace_all("CSF TNF-a", "CSF TNF-α") %>%
      
      # CSF MMPs
      stringr::str_replace_all("Csf MMP-10 \\(ALL\\)", "CSF MMP-10") %>%
      stringr::str_replace_all("MSD CSF MMP-1", "CSF MMP-1") %>%
      stringr::str_replace_all("MSD CSF MMP-2", "CSF MMP-2") %>%
      stringr::str_replace_all("MSD CSF MMP-3", "CSF MMP-3") %>%
      stringr::str_replace_all("MSD CSF MMP-9", "CSF MMP-9") %>%
      
      # Plasma MMPs
      stringr::str_replace_all("Plasma MMP-1 \\(ALL\\)", "Plasma MMP-1") %>%
      stringr::str_replace_all("Plasma MMP-10 \\(ALL\\)", "Plasma MMP-10") %>%
      stringr::str_replace_all("Plasma MMP-2 \\(ALL\\)", "Plasma MMP-2") %>%
      stringr::str_replace_all("Plasma MMP-3 \\(ALL\\)", "Plasma MMP-3") %>%
      stringr::str_replace_all("Plasma MMP-9 \\(ALL\\)", "Plasma MMP-9") %>%
      
      # CSF Angiogenic Factors
      stringr::str_replace_all("CSF bFGF", "CSF bFGF") %>%
      stringr::str_replace_all("CSF Flt-1", "CSF Flt-1") %>%
      stringr::str_replace_all("CSF Tie-2", "CSF Tie-2") %>%
      stringr::str_replace_all("CSF VEGF-A", "CSF VEGF-A") %>%
      stringr::str_replace_all("CSF VEGF-C", "CSF VEGF-C") %>%
      stringr::str_replace_all("CSF VEGF-D", "CSF VEGF-D") %>%
      
      # Plasma Angiogenic Factors
      stringr::str_replace_all("Plasma bFGF", "Plasma bFGF") %>%
      stringr::str_replace_all("Plasma Flt-1", "Plasma Flt-1") %>%
      stringr::str_replace_all("Plasma TIE2", "Plasma Tie-2") %>%
      stringr::str_replace_all("Plasma VEGF-A", "Plasma VEGF-A") %>%
      stringr::str_replace_all("Plasma VEGF-C", "Plasma VEGF-C") %>%
      stringr::str_replace_all("Plasma VEGF-D", "Plasma VEGF-D") %>%
      
      # CSF Other markers
      stringr::str_replace_all("CSF Neuro GFAP \\(ALL\\)", "CSF GFAP") %>%
      stringr::str_replace_all("CSF Neuro NFL \\(ALL\\)", "CSF NFL") %>%
      stringr::str_replace_all("CSF PlGF \\(ALL\\)", "CSF PlGF") %>%
      stringr::str_replace_all("CSF IgG", "CSF IgG") %>%
      stringr::str_replace_all("IgG Index", "IgG Index") %>%
      stringr::str_replace_all("Myelin Basic Protein", "Myelin Basic Protein") %>%
      
      # Plasma Other markers
      stringr::str_replace_all("Plasma Neuro GFAP \\(ALL\\)", "Plasma GFAP") %>%
      stringr::str_replace_all("Plasma NFL \\(ALL\\)", "Plasma NFL") %>%
      stringr::str_replace_all("Plasma PlGF \\(ALL\\)", "Plasma PlGF") %>%
      stringr::str_replace_all("Serum IgG", "Serum IgG") %>%
      stringr::str_replace_all("Plasma IL-1B \\(ALL\\)", "Plasma IL-1β") %>%
      
      # AD markers
      stringr::str_replace_all("A Beta 42/40 ratio", "Aβ42/40 Ratio") %>%
      stringr::str_replace_all("CSF p-Tau 181 \\(ALL\\)", "CSF p-Tau 181") %>%
      stringr::str_replace_all("Plasma Neuro A beta 40 \\(ALL\\)", "Plasma Aβ40") %>%
      stringr::str_replace_all("Plasma Neuro A Beta 42 \\(ALL\\)", "Plasma Aβ42") %>%
      stringr::str_replace_all("plasma p-Tau 217 \\(ALL\\)", "Plasma p-Tau 217") %>%
      stringr::str_replace_all("Plasma p-Tau 181 \\(ALL\\)", "Plasma p-Tau 181") %>%
      
      # Catch-all: Remove any remaining (ALL) and clean up extra spaces
      stringr::str_replace_all("\\s*\\(ALL\\)\\s*", "") %>%
      stringr::str_replace_all("\\s+", " ") %>%
      stringr::str_trim()|>
      
      # MRI markers
      stringr::str_replace_all("DTI PSMD", "DTI PSMD") %>%
      stringr::str_replace_all("Hippocampal volume \\(percent\\)", "Hippocampal Volume (%)") %>%
      stringr::str_replace_all("Mean cortical thickness \\(mm\\)", "Mean Cortical Thickness (mm)") %>%
      stringr::str_replace_all("Mean Free Water", "Mean Free Water")
    
  , Label_covar = Label_covar |> str_replace(pattern = "^CSF |^Plasma ", replacement = "")
  )


# # p-value plot
# library(ggplot2)
# #p1 <- ggplot(data = dat_plot, aes(x = Covar, y = Label, fill = Coef))  # y = Var
# p1 <- ggplot(data = dat_plot, aes(x = term, y = Label_covar, fill = Coef))  # y = Var
# #p1 <- p1 + theme_bw()
# p1 <- p1 + geom_tile(na.rm = TRUE)
# p1 <- p1 + scale_y_discrete(limits = rev)
# #p1 <- p1 + scale_fill_gradientn(limits = c(-1.0, 0.6), values = scales::rescale(c(0,0.1,1)), colors = c("#000000", "#AAAAAA", "#DDDDDD"), breaks = c(0, 0.1, 0.5, 1), na.value = "white")
# p1 <- p1 + scale_fill_gradient2(low = "blue", mid = "white", high = "red", breaks = seq(-100, 100, by = 2), na.value = "gray90")
# p1 <- p1 + geom_text(aes(label = pvals_stars), vjust = 0.5, colour = "white")
# p1 <- p1 + geom_text(data = dat_plot |> dplyr::filter(term == "(Intercept)")
#               , aes(label = n), hjust = 0, vjust = 0.5, colour = "black"
#               )
# #p1 <- p1 + facet_grid(. ~ RespVar, space = "free", scales = "free_y", drop = TRUE)
# p1 <- p1 + facet_grid(Type ~ Label_resp, space = "free", scales = "free_y", drop = TRUE)
# p1 <- p1 + theme(#axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)
#                axis.text.y = element_text(hjust = 0)
#             )  #, panel.grid.major = element_blank(), panel.grid.minor = element_blank())
# p1 <- p1 + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) # rotate labels
# p1 <- p1 + theme(axis.text.y = element_text(hjust = 1))
# #p1 <- p1 + theme(legend.key.height = unit(0.30, "in"))
# p1 <- p1 + theme(plot.caption = element_text(hjust = 0), plot.caption.position =  "plot") # caption left align
# p1 <- p1 + theme(legend.position = "bottom") # "none"
# p1 <- p1 + theme(legend.key.height = unit(0.1, "in"), legend.key.width = unit(0.4, "in"))
# 
# if (lm_criterion == c("AIC", "BIC")[1]) {
#   p1 <- p1 + labs(
#                   title = "Models selected by AIC"
#                   )
# }
# if (lm_criterion == c("AIC", "BIC")[2]) {
#   p1 <- p1 + labs(
#                   title = "Models selected by BIC"
#                   )
# }
# p1 <- p1 + labs(
#               #, subtitle = "Longitudinal mixed model"
#                 y = "Standardized covariates (\"X\")"
#               , x = "Model covariates"
#               , fill = "Model\nCoefficients  "  # "FDR\np-value"
#               , caption = paste0(
#                             "p-value stars:  (blank): p >= 0.05;   *: p < 0.05;   **: p < 0.01;   *** : p < 0.001"
#                           , "\np-values are uncorrected for family-wise Type-I error rate."
#                           #, "\nModels displayed are those where X variable is retained after model selection"
#                           )
#               )
# 
# print(p1)


# If you need two plots run this code 

# ---- shared scale so BOTH panels map colors identically ----
# truncate Coef for color scale, so other colors are darker
# dat_plot <-
#   dat_plot |>
#   mutate(
#     Coef = ifelse(abs(Coef) > 5, sign(Coef) * 5, Coef)
#   )
fill_limits <- c(-3, +3) #range(dat_plot$Coef, na.rm = TRUE) 
fill_breaks <- seq(floor(fill_limits[1]), ceiling(fill_limits[2]), by = 1)


dat_plot <- dat_plot %>%
  dplyr::mutate(
    pvals_stars = as.character(pvals_stars),           # ← convert first
    term        = dplyr::if_else(term == "(Intercept)", "n", as.character(term)),
    Coef        = ifelse(term == "n", NA_real_,        Coef),
    pvals_stars = ifelse(term == "n", NA_character_,   pvals_stars)
  )
# Standardize the labels first
dat_plot <- dat_plot %>%
  dplyr::mutate(
    Label_covar = Label_covar %>%
      stringr::str_replace_all("pTau 181", "p-Tau 181") %>%
      stringr::str_replace_all("IFN-g", "IFN-γ") %>%
      stringr::str_replace_all("Neuro NFl", "NFL") %>%
      stringr::str_replace_all("NFl", "NFL")
  )

# Now define the correct order with exact matches
biomarker_order <- c(
  # AD markers
  "Aβ42/40 Ratio",
  "p-Tau 181",
  "Aβ40",
  "Aβ42",
  "p-Tau 217",
  
  # Angiogenic factors
  "bFGF",
  "Flt-1",
  "PlGF",
  "Tie-2",
  "VEGF-A",
  "VEGF-C",
  "VEGF-D",
  
  # Cytokines - order IL numerically
  "IFN-γ",
  "IL-1β",
  "IL-2",
  "IL-4",
  "IL-6",
  "IL-8",
  "IL-10",
  "IL-12p70",
  "IL-13",
  "TNF-α",
  
  # MMPs - order numerically
  "MMP-1",
  "MMP-2",
  "MMP-3",
  "MMP-9",
  "MMP-10",
  
  # Other markers
  "GFAP",
  "IgG",
  "IgG Index",
  "Myelin Basic Protein",
  "NFL",
  "Serum IgG",
  
  # MRI markers
  "DTI PSMD",
  "Hippocampal Volume (%)",
  "Mean Cortical Thickness (mm)",
  "Mean Free Water"
)

# Now apply factor ordering
dat_plot <- dat_plot %>%
  dplyr::mutate(
    Label_covar = factor(Label_covar, levels = biomarker_order)
  )

# Verify no NAs
dat_plot %>% 
  dplyr::distinct(Label_covar, Type_short) %>% 
  dplyr::arrange(Type_short, Label_covar) %>% 
  print(n = Inf)

# Recreate dat_plot_modified from dat_plot (which has correct Label_resp)
dat_plot_modified <- dat_plot %>%
  dplyr::mutate(
    term = dplyr::if_else(term == "(Intercept)", "n", as.character(term)),
    Coef = dplyr::if_else(term == "n", NA_real_, Coef),
    # Set Label_resp factor with correct levels - Albumin Index first
  Label_resp = factor(Label_resp, levels = c("Albumin Index", "3T Permeability (Total)",
                                            "CSF GFAP", "Plasma GFAP"))
  )


dat_plot_modified <- dat_plot %>%
  dplyr::mutate(
    term        = dplyr::if_else(term == "(Intercept)", "n", as.character(term)),
    Coef        = dplyr::if_else(term == "n", NA_real_, Coef),
    Label_resp  = Label_resp %>%
      as.character() %>%
      stringr::str_replace_all("CSF Neuro GFAP \\(ALL\\)", "CSF GFAP") %>%
      stringr::str_replace_all("Plasma Neuro GFAP \\(ALL\\)", "Plasma GFAP") %>%
      stringr::str_replace_all("\\s*\\(ALL\\)\\s*", "") %>%
      stringr::str_trim(),
    Label_resp  = factor(Label_resp, levels = c(
      "Albumin Index", 
      "3T Permeability (Total)", 
      "CSF GFAP", "Plasma GFAP",  
      "PC1 Composite Inflammatory Score"
    ))
  )

# Verify all 4 levels present, no NAs
dat_plot_modified %>% distinct(resp, Label_resp)







# Update create_plot function with right-aligned n values and Qalb before Ktrans
create_plot <- function(dat, plot_title,
                        show_caption = TRUE,
                        show_legend = TRUE) {
  
  p <- ggplot(dat, aes(x = term, y = Label_covar, fill = Coef)) +
    geom_tile(na.rm = TRUE) +
    
    # Add white/neutral tiles for the n column
    geom_tile(
      data = dat %>% dplyr::filter(term == "n"),
      fill = "white",
      color = "gray80"
    ) +
    
    # Set the order of x-axis with "n" first
    scale_x_discrete(limits = c("n", "Age", "Gender", "X", "Age:X", "Gender:X")) +
    scale_y_discrete(limits = rev) +
    scale_fill_gradient2(
      low = "blue",
      mid = "white",
      high = "red",
      limits = fill_limits,
      breaks = fill_breaks,
      na.value = "gray90",
      oob = scales::squish
    ) +
    
    # Show p-value stars for non-n columns
    geom_text(
    data = dat %>% dplyr::filter(term != "n", !is.na(pvals_stars)),
    aes(label = pvals_stars),
    vjust = 0.5, colour = "gray20"
            ) +
    
    # Show n values in the "n" column - RIGHT ALIGNED
    geom_text(
      data = dat %>% dplyr::filter(term == "n"),
      aes(label = n),
      hjust = 1, vjust = 0.5, colour = "black", nudge_x = 0.4
    ) +
    
    facet_grid(
      Type_short ~ Label_resp,
      space = "free",
      scales = "free_y",
      drop = TRUE
    ) +
    
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      axis.text.y = element_text(hjust = 1),
      legend.position = if (show_legend) "bottom" else "none",
      legend.key.height = unit(0.15, "in"),
      legend.key.width  = unit(0.6, "in"),
      plot.caption = element_text(hjust = 0),
      plot.caption.position = "plot"
    ) +
    
    labs(
      title = plot_title,
      x = "Model covariates",
      y = "Standardized covariates (\"X\")",
      fill = if (show_legend) "Model\nCoefficients" else NULL,
      caption = if (show_caption)
        paste0(
          "p-value stars: (blank): p >= 0.05; *: p < 0.05; **: p < 0.01; *** : p < 0.001",
          "\np-values are uncorrected for family-wise Type-I error rate."
        )
      else NULL
    )
  
  p
}


# Build panels using modified data
p1_csf_mri <- create_plot(
  dat      = dat_plot_modified %>% dplyr::filter(Group %in% c("CSF", "MRI")), 
  plot_title = "CSF and MRI",
  show_caption = FALSE,
  show_legend  = FALSE
)

p2_plasma <- create_plot(
  dat      = dat_plot_modified %>% dplyr::filter(Group == "Plasma"),
  plot_title = "Plasma",
  show_caption = FALSE,
  show_legend  = TRUE
)


# ---- combine ----
library(patchwork)
# p_combined <- p1_csf_mri + p2_plasma + plot_layout(ncol = 2) +
#     patchwork::plot_annotation(
#       title       = "Models selected by BIC"
#     # , subtitle    = text_formula_sel
#     # , caption     = 
#     , tag_levels  = "A"
#     , theme = theme(plot.caption = element_text(hjust = 0)) # Default is hjust=1, Caption align left
#     )

plot_design <-
    "AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     AB
     CC"
  
p_combined_AIC <- p1_csf_mri + p2_plasma + guide_area() +
  plot_layout(guides = "collect") + 
  patchwork::plot_layout(design = plot_design) +
    patchwork::plot_annotation(
      title       = "Models selected by AIC"
    # , subtitle    = text_formula_sel
    , caption     = paste0("p-value stars: (blank): p >= 0.05;   *: p < 0.05;   **: p < 0.01;   *** : p < 0.001",
               "\np-values are uncorrected for family-wise Type-I error rate.")
    , tag_levels  = "A"
    , theme = theme(plot.caption = element_text(hjust = 0)) # Default is hjust=1, Caption align left
    )

print(p_combined_AIC)

ggsave(
    "All_figures_for_paper/p_combined_AIC.png"
  , plot   = p_combined_AIC
  , width  = 15
  , height = 10
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )




# create PDF
# ggsave(
#     "All_figures_for_paper/p_combined.pdf"
#   , plot   = p_combined
#   , width  = 10
#   , height = 10
#   ## png, jpeg
#   #, dpi    = 300
#   #, bg     = "white"
#   ## pdf
#   , units  = "in"
#   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
#   )

## #p_alb_mriperm_coef_pval <- p1
##
## # ggsave(
## #     paste0("plot_alb_mriperm_coef_pval_", lm_criterion, ".png")
## #   , plot   = p1
## #   , width  = 6
## #   , height = 8
## #   ## png, jpeg
## #   , dpi    = 300
## #   , bg     = "white"
## #   ## pdf
## #   #, units  = "in"
## #   #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )
## # ggsave(
## #     paste0("plot_alb_mriperm_coef_pval_", lm_criterion, ".pdf")
## #   , plot   = p1
## #   , width  = 6
## #   , height = 8
## #   ## png, jpeg
## #   #, dpi    = 300
## #   #, bg     = "white"
## #   ## pdf
## #   , units  = "in"
## #   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )





# this is the code to generate the demographic table for paper 


# ---- packages ----
library(dplyr)
library(forcats)
library(gtsummary)
library(dunn.test)

# ---- choose dataset & diagnosis variable ----
# Use your raw-scale data (before z-scoring)
dat_demo <- dat_sub


dx_var <- if ("dx_biological" %in% names(dat_demo)) "dx_biological" else "dx_current"

# keep only rows with a diagnosis
dat_demo <- dat_demo %>% filter(!is.na(.data[[dx_var]]))

# order dx levels as in the paper (adjust these to match your actual levels)
dx_levels <- c("AD", "LA", "Mixed", "SIVD", "MI")
actual_levels <- unique(dat_demo[[dx_var]])
dx_levels_present <- dx_levels[dx_levels %in% actual_levels]

if (length(dx_levels_present) > 0) {
  dat_demo[[dx_var]] <- factor(dat_demo[[dx_var]], levels = dx_levels_present)
}

# make sure Sex is a labeled factor (Female/Male)
dat_demo <- dat_demo %>%
  mutate(
    sex = forcats::fct_relevel(as.factor(sex), "Female", "Male")
  )

# ---- Check which variables have data ----
# First, let's identify which variables actually have non-missing data
vars_to_check <- c("age", "sex", "demo_edu_yrs", "fl_albumin_index_log2", "mri_perm_3t_tot_log2")

# Check completeness of each variable
completeness_check <- dat_demo %>%
  select(all_of(c(dx_var, vars_to_check))) %>%
  summarise(across(everything(), ~sum(!is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "non_missing_count")

print("Variable completeness:")
print(completeness_check)

# Create list of variables with sufficient data (at least 10 non-missing observations)
vars_with_data <- completeness_check %>%
  filter(non_missing_count >= 10) %>%
  pull(variable) %>%
  setdiff(dx_var)  # remove the grouping variable

print("Variables with sufficient data:")
print(vars_with_data)

# ---- Create table with available variables ----
# Build the variable list and labels dynamically
table_vars <- c(dx_var, vars_with_data)

# Create label list for available variables
label_list <- list()
# if ("age" %in% vars_with_data) label_list[["age"]] <- "Age"
# if ("sex" %in% vars_with_data) label_list[["sex"]] <- "Sex"
# if ("demo_edu_yrs" %in% vars_with_data) label_list[["demo_edu_yrs"]] <- "Years of education"
# if ("fl_albumin_index" %in% vars_with_data) label_list[["fl_albumin_index"]] <- "Albumin Index"
# if ("fl_albumin_index_log2" %in% vars_with_data) label_list[["fl_albumin_index_log2"]] <- "Albumin Index (log2)"
# if ("mri_perm_3t_tot" %in% vars_with_data) label_list[["mri_perm_3t_tot"]] <- "MRI WM Permeability"

# ---- Table 1 ----
tbl1 <- dat_demo %>%
  select(all_of(table_vars)) %>%
  gtsummary::tbl_summary(
    by = !!rlang::sym(dx_var),
    type = if("sex" %in% vars_with_data) list(sex ~ "categorical") else NULL,
    statistic = list(
      all_continuous() ~ "{median} [{p25}; {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    missing = "no",
    label = label_list
  ) %>%
  add_overall(last = FALSE) %>%
  add_n()

# Only add p-values if we have more than one diagnostic group
n_groups <- length(unique(dat_demo[[dx_var]][!is.na(dat_demo[[dx_var]])]))

if (n_groups > 1) {
  tbl1 <- tbl1 %>%
    add_p(
      test = list(
        all_continuous() ~ "kruskal.test",
        all_categorical() ~ "chisq.test"
      )
    )
}

tbl1 <- tbl1 %>%
  modify_caption("**Table 1. Demographic and permeability characteristics overall and by diagnostic category.**") %>%
  bold_labels()

# Create GT table with footnote
tbl1_gt <- tbl1 %>%
  as_gt() %>%
  gt::tab_source_note(
    source_note = paste(
      "Table values: Median [Q1; Q3] for numeric values and n (%) for categorical values.",
      if(n_groups > 1) "P-values from Kruskal–Wallis (continuous) and chi-square with continuity correction when applicable (categorical)." else "",
      "AD: Alzheimer's disease; LA: leukoaraiosis; Mixed: mixed dementia; SIVD: subcortical ischemic vascular dementia; MI: myocardial infarction."
    )
  )

# Display the table
tbl1_gt


library(gt)
gtsave(tbl1_gt, "table1.rtf")

#or 

# library(gt)
# 
# # big canvas = crisp when scaled down (≈ 2000–3000 px wide)
# gtsave(
#   data = tbl1_gt,
#   filename = "table1.png",
#   vwidth   = 2400,   # ~8 in at 300 dpi equivalent
#   vheight  = 1600,   # adjust to your table’s height
#   zoom     = 2       # increases pixel density
# )
# 
# # Optional: TIFF (journal-preferred). Convert PNG → TIFF.
# library(magick)
# image_read("table1.png") |>
#   image_write("table1.tiff", format = "tiff")

```




# ╔══════════════════════════════════════════════════════════════════════════════╗
# ║           COHORT CHARACTERIZATION TABLE                                     ║
# ╚══════════════════════════════════════════════════════════════════════════════╝
#
# PURPOSE:
#   Identify which research cohort each participant belongs to and report
#   how many participants per cohort had each key BBB permeability measure
#   available. This is reported in the manuscript Methods section to describe
#   the multi-cohort nature of the dataset.
#
# BACKGROUND — THREE COHORTS IN THIS DATASET:
#   1. MarkVCID  — Multi-center vascular contributions to cognitive impairment
#                  and dementia study. Identified by: markvcid_id
#   2. VCI       — Vascular Cognitive Impairment cohort (UNM local).
#                  Identified by: ptid_vci
#   3. ADRC      — Alzheimer's Disease Research Center cohort (UNM).
#                  Identified by: ptid_adrc (when markvcid_id is NA)
#
# PARTICIPANT ID PREFIXES IN dat_sub:
#   UP-xxx  → VCI cohort
#   VCI-xxx → VCI cohort
#   VDP-xxx → MarkVCID cohort
#
# KEY BIOMARKER AVAILABILITY BEING CHECKED:
#   fl_albumin_index_log2  — Requires lumbar puncture (CSF collection)
#   mri_perm_3t_tot        — Requires contrast-enhanced 3T MRI scan
#   Not all participants had both procedures, so availability varies by cohort
#
# STEPS IN THIS CODE:
#   1. Inspect identifier columns to understand cohort assignment logic
#   2. Assign each participant to exactly one cohort using case_when priority:
#        MarkVCID > VCI > VCID2 > ADRC > Other
#   3. Merge cohort assignment into dat_sub (the analysis dataset)
#   4. Count participants and biomarker availability per cohort
#   5. Display as a formatted GT table for the manuscript
#
# OUTPUT:
#   dat_sub_with_cohort  — dat_sub with cohort column added
#   summary_table        — formatted GT table of biomarker availability by cohort
#
# NOTE:
#   The hardcoded numbers in summary_table (MarkVCID: N=32, VCI: N=117) were
#   verified from the summarise() output above. If the dataset changes,
#   re-run the summarise() step and update these numbers accordingly.
# ─────────────────────────────────────────────────────────────────────────────

```{r}
# Look at the identifier columns to understand cohort structure
dat_all %>%
  select(ptid, ptid_vci, ptid_vcid2, ptid_adrc, markvcid_id) %>%
  head(20)

# Check which identifier columns are populated (non-NA) for different participants
dat_all %>%
  summarise(
    n_total = n(),
    n_with_adrc = sum(!is.na(ptid_adrc)),
    n_with_markvcid = sum(!is.na(markvcid_id)),
    n_with_vci = sum(!is.na(ptid_vci)),
    n_with_vcid2 = sum(!is.na(ptid_vcid2))
  )


# Create a cohort variable based on which specific IDs exist
dat_all_cohort <- dat_all %>%
  mutate(
    cohort = case_when(
      !is.na(markvcid_id) ~ "MarkVCID",
      !is.na(ptid_vci) ~ "VCI",
      !is.na(ptid_vcid2) ~ "VCID2",
      !is.na(ptid_adrc) & is.na(markvcid_id) ~ "ADRC",  # ADRC only if not also MarkVCID
      TRUE ~ "Other"
    )
  )

# Check the distribution
table(dat_all_cohort$cohort)



# Get unique ptid with their cohort from dat_all
ptid_cohort_map <- dat_all_cohort %>%
  select(ptid, cohort) %>%
  distinct() %>%
  # Keep one row per ptid (in case of duplicates, take first)
  group_by(ptid) %>%
  slice(1) %>%
  ungroup()

# Merge with your dat_sub
dat_sub_with_cohort <- dat_sub %>%
  left_join(ptid_cohort_map, by = "ptid")

# Check the results
table(dat_sub_with_cohort$cohort)


# Count how many have albumin index by cohort
dat_sub_with_cohort %>%
  group_by(cohort) %>%
  summarise(
    n_participants = n(),
    n_with_albumin_index = sum(!is.na(fl_albumin_index_log2)),
    n_with_mri_perm = sum(!is.na(mri_perm_3t_tot))

  )
# Your summary table
summary_table <- data.frame(
  cohort = c("MarkVCID", "VCI"),
  n_participants = c(32, 117),
  n_with_albumin_index = c(31, 97),
  n_with_mri_perm = c(23, 50)
)

library(gt)

summary_table %>%
   gt() %>%
  cols_label(
    cohort = "Cohort",
    n_participants = "N",
    n_with_albumin_index = "Albumin Index",
    n_with_mri_perm = "MRI Permeability"
  ) %>%
  tab_spanner(
    label = "Albumin Index"
  ) %>%
  tab_spanner(
    label = "MRI Permeability"
  ) %>%
  tab_header(
    title = "Biomarker Availability by Cohort"
  )

```


# This code is for checking the 3T and 3T log with Diagnosis plots 
```{r}
library(ggplot2)
library(dplyr)

dat_plot_3T_diagnosis <- dat_sub %>%
  filter(!is.na(dx_current), !is.na(mri_perm_3t_tot))


p_perm_dx <- ggplot(dat_plot_3T_diagnosis, aes(x = dx_current, y = mri_perm_3t_tot, fill = dx_current)) +
  geom_violin(trim = FALSE, alpha = 0.35) +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.12, alpha = 0.4, size = 1) +
  labs(x = "Diagnosis", y = "MRI BBB permeability (3T total)",
       title = "BBB permeability by diagnosis") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
p_perm_dx


dat_plot_3Tlog2_diagnosis <- dat_sub %>%
  filter(!is.na(dx_current), !is.na(mri_perm_3t_tot_log2))


p_perm_dx <- ggplot(dat_plot_3Tlog2_diagnosis, aes(x = dx_current, y = mri_perm_3t_tot_log2, fill = dx_current)) +
  geom_violin(trim = FALSE, alpha = 0.35) +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.12, alpha = 0.4, size = 1) +
  labs(x = "Diagnosis", y = "MRI BBB permeability (3Tlog2 total)",
       title = "BBB permeability by diagnosis") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
p_perm_dx

```


#Individual plots associated with significant covariates.
```{r}
dat_plot_sub_indy <-
  dat_plot |>
  dplyr::filter(
    term == "X"
  )

plot_results <- list()

# Interpretations output file
fn_output <-
  file.path(
        "All_figures_for_paper"
      , "interpretations.txt"
      )
output_write <-
  c(
    fn_output
  , "\n\n"
  )
readr::write_lines(x = output_write, file = fn_output, append = FALSE)


for (i_row in seq_len(nrow(dat_plot_sub_indy))) {
  ## i_row = 1

  var_resp  <- dat_plot_sub_indy$resp [i_row]
  var_covar <- dat_plot_sub_indy$covar[i_row]

  plot_results[[ i_row ]] <-
    results[[var_resp]][[var_covar]][["plot_lm_fit_red_BIC"]]
  
  fit_contrasts <-
    e_plot_model_contrasts(
      fit                = results[[var_resp]][[var_covar]]$lm_fit_red_BIC
    , dat_cont           = results[[var_resp]][[var_covar]]$dat_lm
    , sw_print           = FALSE
    , sw_table_in_plot   = FALSE
    , sw_TWI_plots_keep  = "singles"
    , sw_quantile_type   = 1
    , sw_produce_plots = c(TRUE, FALSE)[2]
    )
  #fit_contrasts$tables  # to print tables
  #fit_contrasts$plots   # to print plots
  #fit_contrasts$text    # to print caption text
  fit_contrasts$interp  # to print interpretation text
  
  output_write <-
    c(
      "\n\n"
    , "====="
    , "\nResponse:  "
    , var_resp
    , "Predictor:  "
    , var_covar
    , ""
    , "interpratations:\n"
    , paste(
        fit_contrasts$interp |> names()
      , fit_contrasts$interp |> unlist()
      , sep = "\n"
      , collapse = "\n\n"
      )
    )
    readr::write_lines(x = output_write, file = fn_output, append = TRUE)
} # i_row
    

p_arranged <-
  cowplot::plot_grid(
    plotlist           = plot_results
  , labels             = "AUTO"
  , label_size         = 10
  , label_fontface     = "bold"
  , label_x            = 0.05
  , label_y            = 1.03
  , nrow               = 5
  , ncol               = 10
  )
ggsave(
    paste0("All_figures_for_paper/plot_results_indy_AIC", ".png")
  , plot   = p_arranged
  , width  = 20
  , height = 16
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE
  )
# 
# ── Global panel size — same for every figure in the paper ───────────────────
PANEL_IN <- 4.5

# ── helper: save a plot_grid chunk ────────────────────────────────────────────
save_panel_chunk <- function(plot_list, labels, filename, n_col = 4) {

  # Add top margin so label sits above content, not over it
  plot_list <- lapply(plot_list, function(p) {
    p + ggplot2::theme(
      plot.margin = ggplot2::margin(t = 22, r = 5, b = 5, l = 5, unit = "pt")
    )
  })

  n_plots  <- length(plot_list)
  n_row    <- ceiling(n_plots / n_col)
  w        <- n_col * PANEL_IN
  h        <- n_row * PANEL_IN

  p <- cowplot::plot_grid(
      plotlist         = plot_list
    , labels           = labels
    , label_size       = 14
    , label_fontface   = "bold"
    , label_x          = 0
    , label_y          = 0.995
    , nrow             = n_row
    , ncol             = n_col
    , align            = "hv"
    , axis             = "tblr"
    )

  ggsave(
      filename
    , plot      = p
    , width     = w
    , height    = h
    , dpi       = 300
    , bg        = "white"
    , limitsize = FALSE
    )
  message("Saved: ", filename, " | ", n_plots, " panels | ", w, "x", h, " in")
}

# ── Filter to significant X terms only ───────────────────────────────────────
dat_plot_sub_indy <-
  dat_plot |>
  dplyr::filter(term == "X")

plot_results <- list()

# ── Interpretations output file ───────────────────────────────────────────────
fn_output <- file.path("All_figures_for_paper", "interpretations.txt")
readr::write_lines(x = c(fn_output, "\n\n"), file = fn_output, append = FALSE)

# ── Main loop: collect plots and write interpretations ────────────────────────
for (i_row in seq_len(nrow(dat_plot_sub_indy))) {
  var_resp  <- dat_plot_sub_indy$resp [i_row]
  var_covar <- dat_plot_sub_indy$covar[i_row]

  plot_results[[ i_row ]] <-
    results[[var_resp]][[var_covar]][["plot_lm_fit_red_BIC"]]

  fit_contrasts <-
    e_plot_model_contrasts(
      fit               = results[[var_resp]][[var_covar]]$lm_fit_red_BIC
    , dat_cont          = results[[var_resp]][[var_covar]]$dat_lm
    , sw_print          = FALSE
    , sw_table_in_plot  = FALSE
    , sw_TWI_plots_keep = "singles"
    , sw_quantile_type  = 1
    , sw_produce_plots  = c(TRUE, FALSE)[2]
    )

  output_write <-
    c(
      "\n\n"
    , "====="
    , "\nResponse:  "
    , var_resp
    , "Predictor:  "
    , var_covar
    , ""
    , "interpratations:\n"
    , paste(
        fit_contrasts$interp |> names()
      , fit_contrasts$interp |> unlist()
      , sep      = "\n"
      , collapse = "\n\n"
      )
    )
  readr::write_lines(x = output_write, file = fn_output, append = TRUE)
} # i_row

# ── Split plot_results by response variable ───────────────────────────────────
idx_albumin <- which(dat_plot_sub_indy$resp == "fl_albumin_index_log2")
idx_ktrans  <- which(dat_plot_sub_indy$resp == "mri_perm_3t_tot_log2")

plots_albumin <- plot_results[ idx_albumin ]
plots_ktrans  <- plot_results[ idx_ktrans  ]

n_albumin <- length(plots_albumin)
n_ktrans  <- length(plots_ktrans)

cat("Albumin panels:", n_albumin, "| Ktrans panels:", n_ktrans, "\n")

# ── Fig 3a: Albumin panels A1–A20 ────────────────────────────────────────────
split_at <- 20

save_panel_chunk(
    plot_list = plots_albumin[ 1:min(split_at, n_albumin) ]
  , labels    = paste0("A", 1:min(split_at, n_albumin))
  , filename  = "All_figures_for_paper/Fig3a_Albumin_Index.png"
  , n_col     = 4
  )

# ── Fig 3b: Albumin panels A21–end ───────────────────────────────────────────
if (n_albumin > split_at) {
  idx_rest <- (split_at + 1):n_albumin
  save_panel_chunk(
      plot_list = plots_albumin[ idx_rest ]
    , labels    = paste0("A", idx_rest)
    , filename  = "All_figures_for_paper/Fig3b_Albumin_Index.png"
    , n_col     = 4
    )
}

# ── Fig 4: Ktrans panels B1–end ──────────────────────────────────────────────
save_panel_chunk(
    plot_list = plots_ktrans
  , labels    = paste0("B", 1:n_ktrans)
  , filename  = "All_figures_for_paper/Fig4_Ktrans_3T.png"
  , n_col     = 4
  )

# cowplot::plot_grid(
#     plotlist         = p_list
#   , align            = c("none", "h", "v", "hv")[4]
#   , axis             = c("none", "l", "r", "t", "b", "lr", "tb", "tblr")[1]
#   , nrow             = 1
#   , ncol             = NULL
#   , rel_widths       = 1
#   , rel_heights      = 1
#   , labels           = "AUTO" # "auto", c("A", "B")
#   , label_size       = 14
#   , label_fontfamily = NULL
#   , label_fontface   = "bold"
#   , label_colour     = NULL
#   , label_x          = 0
#   , label_y          = 1
#   , hjust            = -0.5
#   , vjust            = 1.5
#   , scale            = 1
#   , greedy           = TRUE
#   , byrow            = TRUE
# )

```



## Drop extreme outliers

```{r chunk-012, fig.height = 4, fig.width = 10}
var_list_plot <-
  c(
    Var$csf
  , Var$plasma
  , Var$mri
  )
  dat_var_label_matching$Var

par(mfrow=c(1,3))
for (i_plot in 1:length(var_list_plot)) {
  ## i_plot = 1
  #if (i_plot > 3) {next}

  if( all(is.na(dat_sub[[ var_list_plot[i_plot] ]])) ) {
    next
  }

  plot(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , type  = "p"
  , main  = paste0("Before: ", var_list_plot[i_plot])
  , xlab  = ""
  , ylab  = ""
  #, col   = dat_sub$dx_current
  , pch   = 20  #3
  , xlim  = c(1, nrow(dat_sub) * 1.2)
  )
  #text(
  #  x     = 1:nrow(dat_sub)
  #, y     = dat_sub[[ var_list_plot[i_plot] ]]
  #, labels = 1:nrow(dat_sub)
  #, pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
  #)
}
par(mfrow=c(1,1))
```


Drop extreme outlying values.
```{r}
# Drop extreme outlying values
## #dat_sub$fl_albumin_index        [dat_sub$fl_albumin_index         > 20     ]  <- NA   # obs 14
## #dat_sub$fl_ptau                 [dat_sub$fl_ptau                  > 200    ]  <- NA   # obs 140, 275
## dat_sub$fl_msd_csf_mmp_1        [dat_sub$fl_msd_csf_mmp_1         > 80     ]  <- NA   # obs 113, 162
## #dat_sub$fl_msd_csf_mmp_2        [dat_sub$fl_msd_csf_mmp_2         < 5000   ]  <- NA   # obs 260
## dat_sub$fl_msd_csf_mmp_9        [dat_sub$fl_msd_csf_mmp_9         > 1000   ]  <- NA   # obs 162, 181
## dat_sub$fl_msd_csf_mmp_10       [dat_sub$fl_msd_csf_mmp_10        > 250    ]  <- NA   # obs 110
## dat_sub$fl_angio_csf_bfgf       [dat_sub$fl_angio_csf_bfgf        > 4      ]  <- NA   # obs 138, 152
## dat_sub$fl_angio_csf_vegf_a     [dat_sub$fl_angio_csf_vegf_a      > 15     ]  <- NA   # obs 100
## dat_sub$fl_cyto_csf_ifn_g       [dat_sub$fl_cyto_csf_ifn_g        > 1      ]  <- NA   # obs 7
## dat_sub$fl_cyto_csf_il_10       [dat_sub$fl_cyto_csf_il_10        > 0.2    ]  <- NA   # obs 119, 162
## dat_sub$fl_cyto_csf_il_13       [dat_sub$fl_cyto_csf_il_13        > 1.5    ]  <- NA   # obs 85, 92
## dat_sub$fl_cyto_csf_il_1b       [dat_sub$fl_cyto_csf_il_1b        > 1      ]  <- NA   # obs 100
## dat_sub$fl_cyto_csf_il_2        [dat_sub$fl_cyto_csf_il_2         > 0.3    ]  <- NA   # obs 163, 278
## dat_sub$fl_cyto_csf_il_4        [dat_sub$fl_cyto_csf_il_4         > 0.015  ]  <- NA   # obs 268
## dat_sub$fl_cyto_csf_il_6        [dat_sub$fl_cyto_csf_il_6         > 4      ]  <- NA   # obs 268
## dat_sub$fl_msd_plasma_mmp_2     [dat_sub$fl_msd_plasma_mmp_2      > 250000 ]  <- NA   # obs 63, 66, 112
## dat_sub$fl_msd_plasma_mmp_3     [dat_sub$fl_msd_plasma_mmp_3      > 100000 ]  <- NA   # obs 44
## dat_sub$fl_msd_plasma_mmp_9     [dat_sub$fl_msd_plasma_mmp_9      > 500000 ]  <- NA   # obs 168, 171
## dat_sub$fl_angio_plasma_bfgf    [dat_sub$fl_angio_plasma_bfgf     > 300    ]  <- NA   # obs 122
## dat_sub$fl_angio_plasma_flt_1   [dat_sub$fl_angio_plasma_flt_1    > 140    ]  <- NA   # obs 83, 175
## dat_sub$fl_angio_plasma_plgf    [dat_sub$fl_angio_plasma_plgf     > 50     ]  <- NA   # obs 238
## dat_sub$fl_angio_plasma_vegf_a  [dat_sub$fl_angio_plasma_vegf_a   > 1000   ]  <- NA   # obs 12, 168, 171
## dat_sub$fl_angio_plasma_vegf_c  [dat_sub$fl_angio_plasma_vegf_c   > 1000   ]  <- NA   # obs 142
## dat_sub$fl_cyto_plasma_ifn_g    [dat_sub$fl_cyto_plasma_ifn_g     > 80     ]  <- NA   # obs 124
## dat_sub$fl_cyto_plasma_il_10    [dat_sub$fl_cyto_plasma_il_10     > 2.5    ]  <- NA   # obs 33, 177
## dat_sub$fl_cyto_plasma_il_8     [dat_sub$fl_cyto_plasma_il_8      > 40     ]  <- NA   # obs 17, 168, 171
## dat_sub$fl_cyto_plasma_il_6     [dat_sub$fl_cyto_plasma_il_6      > 8      ]  <- NA   # obs 43, 44, 168, 171, 175
## dat_sub$fl_cyto_plasma_il_4     [dat_sub$fl_cyto_plasma_il_4      > 0.25   ]  <- NA   # obs 171, 239 +1more
## dat_sub$fl_cyto_plasma_il_2     [dat_sub$fl_cyto_plasma_il_2      > 1.5    ]  <- NA   # obs +1more
## dat_sub$fl_cyto_plasma_il_1b    [dat_sub$fl_cyto_plasma_il_1b     > 1.5    ]  <- NA   # obs 5, 168, 171
## dat_sub$fl_cyto_plasma_il_12p70 [dat_sub$fl_cyto_plasma_il_12p70  > 1.25   ]  <- NA   # obs 171
##
## # dat_sub$mri_vol_gm                [dat_sub$mri_vol_gm               <  0.22] <- NA   # 1
## # dat_sub$mri_dti_fabnormal_z       [dat_sub$mri_dti_fabnormal_z      < -1.66] <- NA   # 1
## # dat_sub$mri_dti_histpk_z          [dat_sub$mri_dti_histpk_z         < -6.35] <- NA   # 2
## # dat_sub$mri_vol_hippocampal       [dat_sub$mri_vol_hippocampal      <  4000] <- NA   # 1
## # dat_sub$fl_msd_csf_mmp_2          [dat_sub$fl_msd_csf_mmp_2         <  1000] <- NA   # 1
## #dat_sub$mri_perm_combined_tot       [dat_sub$mri_perm_combined_tot    >    35] <- NA   # 1


## # Drop extreme outlying values
## dat_sub[[ "fl_csf_glucose"                     ]] [dat_sub[[ "fl_csf_glucose"                     ]] <      3       ]  <- NA   # obs
## dat_sub[[ "fl_csf_total_protein"               ]] [dat_sub[[ "fl_csf_total_protein"               ]] >      7.5     ]  <- NA   # obs
## dat_sub[[ "fl_csf_rbc"                         ]] [dat_sub[[ "fl_csf_rbc"                         ]] >      8       ]  <- NA   # obs
## #dat_sub[[ "fl_csf_tnc"                         ]] [dat_sub[[ "fl_csf_tnc"                         ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_csf_igg"                         ]] [dat_sub[[ "fl_csf_igg"                         ]] <      0       ]  <- NA   # obs
## #dat_sub[[ "fl_csf_albumin"                     ]] [dat_sub[[ "fl_csf_albumin"                     ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf_igg_synthesis"               ]] [dat_sub[[ "fl_csf_igg_synthesis"               ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_igg_index"                       ]] [dat_sub[[ "fl_igg_index"                       ]] <     -2       ]  <- NA   # obs
## #dat_sub[[ "fl_albumin_index"                   ]] [dat_sub[[ "fl_albumin_index"                   ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_oligoclonal_bands"               ]] [dat_sub[[ "fl_oligoclonal_bands"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_myelin_basic_protein"            ]] [dat_sub[[ "fl_myelin_basic_protein"            ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_serum_igg"                       ]] [dat_sub[[ "fl_serum_igg"                       ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_serum_albumin"                   ]] [dat_sub[[ "fl_serum_albumin"                   ]] <     10       ]  <- NA   # obs
## #dat_sub[[ "fl_zym_plasma_mmp_2_72_kda"         ]] [dat_sub[[ "fl_zym_plasma_mmp_2_72_kda"         ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_zym_plasma_mmp_2_62_kda"         ]] [dat_sub[[ "fl_zym_plasma_mmp_2_62_kda"         ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_zym_mmp_2_index"                 ]] [dat_sub[[ "fl_zym_mmp_2_index"                 ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_zym_plasma_mmp_2_92_kda"         ]] [dat_sub[[ "fl_zym_plasma_mmp_2_92_kda"         ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_zym_plasma_mmp_2_84_kda"         ]] [dat_sub[[ "fl_zym_plasma_mmp_2_84_kda"         ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_msd_plasma_mmp_1"                ]] [dat_sub[[ "fl_msd_plasma_mmp_1"                ]] <      8       ]  <- NA   # obs
## #dat_sub[[ "fl_msd_plasma_mmp_2"                ]] [dat_sub[[ "fl_msd_plasma_mmp_2"                ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_msd_plasma_mmp_3"                ]] [dat_sub[[ "fl_msd_plasma_mmp_3"                ]] <      8       ]  <- NA   # obs
## dat_sub[[ "fl_msd_plasma_mmp_9"                ]] [dat_sub[[ "fl_msd_plasma_mmp_9"                ]] <     12       ]  <- NA   # obs
## #dat_sub[[ "fl_msd_plasma_mmp_10"               ]] [dat_sub[[ "fl_msd_plasma_mmp_10"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_msd_csf_mmp_1"                   ]] [dat_sub[[ "fl_msd_csf_mmp_1"                   ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_msd_csf_mmp_2"                   ]] [dat_sub[[ "fl_msd_csf_mmp_2"                   ]] <     12       ]  <- NA   # obs
## #dat_sub[[ "fl_msd_csf_mmp_3"                   ]] [dat_sub[[ "fl_msd_csf_mmp_3"                   ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_msd_csf_mmp_9"                   ]] [dat_sub[[ "fl_msd_csf_mmp_9"                   ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_msd_csf_mmp_10"                  ]] [dat_sub[[ "fl_msd_csf_mmp_10"                  ]] <      4       ]  <- NA   # obs
## #dat_sub[[ "fl_msd_mmp_2_index"                 ]] [dat_sub[[ "fl_msd_mmp_2_index"                 ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_angio_plasma_bfgf"               ]] [dat_sub[[ "fl_angio_plasma_bfgf"               ]] >      8       ]  <- NA   # obs
## #dat_sub[[ "fl_angio_plasma_flt_1"              ]] [dat_sub[[ "fl_angio_plasma_flt_1"              ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_angio_plasma_tie2"               ]] [dat_sub[[ "fl_angio_plasma_tie2"               ]] <      6       ]  <- NA   # obs
## #dat_sub[[ "fl_angio_plasma_vegf_a"             ]] [dat_sub[[ "fl_angio_plasma_vegf_a"             ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_angio_plasma_vegf_c"             ]] [dat_sub[[ "fl_angio_plasma_vegf_c"             ]] >      8       ]  <- NA   # obs
## dat_sub[[ "fl_angio_plasma_vegf_d"             ]] [dat_sub[[ "fl_angio_plasma_vegf_d"             ]] <      6       ]  <- NA   # obs
## #dat_sub[[ "fl_angio_csf_bfgf"                  ]] [dat_sub[[ "fl_angio_csf_bfgf"                  ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_angio_csf_flt_1"                 ]] [dat_sub[[ "fl_angio_csf_flt_1"                 ]] <      2       ]  <- NA   # obs
## #dat_sub[[ "fl_angio_csf_tie2"                  ]] [dat_sub[[ "fl_angio_csf_tie2"                  ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_angio_csf_vegf_a"                ]] [dat_sub[[ "fl_angio_csf_vegf_a"                ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_angio_csf_vegf_c"                ]] [dat_sub[[ "fl_angio_csf_vegf_c"                ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_angio_csf_vegf_d"                ]] [dat_sub[[ "fl_angio_csf_vegf_d"                ]] <      2       ]  <- NA   # obs
## #dat_sub[[ "fl_abeta_38"                        ]] [dat_sub[[ "fl_abeta_38"                        ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_abeta_4240_ratio"                ]] [dat_sub[[ "fl_abeta_4240_ratio"                ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_ifn_g"                  ]] [dat_sub[[ "fl_cyto_csf_ifn_g"                  ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_il_10"                  ]] [dat_sub[[ "fl_cyto_csf_il_10"                  ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_il_12p70"               ]] [dat_sub[[ "fl_cyto_csf_il_12p70"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_il_13"                  ]] [dat_sub[[ "fl_cyto_csf_il_13"                  ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_cyto_csf_il_1b"                  ]] [dat_sub[[ "fl_cyto_csf_il_1b"                  ]] >      0       ]  <- NA   # obs
## dat_sub[[ "fl_cyto_csf_il_2"                   ]] [dat_sub[[ "fl_cyto_csf_il_2"                   ]] <     -8       ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_il_4"                   ]] [dat_sub[[ "fl_cyto_csf_il_4"                   ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_cyto_csf_il_6"                   ]] [dat_sub[[ "fl_cyto_csf_il_6"                   ]] >      2       ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_csf_il_8"                   ]] [dat_sub[[ "fl_cyto_csf_il_8"                   ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_cyto_csf_tnf_a"                  ]] [dat_sub[[ "fl_cyto_csf_tnf_a"                  ]] <     -5       ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_ifn_g"               ]] [dat_sub[[ "fl_cyto_plasma_ifn_g"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_10"               ]] [dat_sub[[ "fl_cyto_plasma_il_10"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_12p70"            ]] [dat_sub[[ "fl_cyto_plasma_il_12p70"            ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_13"               ]] [dat_sub[[ "fl_cyto_plasma_il_13"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_2"                ]] [dat_sub[[ "fl_cyto_plasma_il_2"                ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_4"                ]] [dat_sub[[ "fl_cyto_plasma_il_4"                ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_cyto_plasma_il_6"                ]] [dat_sub[[ "fl_cyto_plasma_il_6"                ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_cyto_plasma_il_8"                ]] [dat_sub[[ "fl_cyto_plasma_il_8"                ]] >      5       ]  <- NA   # obs
## dat_sub[[ "fl_cyto_plasma_tnf_a"               ]] [dat_sub[[ "fl_cyto_plasma_tnf_a"               ]] <     -1       ]  <- NA   # obs
## dat_sub[[ "fl_simoa_csf_il_1b"                 ]] [dat_sub[[ "fl_simoa_csf_il_1b"                 ]] >     -4.5     ]  <- NA   # obs
## #dat_sub[[ "fl_simoa_edta_plasma_ptau_v2_1"     ]] [dat_sub[[ "fl_simoa_edta_plasma_ptau_v2_1"     ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_simoa_csf_ptau_v2_1"             ]] [dat_sub[[ "fl_simoa_csf_ptau_v2_1"             ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_apoe_ncrad"                      ]] [dat_sub[[ "fl_apoe_ncrad"                      ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__gfap"                  ]] [dat_sub[[ "fl_csf__ALL__gfap"                  ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__gfap"               ]] [dat_sub[[ "fl_plasma__ALL__gfap"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__nfl"                   ]] [dat_sub[[ "fl_csf__ALL__nfl"                   ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__nfl"                ]] [dat_sub[[ "fl_plasma__ALL__nfl"                ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__ptau"                  ]] [dat_sub[[ "fl_csf__ALL__ptau"                  ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__ptau"               ]] [dat_sub[[ "fl_plasma__ALL__ptau"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__abeta_40"              ]] [dat_sub[[ "fl_csf__ALL__abeta_40"              ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__abeta_40"           ]] [dat_sub[[ "fl_plasma__ALL__abeta_40"           ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__abeta_42"              ]] [dat_sub[[ "fl_csf__ALL__abeta_42"              ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__abeta_42"           ]] [dat_sub[[ "fl_plasma__ALL__abeta_42"           ]] > 000000.0000  ]  <- NA   # obs
## dat_sub[[ "fl_csf__ALL__plgf"                  ]] [dat_sub[[ "fl_csf__ALL__plgf"                  ]] <      0       ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__plgf"               ]] [dat_sub[[ "fl_plasma__ALL__plgf"               ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_plasma__ALL__il_1b"              ]] [dat_sub[[ "fl_plasma__ALL__il_1b"              ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__mmp_2"                 ]] [dat_sub[[ "fl_csf__ALL__mmp_2"                 ]] > 000000.0000  ]  <- NA   # obs
## #dat_sub[[ "fl_csf__ALL__mmp_2_index"           ]] [dat_sub[[ "fl_csf__ALL__mmp_2_index"           ]] > 000000.0000  ]  <- NA   # obs

# 1/10/2025
### CSF
#dat_sub[[ "fl_albumin_index_log2"                 ]] [ dat_sub[[ "fl_albumin_index_log2"                 ]] > 0 ] <- NA
#dat_sub[[ "fl_igg_index_log2"                     ]] [ dat_sub[[ "fl_igg_index_log2"                     ]] > 0 ] <- NA
#dat_sub[[ "fl_csf_igg_log2"                       ]] [ dat_sub[[ "fl_csf_igg_log2"                       ]] < 0.2 ] <- NA
#dat_sub[[ "fl_msd_csf_mmp_1_log2"                 ]] [ dat_sub[[ "fl_msd_csf_mmp_1_log2"                 ]] > 0 ] <- NA
#dat_sub[[ "fl_msd_csf_mmp_2_log2"                 ]] [ dat_sub[[ "fl_msd_csf_mmp_2_log2"                 ]] > 0 ] <- NA
#dat_sub[[ "fl_msd_csf_mmp_3_log2"                 ]] [ dat_sub[[ "fl_msd_csf_mmp_3_log2"                 ]] > 0 ] <- NA
dat_sub[[ "fl_msd_csf_mmp_9_log2"                 ]] [ dat_sub[[ "fl_msd_csf_mmp_9_log2"                 ]] < 1 ] <- NA
#dat_sub[[ "fl_csf__ALL__mmp_10_log2"              ]] [ dat_sub[[ "fl_csf__ALL__mmp_10_log2"              ]] > 0 ] <- NA
#dat_sub[[ "fl_angio_csf_bfgf_log2"                ]] [ dat_sub[[ "fl_angio_csf_bfgf_log2"                ]] > 3 ] <- NA
dat_sub[[ "fl_angio_csf_flt_1_log2"               ]] [ dat_sub[[ "fl_angio_csf_flt_1_log2"               ]] < 1 ] <- NA
#dat_sub[[ "fl_angio_csf_tie2_log2"                ]] [ dat_sub[[ "fl_angio_csf_tie2_log2"                ]] > 0 ] <- NA
dat_sub[[ "fl_angio_csf_vegf_a_log2"              ]] [ dat_sub[[ "fl_angio_csf_vegf_a_log2"              ]] < 0.1 ] <- NA
dat_sub[[ "fl_angio_csf_vegf_c_log2"              ]] [ dat_sub[[ "fl_angio_csf_vegf_c_log2"              ]] < 0.1 ] <- NA
dat_sub[[ "fl_angio_csf_vegf_d_log2"              ]] [ dat_sub[[ "fl_angio_csf_vegf_d_log2"              ]] < 1.5 ] <- NA
#dat_sub[[ "fl_cyto_csf_ifn_g_log2"                ]] [ dat_sub[[ "fl_cyto_csf_ifn_g_log2"                ]] > 2.5 ] <- NA
dat_sub[[ "fl_csf__ALL__il_1b_log2"               ]] [ dat_sub[[ "fl_csf__ALL__il_1b_log2"               ]] > 1.2 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_2_log2"                 ]] [ dat_sub[[ "fl_cyto_csf_il_2_log2"                 ]] > 0.65 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_4_log2"                 ]] [ dat_sub[[ "fl_cyto_csf_il_4_log2"                 ]] > 0.025 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_6_log2"                 ]] [ dat_sub[[ "fl_cyto_csf_il_6_log2"                 ]] > 3 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_8_log2"                 ]] [ dat_sub[[ "fl_cyto_csf_il_8_log2"                 ]] > 9 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_10_log2"                ]] [ dat_sub[[ "fl_cyto_csf_il_10_log2"                ]] > 0.55 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_12p70_log2"             ]] [ dat_sub[[ "fl_cyto_csf_il_12p70_log2"             ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_csf_il_13_log2"                ]] [ dat_sub[[ "fl_cyto_csf_il_13_log2"                ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_csf_tnf_a_log2"                ]] [ dat_sub[[ "fl_cyto_csf_tnf_a_log2"                ]] > 0.9 ] <- NA
#dat_sub[[ "fl_csf__ALL__nfl_log2"                 ]] [ dat_sub[[ "fl_csf__ALL__nfl_log2"                 ]] > 0 ] <- NA
dat_sub[[ "fl_csf__ALL__plgf_log2"                ]] [ dat_sub[[ "fl_csf__ALL__plgf_log2"                ]] < 0.5 ] <- NA
#dat_sub[[ "fl_csf__ALL__gfap_log2"                ]] [ dat_sub[[ "fl_csf__ALL__gfap_log2"                ]] > 0 ] <- NA
#dat_sub[[ "fl_abeta_4240_ratio_log2"              ]] [ dat_sub[[ "fl_abeta_4240_ratio_log2"              ]] > 0 ] <- NA
#dat_sub[[ "fl_csf__ALL__ptau_log2"                ]] [ dat_sub[[ "fl_csf__ALL__ptau_log2"                ]] > 0 ] <- NA
#dat_sub[[ "fl_myelin_basic_protein_log2"          ]] [ dat_sub[[ "fl_myelin_basic_protein_log2"          ]] > 0 ] <- NA
### Plasma
#dat_sub[[ "fl_serum_igg_log2"                     ]] [ dat_sub[[ "fl_serum_igg_log2"                     ]] < 8.25 ] <- NA
#dat_sub[[ "fl_plasma__ALL__mmp_1_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__mmp_1_log2"            ]] > 15.5 ] <- NA
#dat_sub[[ "fl_plasma__ALL__mmp_2_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__mmp_2_log2"            ]] > 21 ] <- NA
#dat_sub[[ "fl_plasma__ALL__mmp_3_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__mmp_3_log2"            ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__mmp_9_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__mmp_9_log2"            ]] > 20.5 ] <- NA
#dat_sub[[ "fl_plasma__ALL__mmp_10_log2"           ]] [ dat_sub[[ "fl_plasma__ALL__mmp_10_log2"           ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__bfgf_log2"             ]] [ dat_sub[[ "fl_plasma__ALL__bfgf_log2"             ]] > 10.5 ] <- NA
#dat_sub[[ "fl_plasma__ALL__flt_1_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__flt_1_log2"            ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__tie2_log2"             ]] [ dat_sub[[ "fl_plasma__ALL__tie2_log2"             ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__vegf_a_log2"           ]] [ dat_sub[[ "fl_plasma__ALL__vegf_a_log2"           ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__vegf_c_log2"           ]] [ dat_sub[[ "fl_plasma__ALL__vegf_c_log2"           ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__vegf_d_log2"           ]] [ dat_sub[[ "fl_plasma__ALL__vegf_d_log2"           ]] < 5 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_ifn_g_log2"     ]] [ dat_sub[[ "fl_cyto_heparin_plasma_ifn_g_log2"     ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__il_1b_log2"            ]] [ dat_sub[[ "fl_plasma__ALL__il_1b_log2"            ]] > 1.25 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_2_log2"      ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_2_log2"      ]] > 0 ] <- NA
dat_sub[[ "fl_cyto_heparin_plasma_il_4_log2"      ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_4_log2"      ]] > 1 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_6_log2"      ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_6_log2"      ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_8_log2"      ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_8_log2"      ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_10_log2"     ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_10_log2"     ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_12p70_log2"  ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_12p70_log2"  ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_il_13_log2"     ]] [ dat_sub[[ "fl_cyto_heparin_plasma_il_13_log2"     ]] > 0 ] <- NA
#dat_sub[[ "fl_cyto_heparin_plasma_tnf_a_log2"     ]] [ dat_sub[[ "fl_cyto_heparin_plasma_tnf_a_log2"     ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__nfl_log2"              ]] [ dat_sub[[ "fl_plasma__ALL__nfl_log2"              ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__plgf_log2"             ]] [ dat_sub[[ "fl_plasma__ALL__plgf_log2"             ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__gfap_log2"             ]] [ dat_sub[[ "fl_plasma__ALL__gfap_log2"             ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__abeta_40_log2"         ]] [ dat_sub[[ "fl_plasma__ALL__abeta_40_log2"         ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__abeta_42_log2"         ]] [ dat_sub[[ "fl_plasma__ALL__abeta_42_log2"         ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__ptau_181_log2"         ]] [ dat_sub[[ "fl_plasma__ALL__ptau_181_log2"         ]] > 0 ] <- NA
#dat_sub[[ "fl_plasma__ALL__ptau_217_log2"         ]] [ dat_sub[[ "fl_plasma__ALL__ptau_217_log2"         ]] > 0 ] <- NA
### MRI
#dat_sub[[ "mri_percent_vol_hippo"                 ]] [ dat_sub[[ "mri_percent_vol_hippo"                 ]] > 0 ] <- NA
#dat_sub[[ "mri_ucd_wmhv_percent"                  ]] [ dat_sub[[ "mri_ucd_wmhv_percent"                  ]] > 25 ] <- NA
#dat_sub[[ "mri_mean_cort_thick_mm"                ]] [ dat_sub[[ "mri_mean_cort_thick_mm"                ]] > 0 ] <- NA
#dat_sub[[ "mri_perm_3t_tot"                       ]] [ dat_sub[[ "mri_perm_3t_tot"                       ]] > 0 ] <- NA
#dat_sub[[ "mri_dti_psmd"                          ]] [ dat_sub[[ "mri_dti_psmd"                          ]] > 0.00175 ] <- NA
#dat_sub[[ "mri_dti_freewater_mean"                ]] [ dat_sub[[ "mri_dti_freewater_mean"                ]] > 0 ] <- NA

```


Plot again


```{r chunk-014, fig.height = 4, fig.width = 10}

par(mfrow=c(1,3))
for (i_plot in 1:length(var_list_plot)) {
  ## i_plot = 1
  #if (i_plot > 3) {next}

  if( all(is.na(dat_sub[[ var_list_plot[i_plot] ]])) ) {
    next
  }

  plot(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , type  = "p"
  , main  = paste0("After: ", var_list_plot[i_plot])
  , xlab  = ""
  , ylab  = ""
  #, col   = dat_sub$dx_current
  , pch   = 20  #3
  , xlim  = c(1, nrow(dat_sub) * 1.2)
  )
  #text(
  #  x     = 1:nrow(dat_sub)
  #, y     = dat_sub[[ var_list_plot[i_plot] ]]
  #, labels = 1:nrow(dat_sub)
  #, pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
  #)
}
par(mfrow=c(1,1))
```





# Just the plots that Gary wants for presentation

```{r}

## for (n_group in sort(unique(dat_var_label_matching$Type)) ) {
##   ## n_group = sort(unique(dat_var_label_matching$Type))[1]
##
##   print(paste0("=== n_group: ", n_group))
##
##   # skip some groups
##   if (n_group %in% c("ID", "Demographics")) {
##     next
##   }
##
##   var_temp <-
##     dat_var_label_matching |>
##     dplyr::filter(Type == n_group) |>
##     dplyr::pull(Var)
##   # label_temp <-
##   #   dat_var_label_matching |>
##   #   dplyr::filter(Type == n_group) |>
##   #   dplyr::pull(Label)

#dir.create("out_new/plot_indy_alb")
dir.create("All_figures_for_paper/combined")
var_temp <-
  c(
    Var$demo
  , Var$csf
  , Var$plasma
  , Var$mri
  )

p_list_indy <- list()
for (i_plot in seq_along(var_temp)) {
  ## i_plot = 1

  print(paste0("=== i_plot: ", i_plot, " / ", length(var_temp), ", ", var_temp[i_plot]))


  p_list_indy[[ i_plot ]] <-
    e_plot_scatterplot(
      dat_plot                  = dat_sub
    , var_x                     = var_temp[i_plot]
    , var_y                     = "fl_albumin_index_log2"
    #, var_color                 = "dx_current"
    , text_title                = labelled::var_label(dat_sub[[ var_temp[i_plot] ]] )
    , smooth_all                = c("none", "loess", "lm", "glm", "gam")[3]
    , sw_smooth_all_se          = c(TRUE, FALSE)[1]
    , smooth_by_var_color       = c("none", "loess", "lm", "glm", "gam")[1]
    , sw_smooth_by_var_color_se = c(TRUE, FALSE)[2]
    , sw_corr_in_subtitle       = c(TRUE, FALSE)[1]
    )
  # p_list_indy[[ i_plot ]] <-
  #   p_list_indy[[ i_plot ]] +
  #   #theme(legend.position = "bottom") +
  #   labs(
  #     y = "CSF Albumin Index, log2"  #paste0("log2( ", "Albumin index", " )")
  #   , x = label_temp[i_plot]  #paste0("log2( ", label_temp[i_plot], " )")
  #   )


  # individual plots
  ggsave(
      #paste0("out/plot_scatterplot_albumin_", n_group, ".png")
      paste0("out_new/plot_indy_alb/plot_scatterplot_albumin_", var_temp[i_plot], ".png")
    , plot   = p_list_indy[[ i_plot ]]
    , width  = 6
    , height = 6
    ## png, jpeg
    , dpi    = 300
    , bg     = "white"
    ## pdf
    #, units  = "in"
    #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
    )

} # i_plot

##   # p_arranged <-
##   #   cowplot::plot_grid(
##   #     plotlist = p_list
##   #   , nrow = 1
##   #   , ncol = NULL
##   #   , labels="AUTO"
##   #   )
##   #
##   # p_arranged |> print()
##
##   p_arranged <-
##     patchwork::wrap_plots(
##       p_list
##     , nrow = 3
##     , ncol = 4
##     , tag_level = "new"
##     , guides = "collect"
##     ) +
##     #patchwork::guide_area() +
##     patchwork::plot_annotation(
##     #  title = n_group
##       title = NULL
##     , subtitle = NULL
##     , caption =
##         paste0(
##         #  "Black line is the overall least squares regression line.  Colored lines are by diagnosis group.  Dashed ellipse is a 95% confidence ellipse using the t-distribution."
##           "Black line is the overall least squares regression line.  Dashed ellipse is a 95% confidence ellipse using the t-distribution."
##         )
##     , tag_levels = NULL
##     , tag_prefix = NULL
##     , tag_suffix = NULL
##     , tag_sep = NULL
##     , theme = NULL
##     )
##     #+
##     #theme(plot.caption = element_text(hjust = 0)) # Default is hjust=1, Caption align left
##   #p_arranged |> print()
##
##   ggsave(
##       #paste0("out/plot_scatterplot_albumin_", n_group, ".png")
##       paste0("out/plot_scatterplot_albumin_", ".png")
##     , plot   = p_arranged
##     , width  = 12
##     , height = 12
##     ## png, jpeg
##     , dpi    = 300
##     , bg     = "white"
##     ## pdf
##     #, units  = "in"
##     #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
##     )
##
## } # n_group

```



## Original code for Laura's paper starts here

```{r chunk-01}
#library(erikmisc)
#library(tidyverse)
#library(corrplot)
#library(broom)
#library(dunn.test)
#library(ggpubr)
## devtools::install_github("vincentarelbundock/modelsummary")
#library(modelsummary)
#
#dir.create("out_new")
#
#dat_sheet <- read.csv("PermeabilityAnalysis.Final.csv")
#
#dat_var_label_matching <-
#  tibble::tribble(
#    ~Var                      , ~Label                          , ~Type   # plot facet type labels
#  , "demo_study_id"           , "Study ID"                      , "ID"
#  , "dx_current"              , "Current Dx"                    , "Demographics"
#  , "demo_sex"                , "Gender"                        , "Demographics"
#  , "demo_yrs_edu"            , "Years of Education"            , "Demographics"
#  , "demo_bl_age"             , "Age"                           , "Demographics"
#  #, "fl_abeta_4240_ratio"     , "ABeta Ratio (42/40)"           , "AD" #"Alzheimers Biomarkers"
#  #, "fl_albumin_index"        , "Albumin Index"                 , "Permeability"
#  #, "fl_angio_csf_bfgf"       , "bFgF"                          , "Angiogenic Factors"
#  #, "fl_angio_csf_flt_1"      , "Flt-1"                         , "Angiogenic Factors"
#  #, "fl_cyto_csf_ifn_g"       , "IFN-g"                         , "Cytokines"
#  #, "fl_cyto_csf_il_1b"       , "IL-1b"                         , "Cytokines"
#  #, "fl_cyto_csf_il_2"        , "IL-2"                          , "Cytokines"
#  #, "fl_cyto_csf_il_4"        , "IL-4"                          , "Cytokines"
#  #, "fl_cyto_csf_il_6"        , "IL-6"                          , "Cytokines"
#  #, "fl_cyto_csf_il_8"        , "IL-8"                          , "Cytokines"
#  #, "fl_igg_index"            , "IgG Index"                     , "Other"
#  #, "fl_msd_csf_mmp_1"        , "MMP-1"                         , "MMP" #"Metalloproteinases"
#  #, "fl_msd_csf_mmp_2"        , "MMP-2"                         , "MMP" #"Metalloproteinases"
#  #, "fl_msd_csf_mmp_3"        , "MMP-3"                         , "MMP" #"Metalloproteinases"
#  #, "fl_msd_csf_mmp_9"        , "MMP-9"                         , "MMP" #"Metalloproteinases"
#  #, "fl_msd_csf_mmp_10"       , "MMP-10"                        , "MMP" #"Metalloproteinases"
#  #, "fl_myelin_basic_protein" , "Myelin Basic Protein"          , "Other"
#  #, "fl_angio_csf_plgf"       , "PlGF"                          , "Angiogenic Factors"
#  #, "fl_ptau"                 , "pTau"                          , "AD" #"Alzheimers Biomarkers"
#  #, "fl_cyto_csf_tnf_a"       , "TNF-a"                         , "Cytokines"
#  #, "fl_angio_csf_tie2"       , "Tie-2"                         , "Angiogenic Factors"
#  #, "fl_angio_csf_vegf_a"     , "VEGF-A"                        , "Angiogenic Factors"
#  #, "fl_angio_csf_vegf_c"     , "VEGF-C"                        , "Angiogenic Factors"
#  #, "fl_angio_csf_vegf_d"     , "VEGF-D"                        , "Angiogenic Factors"
#  , "mri_mean_cort_thick_mm"  , "Mean Cortical Thickness"       , "MRI"
#  , "mri_dti_freewater_mean"  , "Mean Free Water"               , "MRI"
#  , "percent_hippo"           , "Percent Hippocampal Volume"    , "MRI"
#  , "mri_dti_psmd"            , "PSMD"                          , "MRI"
#  , "mri_perm_3t_tot"         , "MRI WM Permeability"           , "Permeability"
#  , "percent_wmhv"            , "WMHV"                          , "MRI"
#  )
#
#
##separate columns by intended transformation
#totransform.log2 <-
#  c(
#    "fl_albumin_index"
#  , "mri_dti_freewater_mean"
#  , "mri_dti_psmd"
#  , "fl_myelin_basic_protein"
#  , "fl_igg_index"
#  , "fl_angio_csf_bfgf"
#  , "fl_angio_csf_flt_1"
#  , "fl_angio_csf_plgf"
#  , "fl_angio_csf_tie2"
#  , "fl_angio_csf_vegf_a"
#  , "fl_angio_csf_vegf_c"
#  , "fl_angio_csf_vegf_d"
#  , "fl_cyto_csf_il_1b"
#  , "fl_cyto_csf_il_2"
#  , "fl_cyto_csf_il_4"
#  , "fl_cyto_csf_il_6"
#  , "fl_cyto_csf_il_8"
#  , "fl_cyto_csf_ifn_g"
#  , "fl_cyto_csf_tnf_a"
#  , "fl_msd_csf_mmp_1"
#  , "fl_msd_csf_mmp_2"
#  , "fl_msd_csf_mmp_3"
#  , "fl_msd_csf_mmp_9"
#  , "fl_msd_csf_mmp_10"
#  , "fl_ptau"
#  , "mri_perm_3t_tot"
#  , "percent_wmhv"
#  )
#tokeep <-
#  c(
#    "demo_study_id"
#  , "demo_bl_age"
#  , "demo_sex"
#  , "demo_yrs_edu"
#  , "dx_current"
#  , "percent_hippo"
#  , "mri_mean_cort_thick_mm"
#  , "fl_abeta_4240_ratio"
#  )
#
#dx.tokeep <-
#  c("LA", "MX", "SIVD", "AD")
#
## format data
#dat_sheet <-
#  dat_sheet %>%
#  dplyr::select(
#    all_of(tokeep)
#  , all_of(totransform.log2)
#  ) %>%
#  dplyr::filter(
#    dx_current %in% dx.tokeep
#  ) %>%
#  dplyr::mutate(
#    dx_current    = dx_current
#  , demo_sex      = ifelse(demo_sex == "", NA, demo_sex) %>% factor()
#  , dx_current    = dx_current %>% factor()
#  , demo_study_id = demo_study_id %>% factor()
#  )
```



## Plots for outliers

```{r, fig.height = 4, fig.width = 10}
## var_list_plot <-
##   c(totransform.log2, tokeep)
##
## par(mfrow=c(1,3))
## for (i_plot in seq_along(var_list_plot)) {
##   ## i_plot = 1
##   #if (i_plot > 3) {next}
##
##   plot(
##     x     = 1:nrow(dat_sheet)
##   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
##   , type  = "p"
##   , main  = paste0("Before: ", var_list_plot[i_plot])
##   , xlab  = ""
##   , ylab  = ""
##   , col   = dat_sheet$dx_current
##   , pch   = 20  #3
##   , xlim  = c(1, nrow(dat_sheet) * 1.2)
##   )
##   text(
##     x     = 1:nrow(dat_sheet)
##   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
##   , labels = 1:nrow(dat_sheet)
##   , pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
##   )
## }
## par(mfrow=c(1,1))
```

Determine outliers by IQR.
```{r}
## IQR_dist_max <- 5
##
## tab_outliers <- list()
##
## for (i_var in seq_along(var_list_plot)) {
##   ## i_var = 1
##   ## i_var = 33
##   if (!is.numeric(dat_sheet[[ var_list_plot[i_var] ]])) {
##     next
##   }
##   this_var <- dat_sheet[[ var_list_plot[i_var] ]]
##   this_var_org <- this_var
##   if (var_list_plot[i_var] %in% totransform.log2) {
##     #this_var <- e_log_shift(this_var)
##     this_var <- e_log_shift(this_var, min_add = 1, base_log = 2)
##   }
##   this_med <- median(this_var, na.rm = TRUE)
##   this_iqr <- IQR(   this_var, na.rm = TRUE)
##
##   tab_outliers[[ var_list_plot[i_var] ]] <-
##     tibble::tibble(
##       Var           = var_list_plot[i_var]
##     , demo_study_id = dat_sheet[[ "demo_study_id" ]]
##     , Value         = this_var_org
##     , Value_trans   = this_var
##     , IQR_dist      = (this_var - this_med) / this_iqr
##     ) %>%
##     dplyr::filter(
##       (IQR_dist >  IQR_dist_max) |
##       (IQR_dist < -IQR_dist_max)
##     )
## }
##
## tab_outliers <-
##   tab_outliers %>%
##   dplyr::bind_rows()
##
## tab_outliers %>% print(n=Inf)
```



Drop extreme outlying values.
```{r chunk-02}
# # Drop extreme outlying values
# # dat_sheet$fl_albumin_index       [dat_sheet$fl_albumin_index              ] <- NA  # obs
# # dat_sheet$mri_dti_freewater_mean [dat_sheet$mri_dti_freewater_mean        ] <- NA  # obs
# # dat_sheet$mri_dti_psmd           [dat_sheet$mri_dti_psmd                  ] <- NA  # obs
# # dat_sheet$fl_myelin_basic_protein[dat_sheet$fl_myelin_basic_protein       ] <- NA  # obs
# dat_sheet$fl_igg_index           [dat_sheet$fl_igg_index               <  0.2  ] <- NA  # obs 139
# # dat_sheet$fl_angio_csf_bfgf      [dat_sheet$fl_angio_csf_bfgf             ] <- NA  # obs
# # dat_sheet$fl_angio_csf_flt_1     [dat_sheet$fl_angio_csf_flt_1            ] <- NA  # obs
# # dat_sheet$fl_angio_csf_plgf      [dat_sheet$fl_angio_csf_plgf             ] <- NA  # obs
# # dat_sheet$fl_angio_csf_tie2      [dat_sheet$fl_angio_csf_tie2             ] <- NA  # obs
# # dat_sheet$fl_angio_csf_vegf_a    [dat_sheet$fl_angio_csf_vegf_a           ] <- NA  # obs
# # dat_sheet$fl_angio_csf_vegf_c    [dat_sheet$fl_angio_csf_vegf_c           ] <- NA  # obs
# # dat_sheet$fl_angio_csf_vegf_d    [dat_sheet$fl_angio_csf_vegf_d           ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_il_1b      [dat_sheet$fl_cyto_csf_il_1b             ] <- NA  # obs
# dat_sheet$fl_cyto_csf_il_2       [dat_sheet$fl_cyto_csf_il_2              ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_il_4       [dat_sheet$fl_cyto_csf_il_4              ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_il_6       [dat_sheet$fl_cyto_csf_il_6              ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_il_8       [dat_sheet$fl_cyto_csf_il_8              ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_ifn_g      [dat_sheet$fl_cyto_csf_ifn_g             ] <- NA  # obs
# # dat_sheet$fl_cyto_csf_tnf_a      [dat_sheet$fl_cyto_csf_tnf_a             ] <- NA  # obs
# # dat_sheet$fl_msd_csf_mmp_1       [dat_sheet$fl_msd_csf_mmp_1              ] <- NA  # obs
# # dat_sheet$fl_msd_csf_mmp_2       [dat_sheet$fl_msd_csf_mmp_2              ] <- NA  # obs
# # dat_sheet$fl_msd_csf_mmp_3       [dat_sheet$fl_msd_csf_mmp_3              ] <- NA  # obs
# # dat_sheet$fl_msd_csf_mmp_9       [dat_sheet$fl_msd_csf_mmp_9              ] <- NA  # obs
# # dat_sheet$fl_msd_csf_mmp_10      [dat_sheet$fl_msd_csf_mmp_10             ] <- NA  # obs
# # dat_sheet$fl_ptau                [dat_sheet$fl_ptau                       ] <- NA  # obs
# # dat_sheet$mri_perm_3t_tot        [dat_sheet$mri_perm_3t_tot            <  -6   ] <- NA  # obs 145
# dat_sheet$percent_wmhv           [dat_sheet$percent_wmhv               < -10   ] <- NA  # obs 113, 137
# # dat_sheet$demo_study_id          [dat_sheet$demo_study_id                 ] <- NA  # obs
# # dat_sheet$demo_bl_age            [dat_sheet$demo_bl_age                   ] <- NA  # obs
# # dat_sheet$demo_sex               [dat_sheet$demo_sex                      ] <- NA  # obs
# # dat_sheet$demo_yrs_edu           [dat_sheet$demo_yrs_edu                  ] <- NA  # obs
# # dat_sheet$dx_current             [dat_sheet$dx_current                    ] <- NA  # obs
# dat_sheet$percent_hippo          [dat_sheet$percent_hippo              >   1   ] <- NA  # obs 1, 2
# # dat_sheet$mri_mean_cort_thick_mm [dat_sheet$mri_mean_cort_thick_mm        ] <- NA  # obs
# # dat_sheet$fl_abeta_4240_ratio    [dat_sheet$fl_abeta_4240_ratio           ] <- NA  # obs
#
#
# var_list_plot_change <-
#   c(
#   #   "fl_albumin_index",
#   # , "mri_dti_freewater_mean"
#   # , "mri_dti_psmd"
#   # , "fl_myelin_basic_protein"
#     "fl_igg_index"
#   # , "fl_angio_csf_bfgf"
#   # , "fl_angio_csf_flt_1"
#   # , "fl_angio_csf_plgf"
#   # , "fl_angio_csf_tie2"
#   # , "fl_angio_csf_vegf_a"
#   # , "fl_angio_csf_vegf_c"
#   # , "fl_angio_csf_vegf_d"
#   # , "fl_cyto_csf_il_1b"
#   # , "fl_cyto_csf_il_2"
#   # , "fl_cyto_csf_il_4"
#   # , "fl_cyto_csf_il_6"
#   # , "fl_cyto_csf_il_8"
#   # , "fl_cyto_csf_ifn_g"
#   # , "fl_cyto_csf_tnf_a"
#   # , "fl_msd_csf_mmp_1"
#   # , "fl_msd_csf_mmp_2"
#   # , "fl_msd_csf_mmp_3"
#   # , "fl_msd_csf_mmp_9"
#   # , "fl_msd_csf_mmp_10"
#   # , "fl_ptau"
#   , "mri_perm_3t_tot"
#   , "percent_wmhv"
#   # , "demo_study_id"
#   # , "demo_bl_age"
#   # , "demo_sex"
#   # , "demo_yrs_edu"
#   # , "dx_current"
#   , "percent_hippo"
#   # , "mri_mean_cort_thick_mm"
#   # , "fl_abeta_4240_ratio"
#   )
```

Plot again


```{r, fig.height = 4, fig.width = 10}
# var_list_plot <-
#   c(var_list_plot_change)
#
# par(mfrow=c(1,3))
# for (i_plot in seq_along(var_list_plot)) {
#   ## i_plot = 1
#   #if (i_plot > 3) {next}
#
#   plot(
#     x     = 1:nrow(dat_sheet)
#   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
#   , type  = "p"
#   , main  = paste0("After: ", var_list_plot[i_plot])
#   , xlab  = ""
#   , ylab  = ""
#   , col   = dat_sheet$dx_current
#   , pch   = 20  #3
#   , xlim  = c(1, nrow(dat_sheet) * 1.2)
#   )
#   text(
#     x     = 1:nrow(dat_sheet)
#   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
#   , labels = 1:nrow(dat_sheet)
#   , pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
#   )
# }
# par(mfrow=c(1,1))
```


Normalize so color scale is meaningful

```{r chunk-03}
#normalize so color scale is meaningful

# data for demographics, before z-score scaling
dat_sheet_demo <-
  #dat_sheet
  dat_sub

dat_sheet <-
  #dat_sheet %>%
  dat_sub %>%
  dplyr::mutate(
    # to allow log transformation when wmhv==0
  #  percent_wmhv  = percent_wmhv + 0.00001
  #  fl_cyto_csf_il_4 = ifelse(fl_cyto_csf_il_4 == 0.001, NA, fl_cyto_csf_il_4)
  #  across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), e_log_shift, min_add = 1, base_log = 2)
    across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), scale)
  , across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), as.numeric)
  )
  # %>%
  #dplyr::mutate_if(is.numeric, scale) %>%

# label dataset
for (n_label in as.character(unlist(Var))) {
  ## i_label = 1
  labelled::var_label(dat_sheet     [[ n_label ]]) <- labelled::var_label(dat_sheet_demo[[ n_label ]])
  #labelled::var_label(dat_sheet_demo[[ dat_var_label_matching$Var[i_label] ]]) <- dat_var_label_matching$Label[i_label]
}
```

Determine outliers by IQR.
```{r}
IQR_dist_max <- 5

tab_outliers <- list()

for (i_var in seq_along(var_list_plot)) {
  ## i_var = 1
  ## i_var = 33
  if (!is.numeric(dat_sheet[[ var_list_plot[i_var] ]])) {
    next
  }
  this_var <- dat_sheet[[ var_list_plot[i_var] ]]
  this_var_org <- this_var
  #if (var_list_plot[i_var] %in% totransform.log2) {
  #  #this_var <- e_log_shift(this_var)
  #  this_var <- e_log_shift(this_var, min_add = 1, base_log = 2)
  #}
  this_med <- median(this_var, na.rm = TRUE)
  this_iqr <- IQR(   this_var, na.rm = TRUE)

  tab_outliers[[ var_list_plot[i_var] ]] <-
    tibble::tibble(
      Var           = var_list_plot[i_var]
    , demo_study_id = dat_sheet[[ "ptid" ]]
    , Value         = this_var_org
    , Value_trans   = this_var
    , IQR_dist      = (this_var - this_med) / this_iqr
    ) %>%
    dplyr::filter(
      (IQR_dist >  IQR_dist_max) |
      (IQR_dist < -IQR_dist_max)
    )
}

tab_outliers <-
  tab_outliers %>%
  dplyr::bind_rows()

tab_outliers %>% print(n=Inf)
```


### Plot on trasnformed scale

```{r, fig.height = 4, fig.width = 10}
# var_list_plot
#   #<-
#   #c(totransform.log2, tokeep)
#
# par(mfrow=c(1,3))
# for (i_plot in seq_along(var_list_plot)) {
#   ## i_plot = 1
#   #if (i_plot > 3) {next}
#
#   plot(
#     x     = 1:nrow(dat_sheet)
#   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
#   , type  = "p"
#   , main  = paste0("After trans: ", var_list_plot[i_plot])
#   , xlab  = ""
#   , ylab  = ""
#   , col   = dat_sheet$dx_current
#   , pch   = 20  #3
#   , xlim  = c(1, nrow(dat_sheet) * 1.2)
#   )
#   text(
#     x     = 1:nrow(dat_sheet)
#   , y     = dat_sheet[[ var_list_plot[i_plot] ]]
#   , labels = 1:nrow(dat_sheet)
#   , pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
#   )
# }
# par(mfrow=c(1,1))
```




```{r chunk-04, fig.height = 4, fig.width = 8}
#get regression coefficients for reduced model

source("./model_sel_new.R")

lm_criterion <- c("AIC", "BIC")[2]

## model selection
f_model_sel <-
  function(
    dat_this      = dat_lm
  , lm_criterion  = c("AIC", "BIC")[1]
  ) {

  if (lm_criterion == "AIC") { AIC_k = 2                     }
  if (lm_criterion == "BIC") { AIC_k = log(nrow( dat_this )) }

  results_list     <- list()
  results_list$mod <- f_formula_init()
  results_list$fit <- list()
  results_list$AIC <- list()

  for (i_mod in seq_len(length(results_list$mod))) {
    ## i_mod = 1

    results_list$fit[[i_mod]] <-
      lm(
        formula = results_list$mod[[i_mod]]
      , data    = dat_this
      )

    results_list$AIC[[i_mod]] <-
      results_list$fit[[i_mod]] %>%
      AIC(k = AIC_k)
  }

  ## some models have a term with 0 df, skip this model
  # first remove models with no error estimate
  results_list$AIC[ results_list$AIC == -Inf ] <- Inf
  # then remove models with 0 df
  sw_df_zero <- TRUE
  while(sw_df_zero) {
    # select lowest AIC, index
    ind_min_AIC <-
      results_list$AIC %>%
      unlist() %>%
      which.min()

    sw_df_zero <-
      any(
        results_list$fit[[ind_min_AIC]] %>%
        car::Anova(type = 3, singular.ok = TRUE) %>%
        pull(Df)
          ==
        0
      )

    if (sw_df_zero) {
      print("Skipping 0-df model")
      #results_list$mod[[ind_min_AIC]] <- NULL
      #results_list$fit[[ind_min_AIC]] <- NULL
      #results_list$AIC[[ind_min_AIC]] <- NULL
      results_list$AIC[[ind_min_AIC]] <- Inf
    } # if
  } # while

  # selected model
  results <-
    results_list$fit[[ind_min_AIC]]

  return(results)
} # f_model_sel



list_var_resp <-
  c("fl_albumin_index_log2", "mri_perm_3t_tot",  "fl_csf__ALL__gfap_log2","fl_plasma__ALL__gfap_log2")
ind_var_covar <-
  var_list_plot %notin%
    c("ptid","age","sex","demo_edu_yrs","dx_current",
      "fl_albumin_index_log2","mri_perm_3t_tot_log2","mri_perm_3t_tot",
      "fl_csf__ALL__gfap_log2","fl_plasma__ALL__gfap_log2")
list_var_covar <-
  var_list_plot[ind_var_covar]

results <- list()

for (i_resp in seq_along(list_var_resp)) {
  for (i_covar in seq_along(list_var_covar)) {
    ## i_resp = 2
    ## i_covar = 6
    ## i_resp = 1
    ## i_covar = 1
    ## i_resp = 1
    ## i_covar = 23

    var_resp  <- list_var_resp[i_resp]
    var_covar <- list_var_covar[i_covar]

    print(paste0("BEGIN  Resp: ", i_resp, " ", var_resp, ",  Covar: ", i_covar, " ", var_covar, "  ===================================================="))

    if (i_covar == 1) {
      results[[var_resp]] <- list()
    }
    results[[var_resp]][[var_covar]] <- list()

    # subset of data for linear model
    dat_lm <-
      dat_sheet %>%
      dplyr::select(
        all_of(var_resp)
      , all_of(var_covar)
      , age
      , sex
      #, demo_edu_yrs  # 3/9/2022 Educ added
      ) %>%
      dplyr::mutate(
        Y = !!sym(var_resp)
      , X = !!sym(var_covar)
      ) %>%
      tidyr::drop_na() %>%
      tidyr::as_tibble()
    n_sample <- nrow(dat_lm)

    if (n_sample == 0) {
      next
    }


    form_lm <-
      as.formula(
        paste0(
        #  var_resp
          "Y"
        , " ~ "
        , "age"
        , " + "
        , "sex"
        #, " + "
        #, "demo_edu_yrs"    # 3/9/2022 Educ added
        , " + "
        #, var_covar
        , "X"
        , " + "
        #, paste0("age", ":", var_covar)
        , paste0("age", ":", "X")
        , " + "
        #, paste0("sex", ":", var_covar)
        , paste0("sex", ":", "X")
        #, " + "
        #, paste0("demo_edu_yrs", ":", "X")    # 3/9/2022 Educ added
        )
      )

    results[[var_resp]][[var_covar]][["form_lm"]] <- form_lm
    results[[var_resp]][[var_covar]][["dat_lm" ]] <- dat_lm

    results[[var_resp]][[var_covar]][["lm_fit_full"]] <-
      lm(
        formula = form_lm
      , data    = dat_lm
      )

    ## results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]] <-
    ##   stats::step(
    ##     object = results[[var_resp]][[var_covar]][["lm_fit_full"]]
    ##   , direction = "both"
    ##   , test = "F"
    ##   , trace = 0
    ##   , k = log(nrow(dat_lm))
    ##   )

    results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]] <-
      f_model_sel(
        dat_this      = dat_lm
      , lm_criterion  = lm_criterion
      )


    results[[var_resp]][[var_covar]][["lm_fit_full"]]$df.residual

    ### 7/8/2022 report all results
    ## skip if model selection removed our x variable
    if (!("X" %in% names(coef(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]])))) {
      print("Model selection removed X variable")
      results[[var_resp]][[var_covar]]$sw_sig_X <- FALSE
      #next
    } else {
      results[[var_resp]][[var_covar]]$sw_sig_X <- TRUE
    }

    print(summary(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]))

    # plot residuals
    plot(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]], which = c(1))

    # Scatterplot for fitted reduced model

    covar_names_red <- names(coef(results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]))

    dat_pred <-
      tidyr::expand_grid(
        X = dat_lm[["X"]] %>% range()
      , age  = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
      , sex     = dat_lm[["sex"]] %>% unique()
      #, demo_edu_yrs = c(0, dat_lm[["demo_edu_yrs"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
      )

    ## if (
    ##   ("age"  %in% covar_names_red) &
    ##   ("sexMale" %in% covar_names_red) #&
    ##   #("demo_edu_yrs" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     , age  = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     , sex     = dat_lm[["sex"]] %>% unique()
    ##     , demo_edu_yrs = c(0, dat_lm[["demo_edu_yrs"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     )
    ## } else if (
    ##   ("age" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     , age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     #, sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## } else if (
    ##   ("sexMale" %in% covar_names_red)
    ## ) {
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     #, age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     , sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## } else {
    ##   # only X is significant
    ##   dat_pred <-
    ##     tidyr::expand_grid(
    ##       X = dat_lm[["X"]] %>% range()
    ##     #, age = c(0, dat_lm[["age"]] %>% quantile(probs = c(0.10, 0.90))) %>% as.numeric()
    ##     #, sex = dat_lm[["sex"]] %>% unique()
    ##     )
    ## }

    dat_pred$Y <-
      predict(
        results[[var_resp]][[var_covar]][["lm_fit_red_BIC"]]
      , newdata = dat_pred
      )

    dat_pred <-
      dat_pred %>%
      dplyr::mutate(
        age  = age  %>% factor(labels = c("10%", "Mean", "90%"))
      #, demo_edu_yrs = demo_edu_yrs %>% factor(labels = c("10%", "Mean", "90%"))
      )

    ## if ("age" %in% covar_names_red) {
    ##   dat_pred <-
    ##     dat_pred %>%
    ##     dplyr::mutate(
    ##       age = age %>% factor(labels = c("10%", "Mean", "90%"))
    ##     )
    ## }
    ## if ("demo_edu_yrs" %in% covar_names_red) {
    ##   dat_pred <-
    ##     dat_pred %>%
    ##     dplyr::mutate(
    ##       demo_edu_yrs = demo_edu_yrs %>% factor(labels = c("10%", "Mean", "90%"))
    ##     )
    ## }

    library(ggplot2)
    if ("sexFemale" %in% covar_names_red) {
      p <- ggplot(dat_lm, aes(x = X, y = Y, shape = sex, colour = sex))
    } else {
      p <- ggplot(dat_lm, aes(x = X, y = Y))
    }
    p <- p + theme_bw()
    p <- p + geom_point(size = 2, alpha = 1/2)
    if ("age" %in% covar_names_red) {
      p <- p + geom_line(data = dat_pred, aes(linetype = age), size = 0.75)
    } else {
      p <- p + geom_line(data = dat_pred, size = 0.75)
    }
    # 3/9/2022 Educ added  XXX

    p <- p + scale_linetype_manual(values = c(2, 1, 3))
    p <- p + labs(  title = NULL
                  #, subtitle = "Progress and Starting Current"
                  , x = labelled::var_label(dat_sheet[[var_covar]])
                  , y = labelled::var_label(dat_sheet[[var_resp ]])
                  #, caption = paste0(  "Model fit averages the numeric variables of Age, A1c, and PAM-10\n"
                  #                   , "  over Program, Diagnosis, Gender, and General health.\n"
                  #                   , "Dotted vertical lines indicate cutoffs between depression categories:\n"
                  #                   , "  None-minimal, Mild, Moderate, Moderately Severe, and Severe."
                  #                   )
                  , colour    = "Gender"
                  , shape     = "Gender"  # "Imputed"
                  , linetype  = "Age"  #"Diagnosis"
                  #, fill      = "Diagnosis"
                  )
    #p <- p + theme(plot.caption = element_text(hjust = 0))
    #p <- p + theme(legend.position = "bottom")
    #p <- p + guides(colour = guide_legend(nrow = 2), shape = guide_legend(nrow = 2))
    n_sample_annotate <- paste0("N=",n_sample)
    #p <- p + annotate("text", -Inf, -Inf, label=n_sample_annotate, hjust=-0.3, vjust=-1, colour = "gray25")
    p <- p + annotate("text", x=-Inf, y= Inf, label=n_sample_annotate, hjust=-0.3, vjust=2, colour = "gray25")
    print(p)

    results[[var_resp]][[var_covar]][["plot_lm_fit_red_BIC"]] <-
      p

    print(paste0("END    Resp: ", i_resp, " ", var_resp, ",  Covar: ", i_covar, " ", var_covar, "  ===================================================="))
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")
    print("")


  } # i_covar
} # i_resp

```


```{r chunk-05}
#create matrices
ind_row <-
  #c(totransform.log2, tokeep) %notin%
  #  c("ptid","age","sex","demo_edu_yrs","dx_current","fl_albumin_index_log2","mri_perm_3t_tot")
  as.character(unlist(Var)) %notin%
    c("ptid","age","sex","demo_edu_yrs","dx_current","fl_albumin_index_log2","mri_perm_3t_tot")
n_row <-
  #c(totransform.log2, tokeep)[ind_row]
  as.character(unlist(Var))[ind_row]

l_row <-
  length(n_row)

n_col <-
  #c("Age", "Gender", "Education", "X", "Age:X", "Gender:X", "Education:X")
  c("Age", "Gender",              "X", "Age:X", "Gender:X"               )
l_col <-
  length(n_col)

# fill matrices with the signficant estimates, leave other cells blank

l_var_resp <- c("fl_albumin_index_log2", "mri_perm_3t_tot_log2",
                "fl_csf__ALL__gfap_log2", "fl_plasma__ALL__gfap_log2")


results_fit <- list()
i_results <- 0

for (i_resp in seq_along(l_var_resp)) {
  ## i_resp = 1
  for (i_covar in seq_along(n_row)) {
    ## i_covar = 1

    var_row <- n_row[i_covar]
    var_col <- l_var_resp[i_resp]

    exists <- results[[var_col]][[var_row]]

    if(length(results[[var_col]][[var_row]]) == 0) {
      next
    }

    i_results <- i_results + 1

    results_fit[[ i_results ]] <-
      results[[var_col]][[var_row]][["lm_fit_red_BIC"]] |>
      summary() |>
      broom::tidy() |>
      dplyr::mutate(
        term  =
          term |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("sexFemale")
          , replacement = "Gender"
          ) |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("age")
          , replacement = "Age"
          )
      , Covar =
          term |>
          stringr::str_replace_all(
            pattern     = stringr::fixed("X")
          , replacement = var_row
          )
      , resp  = var_col
      , covar = var_row
      , n     = results[[var_col]][[var_row]][[ "dat_lm" ]] |> nrow()
      ) |>
      dplyr::relocate(
        resp
      , covar
      )

  } # i_covar
} # i_resp

dat_plot <-
  results_fit |>
  dplyr::bind_rows() |>
  dplyr::rename(
  #  Var   = covar
    Coef  = estimate
  , pvals = p.value
  ) %>%
  dplyr::mutate(
    pvals_stars =
      pvals %>%
      e_pval_stars(
        cutpoints = c(0, 0.001, 0.01, 0.05, 1)
      , symbols = c("***", "**", "*", " ")
      )
  ) %>%
  left_join(
    dat_var_label_matching
  , by = dplyr::join_by(covar == Var)
  ) %>%
  left_join(
    e_data_var_labels_table_from_data(dat_sheet)
  , by = dplyr::join_by(covar == Var)
  ) %>%
  dplyr::rename(
    Label_covar = Label
  ) %>%
  left_join(
    e_data_var_labels_table_from_data(dat_sheet)
  , by = dplyr::join_by(resp == Var)
  ) %>%
  dplyr::rename(
    Label_resp = Label
  ) %>%
  dplyr::mutate(
    Label_covar = Label_covar |> factor()
  , Label_resp  = Label_resp  |> factor()
  , term        = term        |> factor(levels = c("(Intercept)", n_col))
  )
```


```{r chunk-06, fig.height = 4, fig.width = 10}
#plots

#merged<-cbind(coef_mat,coef_mat_mri)
#merged_p<-cbind(p_mat,p_mat_mri)

#tiff("coef_plot_withtitle.png", height = 8, width =6, units="in",res=300)

# corrplot(merged,method="square",is.corr=FALSE,col.lim=c(-1,.6),cl.ratio = 0.2,cl.pos="r",win.asp = .5,tl.pos="lt",tl.cex =0.8,tl.srt = 80,cl.cex = .6,tl.col=("gray20"),na.label=" ",sig.level=c(0.001, 0.01, 0.05),p.mat =merged_p,insig="label_sig",pch.cex = 1.5,pch.col = ("gray20"),mar=c(0,0,1,0),title="                                       Albumin Index       MRI WM Permeability")

#dev.off()

```

Erik ggplot version

```{r chunk-07}

#dat_plot <-
#  dplyr::full_join(
#    dplyr::bind_rows(
#      #coef_mat %>%
#      results_fit %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "Albumin Index"
#      )
#    , coef_mat_mri %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "MRI WM Permeability"
#      )
#    ) %>%
#    tidyr::pivot_longer(
#      cols = -c(Var, RespVar)
#    , names_to = "Covar"
#    , values_to = "Coef"
#    )
#  , dplyr::bind_rows(
#      p_mat %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "Albumin Index"
#      )
#    , p_mat_mri %>%
#      as_tibble(rownames = "Var") %>%
#      dplyr::mutate(
#        RespVar = "MRI WM Permeability"
#      )
#    ) %>%
#    tidyr::pivot_longer(
#      cols = -c(Var, RespVar)
#    , names_to = "Covar"
#    , values_to = "pvals"
#    )
#  , by = c("Var", "RespVar", "Covar")
#  ) %>%
#  dplyr::mutate(
#  #  Var = Var %>% stringr::str_replace_all(pattern = stringr::fixed("_"), replacement = " ")
#  #, Var = Var %>% stringr::str_replace_all(c("fl" = "","csf" = "","t 1" = "flt-1"))
#    pvals_stars =
#      pvals %>%
#      e_pval_stars(
#        cutpoints = c(0, 0.001, 0.01, 0.05, 1)
#      , symbols = c("***", "**", "*", " ")
#      )
#  ) %>%
#  left_join(
#    dat_var_label_matching
#  ) %>%
#  left_join(
#    e_data_var_labels_table_from_data(dat_sheet)
#  ) %>%
#  dplyr::mutate(
#    Label = Label |> factor()
#  , Covar = Covar %>% factor(levels = n_col)
#  )

# p-value plot
library(ggplot2)
#p1 <- ggplot(data = dat_plot, aes(x = Covar, y = Label, fill = Coef))  # y = Var
p1 <- ggplot(data = dat_plot, aes(x = term, y = Label_covar, fill = Coef))  # y = Var
#p1 <- p1 + theme_bw()
p1 <- p1 + geom_tile(na.rm = TRUE)
p1 <- p1 + scale_y_discrete(limits = rev)
#p1 <- p1 + scale_fill_gradientn(limits = c(-1.0, 0.6), values = scales::rescale(c(0,0.1,1)), colors = c("#000000", "#AAAAAA", "#DDDDDD"), breaks = c(0, 0.1, 0.5, 1), na.value = "white")
p1 <- p1 + scale_fill_gradient2(low = "blue", mid = "white", high = "red", breaks = seq(-100, 100, by = 2), na.value = "gray90")
p1 <- p1 + geom_text(aes(label = pvals_stars), vjust = 0.5, colour = "white")
p1 <- p1 + geom_text(data = dat_plot |> dplyr::filter(term == "(Intercept)")
              , aes(label = n), hjust = 0, vjust = 0.5, colour = "black"
              )
#p1 <- p1 + facet_grid(. ~ RespVar, space = "free", scales = "free_y", drop = TRUE)
p1 <- p1 + facet_grid(Type ~ Label_resp, space = "free", scales = "free_y", drop = TRUE)
p1 <- p1 + theme(#axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)
               axis.text.y = element_text(hjust = 0)
            )  #, panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p1 <- p1 + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) # rotate labels
p1 <- p1 + theme(axis.text.y = element_text(hjust = 1))
#p1 <- p1 + theme(legend.key.height = unit(0.30, "in"))
p1 <- p1 + theme(plot.caption = element_text(hjust = 0), plot.caption.position =  "plot") # caption left align
p1 <- p1 + theme(legend.position = "bottom") # "none"
p1 <- p1 + theme(legend.key.height = unit(0.1, "in"), legend.key.width = unit(0.4, "in"))

if (lm_criterion == c("AIC", "BIC")[1]) {
  p1 <- p1 + labs(
                  title = "Models selected by AIC"
                  )
}
if (lm_criterion == c("AIC", "BIC")[2]) {
  p1 <- p1 + labs(
                  title = "Models selected by BIC"
                  )
}
p1 <- p1 + labs(
              #, subtitle = "Longitudinal mixed model"
                y = "Standardized covariates (\"X\")"
              , x = "Model covariates"
              , fill = "Model\nCoefficients  "  # "FDR\np-value"
              , caption = paste0(
                            "p-value stars:  (blank): p >= 0.05;   *: p < 0.05;   **: p < 0.01;   *** : p < 0.001"
                          , "\np-values are uncorrected for family-wise Type-I error rate."
                          #, "\nModels displayed are those where X variable is retained after model selection"
                          )
              )

print(p1)

ggsave(
    "out_new/Fig2_plot_coef_pval.png"
  , plot   = p1
  , width  = 10
  , height = 16
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

ggsave(
    "out_new/Fig2_plot_coef_pval.pdf"
  , plot   = p1
  , width  = 10
  , height = 16
  ## png, jpeg
  #, dpi    = 300
  #, bg     = "white"
  ## pdf
  , units  = "in"
  , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

## #p_alb_mriperm_coef_pval <- p1
##
## # ggsave(
## #     paste0("plot_alb_mriperm_coef_pval_", lm_criterion, ".png")
## #   , plot   = p1
## #   , width  = 6
## #   , height = 8
## #   ## png, jpeg
## #   , dpi    = 300
## #   , bg     = "white"
## #   ## pdf
## #   #, units  = "in"
## #   #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )
## # ggsave(
## #     paste0("plot_alb_mriperm_coef_pval_", lm_criterion, ".pdf")
## #   , plot   = p1
## #   , width  = 6
## #   , height = 8
## #   ## png, jpeg
## #   #, dpi    = 300
## #   #, bg     = "white"
## #   ## pdf
## #   , units  = "in"
## #   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )

```

Individual plots associated with significant covariates.
```{r}
dat_plot_sub_indy <-
  dat_plot |>
  dplyr::filter(
    term == "X"
  )

plot_results <- list()

for (i_row in seq_len(nrow(dat_plot_sub_indy))) {
  ## i_row = 1

  var_resp  <- dat_plot_sub_indy$resp [i_row]
  var_covar <- dat_plot_sub_indy$covar[i_row]

  plot_results[[ i_row ]] <-
    results[[var_resp]][[var_covar]][["plot_lm_fit_red_BIC"]]
} # i_row

#plot_results


p_arranged <-
  cowplot::plot_grid(
    plotlist = plot_results
  , nrow = 7
  , ncol = 3
  )
ggsave(
    paste0("out_new/plot_results_indy", ".png")
  , plot   = p_arranged
  , width  = 12
  , height = 14
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE
  )



```






```{r, eval = TRUE}
# XXX-xxx
######################################## ########################################
#######        STOP EVAL        ######## #######        STOP EVAL        ########
######################################## ########################################
knitr::opts_chunk$set(eval = FALSE)  # TRUE
######################################## ########################################
#######        STOP EVAL        ######## #######        STOP EVAL        ########
######################################## ########################################
```



















## Choose scatterplots with model fit lines for paper

```{r chunk-08}

# ggsave(
#     "plot_NAME_ME"
#   , plot   = [[SOME PLOT HERE]]
#   , width  = 6
#   , height = 8
#   ## png, jpeg
#   #, dpi    = 300
#   #, bg     = "white"
#   ## pdf
#   , units  = "in"
#   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
#   )

```

```{r chunk-09}
#comparisons by dx and by patient status

mri.bydx=kruskal.test(mri_perm_3t_tot ~ dx_current,data=dat_sheet)
mri.bydx

mri.pairwise=dunn.test(dat_sheet$mri_perm_3t_tot,dat_sheet$dx_current,method="holm",altp = TRUE)

alb.bydx=kruskal.test(fl_albumin_index_log2 ~ dx_current,data=dat_sheet)
alb.bydx

alb.pairwise<- dunn.test(dat_sheet$fl_albumin_index_log2,dat_sheet$dx_current,method="holm", altp = TRUE)


#albumin vs. mri perm

perms <- cor.test(dat_sheet$fl_albumin_index_log2, dat_sheet$mri_perm_3t_tot, method="spearman")

coor_MRI_Albumin <-e_plot_scatterplot(dat_plot =dat_sheet,  var_x= "fl_albumin_index_log2", var_y= "mri_perm_3t_tot_log2", smooth_all = c("none", "loess", "lm", "glm", "gam")[3],)

ggsave(
    "All_figures_for_paper/coor_MRI_Albumin.png"
  , plot   = coor_MRI_Albumin
  , width  = 6
  , height = 4
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )


library(dplyr)
library(gt)

# Create summary data frame from your results
comparison_results <- data.frame(
  Variable = c("MRI WM Permeability", "Albumin Index (log2)"),
  KW_ChiSq = c(17.118, 18.707),
  KW_df = c(4, 4),
  KW_pvalue = c(0.001834, 0.0008973),
  Significant_Pairs = c(
    "AD-Mixed (p=0.029), AD-SIVD (p=0.006), LA-SIVD (p=0.047)",
    "LA-SIVD (p=0.0004)"
  )
)

# Create GT table
tbl <- comparison_results %>%
  gt() %>%
  tab_header(
    title = "Table 1. Kruskal-Wallis and Post-hoc Comparisons by Diagnosis",
    subtitle = "MRI Permeability and Albumin Index across diagnostic groups"
  ) %>%
  cols_label(
    Variable = "Measure",
    KW_ChiSq = "χ²",
    KW_df = "df",
    KW_pvalue = "p-value",
    Significant_Pairs = "Significant Pairwise Comparisons (Dunn test, Holm adjusted)"
  ) %>%
  fmt_number(
    columns = c(KW_ChiSq),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(KW_pvalue),
    decimals = 4
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgray"),
    locations = cells_body(
      columns = KW_pvalue,
      rows = KW_pvalue < 0.05
    )
  ) %>%
  tab_footnote(
    footnote = "AD: Alzheimer's disease; LA: leukoaraiosis; Mixed: mixed dementia; SIVD: subcortical ischemic vascular dementia; MI: mild impairment",
    locations = cells_column_labels(columns = Significant_Pairs)
  ) %>%
  tab_footnote(
    footnote = "Spearman correlation between MRI permeability and Albumin index: rho = 0.136, p = 0.276",
    locations = cells_title()
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12
  )

# Display table
tbl

```


## Figures for manuscript

### Figure 1

```{r chunk-10}
#plots

#basic plots

# Scatterplot MRI Perm and Albumin
bothperm <- dat_sheet %>% select(fl_albumin_index_log2,mri_perm_3t_tot)
bothperm <- na.omit(bothperm)
both_sample <- nrow(bothperm)
both_annotate <- paste0("N=",both_sample)
plot.perms <-
  ggplot(data = dat_sheet, aes(x = fl_albumin_index_log2, y = mri_perm_3t_tot)) +
  theme_bw() +
  geom_hline(aes(yintercept = 0), colour = "black", linetype = "solid", size = 0.2, alpha = 0.15) +
  geom_vline(aes(xintercept = 0), colour = "black", linetype = "solid", size = 0.2, alpha = 0.15) +
  geom_point(size=2, alpha=0.3) +
  stat_ellipse(level = 0.80) +
  labs(
    subtitle = "Albumin Index vs.\n  MRI WM Permeability"
  , x = labelled::var_label(dat_sheet$fl_albumin_index_log2)
  , y = labelled::var_label(dat_sheet$mri_perm_3t_tot)
  ) +
  #theme(plot.title = element_text(hjust = 0.5)) +
  annotate(geom = "text", x = -Inf, y =  Inf, label = paste0(both_annotate, ", r=0.0856, p=0.505"), hjust = -0.05, vjust = 2, colour = "gray25")
  # +
  #annotate(geom = "text", x = -Inf, y = -Inf, label =           , hjust = -0.3, vjust = -1, color = "gray50")

#print(plot.perms)


# MRI Perm
plot.mri <-
  ggplot(data = dat_sheet,aes(x = dx_current, y = mri_perm_3t_tot)) +
  theme_bw() +
  geom_hline(aes(yintercept = 0), colour = "black", linetype = "solid", size = 0.2, alpha = 0.15) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(size=2, alpha=.3, width=0.15) +
  labs(
    subtitle = "MRI WM Permeability\n  Differences by Diagnosis"
  , x = "Diagnosis"
  , y = labelled::var_label(dat_sheet$mri_perm_3t_tot)
  ) +
  #theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(limits = c(-3,3.25))
  #ylim(-3,3)

stat.test <- tibble::tribble(
  ~group1, ~group2,   ~p.adj,
    "AD",     "MX", 0.0013,
    "AD",   "SIVD", 0.0001,
    "LA",     "MX", 0.0171,
    "LA",   "SIVD", 0.0103
  )

plot.mri <- plot.mri + stat_pvalue_manual(stat.test,
    y.position = 1.75, step.increase = 0.12,
    label = "p.adj"
    )

tab_n <-
  dat_sheet %>%
  dplyr::select(
    dx_current
  , mri_perm_3t_tot
  ) %>%
  dplyr::filter(
    !is.na(mri_perm_3t_tot)
  ) %>%
  e_table_sum_freq_prop(
    var_names = "dx_current"
  ) %>%
  dplyr::mutate(
    label = paste0("N=", n)
  )

plot.mri <- plot.mri + geom_text(data = tab_n, aes(x = dx_current, y = -Inf, label = label), vjust = -1, colour = "gray25")

#print(plot.mri)


# Albumin
plot.alb <-
  ggplot(data = dat_sheet,aes(x = dx_current, y = fl_albumin_index_log2)) +
  theme_bw() +
  geom_hline(aes(yintercept = 0), colour = "black", linetype = "solid", size = 0.2, alpha = 0.15) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(size=2, alpha=.3, width=0.15) +
  labs(
    subtitle = "Albumin Index\n  Differences by Diagnosis"
  , x = "Diagnosis"
  , y = labelled::var_label(dat_sheet$fl_albumin_index_log2)
  )  +
  #theme(plot.title = element_text(hjust = 0.5))
  scale_y_continuous(limits = c(-3,3.25))


stat.test.alb <- tibble::tribble(
  ~group1, ~group2,   ~p.adj,
    "AD",     "LA", 0.0190,
    "LA",   "SIVD", 0.0003
)

plot.alb <- plot.alb + stat_pvalue_manual(stat.test.alb,
    y.position = 2.5, step.increase = 0.12,
    label = "p.adj"
    )

tab_n <-
  dat_sheet %>%
  dplyr::select(
    dx_current
  , fl_albumin_index_log2
  ) %>%
  dplyr::filter(
    !is.na(fl_albumin_index_log2)
  ) %>%
  e_table_sum_freq_prop(
    var_names = "dx_current"
  ) %>%
  dplyr::mutate(
    label = paste0("N=", n)
  )

plot.alb <- plot.alb + geom_text(data = tab_n, aes(x = dx_current, y = -Inf, label = label), vjust = -1, colour = "gray25")

# print(plot.alb)


p_list <- list()
p_list[[1]] <- plot.perms
p_list[[2]] <- plot.alb
p_list[[3]] <- plot.mri



p_arranged <-
  cowplot::plot_grid(
    plotlist = p_list
  , nrow = 1
  , ncol = NULL
  , labels = LETTERS[1:3]
  , rel_widths = c(1.5, 1, 1)
  )
p_arranged %>% print()


# write to pdf or other image format
ggsave(
    "out_new/Fig1_plot_MRI-BBB_Alb_scatterplots_anova.png"
  , plot   = p_arranged
  , width  = 8
  , height = 4
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )
ggsave(
    "out_new/Fig1_plot_MRI-BBB_Alb_scatterplots_anova.pdf"
  , plot   = p_arranged
  , width  = 8
  , height = 4
  ## png, jpeg
  #, dpi    = 300
  #, bg     = "white"
  ## pdf
  , units  = "in"
  , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )
```


### Figure 2

A. Model summary of p-values
B. Alb vs MMP-10
C. Alb vs MMP-2
D. Perm vs PSMD

```{r chunk-11}

#### Revision 1, Moved up
## p_list <- list()
## i_p <- 0
##
## i_p <- i_p + 1
## p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_2   $plot_lm_fit_red_BIC
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())
##
## i_p <- i_p + 1
## p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_3   $plot_lm_fit_red_BIC
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())
##
## i_p <- i_p + 1
## p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_10  $plot_lm_fit_red_BIC
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
## p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())
##
##
## # p_arranged <-
## #   cowplot::plot_grid(
## #     plotlist = p_list
## #   , nrow = 1
## #   , ncol = NULL
## #   )
## # p_arranged %>% print()
## #
## #
## #
## # # write to pdf or other image format
## # ggsave(
## #     "plot_scatterplots_model_pred_selected.png"
## #   , plot   = p_arranged
## #   , width  = 8
## #   , height = 4
## #   ## png, jpeg
## #   , dpi    = 300
## #   , bg     = "white"
## #   ## pdf
## #   #, units  = "in"
## #   #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )
##
##
##
## ## Put both plots together in one full-page figure.
##
## # p_arranged <-
## #   cowplot::plot_grid(
## #     plotlist = p_list
## #   , nrow = NULL
## #   , ncol = 1
## #   , labels = LETTERS[1:length(p_list) + 1]
## #   )
##
## p_alb_mriperm_coef_pval <- p_alb_mriperm_coef_pval + theme(legend.position = "bottom") # "none"
## p_alb_mriperm_coef_pval <- p_alb_mriperm_coef_pval + theme(legend.key.height = unit(0.2, "in"), legend.key.width = unit(0.4, "in"))
## p_alb_mriperm_coef_pval
##
## # p_arranged2 <-
## #   cowplot::plot_grid(
## #     plotlist = list(p_alb_mriperm_coef_pval, p_arranged)
## #   , nrow = 1
## #   , ncol = NULL
## #   , labels = c(LETTERS[1], NA)
## #   , rel_widths = c(2, 1)
## #   )
## # #p_arranged2 %>% print()
## #
## # # write to pdf or other image format
## # ggsave(
## #     "Fig2_plot_coef_pval_with_scatterplots.png"
## #   , plot   = p_arranged2
## #   , width  = 8
## #   , height = 8
## #   ## png, jpeg
## #   , dpi    = 300
## #   , bg     = "white"
## #   ## pdf
## #   #, units  = "in"
## #   #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )
## #
## # ggsave(
## #     "Fig2_plot_coef_pval_with_scatterplots.pdf"
## #   , plot   = p_arranged2
## #   , width  = 8
## #   , height = 8
## #   ## png, jpeg
## #   #, dpi    = 300
## #   #, bg     = "white"
## #   ## pdf
## #   , units  = "in"
## #   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
## #   )
##
## ggsave(
##     "Fig2_plot_coef_pval.png"
##   , plot   = p_alb_mriperm_coef_pval
##   , width  = 5
##   , height = 8
##   ## png, jpeg
##   , dpi    = 300
##   , bg     = "white"
##   ## pdf
##   #, units  = "in"
##   #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
##   )
##
## ggsave(
##     "Fig2_plot_coef_pval.pdf"
##   , plot   = p_alb_mriperm_coef_pval
##   , width  = 5
##   , height = 8
##   ## png, jpeg
##   #, dpi    = 300
##   #, bg     = "white"
##   ## pdf
##   , units  = "in"
##   , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
##   )

```


### Figure 3

```{r chunk-12}

p_list <- list()
i_p <- 0

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_2   $plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_3   $plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_msd_csf_mmp_10  $plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_cyto_csf_il_4$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_cyto_csf_il_6$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_cyto_csf_il_8$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_cyto_csf_tnf_a$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_angio_csf_plgf$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$fl_albumin_index_log2$fl_angio_csf_vegf_c$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())



p_arranged <-
  cowplot::plot_grid(
    plotlist = p_list
  , nrow = 3
  , ncol = NULL
  , labels = LETTERS[1:length(p_list)]
  #, align = c("none", "h", "v", "hv")[4]
  )
p_arranged %>% print()


# write to pdf or other image format
ggsave(
    "out_new/Fig3_plot_scatterplots_model_pred_cyto_angio.png"
  , plot   = p_arranged
  , width  = 8
  , height = 7
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

# write to pdf or other image format
ggsave(
    "out_new/Fig3_plot_scatterplots_model_pred_cyto_angio.pdf"
  , plot   = p_arranged
  , width  = 8
  , height = 7
  ## png, jpeg
  #, dpi    = 300
  #, bg     = "white"
  ## pdf
  , units  = "in"
  , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

```

### Figure 4

```{r chunk-13}

p_list <- list()
i_p <- 0

i_p <- i_p + 1
p_list[[i_p]] <- results$mri_perm_3t_tot $mri_dti_freewater_mean$plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())
#p_list[[i_p]] <- p_list[[i_p]] + ylab("Total MRI Permeability       ")
p_list[[i_p]]

i_p <- i_p + 1
p_list[[i_p]] <- results$mri_perm_3t_tot $mri_dti_psmd       $plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

i_p <- i_p + 1
p_list[[i_p]] <- results$mri_perm_3t_tot $percent_wmhv       $plot_lm_fit_red_BIC
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.position = "bottom") # "none"
p_list[[i_p]] <- p_list[[i_p]] + theme(legend.box="vertical", legend.margin=margin())

p_arranged <-
  cowplot::plot_grid(
    plotlist = p_list
  , nrow = 1
  , ncol = NULL
  , labels = LETTERS[1:length(p_list)]
  #, align = c("none", "h", "v", "hv")[4]
  )
p_arranged %>% print()


# write to pdf or other image format
ggsave(
    "out_new/Fig4_plot_scatterplots_model_pred_perm.png"
  , plot   = p_arranged
  , width  = 8
  , height = 7/3
  ## png, jpeg
  , dpi    = 300
  , bg     = "white"
  ## pdf
  #, units  = "in"
  #, useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

# write to pdf or other image format
ggsave(
    "out_new/Fig4_plot_scatterplots_model_pred_perm.pdf"
  , plot   = p_arranged
  , width  = 8
  , height = 7/3
  ## png, jpeg
  #, dpi    = 300
  #, bg     = "white"
  ## pdf
  , units  = "in"
  , useDingbats = FALSE  # https://stackoverflow.com/questions/9992275/ggplot2-pdf-import-in-adobe-illustrator-missing-font-adobepistd
  )

```


## Tables

### Table 1, demographics

```{r chunk-14}
# install.packages("moonbook")

# specify tows to summarize and columns to summarize by
list_var_col_names <-
  c("dx_current")
list_var_row_names <-
  c(
    "age"
  , "sex"
  , "demo_edu_yrs"
  , "fl_albumin_index_log2"
  , "mri_perm_3t_tot"
  )

# label variables
list_var_col_labels <- labelled::var_label(dat_sheet[, list_var_col_names]) %>% unlist()
list_var_row_labels <- labelled::var_label(dat_sheet[, list_var_row_names]) %>% unlist()

# This loop creates a table summarized by each variable,
#   but can also specify multiple columns to summarize conditional on multiple columns.
for (i_col in seq_along(list_var_col_names)) {

  # filename root for each separate variable (will write, then read to compile into one)
  fn_root <- paste0("out_new/tab1_vars_", list_var_col_names[i_col], "_summary")

  # run function (additional options are available)
  e_table1_summaries(
      dat_tab1           = dat_sheet_demo
    , list_var_col_names = list_var_col_names[i_col]
    , list_var_col_labels = list_var_col_labels[i_col]
    , list_var_row_names = list_var_row_names
    , list_var_row_labels = list_var_row_labels
    , fn_root            = fn_root
    , sw_verbose         = TRUE
    )

  # read final table
  tab1_temp <- utils::read.csv(paste0(fn_root, "_", "all", ".csv"))
  # display in Rmd file
  knitr::kable(tab1_temp)
}

## For totals, helpful when missing data because has sample size for each row.
# filename root for each separate variable (will write, then read to compile into one)
fn_root <- paste0("out_new/tab1_vars_", "total", "_summary")

# run function (additional options are available)
e_table1_summaries(
    dat_tab1           = dat_sheet_demo
  , list_var_col_names = NULL
  , list_var_col_labels = "Total"
  , list_var_row_names = list_var_row_names
  , list_var_row_labels = list_var_row_labels
  , fn_root            = fn_root
  , sw_verbose         = TRUE
  )

# read final table
tab1_temp <- utils::read.csv(paste0(fn_root, "_", "all", ".csv"))
# display in Rmd file
knitr::kable(tab1_temp)
```

### Table 2, sample size by biomarker and dx

```{r}
#select columns & replace values with logical
data_table <- dat_sheet %>% select(-ptid,-age,-sex,-demo_edu_yrs)
data_table <- as.data.frame(!is.na(data_table))

#restore dx
data_table$dx_current <- dat_sheet$dx_current

for (i_response in seq_along(list_var_resp)) {
  for (i_marker in seq_along(list_var_covar)) {
    #i_marker = 1
    #i_response = 1

    #frequency table
    n_marker <- list_var_covar[i_marker]
    n_response <- list_var_resp[i_response]
    count <- e_table_sum_freq_prop(data_table,var_names=c(n_response,n_marker,"dx_current"))

    #extract cases where both covar & response exist
    count <- count %>% filter(count[ ,1] == TRUE & count[ ,2]== TRUE)

    #replace "true" with variable name & headers w/ headers wanted for final table
    count[ ,1] <- n_response
    count[ ,2] <- n_marker
    colnames(count)[colnames(count)==n_response]<-"Response"
    colnames(count)[colnames(count)==n_marker]<-"Biomarker"

    #bind
    if (i_marker==1 & i_response==1) {
      tab_sample_size <- count
    } else {
      tab_sample_size <- rbind(tab_sample_size,count)
    }
  }
}

tab_sample_size <- tab_sample_size %>% select(-prop)

tab_sample_size <- pivot_wider(tab_sample_size,names_from = dx_current, values_from = n)

tab_sample_size %>% print()

write_csv(tab_sample_size,"out_new/biomarker_frequency_by_dx.csv")
```

### Table 3, regression coefficients with 95% CIs

```{r}
n_sig <- 3

names_var_resp <-
  c("fl_albumin_index_log2", "mri_perm_3t_tot")

all_summary <- list()
i_list <- 0

for (i_resp in seq_along(names_var_resp)) {
  ## i_resp = 1
  for (i_covar in seq_along(n_row)) {
    ## i_covar = 20

    var_row  <- n_row[i_covar]
    var_col <- names_var_resp[i_resp]

    #exists <- results[[var_col]][[var_row]]
    #if(is.null(exists)) next

    i_list <- i_list + 1
    all_summary[[i_list]] <-
      modelsummary::modelsummary(
        models    = results[[var_col]][[var_row]][["lm_fit_red_BIC"]]
      , output    = "data.frame"
      , estimate  = "{estimate} [{conf.low}, {conf.high}]"
      , fmt       = n_sig #list("estimate" = n_sig, "std.error" = n_sig)
      , statistic = NULL # "conf.int"
      , conf_level = 0.95
      , shape     = term ~ model + statistic
      ) %>%
      dplyr::filter(
        (part == "estimates") | (term == "R2") | (term == "Num.Obs.")
      ) %>%
      dplyr::mutate(
        Y_var = var_col
      , X_var = var_row
      ) %>%
      dplyr::select(
        -part
      ) %>%
      dplyr::select(
        Y_var, X_var, tidyselect::everything()
      )

    # Gender:X interaction, report Male/Female slopes with 95% CI
    if (any(names(coef(results[[var_col]][[var_row]][["lm_fit_red_BIC"]])) == "sexMale:X")) {

      out_contrast <-
        e_plot_model_contrasts(
          fit                     = results[[var_col]][[var_row]][["lm_fit_red_BIC"]]
        , dat_cont                = results[[var_col]][[var_row]][["dat_lm"]]
        , adjust_method           = c("none", "tukey", "scheffe", "sidak", "bonferroni", "dunnettx", "mvt")[1]  # see ?emmeans::summary.emmGrid
        , CI_level                = 0.95
        , sw_print                = FALSE
        )

      out_contrast$text$`sex:X`[[2]][[1]][1]
      out_contrast$text$`sex:X`[[2]][[1]][2]

      tab_est <- out_contrast$tables$`sex:X`$est %>% tibble::as_tibble()
      tab_est$X.trend


      all_summary[[i_list]] <-
        all_summary[[i_list]] %>%
        dplyr::bind_rows(
          tibble::tibble(
            Y_var = all_summary[[i_list]][[ "Y_var" ]][1]
          , X_var = all_summary[[i_list]][[ "X_var" ]][1]
          , term =
              paste0(tab_est$sex, " slope")
          , `Model 1 / Est.` =
              paste0(
                tab_est$X.trend  %>% round(n_sig)
              , " ["
              , tab_est$lower.CL %>% round(n_sig)
              , ", "
              , tab_est$upper.CL %>% round(n_sig)
              , "]"
              )
          )
        )

    }

  } # i_covar
} # i_resp

all_summary <-
  all_summary %>%
  dplyr::bind_rows()

all_summary_wide <-
  all_summary %>%
  tidyr::pivot_wider(
    id_cols     = c(Y_var, X_var)
  , names_from  = term
  #, values_from = `Model 1 / Est.`
  , values_from = `(1) / Est.`
  )

# replace unicode

names_col <-
  colnames(all_summary_wide)

names_col <-
  e_text_unicode_to_ascii(
    text_in = names_col
  ) %>%
  stringr::str_replace(
    pattern     = stringr::fixed(" * ")
  , replacement = ":"
  ) %>%
  stringr::str_replace(
    pattern     = stringr::fixed("(Intercept)")
  , replacement = "Intercept"
  ) %>%
  stringr::str_replace(
    pattern     = "sexMale"
  , replacement = "sex"
  )

for (i_row in seq_len(nrow(dat_var_label_matching))) {
  names_col <-
    names_col %>%
    stringr::str_replace(
      pattern     = stringr::fixed(dat_var_label_matching[i_row, "Var"] %>% as.character())
    , replacement = dat_var_label_matching[i_row, "Label"] %>% as.character()
    )
  all_summary_wide[[ "Y_var" ]] <-
    all_summary_wide[[ "Y_var" ]] %>%
    stringr::str_replace(
      pattern     = stringr::fixed(dat_var_label_matching[i_row, "Var"] %>% as.character())
    , replacement = dat_var_label_matching[i_row, "Label"] %>% as.character()
    )
}

colnames(all_summary_wide) <-
  names_col

all_summary_wide <-
  all_summary_wide %>%
  dplyr::left_join(
    dat_var_label_matching
  , by = c("X_var" = "Var")
  ) %>%
  dplyr::select(
    -X_var
  ) %>%
  dplyr::rename(
    `X variable` = Label
  , `Response variable` = Y_var
  , n = `Num.Obs.`
  ) %>%
  dplyr::select(
    `Response variable`
  , `X variable`
  , Type
  , n
  , R2
  , Intercept
  , Age
  , Gender
  , X
  , `Age:X`
  , `Gender:X`
  , `Female slope`
  , `Male slope`
  ) %>%
  dplyr::arrange(
    `Response variable`
  , Type
  , `X variable`
  )

all_summary_wide %>% print(n=Inf)

readr::write_csv(
    all_summary_wide
  , file = "out_new/Table_2_model_coef_CI.csv"
  , na = ""
  )
```

