---
title: "Multivariate,Triple_dichotomy_refinedaxes_Only_Albuminaxes"
author: "Azadeh"
format: html
editor: visual
---

#Package

```{r}

# remotes::install_github("erikerhardt/erikmisc")

library(ggplot2)
library(dplyr)
library(tidyr)
library(car)
library(broom)
library(emmeans)
library(pROC)
library(rlang)
library(patchwork)
library(erikmisc)
```

#load data

```{r}
 # If needed load dat_all_combine
# load("../../../Data/ADRC_VCI_MarkVCID2_ALL_combine.RData")

load("dat_sub_Data_inflammation_analysis_paper_06052025.RData")
load("dat_sub_and_Var_for_inflammation_analysis.RData")


# 1. outcome variable
y <- c("fl_albumin_index_log2", "mri_perm_3t_tot")

# 2. predictors to scale (everything in CSF / plasma **except** the outcome)
vars_to_scale <- setdiff(c(Var$csf, Var$plasma, Var$mri), y)

dat_sheet <- dat_sub %>% 
  mutate(
    across(all_of(vars_to_scale),
           ~ as.numeric(scale(.x)))
  )        # y stays on its original log2 scale

# 176 subjects , filter on baseline
```

##Do any fluid biomarkers correlate with the VD axis?Do any fluid biomarkers correlate with the AD axis?

Note: We will associate plasma biomarkers or their ratios with CSF as the "ground truth" using the model: CSF AD \~ Plasma x1 + Plasma x2 + Plasma x3

Additionally, we filter data based on the Baseline visit in the RedCap event name. Good news---there is still a substantial amount of data available.

```{r}
library(dplyr)

# Define the baseline visit categories
baseline_visits <- c(
  "Baseline (Arm 1: Patients)",
  "Baseline (Arm 2: Control)",
  "Baseline (Arm 3: CSF controls)",
  "Baseline (Arm 4: Diabetes)",
  "baseline"
)

# Filter dataset based on baseline visits
dat_all_combine_AD_VD <- dat_all_combine |>
  filter(redcap_event_name %in% baseline_visits) |>  # Keep only baseline visits
  mutate(
    TripleDichotomy_VD = mri_dti_psmd_z,
    TripleDichotomy_VD_log2 = e_log_shift(TripleDichotomy_VD),
    TripleDichotomy_AD = (fl_csf__ALL__ptau / fl_csf__ALL__abeta_42),
    TripleDichotomy_AD_log2 = log2(TripleDichotomy_AD),
    fl_plasma__ALL__abeta42_40_ratio_log2 = log2(fl_plasma__ALL__abeta_42/fl_plasma__ALL__abeta_40)
  )
# nrow(dat_all_combine_AD_VD)
# Check unique values after filtering on Baseline 
# unique(dat_all_combine_AD_VD$redcap_event_name)
 # hist(dat_all_combine_AD_VD$TripleDichotomy_VD_log2)
### YES YES YES Now 630 rows!!!! YAY!!!!!!!
 # hist(dat_all_combine_AD_VD$TripleDichotomy_AD, 100)



```

# Find all plasma biomarker combinations with albumin

 Note: we check the  Multivariate
1. CSF: Remove IGG.  "IGG is confusing, the other biomarkers are more interesting."
2. CSF-only, add MMP-3 (even if n < 35)

Still the R-square was low 


Running albumin \~ plasma_only

```{r}

library(randomForest)
library(caret)

# Get only plasma biomarkers ending in _log2
find_optimal_log2_groups <- function(data, min_cases = 30) {
  # Get only plasma markers ending in _log2
  plasma_markers <- grep("^fl_plasma.*_log2$", names(data), value = TRUE)
  
  # Start with outcome variable
  current_vars <- c("fl_albumin_index_log2")
  results <- list()
  
  for(marker in plasma_markers) {
    temp_vars <- c(current_vars, marker)
    complete_cases <- sum(complete.cases(data[, temp_vars]))
    
    if(complete_cases >= min_cases) {
      current_vars <- temp_vars
      results[[length(results) + 1]] <- list(
        variables = current_vars,
        complete_cases = complete_cases
      )
    }
  }
  return(results)
}

# Apply the function
log2_groups <- find_optimal_log2_groups(dat_sub)

# Print summary of available biomarkers and sample sizes
for(i in seq_along(log2_groups)) {
  cat("\nGroup", i, ":\n")
  cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
  cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
}

# Get the final set of all log2 biomarkers
final_log2_vars <- log2_groups[[length(log2_groups)]]$variables

# Print final number of biomarkers and sample size
cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")

# Function to print breakpoint summary
summarize_breakpoints <- function(groups) {
  prev_vars <- NULL
  prev_n <- NULL
  
  cat("Sample Size Breakpoints:\n")
  
  for(i in seq_along(groups)) {
    current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
    current_n <- groups[[i]]$complete_cases
    
    if(is.null(prev_n) || current_n != prev_n) {
      new_vars <- if(is.null(prev_vars)) {
        current_vars
      } else {
        setdiff(current_vars, prev_vars)
      }
      
      # Clean up variable names for display
      clean_vars <- gsub("fl_plasma__ALL__", "", new_vars)
      clean_vars <- gsub("_log2", "", clean_vars)
      
      cat(sprintf("%d cases: %s\n", 
                  current_n,
                  if(i == 1) {
                    sprintf("starting with %s", paste(clean_vars, collapse = " and "))
                  } else {
                    sprintf("adding %s", paste(clean_vars, collapse = " and "))
                  }))
      
      prev_vars <- current_vars
      prev_n <- current_n
    }
  }
}

# Apply the function to your groups
summarize_breakpoints(log2_groups)

# Also show the complete list of final variables
cat("\nFinal set of biomarkers:\n")
final_vars <- gsub("fl_plasma__ALL__", "", 
                  gsub("_log2", "", 
                       log2_groups[[length(log2_groups)]]$variables[-1]))
cat(paste(final_vars, collapse = "\n"))



# Apply RandomForest to check the important varibels for Albumin index with 13 Var and 42 Obervation

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",         
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
 "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,     
 "fl_plasma__ALL__il_1b_log2",
 "fl_plasma__ALL__mmp_2_log2" ,
 "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 "fl_plasma__ALL__mmp_9_log2" ,
 "fl_plasma__ALL__mmp_10_log2",
 # "fl_csf__ALL__mmp_10_log2" ,
 "fl_plasma__ALL__bfgf_log2",
"fl_plasma__ALL__flt_1_log2",
 "fl_plasma__ALL__tie2_log2" ,
 "fl_plasma__ALL__vegf_a_log2",
 "fl_plasma__ALL__vegf_c_log2",
 "fl_plasma__ALL__vegf_d_log2"

)


n_complete <- sum(complete.cases(dat_sheet[, c("fl_albumin_index_log2", biomarkers)]))
print(n_complete) # Only one remained




# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
  var_imp_df <- data.frame(
    variable = gsub("fl_plasma__ALL__", "", gsub("_log2", "", rownames(var_imp))),
    importance = var_imp[, "%IncMSE"],  # Using %IncMSE as importance measure
    stringsAsFactors = FALSE
  )
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub,
  outcome_var = "fl_albumin_index_log2",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  top_vars <- paste0("fl_plasma__ALL__", 
                    importance_df$variable[1:n_vars], 
                    "_log2")
  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

    fml_upper <- as.formula(
  paste("outcome_var ~ (", paste(top_vars, collapse = " + "), ")^2")
)
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 5 variables
top5_model <- create_top_n_model(
  data = dat_sub,
  outcome_var = "fl_albumin_index_log2",
  importance_df = rf_results$importance,
  n_vars = 5
)

# Print results of top 4 model
cat("\nModel with top 5 variables from Random Forest:\n")
summary(top5_model)


# Now base on the result of Random-forest we create new albumin




 
# if you choose this you will get lots of interaction but Rsquare still low around 40%
# 5 Important varibels base on the result of RF IMportance
biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",         
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,     
 # "fl_plasma__ALL__il_1b_log2",
 # "fl_plasma__ALL__mmp_2_log2" ,
 # "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 # "fl_plasma__ALL__mmp_9_log2" ,
 "fl_plasma__ALL__mmp_10_log2",
 # "fl_csf__ALL__mmp_10_log2" ,
 "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
 # "fl_plasma__ALL__tie2_log2" ,
 "fl_plasma__ALL__vegf_a_log2",
 "fl_plasma__ALL__vegf_c_log2"
 # "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("fl_albumin_index_log2", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
albumin_form <- as.formula(
  paste("fl_albumin_index_log2 ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
albumin_main <- lm(albumin_form, data = dat_sheet)

summary(albumin_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_albumin <- as.formula(
  paste("fl_albumin_index_log2 ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_albumin <- as.formula("fl_albumin_index_log2 ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = albumin_main,
  scope   = list(lower = fml_lower_albumin, upper = fml_upper_albumin),
  direction = "both"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)




# Do I need to add plot contrast? regression assumption? add later? 


#─ tidy helpers --------------------------------------------------------------
library(tidyverse)
library(glue)
library(gt)
library(patchwork)
library(scales)

pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")

# 1 Break-point data → nice tibble


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display


# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like

# save
ggsave("albumin_analysis_plasmaOnly.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf

```

Running albumin \~ csf_only …

```{r}

dat_sub|> dplyr::select(-fl_csf_igg_log2) -> dat_sub_without_igg

csf_markers <- dat_sub_without_igg %>%                 # start with your data
  select(                                             # pick columns …
    contains("csf"), 
    fl_albumin_index_log2                             # … and this one explicit name
  ) %>% 
  names()   

result <- e_data_complete_by_variable_subset(dat_sub_without_igg, var_list = csf_markers, var_resp = "fl_albumin_index_log2")
# See all variable names for each row
result$var_names


# OLD WAY also not include everything
# # Get only plasma biomarkers ending in _log2
# find_optimal_log2_groups <- function(data, min_cases = 30) {
#   # Get only plasma markers ending in _log2
#   csf_markers <- grep("^fl_csf.*_log2$", names(data), value = TRUE)
#   
#   # Start with outcome variable
#   current_vars <- c("fl_albumin_index_log2")
#   results <- list()
#   
#   for(marker in csf_markers) {
#     temp_vars <- c(current_vars, marker)
#     complete_cases <- sum(complete.cases(data[, temp_vars]))
#     
#     if(complete_cases >= min_cases) {
#       current_vars <- temp_vars
#       results[[length(results) + 1]] <- list(
#         variables = current_vars,
#         complete_cases = complete_cases
#       )
#     }
#   }
#   return(results)
# }
# 
# # Apply the function
# log2_groups <- find_optimal_log2_groups(dat_sub_without_igg)
# 
# # Print summary of available biomarkers and sample sizes
# for(i in seq_along(log2_groups)) {
#   cat("\nGroup", i, ":\n")
#   cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
#   cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
# }
# 
# # Get the final set of all log2 biomarkers
# final_log2_vars <- log2_groups[[length(log2_groups)]]$variables
# 
# # Print final number of biomarkers and sample size
# cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
# cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")
# 
# # Function to print breakpoint summary
# summarize_breakpoints <- function(groups) {
#   prev_vars <- NULL
#   prev_n <- NULL
#   
#   cat("Sample Size Breakpoints:\n")
#   
#   for(i in seq_along(groups)) {
#     current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
#     current_n <- groups[[i]]$complete_cases
#     
#     if(is.null(prev_n) || current_n != prev_n) {
#       new_vars <- if(is.null(prev_vars)) {
#         current_vars
#       } else {
#         setdiff(current_vars, prev_vars)
#       }
#       
#       # Clean up variable names for display
#       clean_vars <- gsub("fl_csf__ALL__", "", new_vars)
#       clean_vars <- gsub("_log2", "", clean_vars)
#       
#       cat(sprintf("%d cases: %s\n", 
#                   current_n,
#                   if(i == 1) {
#                     sprintf("starting with %s", paste(clean_vars, collapse = " and "))
#                   } else {
#                     sprintf("adding %s", paste(clean_vars, collapse = " and "))
#                   }))
#       
#       prev_vars <- current_vars
#       prev_n <- current_n
#     }
#   }
# }
# 
# # Apply the function to your groups
# summarize_breakpoints(log2_groups)
# 
# # Also show the complete list of final variables
# cat("\nFinal set of biomarkers:\n")
# final_vars <- gsub("fl_csf__ALL__", "", 
#                   gsub("_log2", "", 
#                        log2_groups[[length(log2_groups)]]$variables[-1]))
# cat(paste(final_vars, collapse = "\n"))



# Apply RandomForest to check the important varibels for Albumin index with 7 cases and 61 Observation
# 6        108     5 "fl_albumin_index_log2 fl_msd_csf_mmp_1_log2                       fl_msd_csf_mmp_3_log2 fl_msd_csf_mmp_9_log2 fl_csf__ALL__mmp_10_log2                                                                                                                                                                                          … 
# [1] "fl_albumin_index_log2"    "fl_msd_csf_mmp_1_log2"    "fl_msd_csf_mmp_2_log2"    "fl_msd_csf_mmp_3_log2"    "fl_msd_csf_mmp_9_log2"    "fl_csf__ALL__mmp_10_log2" "fl_csf__ALL__nfl_log2"    "fl_csf__ALL__gfap_log2"  
# [9] "fl_csf__ALL__ptau_log2"   74

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 "fl_csf__ALL__nfl_log2",
 # "fl_plasma__ALL__nfl_log2",
 "fl_csf__ALL__ptau_log2",
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
#  "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,
#  "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2",
 "fl_msd_csf_mmp_1_log2",
 "fl_msd_csf_mmp_3_log2",
 "fl_msd_csf_mmp_9_log2",
 "fl_msd_csf_mmp_2_log2" 
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
# "fl_plasma__ALL__tie2_log2" ,
# "fl_plasma__ALL__vegf_a_log2",
# "fl_plasma__ALL__vegf_c_log2",
# "fl_plasma__ALL__vegf_d_log2"

)




library(randomForest)
library(caret)
library(ggplot2)

# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
 var_imp_df <- data.frame(
  variable = rownames(var_imp),  # Keep original variable names
  variable_clean = gsub("fl_csf__ALL__", "", gsub("_log2", "", rownames(var_imp))),
  importance = var_imp[, "%IncMSE"],
  stringsAsFactors = FALSE
)
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub_without_igg,
  outcome_var = "fl_albumin_index_log2",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  # Use the original variable names from the importance dataframe
  top_vars <- importance_df$variable[1:n_vars]
  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

    fml_upper <- as.formula(
  paste("outcome_var ~ (", paste(top_vars, collapse = " + "), ")^2")
)
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 4 variables
top7_model <- create_top_n_model(
  data = dat_sub_without_igg,
  outcome_var = "fl_albumin_index_log2",
  importance_df = rf_results$importance,
  n_vars = 7
)

# Print results of top 4 model
cat("\nModel with top 7 variables from Random Forest:\n")
summary(top7_model)


# Now base on the result of Random-forest we create new albumin with 4 Var

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",
 # "fl_plasma__ALL__nfl_log2",
 "fl_csf__ALL__ptau_log2",
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
#  "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,
#  "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" 
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" ,
#  "fl_plasma__ALL__vegf_a_log2",
#  "fl_plasma__ALL__vegf_c_log2",
#  "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("fl_albumin_index_log2", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
albumin_form <- as.formula(
  paste("fl_albumin_index_log2 ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
albumin_main <- lm(albumin_form, data = dat_sheet)

summary(albumin_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_albumin <- as.formula(
  paste("fl_albumin_index_log2 ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_albumin <- as.formula("fl_albumin_index_log2 ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = albumin_main,
  scope   = list(lower = fml_lower_albumin, upper = fml_upper_albumin),
  direction = "forward"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)




e_plot_scatterplot( dat_plot = dat_sheet,
  var_x ="fl_csf_igg_log2",
  var_y = "fl_albumin_index_log2",
  smooth_all = c("lm")
  )



pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display



# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display

# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like

# save
ggsave("albumin_analys_onlycsf_overview.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf

```

Running albumin \~ plasma_csf …

```{r}
# 
# # Get only plasma biomarkers ending in _log2
# find_optimal_log2_groups <- function(data, min_cases = 30) {
#   # Get only plasma markers ending in _log2
#   plasma_csf_markers <- grep("^fl_(plasma|csf).*_log2$", names(data), value = TRUE)
#    
#   # Start with outcome variable
#   current_vars <- c("fl_albumin_index_log2")
#   results <- list()
#   
#   for(marker in plasma_csf_markers) {
#     temp_vars <- c(current_vars, marker)
#     complete_cases <- sum(complete.cases(data[, temp_vars]))
#     
#     if(complete_cases >= min_cases) {
#       current_vars <- temp_vars
#       results[[length(results) + 1]] <- list(
#         variables = current_vars,
#         complete_cases = complete_cases
#       )
#     }
#   }
#   return(results)
# }
# 
# # Apply the function
# log2_groups <- find_optimal_log2_groups(dat_sub)
# 
# # Print summary of available biomarkers and sample sizes
# for(i in seq_along(log2_groups)) {
#   cat("\nGroup", i, ":\n")
#   cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
#   cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
# }
# 
# # Get the final set of all log2 biomarkers
# final_log2_vars <- log2_groups[[length(log2_groups)]]$variables
# 
# # Print final number of biomarkers and sample size
# cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
# cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")
# 
# # Function to print breakpoint summary
# summarize_breakpoints <- function(groups) {
#   prev_vars <- NULL
#   prev_n <- NULL
#   
#   cat("Sample Size Breakpoints:\n")
#   
#   for(i in seq_along(groups)) {
#     current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
#     current_n <- groups[[i]]$complete_cases
#     
#     if(is.null(prev_n) || current_n != prev_n) {
#       new_vars <- if(is.null(prev_vars)) {
#         current_vars
#       } else {
#         setdiff(current_vars, prev_vars)
#       }
#       
#       # Separate CSF and plasma variables
#       csf_vars <- grep("csf", new_vars, value = TRUE)
#       plasma_vars <- grep("plasma", new_vars, value = TRUE)
#       
#       # Clean up variable names for display but keep CSF/Plasma indicator
#       clean_vars <- gsub("^fl_(plasma|csf)__ALL__", "\\1: ", new_vars)
#       clean_vars <- gsub("^fl_csf_", "csf: ", clean_vars)
#       clean_vars <- gsub("^fl_plasma_", "plasma: ", clean_vars)
#       clean_vars <- gsub("_log2", "", clean_vars)
#       
#       cat(sprintf("%d cases: %s\n", 
#                   current_n,
#                   if(i == 1) {
#                     sprintf("starting with %s", paste(clean_vars, collapse = " and "))
#                   } else {
#                     sprintf("adding %s", paste(clean_vars, collapse = " and "))
#                   }))
#       
#       prev_vars <- current_vars
#       prev_n <- current_n
#     }
#   }
# }
# 
# # Apply the function to your groups
# summarize_breakpoints(log2_groups)
# 
# # Also show the complete list of final variables
# cat("\nFinal set of biomarkers:\n")
# 
# # Separate CSF and plasma for final display
# final_vars <- log2_groups[[length(log2_groups)]]$variables[-1]
# csf_vars <- grep("csf", final_vars, value = TRUE)
# plasma_vars <- grep("plasma", final_vars, value = TRUE)
# 
# cat("\nCSF biomarkers (", length(csf_vars), "):\n", sep = "")
# csf_clean <- gsub("^fl_(csf)__ALL__", "", csf_vars)
# csf_clean <- gsub("^fl_csf_", "", csf_clean)
# csf_clean <- gsub("_log2", "", csf_clean)
# cat(paste(csf_clean, collapse = "\n"))
# 
# cat("\n\nPlasma biomarkers (", length(plasma_vars), "):\n", sep = "")
# plasma_clean <- gsub("^fl_(plasma)__ALL__", "", plasma_vars)
# plasma_clean <- gsub("^fl_plasma_", "", plasma_clean)
# plasma_clean <- gsub("_log2", "", plasma_clean)
# cat(paste(plasma_clean, collapse = "\n"))



dat_sub|> dplyr::select(-fl_csf_igg_log2) -> dat_sub_without_igg

csf_plasma_markers <- dat_sub_without_igg |>
  select(
    contains("csf"),
    contains("plasma"),
    fl_albumin_index_log2
  ) %>%
  names()
result <- e_data_complete_by_variable_subset(dat_sub_without_igg, var_list = csf_plasma_markers, var_resp = "fl_albumin_index_log2")
# See all variable names for each row
result$var_names[8]


# Apply RandomForest to check the important varibels for Albumin index with 40 cases both plasma and csf
# [1] "fl_albumin_index_log2"      "fl_msd_csf_mmp_1_log2"      "fl_msd_csf_mmp_3_log2"      "fl_msd_csf_mmp_9_log2"      "fl_csf__ALL__mmp_10_log2"   "fl_plasma__ALL__mmp_2_log2"


 biomarkers<- c(
  "fl_msd_csf_mmp_1_log2" ,
  "fl_msd_csf_mmp_3_log2" ,
  "fl_msd_csf_mmp_9_log2"  ,
  "fl_csf__ALL__mmp_10_log2",
  "fl_plasma__ALL__mmp_2_log2"
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",
 # # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
#  "fl_csf__ALL__plgf_log2",
#  # "fl_plasma__ALL__plgf_log2" ,
#  "fl_csf__ALL__il_1b_log2" ,
#  # "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
#  "fl_csf__ALL__mmp_10_log2" ,
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" 
 # "fl_plasma__ALL__vegf_a_log2",
 # "fl_plasma__ALL__vegf_c_log2",
 # "fl_plasma__ALL__vegf_d_log2"

)



n_complete <- sum(complete.cases(dat_sub_without_igg[, c("fl_albumin_index_log2", biomarkers)]))
print(n_complete) # Only one remained

library(randomForest)
library(caret)
library(ggplot2)

# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
  var_imp_df <- data.frame(
  variable = rownames(var_imp),  # CHANGE: Keep original names
  variable_clean = gsub("^fl_(plasma|csf)__ALL__", "", gsub("_log2", "", rownames(var_imp))), # For display
  importance = var_imp[, "%IncMSE"],
  stringsAsFactors = FALSE
)
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub_without_igg,
  outcome_var = "fl_albumin_index_log2",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  top_vars <- top_vars <- importance_df$variable[1:n_vars]

  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

  fml_upper <- as.formula(
  paste(outcome_var, "~ (", paste(top_vars, collapse = " + "), ")^2")
  )
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 7 variables
top3_model <- create_top_n_model(
  data = dat_sub_without_igg,
  outcome_var = "fl_albumin_index_log2",
  importance_df = rf_results$importance,
  n_vars = 3
)

# Print results of top 4 model
cat("\nModel with top 3 variables from Random Forest:\n")
summary(top3_model)


# Now base on the result of Random-forest we create new albumin



library(dplyr)

# 1. Extract columns that start with either "fl_plasma_" or "fl_csf_"
# biomarker_cols <- dat_all_combine %>%
#   dplyr::select(matches("^(fl_plasma_|fl_csf_|fl_simoa_|fl_msd_|fl_abeta_).*log2$")) %>%
#   names()


# 
# biomarker_cols <- dat_all_combine_AD_VD %>%
#   dplyr::select(matches("^(fl_plasma_|fl_csf_|fl_abeta_38).*log2$")) %>%
#   names()
# 
# biomarker_cols_final <- biomarker_cols[ !grepl("original", biomarker_cols) ]
# 
# # Number of complete rows for the variables of interest:
# # n_complete <- sum(complete.cases(dat_all_combine[, c("TripleDichotomy_VD_log2", biomarker_cols)]))
# # print(n_complete) # Only one remained
# 
# # how many rows remain for each biomarker individually:
# sapply(biomarker_cols, function(x) sum(!is.na(dat_all_combine_AD_VD[[x]])))
# 
# e_data_complete_by_variable_subset(dat_all_combine_AD_VD, var_list = c(some_biomarkers_vd, "fl_albumin_index_log2"), var_resp = "fl_albumin_index_log2")
 
# if you choose this you will get lots of interaction but Rsquare still low around 40%
# 5 Important varibels base on the result of RF IMportance
biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",         
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 "fl_csf__ALL__il_1b_log2" ,
 # "fl_plasma__ALL__il_1b_log2",
 # "fl_plasma__ALL__mmp_2_log2" ,
 # "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 # "fl_plasma__ALL__mmp_9_log2" ,
 # "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" ,
 # "fl_plasma__ALL__bfgf_log2",
"fl_plasma__ALL__flt_1_log2",
 "fl_plasma__ALL__tie2_log2" 
 # "fl_plasma__ALL__vegf_a_log2",
 # "fl_plasma__ALL__vegf_c_log2"
 # "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("fl_albumin_index_log2", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
albumin_form <- as.formula(
  paste("fl_albumin_index_log2 ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
albumin_main <- lm(albumin_form, data = dat_sheet)

summary(albumin_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_albumin <- as.formula(
  paste("fl_albumin_index_log2 ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_albumin <- as.formula("fl_albumin_index_log2 ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = albumin_main,
  scope   = list(lower = fml_lower_albumin, upper = fml_upper_albumin),
  direction = "forward"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)



pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")

# 1 Break-point data → nice tibble


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display


# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like

# save
ggsave("albumin_analys_csf&plasma_overview.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf




```

Running mri3t \~ plasma_only …

```{r}

# Get only plasma biomarkers ending in _log2
find_optimal_log2_groups <- function(data, min_cases = 30) {
  # Get only plasma markers ending in _log2
  plasma_markers <- grep("^fl_plasma.*_log2$", names(data), value = TRUE)
  
  # Start with outcome variable
  current_vars <- c("mri_perm_3t_tot")
  results <- list()
  
  for(marker in plasma_markers) {
    temp_vars <- c(current_vars, marker)
    complete_cases <- sum(complete.cases(data[, temp_vars]))
    
    if(complete_cases >= min_cases) {
      current_vars <- temp_vars
      results[[length(results) + 1]] <- list(
        variables = current_vars,
        complete_cases = complete_cases
      )
    }
  }
  return(results)
}

# Apply the function
log2_groups <- find_optimal_log2_groups(dat_sub)

# Print summary of available biomarkers and sample sizes
for(i in seq_along(log2_groups)) {
  cat("\nGroup", i, ":\n")
  cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
  cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
}

# Get the final set of all log2 biomarkers
final_log2_vars <- log2_groups[[length(log2_groups)]]$variables

# Print final number of biomarkers and sample size
cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")

# Function to print breakpoint summary
summarize_breakpoints <- function(groups) {
  prev_vars <- NULL
  prev_n <- NULL
  
  cat("Sample Size Breakpoints:\n")
  
  for(i in seq_along(groups)) {
    current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
    current_n <- groups[[i]]$complete_cases
    
    if(is.null(prev_n) || current_n != prev_n) {
      new_vars <- if(is.null(prev_vars)) {
        current_vars
      } else {
        setdiff(current_vars, prev_vars)
      }
      
      # Clean up variable names for display
      clean_vars <- gsub("fl_plasma__ALL__", "", new_vars)
      clean_vars <- gsub("_log2", "", clean_vars)
      
      cat(sprintf("%d cases: %s\n", 
                  current_n,
                  if(i == 1) {
                    sprintf("starting with %s", paste(clean_vars, collapse = " and "))
                  } else {
                    sprintf("adding %s", paste(clean_vars, collapse = " and "))
                  }))
      
      prev_vars <- current_vars
      prev_n <- current_n
    }
  }
}

# Apply the function to your groups
summarize_breakpoints(log2_groups)

# Also show the complete list of final variables
cat("\nFinal set of biomarkers:\n")
final_vars <- gsub("fl_plasma__ALL__", "", 
                  gsub("_log2", "", 
                       log2_groups[[length(log2_groups)]]$variables[-1]))
cat(paste(final_vars, collapse = "\n"))



# Apply RandomForest to check the important Var base on 10 var and 54 Observation



 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",         
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,     
 # "fl_plasma__ALL__il_1b_log2",
 "fl_plasma__ALL__mmp_2_log2" ,
 "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 "fl_plasma__ALL__mmp_9_log2" ,
 "fl_plasma__ALL__mmp_10_log2",
 # "fl_csf__ALL__mmp_10_log2" ,
 "fl_plasma__ALL__bfgf_log2",
"fl_plasma__ALL__flt_1_log2",
 "fl_plasma__ALL__tie2_log2" ,
 "fl_plasma__ALL__vegf_a_log2",
 "fl_plasma__ALL__vegf_c_log2"
 # "fl_plasma__ALL__vegf_d_log2"

)




library(randomForest)
library(caret)
library(ggplot2)

# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
  var_imp_df <- data.frame(
    variable = gsub("fl_plasma__ALL__", "", gsub("_log2", "", rownames(var_imp))),
    importance = var_imp[, "%IncMSE"],  # Using %IncMSE as importance measure
    stringsAsFactors = FALSE
  )
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub,
  outcome_var = "mri_perm_3t_tot",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  top_vars <- paste0("fl_plasma__ALL__", 
                    importance_df$variable[1:n_vars], 
                    "_log2")
  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

    fml_upper <- as.formula(
  paste("outcome_var ~ (", paste(top_vars, collapse = " + "), ")^2")
)
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 4 variables
top6_model <- create_top_n_model(
  data = dat_sub,
  outcome_var = "mri_perm_3t_tot",
  importance_df = rf_results$importance,
  n_vars = 6
)

# Print results of top 4 model
cat("\nModel with top 5 variables from Random Forest:\n")
summary(top6_model)


# Now base on the result of Random-forest we create new albumin



# if you choose this you will get lots of interaction but Rsquare still low around 40%
# 5 Important varibels base on the result of RF IMportance
biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",         
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,     
 # "fl_plasma__ALL__il_1b_log2",
 "fl_plasma__ALL__mmp_2_log2" ,
 "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 # "fl_plasma__ALL__mmp_9_log2" ,
 # "fl_plasma__ALL__mmp_10_log2",
 # "fl_csf__ALL__mmp_10_log2" ,
 "fl_plasma__ALL__bfgf_log2",
"fl_plasma__ALL__flt_1_log2",
 # "fl_plasma__ALL__tie2_log2" ,
 # "fl_plasma__ALL__vegf_a_log2",
 "fl_plasma__ALL__vegf_c_log2"
 # "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("mri_perm_3t_tot", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
mri_form <- as.formula(
  paste("mri_perm_3t_tot ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
mri_main <- lm(mri_form, data = dat_sheet)

summary(mri_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_mri <- as.formula(
  paste("mri_perm_3t_tot ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_mri <- as.formula("mri_perm_3t_tot ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = mri_main,
  scope   = list(lower = fml_lower_mri, upper = fml_upper_mri),
  direction = "forward"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)




pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")

# 1 Break-point data → nice tibble


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display


# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like

# save
ggsave("mri_analys_onlyplasma_overview.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf
```

Running mri3t \~ csf_only …

```{r}
# Get only plasma biomarkers ending in _log2
# find_optimal_log2_groups <- function(data, min_cases = 30) {
#   # Get only plasma markers ending in _log2
#   csf_markers <- grep("^fl_csf.*_log2$", names(data), value = TRUE)
#   
#   # Start with outcome variable
#   current_vars <- c("mri_perm_3t_tot")
#   results <- list()
#   
#   for(marker in csf_markers) {
#     temp_vars <- c(current_vars, marker)
#     complete_cases <- sum(complete.cases(data[, temp_vars]))
#     
#     if(complete_cases >= min_cases) {
#       current_vars <- temp_vars
#       results[[length(results) + 1]] <- list(
#         variables = current_vars,
#         complete_cases = complete_cases
#       )
#     }
#   }
#   return(results)
# }
# 
# # Apply the function
# log2_groups <- find_optimal_log2_groups(dat_sub)
# 
# # Print summary of available biomarkers and sample sizes
# for(i in seq_along(log2_groups)) {
#   cat("\nGroup", i, ":\n")
#   cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
#   cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
# }
# 
# # Get the final set of all log2 biomarkers
# final_log2_vars <- log2_groups[[length(log2_groups)]]$variables
# 
# # Print final number of biomarkers and sample size
# cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
# cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")
# 
# # Function to print breakpoint summary
# summarize_breakpoints <- function(groups) {
#   prev_vars <- NULL
#   prev_n <- NULL
#   
#   cat("Sample Size Breakpoints:\n")
#   
#   for(i in seq_along(groups)) {
#     current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
#     current_n <- groups[[i]]$complete_cases
#     
#     if(is.null(prev_n) || current_n != prev_n) {
#       new_vars <- if(is.null(prev_vars)) {
#         current_vars
#       } else {
#         setdiff(current_vars, prev_vars)
#       }
#       
#       # Clean up variable names for display
#       clean_vars <- gsub("fl_csf__ALL__", "", new_vars)
#       clean_vars <- gsub("_log2", "", clean_vars)
#       
#       cat(sprintf("%d cases: %s\n", 
#                   current_n,
#                   if(i == 1) {
#                     sprintf("starting with %s", paste(clean_vars, collapse = " and "))
#                   } else {
#                     sprintf("adding %s", paste(clean_vars, collapse = " and "))
#                   }))
#       
#       prev_vars <- current_vars
#       prev_n <- current_n
#     }
#   }
# }
# 
# # Apply the function to your groups
# summarize_breakpoints(log2_groups)
# 
# # Also show the complete list of final variables
# cat("\nFinal set of biomarkers:\n")
# final_vars <- gsub("fl_csf__ALL__", "", 
#                   gsub("_log2", "", 
#                        log2_groups[[length(log2_groups)]]$variables[-1]))
# cat(paste(final_vars, collapse = "\n"))
# 



dat_sub|> dplyr::select(-fl_csf_igg_log2) -> dat_sub_without_igg

csf_mri_markers <- dat_sub_without_igg |>
  select(
    contains("csf"),
    mri_perm_3t_tot
  ) %>%
  names()
result <- e_data_complete_by_variable_subset(dat_sub_without_igg, var_list = csf_mri_markers, var_resp = "mri_perm_3t_tot")
# See all variable names for each row
result$var_names[5]


# Apply RandomForest to check the important varibels for mri, 7 cases and 38 Observation

# [[1]]
# [1] "mri_perm_3t_tot"          "fl_msd_csf_mmp_1_log2"    "fl_msd_csf_mmp_2_log2"    "fl_msd_csf_mmp_3_log2"    "fl_msd_csf_mmp_9_log2"    "fl_csf__ALL__mmp_10_log2" "fl_csf__ALL__ptau_log2"  
# 51

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
#  "fl_csf_igg_log2",
#  # "fl_csf_albumin_log2",           
#  # "fl_csf_igg_synthesis_log2",     
#  # "fl_abeta_38_log2",
#  "fl_csf__ALL__gfap_log2",
#  # "fl_plasma__ALL__gfap_log2",    
#  "fl_csf__ALL__nfl_log2",
#  # "fl_plasma__ALL__nfl_log2",
#  "fl_csf__ALL__ptau_log2",
#  # "fl_plasma__ALL__ptau_181_log2",
#  # "fl_plasma__ALL__abeta42_40_ratio_log2",
#  # "fl_plasma__ALL__ptau_217_log2",
#  # "fl_csf__ALL__abeta_40_log2",   
#  # "fl_plasma__ALL__abeta_40_log2",
#  # # "fl_csf__ALL__abeta_42_log2" ,   
#  # "fl_plasma__ALL__abeta_42_log2",
#  "fl_csf__ALL__plgf_log2",
# #  "fl_plasma__ALL__plgf_log2" ,
#  "fl_csf__ALL__il_1b_log2" ,
#  "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" ,
 "fl_msd_csf_mmp_1_log2",
 "fl_msd_csf_mmp_2_log2",
 "fl_msd_csf_mmp_3_log2",
 "fl_msd_csf_mmp_9_log2",
 "fl_csf__ALL__ptau_log2"
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" ,
#  "fl_plasma__ALL__vegf_a_log2",
#  "fl_plasma__ALL__vegf_c_log2",
#  "fl_plasma__ALL__vegf_d_log2"

)




library(randomForest)
library(caret)
library(ggplot2)

# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
 var_imp_df <- data.frame(
  variable = rownames(var_imp),  # Keep original variable names
  variable_clean = gsub("fl_csf__ALL__", "", gsub("_log2", "", rownames(var_imp))),
  importance = var_imp[, "%IncMSE"],
  stringsAsFactors = FALSE
)
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub_without_igg,
  outcome_var = "mri_perm_3t_tot",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  # Use the original variable names from the importance dataframe
  top_vars <- importance_df$variable[1:n_vars]
  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

    fml_upper <- as.formula(
  paste("outcome_var ~ (", paste(top_vars, collapse = " + "), ")^2")
)
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 4 variables
top2_model <- create_top_n_model(
  data = dat_sub_without_igg,
  outcome_var = "mri_perm_3t_tot",
  importance_df = rf_results$importance,
  n_vars = 2
)

# Print results of top 4 model
cat("\nModel with top 2 variables from Random Forest:\n")
summary(top2_model)


# Now base on the result of Random-forest we create new albumin



library(dplyr)

# 1. Extract columns that start with either "fl_plasma_" or "fl_csf_"
# biomarker_cols <- dat_all_combine %>%
#   dplyr::select(matches("^(fl_plasma_|fl_csf_|fl_simoa_|fl_msd_|fl_abeta_).*log2$")) %>%
#   names()




 
# if you choose this you will get lots of interaction but Rsquare still low around 40%
# 5 Important varibels base on the result of RF IMportance

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 "fl_csf__ALL__nfl_log2",
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
#  "fl_plasma__ALL__plgf_log2" ,
 "fl_csf__ALL__il_1b_log2" ,
#  "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" 
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" ,
#  "fl_plasma__ALL__vegf_a_log2",
#  "fl_plasma__ALL__vegf_c_log2",
#  "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("mri_perm_3t_tot", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
mri_form <- as.formula(
  paste("mri_perm_3t_tot ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
mri_main <- lm(mri_form, data = dat_sheet)

summary(mri_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_mri <- as.formula(
  paste("mri_perm_3t_tot ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_mri <- as.formula("mri_perm_3t_tot ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = mri_main,
  scope   = list(lower = fml_lower_mri, upper = fml_upper_mri),
  direction = "forward"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)


pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")

# 1 Break-point data → nice tibble


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display


# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like

# save
ggsave("mri_analys_onlycsf_overview.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf

```

Running mri3t \~ plasma_csf …

```{r}


# Get only plasma biomarkers ending in _log2
# find_optimal_log2_groups <- function(data, min_cases = 30) {
#   # Get only plasma markers ending in _log2
#   plasma_csf_markers <- grep("^fl_(plasma|csf).*_log2$", names(data), value = TRUE)
#    
#   # Start with outcome variable
#   current_vars <- c("mri_perm_3t_tot")
#   results <- list()
#   
#   for(marker in plasma_csf_markers) {
#     temp_vars <- c(current_vars, marker)
#     complete_cases <- sum(complete.cases(data[, temp_vars]))
#     
#     if(complete_cases >= min_cases) {
#       current_vars <- temp_vars
#       results[[length(results) + 1]] <- list(
#         variables = current_vars,
#         complete_cases = complete_cases
#       )
#     }
#   }
#   return(results)
# }
# 
# # Apply the function
# log2_groups <- find_optimal_log2_groups(dat_sub)
# 
# # Print summary of available biomarkers and sample sizes
# for(i in seq_along(log2_groups)) {
#   cat("\nGroup", i, ":\n")
#   cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
#   cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
# }
# 
# # Get the final set of all log2 biomarkers
# final_log2_vars <- log2_groups[[length(log2_groups)]]$variables
# 
# # Print final number of biomarkers and sample size
# cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
# cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")
# 
# # Function to print breakpoint summary
# summarize_breakpoints <- function(groups) {
#   prev_vars <- NULL
#   prev_n <- NULL
#   
#   cat("Sample Size Breakpoints:\n")
#   
#   for(i in seq_along(groups)) {
#     current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
#     current_n <- groups[[i]]$complete_cases
#     
#     if(is.null(prev_n) || current_n != prev_n) {
#       new_vars <- if(is.null(prev_vars)) {
#         current_vars
#       } else {
#         setdiff(current_vars, prev_vars)
#       }
#       
#       # Separate CSF and plasma variables
#       csf_vars <- grep("csf", new_vars, value = TRUE)
#       plasma_vars <- grep("plasma", new_vars, value = TRUE)
#       
#       # Clean up variable names for display but keep CSF/Plasma indicator
#       clean_vars <- gsub("^fl_(plasma|csf)__ALL__", "\\1: ", new_vars)
#       clean_vars <- gsub("^fl_csf_", "csf: ", clean_vars)
#       clean_vars <- gsub("^fl_plasma_", "plasma: ", clean_vars)
#       clean_vars <- gsub("_log2", "", clean_vars)
#       
#       cat(sprintf("%d cases: %s\n", 
#                   current_n,
#                   if(i == 1) {
#                     sprintf("starting with %s", paste(clean_vars, collapse = " and "))
#                   } else {
#                     sprintf("adding %s", paste(clean_vars, collapse = " and "))
#                   }))
#       
#       prev_vars <- current_vars
#       prev_n <- current_n
#     }
#   }
# }
# 
# # Apply the function to your groups
# summarize_breakpoints(log2_groups)
# 
# # Also show the complete list of final variables
# cat("\nFinal set of biomarkers:\n")
# 
# # Separate CSF and plasma for final display
# final_vars <- log2_groups[[length(log2_groups)]]$variables[-1]
# csf_vars <- grep("csf", final_vars, value = TRUE)
# plasma_vars <- grep("plasma", final_vars, value = TRUE)
# 
# cat("\nCSF biomarkers (", length(csf_vars), "):\n", sep = "")
# csf_clean <- gsub("^fl_(csf)__ALL__", "", csf_vars)
# csf_clean <- gsub("^fl_csf_", "", csf_clean)
# csf_clean <- gsub("_log2", "", csf_clean)
# cat(paste(csf_clean, collapse = "\n"))
# 
# cat("\n\nPlasma biomarkers (", length(plasma_vars), "):\n", sep = "")
# plasma_clean <- gsub("^fl_(plasma)__ALL__", "", plasma_vars)
# plasma_clean <- gsub("^fl_plasma_", "", plasma_clean)
# plasma_clean <- gsub("_log2", "", plasma_clean)
# cat(paste(plasma_clean, collapse = "\n"))
# 

dat_sub|> dplyr::select(-fl_csf_igg_log2) -> dat_sub_without_igg

csf_plasma_mri_markers <- dat_sub_without_igg |>
  select(
    contains("csf"),
    contains("plasma"),
    mri_perm_3t_tot
  ) %>%
  names()
result <- e_data_complete_by_variable_subset(dat_sub_without_igg, var_list = csf_plasma_mri_markers, var_resp = "mri_perm_3t_tot")
# See all variable names for each row
result$var_names[2]



# Apply RandomForest to check the important varibels for mri index with 35 cases both plasma and csf

# [[1]]
# [1] "mri_perm_3t_tot"            "fl_msd_csf_mmp_1_log2"      "fl_msd_csf_mmp_3_log2"      "fl_msd_csf_mmp_9_log2"      "fl_plasma__ALL__mmp_1_log2" "fl_plasma__ALL__mmp_3_log2" "fl_plasma__ALL__mmp_9_log2"
# 60

 biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 # "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 # "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 # "fl_csf__ALL__nfl_log2",
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 # "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 # "fl_csf__ALL__il_1b_log2" ,
 # "fl_plasma__ALL__il_1b_log2",
 # "fl_plasma__ALL__mmp_2_log2" ,
 "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 "fl_plasma__ALL__mmp_9_log2" ,
"fl_msd_csf_mmp_1_log2",
"fl_msd_csf_mmp_3_log2",
"fl_msd_csf_mmp_9_log2"
#  "fl_plasma__ALL__mmp_10_log2",
 # "fl_csf__ALL__mmp_10_log2" 
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" 
 # "fl_plasma__ALL__vegf_a_log2",
 # "fl_plasma__ALL__vegf_c_log2",
 # "fl_plasma__ALL__vegf_d_log2"

)




library(randomForest)
library(caret)
library(ggplot2)

# Function to run Random Forest analysis
run_rf_analysis <- function(data, outcome_var, predictors, n_trees = 500) {
  # Prepare data
  model_data <- na.omit(data[, c(outcome_var, predictors)])
  
  # Create formula
  rf_formula <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))
  
  # Run Random Forest
  set.seed(123)  # for reproducibility
  rf_model <- randomForest(
    formula = rf_formula,
    data = model_data,
    ntree = n_trees,
    importance = TRUE
  )

    # Get variable importance
  var_imp <- importance(rf_model)
  var_imp_df <- data.frame(
  variable = rownames(var_imp),  # CHANGE: Keep original names
  variable_clean = gsub("^fl_(plasma|csf)__ALL__", "", gsub("_log2", "", rownames(var_imp))), # For display
  importance = var_imp[, "%IncMSE"],
  stringsAsFactors = FALSE
   )
  var_imp_df <- var_imp_df[order(-var_imp_df$importance), ]
  
  # Calculate model performance
  predictions <- predict(rf_model) #If no new data is provided, it predicts for the training data.
  rmse <- sqrt(mean((model_data[[outcome_var]] - predictions)^2))
  r2 <- 1 - sum((model_data[[outcome_var]] - predictions)^2) / 
            sum((model_data[[outcome_var]] - mean(model_data[[outcome_var]]))^2)
  
  return(list(
    model = rf_model,
    importance = var_imp_df,
    performance = list(
      rmse = rmse,
      r2 = r2,
      n_obs = nrow(model_data)
    )
  ))
}

# Run Random Forest analysis
rf_results <- run_rf_analysis(
  data = dat_sub_without_igg,
  outcome_var = "mri_perm_3t_tot",
  predictors = biomarkers
)

# Print results
cat("\nRandom Forest Model Performance:\n")
cat("R-squared:", round(rf_results$performance$r2, 3), "\n")
cat("RMSE:", round(rf_results$performance$rmse, 3), "\n")
cat("Number of observations:", rf_results$performance$n_obs, "\n")

# Print variable importance
cat("\nVariable Importance (top 10):\n")
print(head(rf_results$importance, 10))

# Plot variable importance
ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# Function to create models with top N variables
create_top_n_model <- function(data, outcome_var, importance_df, n_vars) {
  top_vars <- top_vars <- importance_df$variable[1:n_vars]

  
  # Create formula with interactions
  formula <- as.formula(paste(outcome_var, "~", 
                            paste(top_vars, collapse = " + ")))
  
  # Fit model
  model_data <- na.omit(data[, c(outcome_var, top_vars)])
  model <- lm(formula, data = model_data)

    fml_upper <- as.formula(
  paste(outcome_var, "~ (", paste(top_vars, collapse = " + "), ")^2")
)
 
     fml_lower <- as.formula("outcome_var ~ 1")
  
  # 6. Perform stepwise selection

     model_selected <- step(
  object  = model,
  scope   = list(lower = fml_lower, upper = fml_upper),
  direction = "both",
  trace = 0  # Suppress stepwise output
)

  return(model_selected)
}

# Create models with top 4 variables
top5_model <- create_top_n_model(
  data = dat_sub_without_igg,
  outcome_var = "mri_perm_3t_tot",
  importance_df = rf_results$importance,
  n_vars = 5
)

# Print results of top 4 model
cat("\nModel with top 5 variables from Random Forest:\n")
summary(top5_model)


# Now base on the result of Random-forest we create new albumin



library(dplyr)

# 1. Extract columns that start with either "fl_plasma_" or "fl_csf_"
# biomarker_cols <- dat_all_combine %>%
#   dplyr::select(matches("^(fl_plasma_|fl_csf_|fl_simoa_|fl_msd_|fl_abeta_).*log2$")) %>%
#   names()


# 
# biomarker_cols <- dat_all_combine_AD_VD %>%
#   dplyr::select(matches("^(fl_plasma_|fl_csf_|fl_abeta_38).*log2$")) %>%
#   names()
# 
# biomarker_cols_final <- biomarker_cols[ !grepl("original", biomarker_cols) ]
# 
# # Number of complete rows for the variables of interest:
# # n_complete <- sum(complete.cases(dat_all_combine[, c("TripleDichotomy_VD_log2", biomarker_cols)]))
# # print(n_complete) # Only one remained
# 
# # how many rows remain for each biomarker individually:
# sapply(biomarker_cols, function(x) sum(!is.na(dat_all_combine_AD_VD[[x]])))
# 
# e_data_complete_by_variable_subset(dat_all_combine_AD_VD, var_list = c(some_biomarkers_vd, "fl_albumin_index_log2"), var_resp = "fl_albumin_index_log2")
 
# if you choose this you will get lots of interaction but Rsquare still low around 40%
# 5 Important varibels base on the result of RF IMportance
biomarkers<- c(
 # "fl_csf_glucose_log2",
 # "fl_csf_total_protein_log2",
 # "fl_csf_rbc_log2",
 # "fl_csf_tnc_log2",
 "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",           
 # "fl_csf_igg_synthesis_log2",     
 # "fl_abeta_38_log2",
 "fl_csf__ALL__gfap_log2",
 # "fl_plasma__ALL__gfap_log2",    
 "fl_csf__ALL__nfl_log2",
 # "fl_plasma__ALL__nfl_log2",
 # "fl_csf__ALL__ptau_log2",        
 # "fl_plasma__ALL__ptau_181_log2",
 # "fl_plasma__ALL__abeta42_40_ratio_log2",
 # "fl_plasma__ALL__ptau_217_log2",
 # "fl_csf__ALL__abeta_40_log2",   
 # "fl_plasma__ALL__abeta_40_log2",
 # # "fl_csf__ALL__abeta_42_log2" ,   
 # "fl_plasma__ALL__abeta_42_log2",
 "fl_csf__ALL__plgf_log2",
 # "fl_plasma__ALL__plgf_log2" ,
 "fl_csf__ALL__il_1b_log2" ,
 # "fl_plasma__ALL__il_1b_log2",
 "fl_plasma__ALL__mmp_2_log2" ,
 # "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
 # "fl_plasma__ALL__mmp_9_log2" ,
 # "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" 
 # "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" 
 # "fl_plasma__ALL__vegf_a_log2",
 # "fl_plasma__ALL__vegf_c_log2"
 # "fl_plasma__ALL__vegf_d_log2"

)




n_complete <- sum(complete.cases(dat_sheet[, c("mri_perm_3t_tot", biomarkers)]))
print(n_complete) # Only one remained

# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_sheet[[x]])))



# 2. Build a formula: TripleDichotomy_VD ~ biomarker1 + biomarker2 + ...
mri_form <- as.formula(
  paste("mri_perm_3t_tot ~", paste(biomarkers, collapse = " + "))
)

# Fit main-effects model
mri_main <- lm(mri_form, data = dat_sheet)

summary(mri_main)


# car::Anova(mod_main, type = 3)
# # 
# e_plot_lm_diagnostics(mod_main, sw_plot_set = "simple")
# 
# # Perform Shapiro-Wilk test on the residuals
# shapiro_test <- shapiro.test(residuals(mod_main))


## -- 2. Define the 'upper' and 'lower' scopes for stepwise selection
#    Upper includes two-way interactions: (x1 + x2 + ...)^2
#    Lower is just an intercept-only model: ~ 1

fml_upper_mri <- as.formula(
  paste("mri_perm_3t_tot ~ (", paste(biomarkers, collapse = " + "), ")^2")
)
fml_lower_mri <- as.formula("mri_perm_3t_tot ~ 1")

## -- 3. Run stepwise selection
model_selected_albumin <- step(
  object  = mri_main,
  scope   = list(lower = fml_lower_mri, upper = fml_upper_mri),
  direction = "forward"
)



## -- 4. Inspect the final model
summary(model_selected_albumin)



pal_okabe <- c("#E69F00", "#56B4E9", "#009E73",
               "#F0E442", "#0072B2", "#D55E00",
               "#CC79A7", "#999999")

# 1 Break-point data → nice tibble


# turn your `log2_groups` list into a tidy frame
break_df <- imap_dfr(log2_groups, ~{
  tibble(
    order         = .y,
    n_cases       = .x$complete_cases,
    added_vars    = setdiff(.x$variables[-1],
                            if (.y == 1) character(0) else log2_groups[[.y-1]]$variables[-1])
  )
}) %>%
  unnest_longer(added_vars, keep_empty = TRUE) %>%
  mutate(added_vars = str_remove_all(added_vars,
                                     "^fl_(plasma|csf)__ALL__|_log2$"))

# 2 Figure 1: attrition curve
break_plot <- ggplot(break_df,
                     aes(x = order, y = n_cases)) +
  geom_line(linewidth = .8, colour = pal_okabe[2]) +
  geom_point(size = 2, colour  = pal_okabe[1]) +
  geom_text(aes(label = added_vars),
            hjust = -0.05, vjust = 0.5, size = 3,
            family = "mono") +
  scale_x_continuous(breaks = break_df$order) +
  scale_y_continuous(labels = comma) +
  labs(y = "complete cases",
       x = "biomarker-addition step",
       title = "Sample-size attrition as biomarkers are added") +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 60, hjust = 1))

# 3 Figure 2: improved RF-importance plot


imp_plot <- ggplot(rf_results$importance, aes(x = reorder(variable, importance), y = importance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Variable Importance in Random Forest Model",
    x = "Variables",
    y = "Importance (%IncMSE)"
  )

# 4 Figure 3: coefficient plot with model stats in the strip
# tidy & label the final model  (albumin example shown)
coef_df <- tidy(model_selected_albumin, conf.int = TRUE) %>% 
  filter(term != "(Intercept)") %>%                # ← NEW: remove intercept
  mutate(
    term    = str_remove_all(term, "^fl_(plasma|csf)__ALL__|_log2$"),
    term    = if_else(str_detect(term, ":"),         # italicise interactions
                      paste0("*", term, "*"), term),
    t_stat  = abs(statistic)                         # keep the t-value for plot 4
  )

model_stats <- glance(model_selected_albumin)

coef_plot <- ggplot(coef_df,
                    aes(y = fct_reorder(term, estimate),
                        x = estimate)) +
  geom_point(size = 2, colour = pal_okabe[5]) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.15, colour = pal_okabe[5]) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(y = NULL, x = "β  (95 % CI)",
       title = glue("Linear model – Albumin (R² {round(model_stats$r.squared,3)}, n = {model_stats$nobs})")) +
  theme_minimal(base_size = 11)

# t-stat bar 

# Option 4: Combined plot with effect size context
t_plot <- coef_df %>%
  mutate(
    sig_level = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    sig_color = case_when(
      p.value < 0.05 ~ "Significant",
      p.value < 0.1  ~ "Marginal",
      TRUE ~ "Not significant"
    )
  ) %>%
  ggplot(aes(y = fct_reorder(term, estimate),
             x = statistic)) +
  geom_col(aes(fill = sig_color)) +
  geom_text(aes(label = sig_level, 
                x = statistic + sign(statistic) * 0.3),
            size = 4, vjust = 0.5) +
  scale_fill_manual(values = c("Significant" = "#009E73",
                               "Marginal" = "#F0E442",
                               "Not significant" = "#999999"),
                    name = "Significance") +
  geom_vline(xintercept = 0, linetype = "solid") +
  labs(y = NULL, x = "t-statistic",
       title = "T-statistics with significance levels",
       subtitle = "*** p<0.001, ** p<0.01, * p<0.05, . p<0.1") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")

# 5 Table 1: breakpoint narrative

tbl_breaks <- break_df %>%
  group_by(order) %>%
  summarise(`complete cases` = first(n_cases),
            `added biomarkers` = paste(added_vars, collapse = ", ")) %>%
  mutate(step = row_number()) %>%
  select(step, everything()) %>%
  gt() %>%
  tab_header(title = md("**Sample-size breakpoints**")) %>%
  fmt_number(columns = `complete cases`, sep_mark = ",") %>%
  tab_style(cell_borders(sides = "top", weight = px(1)),
            locations = cells_body(rows = 1)) %>%
  opt_row_striping()

# 6 Table 2: model performance
tbl_perf <- tibble(
  outcome = "Albumin index",
  adj_R2  = round(model_stats$r.squared, 3),
  RMSE    = round(model_stats$sigma, 3),
  n       = model_stats$nobs
) %>%
  gt() %>%
  tab_header(title = md("**Final model performance**"))

# 7 Arrange and save / display


# side-by-side figures, adjust widths to taste
(break_plot | imp_plot | coef_plot | t_plot) +
  plot_layout(widths = c(1.2, 1, 1, 0.9))   # tweak widths as you like
# save
ggsave("mri_analys_csfplasma_overview.png", width = 17, height = 4, dpi = 300)

# view tables in the Viewer or knit to Quarto
tbl_breaks
tbl_perf


```

# 4/30/25 Arvind, try classification instead of inference for albumin index

how? Find a cut point for Albumin Index for positive/negative from literature (maybe 5.0 or 5.5) If can't find, then plot by DX. Use the 95th-percentile of Controls as an upper bound cut-off. Above cut-off is positive. Predict positive/negative using fluid biomarkers RF Logistic regression Assess ROC Curve Note: I am using the dat_all_combine without alignment

```{r}
library(gridExtra) 
# library(grid.text )# for grid.arrange()
library(grid)        # for textGrob(), gpar()
library(randomForest)

# Define the baseline visit categories
baseline_visits <- c(
  "Baseline (Arm 1: Patients)",
  "Baseline (Arm 2: Control)",
  "Baseline (Arm 3: CSF controls)",
  "Baseline (Arm 4: Diabetes)",
  "baseline"
)

# Filter dataset based on baseline visits
dat_baseline_AD_VD <- dat_all_combine |>
  filter(redcap_event_name %in% baseline_visits) |>  # Keep only baseline visits
  mutate(
    TripleDichotomy_VD = mri_dti_psmd_z,
    TripleDichotomy_VD_log2 = e_log_shift(TripleDichotomy_VD),
    TripleDichotomy_AD = (fl_csf__ALL__ptau / fl_csf__ALL__abeta_42),
    TripleDichotomy_AD_log2 = log2(TripleDichotomy_AD),
    fl_plasma__ALL__abeta42_40_ratio_log2 = log2(fl_plasma__ALL__abeta_42/fl_plasma__ALL__abeta_40)
  )

# find cut point base on 95th-percentile of Controls
cut_75 <- quantile (dat_baseline_AD_VD$fl_albumin_index_log2
                                      [dat_baseline_AD_VD$dx_biological == "LA"], probs = 0.75,
                                      na.rm = TRUE) 

# create a factor base on the 95th-percentile of Controls which is coming from the cut_95

# dat_baseline_AD_VD <- dat_baseline_AD_VD |> mutate(
#   
#     AlbuminIndex_status= if_else (fl_albumin_index > cut_95, 1,0), # 1 = Positive
#     AlbuminIndex_status= factor(AlbuminIndex_status, levels=c(0,1))
# )
# table(dat_all_combine_AD_VD$AlbuminIndex_status)
# 
#   0   1 
# 276   8 


## --- 2. quick plot by diagnosis --------------------------------
# ggplot(dat_baseline_AD_VD, aes(x = dx_biological, y = fl_albumin_index)) +
#   geom_boxplot(outlier.shape = NA, fill = "grey90") +
#   geom_jitter(width = 0.25, alpha = 0.3) +
#   geom_hline(yintercept = cut_point_95thpercentile, colour = "red", linetype = 2, linewidth = 1) +
#   labs(title = "Albumin Index by dx_biological (baseline)",
#        y = "Albumin Index", x = "Diagnosis group") +
#   theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))


# It is really an unbalanced result, I also use cut point from literture

lit9_log2  <- log2(  9    )  # literature cut-point 9
lit10.2_l2 <- log2( 10.2    )  # literature 10.2
lit5.5_l2  <- log2(  5.5     )  # literature 5.5


cut_pts <- tibble(
  cutoff = c(cut_75, lit9_log2, lit10.2_l2, lit5.5_l2),
  label  = c("75th % LA", "Literature 9", "Literature 10.2", "Literature 5.5")
)

## ------------------------------------------------------------------ 
## 3. (optional) add a status column for EACH cut‑point -------------
dat_baseline_AD_VD <- dat_baseline_AD_VD %>%
  mutate(
    AlbumIdx_pos_75      = factor(if_else(fl_albumin_index_log2 > cut_75, 1, 0),
                                  levels = c(0, 1)),
    AlbumIdx_pos_9       = factor(if_else(fl_albumin_index_log2 > lit9_log2,  1, 0),
                                  levels = c(0, 1)),
    AlbumIdx_pos_10_2    = factor(if_else(fl_albumin_index_log2 > lit10.2_l2, 1, 0),
                                  levels = c(0, 1)),
    AlbumIdx_pos_5_5     = factor(if_else(fl_albumin_index_log2 > lit5.5_l2, 1, 0),
                                  levels = c(0, 1))
  )

## ------------------------------------------------------------------ 
## 4. plot with one horizontal line per cut‑point -------------------
ggplot(dat_baseline_AD_VD,
       aes(x = dx_biological, y = fl_albumin_index_log2)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90") +
  geom_jitter(width = 0.25, alpha = 0.3, size = 1) +
  geom_hline(data = cut_pts,
             aes(yintercept = cutoff, colour = label),
             linetype = "dashed", linewidth = 1) +
  scale_colour_manual(values = c("red", "blue", "darkgreen", "purple"),
                      name   = "Cut‑points") +
  labs(title = "Albumin Index by dx_biological (baseline)",
       y = "Albumin Index", x = "Diagnosis group") +
  theme_bw() +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1),
        legend.title = element_text(size = 10),
        legend.text  = element_text(size = 9))


# 
# # ------------ 2. biomarker list --------------------------------------------
# some_biomarkers_albumin <- c(
#  "fl_csf_glucose_log2",
#  "fl_csf_total_protein_log2",
#  "fl_csf_rbc_log2",
#  "fl_csf_tnc_log2",
#  "fl_csf_igg_log2",
#  # "fl_csf_albumin_log2",
#  "fl_csf_igg_synthesis_log2",
#  "fl_abeta_38_log2",
#  "fl_csf__ALL__gfap_log2",
#  "fl_plasma__ALL__gfap_log2",
#  "fl_csf__ALL__nfl_log2",
#  "fl_plasma__ALL__nfl_log2",
#  "fl_csf__ALL__ptau_log2",
#  "fl_plasma__ALL__ptau_181_log2",
#  "fl_plasma__ALL__abeta42_40_ratio_log2",
#  "fl_plasma__ALL__ptau_217_log2",
#  "fl_csf__ALL__abeta_40_log2",
#  "fl_plasma__ALL__abeta_40_log2",
#  "fl_csf__ALL__abeta_42_log2" ,
#  "fl_plasma__ALL__abeta_42_log2",
#  "fl_csf__ALL__plgf_log2",
#  "fl_plasma__ALL__plgf_log2" ,
#  "fl_csf__ALL__il_1b_log2" ,
#  "fl_plasma__ALL__il_1b_log2",
#  "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
#  "fl_csf__ALL__mmp_10_log2" ,
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" ,
#  "fl_plasma__ALL__vegf_a_log2",
#  "fl_plasma__ALL__vegf_c_log2",
#  "fl_plasma__ALL__vegf_d_log2"
# )
# 
# 
# 
# 
# 
# e_data_complete_by_variable_subset(dat_baseline_AD_VD, var_list = c(some_biomarkers_albumin, "fl_albumin_index_log2"), var_resp = "fl_albumin_index_log2")
# 
# ## 3. Jack-knife CV helper for single-predictor GLM -------------------
# jackknife_roc_cv <- function(data, outcome_var, predictor_var) {
#   # data = dat_i
#   # outcome_var = "AlbuminIndex_status"
#   # predictor_var =  bm
#   n <- nrow(data)
#   labels <- integer(n)
#   probs  <- numeric(n)
#   for (i in seq_len(n)) {
#     # i=1
#     train <- data[-i, ]
#     test  <- data[i, , drop=FALSE]
#     fit   <- glm(
#       formula(paste(outcome_var, "~", predictor_var)),
#       data = train, family = binomial
#     )
#     labels[i] <- as.numeric(as.character(data[[outcome_var]][i]))
#     probs[i]  <- predict(fit, newdata = test, type = "response")
#   }
#   list(true = labels, pred = probs)
# }
# 
# library(patchwork)   # wrap_plots() + plot_annotation()
# 
# # ───────────────────────────────
# # 1.  helpers ---------------------------------------------------------
# # ───────────────────────────────
# # Create the binary outcome for a single cut‑point
# add_status <- function(data, cut) {
#   data %>% 
#     mutate(
#       # status = factor(if_else(fl_albumin_index_log2 > cut, 1, 0),
#       #                 levels = c(0, 1))
#        status = if_else(fl_albumin_index_log2 > cut, 1, 0)
#     )
# }
# 
# # Build one ROC plot for one biomarker
# make_roc_plot <- function(df, biomarker) {
#   tmp <- df %>% select(status, all_of(biomarker)) %>% drop_na()
# 
#   jk  <- jackknife_roc_cv(data= tmp, 
#                           outcome_var = "status",                              predictor_var = biomarker)
# 
#   e_plot_roc(
#     labels_true     = jk$true,
#     pred_values_pos = jk$pred,
#     label_neg_pos   = c(0, 1),
#     sw_plot         = TRUE,
#     sw_caption_desc = FALSE
#   )$plot_roc +
#     ggtitle(biomarker) +
#     theme_minimal(base_size = 14) +
#     theme(
#       plot.title  = element_text(size = 16, face = "bold"),
#       axis.title  = element_text(size = 14),
#       axis.text   = element_text(size = 12),
#       legend.text = element_text(size = 12)
#     )
# }
# 
# # Split any vector into chunks of size n
# chunk <- function(x, n) split(x, ceiling(seq_along(x) / n))
# 
# # ───────────────────────────────
# # 2.  main pipeline ---------------------------------------------------
# # ───────────────────────────────
# plots_per_page <- 4   # change to 6 for 3×2 pages, etc.
# 
# pdf("Albumin_ROCs_by_cutpoint.pdf", onefile = TRUE, width = 20, height = 20)
# 
# cut_pts %>% 
#   mutate(data  = map(cutoff, ~add_status(dat_baseline_AD_VD, .x))) %>%  # add ‘status’
#   mutate(plots = map(data,  ~map(some_biomarkers_albumin, make_roc_plot, df = .x))) %>% 
#   pwalk(function(label, plots, ...) {                       # one “cut‑point” at a time
#     
#     # lay out pages for this cut‑point
#     chunk(plots, plots_per_page) %>% 
#       map(~wrap_plots(.x, ncol = ceiling(sqrt(plots_per_page))) +
#             plot_annotation(
#               title = paste("Albumin Index cut‑point:", label),
#               theme = theme(plot.title = element_text(size = 18, face = "bold"))
#             )) %>% 
#       walk(print)                                           # send pages to the pdf device
#   })
# 
# dev.off()
# 
# # 2. count non‐missing fl_csf_rbc_log2 by above/below cut_95
# # dat_baseline_AD_VD %>%
# #   mutate(
# #     status = if_else(fl_albumin_index > cut_95, "Above_95", "Below_95")
# #   ) %>%
# #   group_by(status) %>%
# #   summarize(
# #     n_with_rbc = sum(!is.na(fl_csf_rbc_log2)),
# #     total      = n()
# #   )
# 
# 
# 
# # # the 2nd dataset (Literature 9):
# # df_lit9 <- cut_pts$data[[2]]
# # 
# # # a quick peek:
# # df_lit9 %>% dplyr::select(fl_albumin_index, status) %>% head()
# 
# 
# 
# 
# ## ------------------------------------------------------------------
# ## 1.  Helper to pick important variables from a binary outcome
# ## ------------------------------------------------------------------
# identify_important_variables <- function(data, status_var, predictors,
#                                          n_trees = 500) {
#   model_data <- data[, c(status_var, predictors)] |>
#                 tidyr::drop_na()        # fast NA fill
# 
#   rf_formula <- reformulate(predictors, response = status_var)
# 
#   set.seed(123)
#   rf <- randomForest(
#           formula   = rf_formula,
#           data      = model_data,
#           ntree     = n_trees,
#           importance= TRUE
#         )
# 
#   var_imp <- importance(rf, type = 2)[, "MeanDecreaseGini"]
#   names(var_imp)[var_imp > 0]
# }
# 
# ## ------------------------------------------------------------------
# ## 2.  Final RF with those variables  +  accuracy / confusion
# ## ------------------------------------------------------------------
# run_rf_analysis <- function(data, status_var, predictors, n_trees = 500) {
# 
#   model_data <- data[, c(status_var, predictors)] |>
#                 tidyr::drop_na()
# 
#   rf_formula <- reformulate(predictors, response = status_var)
# 
#   set.seed(123)
#   rf_model <- randomForest(
#                formula   = rf_formula,
#                data      = model_data,
#                ntree     = n_trees,
#                importance= TRUE,
#                strata    = model_data[[status_var]],
#                sampsize  = rep(min(table(model_data[[status_var]])), 2)
#              )
# 
#   pred  <- predict(rf_model)
#   cm    <- table(Predicted = pred, Actual = model_data[[status_var]])
#   acc   <- sum(diag(cm)) / sum(cm)
# 
#   imp_df <- importance(rf_model, type = 2)[, "MeanDecreaseGini"] |>
#             sort(decreasing = TRUE) |>
#             stack() |>
#             setNames(c("importance", "variable"))
# 
#   list(model = rf_model,
#        importance = imp_df,
#        performance = list(confusion_matrix = cm,
#                           accuracy = acc,
#                           n_obs = nrow(model_data)))
# }
# 
# ## ------------------------------------------------------------------
# ## 3.  Apply once for each albumin‑index cut‑point
# ## ------------------------------------------------------------------
# for (i in seq_len(nrow(cut_pts))) {
# 
#   cp  <- cut_pts$cutoff[i]
#   lbl <- cut_pts$label[i]
# 
#   dat_i <- dat_baseline_AD_VD |>
#            mutate(status = factor(if_else(fl_albumin_index_log2 > cp, 1, 0),
#                                   levels = c(0, 1)))
# model_data <- dat_i[, c("status", some_biomarkers_albumin)] %>% drop_na()
# if (n_distinct(model_data$status) < 2) {
#   cat(lbl, ": only one class present, skipping RF\n")
#   next
# }
#   # 3A – get variables with positive Gini gain
#   important_vars <- identify_important_variables(
#                       data       = dat_i,
#                       status_var = "status",
#                       predictors = some_biomarkers_albumin
#                     )
# 
#   cat("\n", lbl, ": important variables\n")
#   print(important_vars)
# 
#   # 3B – final RF on those variables
#   final_rf <- run_rf_analysis(
#                 data       = dat_i,
#                 status_var = "status",
#                 predictors = important_vars
#               )
# 
#   cat("  Accuracy:", round(final_rf$performance$accuracy, 3),
#       " |  n =", final_rf$performance$n_obs, "\n")
#   print(final_rf$performance$confusion_matrix)
# 
#   # 3C – quick importance barplot 
#   ggplot(final_rf$importance,
#          aes(reorder(variable, importance), importance)) +
#     geom_col(fill = "steelblue") + coord_flip() +
#     labs(title = paste("RF importance –", lbl),
#          x = NULL, y = "Mean Decrease Gini") +
#     theme_minimal(base_size = 12)
# }
# 
# 
# # Only one class left so I have to see what I can do like check the number of observation and then add the modeling
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# 
# # combination of plasma biomarkers
# 
# 
# # Get only plasma biomarkers ending in _log2
# find_optimal_log2_groups <- function(data, min_cases = 30) {
#   # Get only plasma markers ending in _log2
#   plasma_markers <- grep("^fl_plasma.*_log2$", names(data), value = TRUE)
#   
#   # Start with outcome variable
#   current_vars <- c("fl_albumin_index_log2")
#   results <- list()
#   
#   for(marker in plasma_markers) {
#     temp_vars <- c(current_vars, marker)
#     complete_cases <- sum(complete.cases(data[, temp_vars]))
#     
#     if(complete_cases >= min_cases) {
#       current_vars <- temp_vars
#       results[[length(results) + 1]] <- list(
#         variables = current_vars,
#         complete_cases = complete_cases
#       )
#     }
#   }
#   return(results)
# }
# 
# # Apply the function
# log2_groups <- find_optimal_log2_groups(dat_baseline_AD_VD)
# 
# # Print summary of available biomarkers and sample sizes
# for(i in seq_along(log2_groups)) {
#   cat("\nGroup", i, ":\n")
#   cat("Variables:", paste(log2_groups[[i]]$variables[-1], collapse = ", "), "\n")
#   cat("Complete cases:", log2_groups[[i]]$complete_cases, "\n")
# }
# 
# # Get the final set of all log2 biomarkers
# final_log2_vars <- log2_groups[[length(log2_groups)]]$variables
# 
# # Print final number of biomarkers and sample size
# cat("\nFinal model will include:", length(final_log2_vars)-1, "biomarkers\n")
# cat("With sample size:", log2_groups[[length(log2_groups)]]$complete_cases, "\n")
# 
# # Function to print breakpoint summary
# summarize_breakpoints <- function(groups) {
#   prev_vars <- NULL
#   prev_n <- NULL
#   
#   cat("Sample Size Breakpoints:\n")
#   
#   for(i in seq_along(groups)) {
#     current_vars <- groups[[i]]$variables[-1]  # exclude outcome variable
#     current_n <- groups[[i]]$complete_cases
#     
#     if(is.null(prev_n) || current_n != prev_n) {
#       new_vars <- if(is.null(prev_vars)) {
#         current_vars
#       } else {
#         setdiff(current_vars, prev_vars)
#       }
#       
#       # Clean up variable names for display
#       clean_vars <- gsub("fl_plasma__ALL__", "", new_vars)
#       clean_vars <- gsub("_log2", "", clean_vars)
#       
#       cat(sprintf("%d cases: %s\n", 
#                   current_n,
#                   if(i == 1) {
#                     sprintf("starting with %s", paste(clean_vars, collapse = " and "))
#                   } else {
#                     sprintf("adding %s", paste(clean_vars, collapse = " and "))
#                   }))
#       
#       prev_vars <- current_vars
#       prev_n <- current_n
#     }
#   }
# }
# 
# # Apply the function to your groups
# summarize_breakpoints(log2_groups)
# 
# # Also show the complete list of final variables
# cat("\nFinal set of biomarkers:\n")
# final_vars <- gsub("fl_plasma__ALL__", "", 
#                   gsub("_log2", "", 
#                        log2_groups[[length(log2_groups)]]$variables[-1]))
# cat(paste(final_vars, collapse = "\n"))
# 
# 
# # Or 
# e_data_complete_by_variable_subset(dat_all_combine_AD_VD, var_list = c(some_biomarkers_vd, "fl_albumin_index_log2"), var_resp = "fl_albumin_index_log2")
# 
# # now define base on the result of the number of observation and Importance one 
# some_biomarkers_albumin <- c(
#  # "fl_csf_glucose_log2",
#  # "fl_csf_total_protein_log2",
#  # "fl_csf_rbc_log2",
#  # "fl_csf_tnc_log2",
#  # "fl_csf_igg_log2",
#  # "fl_csf_albumin_log2",           
#  # "fl_csf_igg_synthesis_log2",     
#  # "fl_abeta_38_log2",
#  "fl_csf__ALL__gfap_log2",
#  "fl_plasma__ALL__gfap_log2",
#  "fl_csf__ALL__nfl_log2",
#  "fl_plasma__ALL__nfl_log2",
#  # "fl_csf__ALL__ptau_log2",        
#  "fl_plasma__ALL__ptau_181_log2",
#  "fl_plasma__ALL__abeta42_40_ratio_log2",
#  # "fl_plasma__ALL__ptau_217_log2",
#  # "fl_csf__ALL__abeta_40_log2",   
#  # "fl_plasma__ALL__abeta_40_log2",
#  # # "fl_csf__ALL__abeta_42_log2" ,   
#  # "fl_plasma__ALL__abeta_42_log2",
#  # "fl_csf__ALL__plgf_log2",
#  "fl_plasma__ALL__plgf_log2" ,
#  # "fl_csf__ALL__il_1b_log2" ,     
#  # "fl_plasma__ALL__il_1b_log2", 
#  # "fl_plasma__ALL__mmp_2_log2" ,
#  "fl_plasma__ALL__mmp_1_log2",
# "fl_plasma__ALL__mmp_3_log2" ,
#  "fl_plasma__ALL__mmp_9_log2" ,
#  "fl_plasma__ALL__mmp_10_log2",
#  "fl_csf__ALL__mmp_10_log2" ,
#  "fl_plasma__ALL__bfgf_log2",
# "fl_plasma__ALL__flt_1_log2",
#  "fl_plasma__ALL__tie2_log2" ,
#  "fl_plasma__ALL__vegf_a_log2",
#  "fl_plasma__ALL__vegf_c_log2",
#  "fl_plasma__ALL__vegf_d_log2"
# 
# )
# 
# 
# n_complete <- sum(complete.cases(dat_baseline_AD_VD[, c("fl_albumin_index_log2", some_biomarkers_albumin)]))
# print(n_complete) # Only one remained
# 
# sapply(some_biomarkers_vd, function(x) sum(!is.na(dat_baseline_AD_VD[[x]])))
# 
# # your 4 cut-points & labels
# cut_pts <- tibble(
#   cutoff = c(cut_75, lit9_log2, lit10.2_l2, lit5.5_l2),
#   label  = c("75th % Control", "Literature 9", "Literature 10.2", "Literature 5.5")
# )
# 
# # open PDF device: one page per cut‐point
# pdf("RF_VarImp_by_cutpoint.pdf", onefile = TRUE, width = 8, height = 6)
# 
# for (i in seq_len(nrow(cut_pts))) {
#   cp  <- cut_pts$cutoff[i]
#   lbl <- cut_pts$label[i]
# 
#   # 1) build the status + keep only the 11 plasma markers
#   df_rf <- dat_baseline_AD_VD %>%
#     mutate(status = factor(if_else(fl_albumin_index > cp, 1, 0),
#                            levels = c(0, 1))) %>%
#     select(status, all_of(some_biomarkers_albumin)) %>%
#     drop_na()
# 
#   # skip if only one class remains
#   if (n_distinct(df_rf$status) < 2) {
#     message(lbl, ": only one class present—skipping")
#     next
#   }
# 
#   # 2) train a balanced RF
#   set.seed(123)
#   rf_fit <- randomForest(
#     status ~ .,
#     data       = df_rf,
#     ntree      = 1000,
#     importance = TRUE
#     # strata     = df_rf$status,
#     # sampsize   = rep(min(table(df_rf$status)), 2)
#   )
# 
#   # 3) extract and tidy the Gini importances
#   imp_df <- importance(rf_fit, type = 2)[, "MeanDecreaseGini"] %>%
#     enframe(name = "variable", value = "importance") %>%
#     arrange(importance)
# 
#   # 4) plot horizontal bar chart
#   p <- ggplot(imp_df, aes(x = reorder(variable, importance), y = importance)) +
#     geom_col(fill = "steelblue") +
#     coord_flip() +
#     labs(
#       title = paste("RF Variable Importance —", lbl),
#       x     = NULL,
#       y     = "Mean Decrease Gini"
#     ) +
#     theme_minimal(base_size = 12)
# 
#   print(p)
# }
# 
# dev.off()
# 
# 
# 
# 
# 
# 
# 
# 
# ## -----------------------------------------------------------
# ##  ROC curves for albumin-index status at each cut‑point
# ## -----------------------------------------------------------
# 
# for (i in seq_len(nrow(cut_pts))) {
#   
#   cp  <- cut_pts$cutoff[i]
#   lbl <- cut_pts$label[i]
# 
#   ## 1. create the binary outcome for this cut point
#   df_i <- dat_baseline_AD_VD %>%
#             mutate(status = factor(if_else(fl_albumin_index > cp, 1, 0),
#                                    levels = c(0, 1))) %>%
#             select(status, all_of(some_biomarkers_albumin)) %>%
#             drop_na()                              # complete cases
# 
#   ## need both classes
#   if (dplyr::n_distinct(df_i$status) < 2) {
#     message(lbl, " → only one class; skipping")
#     next
#   }
# 
#   ## 2. fit a simple logistic model with all biomarkers
#   fit <- glm(status ~ ., family = binomial, data = df_i)
# 
#   ## 3. predicted probabilities
#   df_i$pred <- predict(fit, type = "response")
# 
#   ## 4. ROC plot
#   roc_plot <- e_plot_roc(
#                 labels_true     = df_i$status,
#                 pred_values_pos = df_i$pred,
#                 label_neg_pos   = c(0, 1),
#                 sw_plot         = TRUE,
#                 sw_caption_desc = FALSE
#               )$plot_roc +
#               ggtitle(paste("Albumin cut‑point:", lbl))
# 
#   print(roc_plot)     
#   
# }
# 
# # add the glm for the all plasma and csf combination but base on the Imp pdf and also this is glm 
# 
# # drop number of varb base on the imp pdf then again apply the rF 
# 
# # larger sample size choose Vb , keep only IMP VB maybe RM the baseline for first pass and then later filter on baseline
# 
# # which featuers are more varibles visiti to visit if albumin chnage a lot depedn time of daya maybe albumin is highly variabels , albumin is consisitent and one varies a lot that varibles will be less important 
# 
# # so working with larger sample size base on not only baseline, it is okay to do that ,
# # how much info is more valuable for us , since we do prediction is m=not matter that much how much obs are dependent from each other
# 
# # e-randomforest option for imputation look for missing NA and include only one have lower NA and include and have RF impute those NA sw_NA_action omit or impute sw quick full only does not do model selection
```

# AFter meeting with Erik and Run the above code these are the changes I need to make

"Use more data" → Don't limit to baseline visits only "Let the computer fill gaps" → Use Random Forest to handle missing values "Focus on what matters" → Use importance rankings to pick best biomarkers "Check consistency" → Make sure biomarkers are stable over time Note here I did not use the cutpoint for Albumin

```{r}
# ===================================================================
# Expanding beyond baseline + RF imputation + Variable importance
# ===================================================================
library(randomForest)
library(patchwork)
library(VIM)  # for missing data visualization
library(purrr)

# ===================================================================
# 1. EXPAND BEYOND BASELINE (Address sample size issues)
# ===================================================================

# Use ALL visits, not just baseline
dat_expanded <- dat_all_combine |>
  mutate(
    TripleDichotomy_VD = mri_dti_psmd_z,
    TripleDichotomy_VD_log2 = e_log_shift(TripleDichotomy_VD),
    TripleDichotomy_AD = (fl_csf__ALL__ptau / fl_csf__ALL__abeta_42),
    TripleDichotomy_AD_log2 = log2(TripleDichotomy_AD),
    fl_plasma__ALL__abeta42_40_ratio_log2 = log2(fl_plasma__ALL__abeta_42/fl_plasma__ALL__abeta_40)
  )

cat("Expanded dataset size:", nrow(dat_expanded), "observations\n")
# cat("Baseline-only dataset size:", nrow(dat_all_combine_AD_VD), "observations\n")

# ===================================================================
# 2. SMART VARIABLE SELECTION BASED ON IMPORTANCE
# ===================================================================

# Define ALL potential biomarkers
all_potential_biomarkers <-  c(
 "fl_csf_glucose_log2",
 "fl_csf_total_protein_log2",
 "fl_csf_rbc_log2",
 "fl_csf_tnc_log2",
 "fl_csf_igg_log2",
 # "fl_csf_albumin_log2",
 "fl_csf_igg_synthesis_log2",
 "fl_abeta_38_log2",
 "fl_csf__ALL__gfap_log2",
 "fl_plasma__ALL__gfap_log2",
 "fl_csf__ALL__nfl_log2",
 "fl_plasma__ALL__nfl_log2",
 "fl_csf__ALL__ptau_log2",
 "fl_plasma__ALL__ptau_181_log2",
 "fl_plasma__ALL__abeta42_40_ratio_log2",
 "fl_plasma__ALL__ptau_217_log2",
 "fl_csf__ALL__abeta_40_log2",
 "fl_plasma__ALL__abeta_40_log2",
 "fl_csf__ALL__abeta_42_log2" ,
 "fl_plasma__ALL__abeta_42_log2",
 "fl_csf__ALL__plgf_log2",
 "fl_plasma__ALL__plgf_log2" ,
 "fl_csf__ALL__il_1b_log2" ,
 "fl_plasma__ALL__il_1b_log2",
 "fl_plasma__ALL__mmp_2_log2" ,
 "fl_plasma__ALL__mmp_1_log2",
"fl_plasma__ALL__mmp_3_log2" ,
 "fl_plasma__ALL__mmp_9_log2" ,
 "fl_plasma__ALL__mmp_10_log2",
 "fl_csf__ALL__mmp_10_log2" ,
 "fl_plasma__ALL__bfgf_log2",
"fl_plasma__ALL__flt_1_log2",
 "fl_plasma__ALL__tie2_log2" ,
 "fl_plasma__ALL__vegf_a_log2",
 "fl_plasma__ALL__vegf_c_log2",
 "fl_plasma__ALL__vegf_d_log2"
)

# Check missing data patterns
missing_summary <- dat_expanded %>%
  select(fl_albumin_index_log2, all_of(all_potential_biomarkers)) %>%
  summarise_all(~sum(is.na(.))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_count") %>%
  mutate(
    total_obs = nrow(dat_expanded),
    missing_pct = (missing_count / total_obs) * 100,
    available_pct = 100 - missing_pct
  ) %>%
  arrange(missing_pct)

# Visualize missing data pattern
ggplot(missing_summary, aes(x = reorder(variable, -missing_pct), y = missing_pct)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Missing Data Percentage by Biomarker",
       x = "Biomarker", y = "Missing Data %") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 8))

# 3. GET INITIAL IMPORTANCE (one RF per variable with available data)
get_univariate_importance <- function(data,
                                      outcome,
                                      predictors) {
  imp_scores <- sapply(predictors, function(var) {
    temp_data <- data[, c(outcome, var)] |> na.omit()
    if(nrow(temp_data) < 50) return(0)
    
    rf <- randomForest(as.formula(paste(outcome, "~", var)), 
                      data = temp_data,
                      ntree = 100)
    tail(rf$rsq, 1)
  })
  
  data.frame(
    variable = names(imp_scores),
    importance = imp_scores
  ) |> arrange(desc(importance))
}

initial_importance <- get_univariate_importance(
  data = dat_expanded,
  outcome = "fl_albumin_index_log2",
  predictors = all_potential_biomarkers)
print(head(initial_importance, 10))


# 4. ITERATIVE RF WITH TOP VARIABLES
build_progressive_rf <- function(data, outcome, importance_df, max_vars = 15) {
  results <- list()
  
  for(n in c(3, 5, 8, 10, 12, 15)) {
    if(n > nrow(importance_df)) n <- nrow(importance_df)
    
    top_vars <- importance_df$variable[1:n]
    temp_data <- data[, c(outcome, top_vars)]
    
    # Two approaches: complete cases vs imputation
    # Approach 1: Complete cases
    complete_data <- na.omit(temp_data)
    if(nrow(complete_data) > 50) {
      rf_complete <- randomForest(as.formula(paste(outcome, "~ .")), 
                                 data = complete_data, ntree = 500)
      r2_complete <- tail(rf_complete$rsq, 1)
    } else {
      r2_complete <- NA
    }
    
    # Approach 2: With imputation
    if(nrow(temp_data[!is.na(temp_data[[outcome]]), ]) > 50) {
      rf_imputed <- randomForest(as.formula(paste(outcome, "~ .")), 
                                data = temp_data[!is.na(temp_data[[outcome]]), ],
                                ntree = 500, na.action = na.roughfix)
      r2_imputed <- tail(rf_imputed$rsq, 1)  # RF's built-in imputation method
    } else {
      r2_imputed <- NA
    }
    
    results[[as.character(n)]] <- list(
      n_vars = n,
      n_complete = nrow(complete_data),
      n_imputed = nrow(temp_data[!is.na(temp_data[[outcome]]), ]),
      r2_complete = r2_complete,
      r2_imputed = r2_imputed,
      variables = top_vars
    )
  }
  
  results
}

rf_results <- build_progressive_rf(
  data = dat_expanded,
  outcome = "fl_albumin_index_log2", 
   importance_df = initial_importance)



 #Plot R-squared comparison
# library(tidyr)
# results_long <- results_df |>
#   select(n_vars, Complete_Cases, With_Imputation) |>
#   pivot_longer(-n_vars, names_to = "Method", values_to = "R_squared")
# 
# ggplot(results_long, aes(x = n_vars, y = R_squared, color = Method)) +
#   geom_line(size = 1.2) +
#   geom_point(size = 3) +
#   scale_color_manual(values = c("Complete_Cases" = "blue", "With_Imputation" = "red")) +
#   labs(title = "Model Performance: Complete Cases vs Imputation",
#        x = "Number of Variables", y = "R-squared") +
#   theme_minimal() +
#   theme(legend.position = "bottom")




# 6. FINAL MODEL WITH OPTIMAL VARIABLES (3-variable complete cases)
optimal_n <- 3  # Use 3 variables instead of max imputation
final_vars <- rf_results[["3"]]$variables
cat("\nOPTIMAL MODEL:")
cat("\nVariables:", optimal_n)
cat("\nR² (complete cases):", round(results_df$Complete_Cases[results_df$n_vars == 3], 3))
cat("\nSample size:", results_df$n_obs_complete[results_df$n_vars == 3])

# Final model with complete cases (no imputation)
final_data <- dat_expanded[, c("fl_albumin_index_log2", final_vars)] |>
  na.omit()  # Complete cases only

final_rf <- randomForest(
  fl_albumin_index_log2 ~ .,
  data = final_data,
  ntree = 1000,
  importance = TRUE
)

# Variable importance plot
imp_final <- importance(final_rf) |>
  as.data.frame() |>
  tibble::rownames_to_column("variable") |>
  mutate(clean_name = gsub(".*__ALL__|_log2", "", variable)) |>
  arrange(desc(`%IncMSE`))

ggplot(imp_final, aes(x = reorder(clean_name, `%IncMSE`), y = `%IncMSE`)) +
  geom_col(fill = "darkgreen") +
  coord_flip() +
  labs(title = "Final Model: Variable Importance",
       x = NULL, y = "% Increase in MSE") +
  theme_minimal()

print(final_rf)



# need to work on this tom



library(tidymodels)
library(purrr)
library(tidymodels)
library(forcats)
# ------------------------------------------------------------------
# 0.  DATA  (continuous outcome + 3 predictors + true binary label)
# ------------------------------------------------------------------
vars <- c("fl_albumin_index_log2", final_vars)
d_raw <- dat_expanded |>
         select(all_of(vars), dx_biological, age) |>   # keep age+dx for strat rule
         drop_na()                                     # complete cases only

# build the binary “ground-truth” label once,
# here: 95th percentile of controls <65 y
cut_true <- quantile(
  d_raw$fl_albumin_index_log2[
    d_raw$dx_biological == "LA"], probs= 0.75, na.rm = TRUE
)

# cut_75 <- quantile (dat_baseline_AD_VD$fl_albumin_index_log2
#                                       [dat_baseline_AD_VD$dx_biological == "LA"], probs = 0.75,
#                                       na.rm = TRUE) 


d <- d_raw |>
     mutate(albumin_hi = factor(fl_albumin_index_log2 > cut_true,
                                levels = c(FALSE, TRUE),
                                labels = c("normal", "elevated"))) |>
     select(-dx_biological, -age)                      # not used downstream

# ------------------------------------------------------------------
# 1.  80 / 20 split  (stratified)
# ------------------------------------------------------------------
set.seed(2025)
split_obj <- initial_split(d, prop = 0.80, strata = albumin_hi)
d_train   <- training(split_obj)
d_test    <- testing(split_obj)        # remains completely untouched

# ------------------------------------------------------------------
# 2.  Inner 10-fold CV on the 80 %
# ------------------------------------------------------------------
folds <- vfold_cv(d_train, v = 10, strata = albumin_hi)

# Helper to choose a cut-point on a training slice
choose_threshold <- function(truth, pred_cont, rule = "youden") {

  df <- tibble(
    truth = fct_relevel(truth, "normal", "elevated"),
    .pred = pred_cont
  )

  roc_tbl <- roc_curve(df, truth = truth, .pred, event_level = "second")

  if (rule == "youden") {
    roc_tbl %>%
      mutate(j = sensitivity + specificity - 1) %>%
      slice_max(j, n = 1) %>%
      pull(.threshold)
  } else {
    stop("only 'youden' implemented")
  }
}


# run across folds
cv_stats <- map_dfr(folds$splits, function(sp) {
  tr <- analysis(sp)
  va <- assessment(sp)

  fit <- glm(fl_albumin_index_log2 ~ . - albumin_hi,
             data   = tr,
             family = gaussian())

  pred_tr <- predict(fit, newdata = tr)
  thr     <- choose_threshold(tr$albumin_hi, pred_tr)   # fold-specific

  pred_va <- predict(fit, newdata = va)
  class_va<- factor(pred_va > thr, levels = c(FALSE, TRUE),
                                  labels  = c("normal", "elevated"))

  tibble(
    id   = sp$id,
    thr  = thr,
    sens = sensitivity_vec(va$albumin_hi, class_va, event_level = "second"),
    spec = specificity_vec(va$albumin_hi, class_va, event_level = "second"),
    auc  = roc_auc_vec     (va$albumin_hi, pred_va, event_level = "second")
  )
})

print(cv_stats)
cv_stats |>
  summarise(across(c(thr, sens, spec, auc),
                   list(mean = mean, sd = sd), .names = "{.col}_{.fn}"))

# ------------------------------------------------------------------
# 3.  Fit final model on 100 % of the training data
# ------------------------------------------------------------------
final_glm <- glm(fl_albumin_index_log2 ~ . - albumin_hi,
                 data   = d_train,
                 family = gaussian())

# Derive ONE frozen cut-point on the full training set
train_pred <- predict(final_glm, newdata = d_train)
thr_final  <- choose_threshold(d_train$albumin_hi, train_pred)  # still Youden J

# optionally round or publish as is
cat("Frozen threshold (log2 scale):", round(thr_final, 3), "\n")

# ------------------------------------------------------------------
# 4.  Evaluate on the untouched 20 % test set
# ------------------------------------------------------------------
test_pred  <- predict(final_glm, newdata = d_test)

test_res   <- tibble(
  truth    = d_test$albumin_hi |> forcats::fct_relevel("normal", "elevated"),
  pred_num = test_pred,                                    # continuous
  pred_cls = factor(test_pred > thr_final,                 # binary
                    levels = c(FALSE, TRUE),
                    labels  = c("normal", "elevated"))
)

cat("\n=== TEST-SET PERFORMANCE ===\n")

## 4·1  confusion matrix
print(conf_mat(test_res, truth = truth, estimate = pred_cls))

acc <- accuracy_vec(truth = test_res$truth,
                    estimate = test_res$pred_cls)
sens <- sensitivity_vec(truth = test_res$truth,
                        estimate = test_res$pred_cls,
                        event_level = "second")
spec <- specificity_vec(truth = test_res$truth,
                        estimate = test_res$pred_cls,
                        event_level = "second")

tibble(metric = c("accuracy", "sensitivity", "specificity"),
       value  = c(acc, sens, spec))


# ------------------------------------------------------------------
# 5.  Save model & threshold for deployment
# ------------------------------------------------------------------
saveRDS(final_glm, "albumin_glm_v1.rds")
saveRDS(thr_final, "albumin_thr_v1.rds")

# l_csf_total_protein_log2 has albumin

```

for now this is good i belive there is no realtionship exist btw then we applied the glm in the poster too anyway let go for now and need to read my way again

# This code is for myself to investigate baseline vs non baseline in DX and Investigate more I did talk about this to Erik , also to see age effect on labumin index

```{r}
# This code is for myself to investigate baseline vs non baseline in DX 
# Critical Finding:
# 
# ALL diagnosis groups have NA for non-baseline observations
# Only the NA diagnosis group has any non-baseline data (8 observations)
# Every other group is 100% baseline data
# ===================================================================
# INVESTIGATE BASELINE vs NON-BASELINE MYSTERY
# Why do filtered and unfiltered datasets look so similar?
# ===================================================================
library(dplyr)
library(ggplot2)
library(tidyr)
library(patchwork)


# cat("\n=== DIAGNOSIS BREAKDOWN BY VISIT TYPE ===\n")
# 
# # For subjects with albumin data
# dx_visit_breakdown <- dat_all_combine %>%
#   filter(!is.na(fl_albumin_index_log2)) %>%
#   mutate(
#     visit_type = if_else(redcap_event_name %in% baseline_visits, 
#                         "Baseline", "Non_Baseline")  # Use underscore instead of hyphen
#   ) %>%
#   group_by(dx_biological, visit_type) %>%
#   summarise(
#     n_observations = n(),
#     n_subjects = n_distinct(ptid),  # Fixed: use subject_id instead of ptid
#     median_albumin = median(fl_albumin_index_log2, na.rm = TRUE),
#     mean_albumin = mean(fl_albumin_index_log2, na.rm = TRUE),
#     sd_albumin = sd(fl_albumin_index_log2, na.rm = TRUE),
#     .groups = 'drop'
#   ) %>%
#   pivot_wider(
#     names_from = visit_type,
#     values_from = c(n_observations, n_subjects, median_albumin, mean_albumin, sd_albumin),
#     names_sep = "_"
#   ) %>%
#   mutate(
#     # Calculate differences - now using proper column names
#     median_diff = median_albumin_Baseline - median_albumin_Non_Baseline,
#     mean_diff = mean_albumin_Baseline - mean_albumin_Non_Baseline
#   )
# 
# print(dx_visit_breakdown)



#Investigates Age Effect:

# Tests correlation between age and albumin in controls
# Shows if older "controls" have higher albumin (BBB dysfunction)

# Discovered:
# Control group heterogeneity is explained by age
# ===================================================================
# 1. INVESTIGATE AGE EFFECT IN CONTROLS
# ==================================================================
cat("=== AGE EFFECT IN CONTROLS ===\n")

# Focus on controls only
controls_data <- dat_baseline_AD_VD %>%
  filter(dx_biological == "Control", 
         !is.na(fl_albumin_index_log2), 
         !is.na(age))

cat("Total controls with albumin and age data:", 
    nrow(controls_data), "\n")
cat("Age range:", round(min(controls_data$age), 1), 
    "-", 
    round(max(controls_data$age), 1), 
    "years\n")
cat("Mean age:",
    round(mean(controls_data$age), 1), 
    "years\n")

# Check age-albumin correlation in controls
age_albumin_cor <- cor(controls_data$age, controls_data$fl_albumin_index_log2, use = "complete.obs")
cat("Age-albumin correlation in controls:", round(age_albumin_cor, 3), "\n")

# # Test statistical significance
# age_albumin_test <- cor.test(controls_data$age, controls_data$fl_albumin_index_log2)
# cat("Correlation p-value:", round(age_albumin_test$p.value, 4), "\n")

e_plot_scatterplot (
  dat_plot = controls_data,
  var_x = "age",
  var_y = "fl_albumin_index_log2",
  smooth_all = c("lm")
)

# shapiro.test(controls_data$age)
# shapiro.test(controls_data$fl_albumin_index_log2)

# ===================================================================
# 2. VISUALIZE AGE EFFECT
# ===================================================================

# Age vs albumin scatter plot for controls
# age_albumin_plot <- ggplot(controls_data, aes(x = age, y = fl_albumin_index_log2)) +
#   geom_point(alpha = 0.6, size = 2) +
#   geom_smooth(method = "lm", se = TRUE, color = "blue") +
#   geom_smooth(method = "loess", se = FALSE, color = "red", linetype = "dashed") +
#   labs(title = "Age vs Albumin Index in Controls",
#        subtitle = paste("r =", round(age_albumin_cor, 3), 
#                         ", p =", round(age_albumin_test$p.value, 4)),
#        x = "Age (years)", y = "Albumin Index (log2)") +
#   theme_bw()
# 
# print(age_albumin_plot)




# ===================================================================
# 3. AGE-STRATIFIED CONTROL GROUPS
# ===================================================================

cat("\n=== AGE-STRATIFIED CONTROL ANALYSIS ===\n")

# Define age groups for controls
controls_age_stratified <- controls_data %>%
  mutate(
    age_group = case_when(
      age < 50 ~ "Young (<50)",
      age >= 50 & age < 65 ~ "Middle (50-64)", 
      age >= 65 & age < 75 ~ "Older (65-74)",
      age >= 75 ~ "Elderly (75+)",
      TRUE ~ "Unknown"
    ),
    age_group = factor(age_group, 
                       levels = c("Young (<50)",
                                  "Middle (50-64)", 
                                  "Older (65-74)", 
                                  "Elderly (75+)"))
  )

# Summary by age group
age_group_summary <- controls_age_stratified %>%
  group_by(age_group) %>%
  summarise(
    n = n(),
    mean_age = round(mean(age), 1),
    mean_albumin = round(mean(fl_albumin_index_log2), 3),
    sd_albumin = round(sd(fl_albumin_index_log2), 3),
    median_albumin = round(median(fl_albumin_index_log2), 3),
    q75_albumin = round(quantile(fl_albumin_index_log2, 0.75), 3),
    q90_albumin = round(quantile(fl_albumin_index_log2, 0.90), 3),
    q95_albumin = round(quantile(fl_albumin_index_log2, 0.95), 3),
    .groups = 'drop'
  )

cat("Controls by age group:\n")
print(age_group_summary)

# ===================================================================
# 4. CHOOSE OPTIMAL AGE-STRATIFIED CUTPOINT
# ===================================================================

#  Use younger, healthier controls for cut-point
#Older controls may have subclinical BBB dysfunction, or not sure need to ask Erik

# Option 1: Young controls only (<50 years)
young_controls <- controls_age_stratified %>% filter(age < 50)

# Option 2: Young + middle-aged controls (<65 years)
youngmid_controls <- controls_age_stratified %>% filter(age < 65)

# Option 3: Exclude elderly (< 75 years)
non_elderly_controls <- controls_age_stratified %>% filter(age < 75)

cat("\n=== AGE-STRATIFIED CUT-POINT OPTIONS ===\n")

# Calculate cut-points for each age strategy
if(nrow(young_controls) >= 20) {
  cut_75_young <- quantile(young_controls$fl_albumin_index_log2, 0.75, na.rm = TRUE)
  cat("Young controls (<50y, n=", nrow(young_controls), ") - 75th percentile:", round(cut_75_young, 3), "\n")
} else {
  cat("Young controls (<50y) - insufficient sample (n=", nrow(young_controls), ")\n")
  cut_75_young <- NA
}

if(nrow(youngmid_controls) >= 30) {
  cut_75_youngmid <- quantile(youngmid_controls$fl_albumin_index_log2, 0.75, na.rm = TRUE)
  cat("Young+Mid controls (<65y, n=", nrow(youngmid_controls), ") - 95th percentile:", round(cut_95_youngmid, 3), "\n")
} else {
  cat("Young+Mid controls (<65y) - insufficient sample (n=", nrow(youngmid_controls), ")\n")
  cut_75_youngmid <- NA
}

if(nrow(non_elderly_controls) >= 40) {
  cut_75_nonelderly <- quantile(non_elderly_controls$fl_albumin_index_log2, 0.75, na.rm = TRUE)
  cat("Non-elderly controls (<75y, n=", nrow(non_elderly_controls), ") - 75th percentile:", round(cut_95_nonelderly, 3), "\n")
} else {
  cat("Non-elderly controls (<75y) - insufficient sample (n=", nrow(non_elderly_controls), ")\n")
  cut_75_nonelderly <- NA
}

# For comparison: all controls
cut_75_all_controls <- quantile(controls_data$fl_albumin_index_log2, 0.75, na.rm = TRUE)
cat("All controls (n=", nrow(controls_data), ") - 75th percentile:", round(cut_75_all_controls, 3), "\n")

# ===================================================================
# 5. VISUALIZE AGE DISTRIBUTIONS
# ===================================================================

# Boxplot by age group
  age_boxplot <- ggplot(controls_age_stratified, aes(x = age_group, y = fl_albumin_index_log2)) +
    geom_boxplot(fill = "lightblue", alpha = 0.7) +
    geom_jitter(width = 0.2, alpha = 0.5) +
    stat_summary(fun = mean, geom = "point", shape = 23, size = 3, fill = "red") +
    labs(title = "Albumin Index Distribution by Age Group (Controls Only)",
         subtitle = "Red diamonds = group means",
         x = "Age Group", y = "Albumin Index (log2)") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  print(age_boxplot)



```




Add these pairwise correlation matrix
NFL, GFAP and others (ILs, etc) vs MMPs
That is, for each inflammatory biomarkers, relate them to others to see if there are any relationships -- maybe summarize in a heatmap with corr values overlaid.
Maybe a big correlation plot for all fluid variables together. I think there's a e_plot_correlation function.
CSF-only, Plasma-only


```{r}
# Load required libraries
library(dplyr)
library(corrplot)
library(Hmisc)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(tidyselect)

# ====== 1. IDENTIFY VARIABLE GROUPS ======

# CSF variables
csf_vars <- names(dat_sub_without_igg)[grepl("^fl_.*csf|^fl_msd_csf|^fl_cyto_csf|^fl_angio_csf|fl_albumin_index_log2|fl_abeta_4240_ratio_log2|fl_myelin_basic_protein_log2", names(dat_sub_without_igg))]

# Plasma variables  
plasma_vars <- names(dat_sub_without_igg)[grepl("^fl_plasma|^fl_cyto_heparin_plasma", names(dat_sub_without_igg))]

# MMPs (both CSF and plasma)
mmp_vars <- names(dat_sub_without_igg)[grepl("mmp", names(dat_sub_without_igg), ignore.case = TRUE)]

# Inflammatory markers (ILs, TNF, IFN)
inflammatory_vars <- names(dat_sub_without_igg)[grepl("il_|tnf_|ifn_", names(dat_sub_without_igg), ignore.case = TRUE)]

# Neurological markers (NFL, GFAP, etc.)
neuro_vars <- names(dat_sub_without_igg)[grepl("nfl_|gfap_|ptau_|abeta", names(dat_sub_without_igg), ignore.case = TRUE)]

# Print variable groups for verification
cat("CSF variables (", length(csf_vars), "):\n")
cat(paste(csf_vars, collapse = "\n"), "\n\n")

cat("Plasma variables (", length(plasma_vars), "):\n")
cat(paste(plasma_vars, collapse = "\n"), "\n\n")

cat("MMP variables (", length(mmp_vars), "):\n")
cat(paste(mmp_vars, collapse = "\n"), "\n\n")

cat("Inflammatory variables (", length(inflammatory_vars), "):\n")
cat(paste(inflammatory_vars, collapse = "\n"), "\n\n")


cat("Neurological variables (", length(neuro_vars), "):\n")
cat(paste(neuro_vars, collapse = "\n"), "\n\n")

# ====== 2. CORRELATION ANALYSIS FUNCTION ======

create_correlation_plot <- function(data, var_list, title_text, method = "pearson", 
                                   plot_method = "ellipse", order_method = "AOE") {
  
  # Select and prepare data
  dat_sel_mat <- data %>%
    select(all_of(var_list)) %>%
    select(where(is.numeric)) %>%
    as.matrix()
  
  # Remove variables with too many missing values (>50%)
  missing_prop <- colSums(is.na(dat_sel_mat)) / nrow(dat_sel_mat)
  dat_sel_mat <- dat_sel_mat[, missing_prop < 0.5, drop = FALSE]
  
  if (ncol(dat_sel_mat) < 2) {
    cat("Not enough variables for correlation analysis\n")
    return(NULL)
  }
  
  # Calculate correlation matrix
  tab_cor <- rcorr(dat_sel_mat, type = method)
  
  # Correlation p-values and confidence intervals
  cor_mtest_out <- cor.mtest(mat = dat_sel_mat, conf.level = 0.95)
  
  # Create correlation plot
  corrplot(
    corr = tab_cor$r,
    # is.corr = TRUE,
    type = "upper",
    # p.mat = cor_mtest_out$p,
    method = plot_method,
    # tl.pos = "d",
    # tl.col = "black",
    # tl.cex = 0.8,
    # bg = "white",
    # addgrid.col = "grey",
    # addCoef.col = "black",
    # number.cex = 0.6,
    # mar = c(0, 0, 2, 0),
    # title = title_text,
    # insig = "pch",
    order = order_method
    # sig.level = 0.05
  )
  
  return(list(correlation = tab_cor$r, p_values = tab_cor$P, data_matrix = dat_sel_mat))
}

# ====== 3. CSF-ONLY CORRELATION ANALYSIS ======

cat("=== CSF-ONLY CORRELATION ANALYSIS ===\n")
par(mfrow = c(1, 1))

csf_result <- create_correlation_plot(
  data = dat_sub_without_igg,
  var_list = csf_vars,
  title_text = "CSF Biomarkers Correlation Matrix",
  method = "pearson",
  plot_method = "ellipse"
)

# ====== 4. PLASMA-ONLY CORRELATION ANALYSIS ======

cat("\n=== PLASMA-ONLY CORRELATION ANALYSIS ===\n")

plasma_result <- create_correlation_plot(
  data = dat_sub_without_igg,
  var_list = plasma_vars,
  title_text = "Plasma Biomarkers Correlation Matrix",
  method = "pearson",
  plot_method = "ellipse"
)

# ====== 5. NFL, GFAP, INFLAMMATORY vs MMPs ANALYSIS ======

cat("\n=== NFL, GFAP, INFLAMMATORY vs MMPs ANALYSIS ===\n")

# Combine neurological and inflammatory markers
neuro_inflam_vars <- c(neuro_vars, inflammatory_vars)
neuro_inflam_mmp_vars <- c(neuro_inflam_vars, mmp_vars)

neuro_mmp_result <- create_correlation_plot(
  data = dat_sub_without_igg,
  var_list = neuro_inflam_mmp_vars,
  title_text = "NFL, GFAP, Inflammatory vs MMPs Correlation",
  method = "pearson",
  plot_method = "ellipse"
)

# ====== 6. MIXED CORRELATION PLOT (using corrplot.mixed) ======

cat("\n=== MIXED CORRELATION PLOT ===\n")

# Create mixed correlation plot for CSF variables
if (!is.null(csf_result)) {
  corrplot.mixed(
    csf_result$correlation,
    order = 'AOE',
    tl.pos = 'lt',
    tl.col = 'black',
    tl.cex = 0.7,
    number.cex = 0.6,
    title = "CSF Biomarkers - Mixed Correlation Plot",
    mar = c(0, 0, 2, 0)
  )
}

# Create mixed correlation plot for Plasma variables
if (!is.null(plasma_result)) {
  corrplot.mixed(
    plasma_result$correlation,
    order = 'AOE',
    tl.pos = 'lt',
    tl.col = 'black',
    tl.cex = 0.7,
    number.cex = 0.6,
    title = "Plasma Biomarkers - Mixed Correlation Plot",
    mar = c(0, 0, 2, 0)
  )
}
# ====== 7. IDENTIFY STRONG CORRELATIONS ======

identify_strong_correlations <- function(cor_matrix, threshold = 0.6) {
  # Convert to data frame for easier manipulation
  cor_df <- as.data.frame(as.table(cor_matrix))
  names(cor_df) <- c("Var1", "Var2", "Correlation")
  
  # Remove self-correlations and duplicates
  cor_df <- cor_df %>%
    filter(Var1 != Var2) %>%
    filter(abs(Correlation) >= threshold) %>%
    arrange(desc(abs(Correlation)))
  
  # Remove duplicate pairs (A-B is same as B-A)
  cor_df$pair <- paste(pmin(as.character(cor_df$Var1), as.character(cor_df$Var2)),
                       pmax(as.character(cor_df$Var1), as.character(cor_df$Var2)),
                       sep = " - ")
  cor_df <- cor_df[!duplicated(cor_df$pair), ]
  
  return(cor_df)
}

# Find strong correlations in CSF data
if (!is.null(csf_result)) {
  strong_csf_cors <- identify_strong_correlations(csf_result$correlation, threshold = 0.6)
  cat("Strong CSF correlations (|r| >= 0.6):\n")
  print(strong_csf_cors[, c("Var1", "Var2", "Correlation")])
}

# Find strong correlations in plasma data
if (!is.null(plasma_result)) {
  strong_plasma_cors <- identify_strong_correlations(plasma_result$correlation, threshold = 0.6)
  cat("\nStrong Plasma correlations (|r| >= 0.6):\n")
  print(strong_plasma_cors[, c("Var1", "Var2", "Correlation")])
}

# Find strong correlations in plasma data
if (!is.null(neuro_mmp_result)) {
  strong_neuro_cors <- identify_strong_correlations(neuro_mmp_result$correlation, threshold = 0.6)
  cat("\nStrong neuro correlations (|r| >= 0.6):\n")
  print(strong_neuro_cors[, c("Var1", "Var2", "Correlation")])
}

# ====== 8. SCATTERPLOT GRID FOR STRONG CORRELATIONS ======

create_scatterplot_grid <- function(data, correlation_df, max_plots = 12) {
  # Limit to top correlations
  correlation_df <- head(correlation_df, max_plots)
  
  plot_list <- list()
  
  for (i in 1:nrow(correlation_df)) {
    var1 <- as.character(correlation_df$Var1[i])
    var2 <- as.character(correlation_df$Var2[i])
    cor_val <- round(correlation_df$Correlation[i], 3)
    
    # Create scatterplot
    p <- ggplot(data, aes_string(x = var1, y = var2)) +
      geom_point(alpha = 0.6, color = "steelblue") +
      geom_smooth(method = "lm", se = TRUE, color = "red", size = 0.8) +
      labs(
        x = gsub("fl_|_log2", "", var1),
        y = gsub("fl_|_log2", "", var2),
        title = paste("r =", cor_val)
      ) +
      theme_minimal() +
      theme(
        axis.title = element_text(size = 8),
        plot.title = element_text(size = 9, hjust = 0.5),
        axis.text = element_text(size = 7)
      )
    
    plot_list[[i]] <- p
  }
  
  return(plot_list)
}

# Create scatterplot grids for strong correlations
if (!is.null(csf_result) && nrow(strong_csf_cors) > 0) {
  cat("\n=== CSF STRONG CORRELATIONS SCATTERPLOTS ===\n")
  csf_plots <- create_scatterplot_grid(dat_sub_without_igg, strong_csf_cors, max_plots = 9)
  grid.arrange(grobs = csf_plots, ncol = 3, top = "Strong CSF Correlations")
}

if (!is.null(plasma_result) && nrow(strong_plasma_cors) > 0) {
  cat("\n=== PLASMA STRONG CORRELATIONS SCATTERPLOTS ===\n")
  plasma_plots <- create_scatterplot_grid(dat_sub_without_igg, strong_plasma_cors, max_plots = 9)
  grid.arrange(grobs = plasma_plots, ncol = 3, top = "Strong Plasma Correlations")
}

if (!is.null(neuro_mmp_result) && nrow(strong_neuro_cors) > 0) {
  cat("\n=== Nuero STRONG CORRELATIONS SCATTERPLOTS ===\n")
  neuro_plots <- create_scatterplot_grid(dat_sub_without_igg, strong_neuro_cors, max_plots = 9)
  grid.arrange(grobs = neuro_plots, ncol = 3, top = "Strong neuro_mmp_result Correlations")
}

# ====== 9. SUMMARY HEATMAP WITH VALUES ======

cat("\n=== SUMMARY HEATMAP WITH CORRELATION VALUES ===\n")

# Create summary heatmap for all fluid variables
all_fluid_vars <- c(csf_vars, plasma_vars)

summary_result <- create_correlation_plot(
  data = dat_sub_without_igg,
  var_list = all_fluid_vars,
  title_text = "All Fluid Biomarkers - Correlation Heatmap",
  method = "pearson",
  plot_method = "number"  # Shows correlation values
)

# ====== 10. SPECIFIC NFL/GFAP vs MMP ANALYSIS ======

cat("\n=== SPECIFIC NFL/GFAP vs MMP CORRELATION ===\n")

# Create specific correlation matrix for NFL/GFAP vs MMPs
nfl_gfap_vars <- grep("nfl|gfap", names(dat_sub_without_igg), value = TRUE, ignore.case = TRUE)
specific_vars <- c(nfl_gfap_vars, mmp_vars, inflammatory_vars)

specific_result <- create_correlation_plot(
  data = dat_sub_without_igg,
  var_list = specific_vars,
  title_text = "NFL/GFAP vs MMPs and Inflammatory Markers",
  method = "pearson",
  plot_method = "ellipse"
)

cat("Analysis complete!\n")
cat("Key outputs:\n")
cat("1. CSF-only correlation matrix\n")
cat("2. Plasma-only correlation matrix\n") 
cat("3. NFL/GFAP/Inflammatory vs MMPs correlation\n")
cat("4. Mixed correlation plots\n")
cat("5. Strong correlations identified\n")
cat("6. Scatterplot grids for strong correlations\n")
cat("7. Summary heatmap with correlation values\n")
```


# make pdf of all corr result

Note: some correlation found 

```{r}
# 1. Open a PDF device; onefile=TRUE (the default) means multiple pages
pdf("all_correlation_plots.pdf", width = 7, height = 7)

# 2. CSF-only correlation matrix
if (!is.null(csf_result)) {
  corrplot(
    corr  = csf_result$correlation,
    type  = "upper",
    method= "ellipse",
    order = "AOE",
    title = "CSF Biomarkers Correlation Matrix",
    mar   = c(0,0,2,0)
  )
}

# 3. Plasma-only correlation matrix
if (!is.null(plasma_result)) {
  corrplot(
    corr  = plasma_result$correlation,
    type  = "upper",
    method= "ellipse",
    order = "AOE",
    title = "Plasma Biomarkers Correlation Matrix",
    mar   = c(0,0,2,0)
  )
}

# 4. Mixed correlation plots
if (!is.null(csf_result)) {
  corrplot.mixed(
    csf_result$correlation,
    order  = "AOE",
    tl.pos = "lt",
    tl.cex = 0.7,
    number.cex = 0.6,
    title = "CSF Biomarkers - Mixed",
    mar   = c(0,0,2,0)
  )
}
if (!is.null(plasma_result)) {
  corrplot.mixed(
    plasma_result$correlation,
    order  = "AOE",
    tl.pos = "lt",
    tl.cex = 0.7,
    number.cex = 0.6,
    title = "Plasma Biomarkers - Mixed",
    mar   = c(0,0,2,0)
  )
}

# 5. Scatter‐plots for strong CSF correlations (one per page)
if (exists("csf_plots")) {
  for (p in csf_plots) {
    print(p)
  }
}

# 6. Scatter‐plots for strong Plasma correlations
if (exists("plasma_plots")) {
  for (p in plasma_plots) {
    print(p)
  }
}

# 7. Scatter‐plots for strong Neuro/MMP correlations
if (exists("neuro_plots")) {
  for (p in neuro_plots) {
    print(p)
  }
}

# 8. Close the device
dev.off()

```


