---
title:        "[CSF] ~ [Plasma] for all biomarkers, scatterplots colored by DX group"
subtitle:     "Based on Erhardt, et al., (2021). *Inflammatory biomarkers aid in diagnosis of dementia Frontiers in Aging Neuroscience*"
author:       Azadeh Golduzian
date:         last-modified # today, now, last-modified
date-format:  long          # full, long, medium, short, iso,   https://quarto.org/docs/reference/dates.html
format:
  html:
    theme:                  litera
    highlight-style:        atom-one
    page-layout:            full      # article, full   # https://quarto.org/docs/output-formats/page-layout.html
    toc:                    true
    toc-depth:              4
    toc-location:           body      # left, body, right
    number-sections:        true      # true, false
    number-depth:           3
    code-fold:              show      # true (initially hidden), false, show (initially shown)
    code-tools:                       # menu top-right to show/hide all code
      toggle:               true
      caption:              "Code"    # none
      source:               false
    code-overflow:          scroll    # scroll, wrap
    code-block-bg:          true
    code-block-border-left: "#30B0E0"
    df-print:               paged     # default, kable, tibble, paged   # https://quarto.org/docs/computations/r.html
fig-width:                  6
fig-height:                 4
execute: # https://quarto.org/docs/computations/execution-options.html, https://quarto.org/docs/computations/r.html
  cache:    false   # false, true
  eval:     true    # true, false  Evaluate the code chunk (if false, just echos the code into the output).
  echo:     true    # true, false  Include the source code in output
---

# Azadeh code

```{r}

load("dat_sub_Data_inflammation_analysis_paper_06052025.RData")
load("dat_sub_and_Var_for_inflammation_analysis.RData")


# library(tidyverse)
# library(ggplot2)
# library(patchwork)
# library(scales)

dat_var_label_matching <-
  tibble::tribble(
    ~Var                                    , ~Group  , ~Type                          , ~Scale
  #, "ptid"                                 ,         ,                                ,
  #, "dx_current"                           ,         ,                                ,
  #, "sex"                                  ,         ,                                ,
  #, "demo_edu_yrs"                         ,         ,                                ,
  #, "age"                                  ,         ,                                ,
  ##### CSF                                 ,         ,                                ,
  , "fl_albumin_index_log2"                 , "CSF"   , "CSF Permeability"             , "log2"
  , "fl_igg_index_log2"                     , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf_igg_log2"                       , "CSF"   , "CSF Other"                    , "log2"
  , "fl_msd_csf_mmp_1_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_2_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_3_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_msd_csf_mmp_9_log2"                 , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_csf__ALL__mmp_10_log2"              , "CSF"   , "CSF MMP"                      , "log2"
  , "fl_angio_csf_bfgf_log2"                , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_flt_1_log2"               , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_tie2_log2"                , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_a_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_c_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_angio_csf_vegf_d_log2"              , "CSF"   , "CSF Angio"       , "log2"
  , "fl_cyto_csf_ifn_g_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_csf__ALL__il_1b_log2"               , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_2_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_4_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_6_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_8_log2"                 , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_10_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_12p70_log2"             , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_il_13_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_cyto_csf_tnf_a_log2"                , "CSF"   , "CSF Cytokines"                , "log2"
  , "fl_csf__ALL__nfl_log2"                 , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf__ALL__plgf_log2"                , "CSF"   , "CSF Other"                    , "log2"
  , "fl_csf__ALL__gfap_log2"                , "CSF"   , "CSF Other"                    , "log2"
  , "fl_abeta_4240_ratio_log2"              , "CSF"   , "CSF AD"                       , "log2"
  , "fl_csf__ALL__ptau_log2"                , "CSF"   , "CSF AD"                       , "log2"
  , "fl_myelin_basic_protein_log2"          , "CSF"   , "CSF Other"                    , "log2"
  ##### Plasma                              ,         ,                                ,
  , "fl_serum_igg_log2"                     , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__mmp_1_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_2_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_3_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_9_log2"            , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__mmp_10_log2"           , "Plasma", "Plasma MMP"                   , "log2"
  , "fl_plasma__ALL__bfgf_log2"             , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__flt_1_log2"            , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__tie2_log2"             , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_a_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_c_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_plasma__ALL__vegf_d_log2"           , "Plasma", "Plasma Angio"    , "log2"
  , "fl_cyto_heparin_plasma_ifn_g_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_plasma__ALL__il_1b_log2"            , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_2_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_4_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_6_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_8_log2"      , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_10_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_12p70_log2"  , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_il_13_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_cyto_heparin_plasma_tnf_a_log2"     , "Plasma", "Plasma Cytokines"             , "log2"
  , "fl_plasma__ALL__nfl_log2"              , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__plgf_log2"             , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__gfap_log2"             , "Plasma", "Plasma Other"                 , "log2"
  , "fl_plasma__ALL__abeta_40_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__abeta_42_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__ptau_181_log2"         , "Plasma", "Plasma AD"                    , "log2"
  , "fl_plasma__ALL__ptau_217_log2"         , "Plasma", "Plasma AD"                    , "log2"
  ##### MRI                                 ,         ,                                ,
  , "mri_percent_vol_hippo"                 , "MRI"   , "MRI"                          , "org"
  # , "mri_ucd_wmhv_percent"                  , "MRI"   , "MRI Other"                    , "org"  # 2/19/2025 too few observations
  , "mri_mean_cort_thick_mm"                , "MRI"   , "MRI"                          , "org"
  , "mri_perm_3t_tot"                       , "MRI"   , "MRI"                          , "org"
  , "mri_dti_psmd"                          , "MRI"   , "MRI"                          , "org"
  , "mri_dti_freewater_mean"                , "MRI"   , "MRI"                          , "org"
  )

 # this code generate pairs by pairs 
# # Create matching pairs of CSF and Plasma biomarkers based on naming patterns
create_biomarker_pairs <- function() {
  # Extract biomarker names and create pairs
  pairs <- tibble::tribble(
    ~CSF, ~Plasma, ~Biomarker,
    # MMP biomarkers
    "fl_msd_csf_mmp_1_log2", "fl_plasma__ALL__mmp_1_log2", "MMP-1",
    "fl_msd_csf_mmp_2_log2", "fl_plasma__ALL__mmp_2_log2", "MMP-2",
    "fl_msd_csf_mmp_3_log2", "fl_plasma__ALL__mmp_3_log2", "MMP-3",
    "fl_msd_csf_mmp_9_log2", "fl_plasma__ALL__mmp_9_log2", "MMP-9",
    "fl_csf__ALL__mmp_10_log2", "fl_plasma__ALL__mmp_10_log2", "MMP-10",

    # Angiogenic factors
    "fl_angio_csf_bfgf_log2", "fl_plasma__ALL__bfgf_log2", "bFGF",
    "fl_angio_csf_flt_1_log2", "fl_plasma__ALL__flt_1_log2", "FLT-1",
    "fl_angio_csf_tie2_log2", "fl_plasma__ALL__tie2_log2", "TIE2",
    "fl_angio_csf_vegf_a_log2", "fl_plasma__ALL__vegf_a_log2", "VEGF-A",
    "fl_angio_csf_vegf_c_log2", "fl_plasma__ALL__vegf_c_log2", "VEGF-C",
    "fl_angio_csf_vegf_d_log2", "fl_plasma__ALL__vegf_d_log2", "VEGF-D",

    # Cytokines
    "fl_cyto_csf_ifn_g_log2", "fl_cyto_heparin_plasma_ifn_g_log2", "IFN-γ",
    "fl_csf__ALL__il_1b_log2", "fl_plasma__ALL__il_1b_log2", "IL-1β",
    "fl_cyto_csf_il_2_log2", "fl_cyto_heparin_plasma_il_2_log2", "IL-2",
    "fl_cyto_csf_il_4_log2", "fl_cyto_heparin_plasma_il_4_log2", "IL-4",
    "fl_cyto_csf_il_6_log2", "fl_cyto_heparin_plasma_il_6_log2", "IL-6",
    "fl_cyto_csf_il_8_log2", "fl_cyto_heparin_plasma_il_8_log2", "IL-8",
    "fl_cyto_csf_il_10_log2", "fl_cyto_heparin_plasma_il_10_log2", "IL-10",
    "fl_cyto_csf_il_12p70_log2", "fl_cyto_heparin_plasma_il_12p70_log2", "IL-12p70",
    "fl_cyto_csf_il_13_log2", "fl_cyto_heparin_plasma_il_13_log2", "IL-13",
    "fl_cyto_csf_tnf_a_log2", "fl_cyto_heparin_plasma_tnf_a_log2", "TNF-α",

    # Other markers
    "fl_csf__ALL__nfl_log2", "fl_plasma__ALL__nfl_log2", "NFL",
    "fl_csf__ALL__plgf_log2", "fl_plasma__ALL__plgf_log2", "PLGF",
    "fl_csf__ALL__gfap_log2", "fl_plasma__ALL__gfap_log2", "GFAP",

    # AD markers
    "fl_csf__ALL__ptau_log2", "fl_plasma__ALL__ptau_181_log2", "pTau-181",
    # "fl_csf__ALL__ptau_log2", "fl_plasma__ALL__ptau_217_log2", "pTau-217" # wrong I could not find CSF for 217
    
  
    
  )

  # Add group information based on dat_var_label_matching
  pairs <- pairs %>%
    left_join(
      dat_var_label_matching %>% select(Var, Type) %>% rename(CSF = Var, CSF_Type = Type),
      by = "CSF"
    ) %>%
    left_join(
      dat_var_label_matching %>% select(Var, Type) %>% rename(Plasma = Var, Plasma_Type = Type),
      by = "Plasma"
    ) %>%
    mutate(
      Group = case_when(
        str_detect(CSF_Type, "MMP") ~ "Matrix Metalloproteinases",
        str_detect(CSF_Type, "Angiogenic") ~ "Angiogenic Factors",
        str_detect(CSF_Type, "Cytokines") ~ "Cytokines",
        str_detect(CSF_Type, "AD") ~ "AD Biomarkers",
        Biomarker %in% c("NFL", "PLGF", "GFAP") ~ "Neurodegeneration Markers",
        TRUE ~ "Other Biomarkers"
      )
    )

  return(pairs)
}
#
# # Define color palette for diagnosis groups
dx_colors <- c(
  "AD" = "#E41A1C",        # Red
  "SIVD" = "#377EB8",      # Blue
  "Mixed" = "#4DAF4A",     # Green
  "LA" = "#984EA3",        # Purple
  "MI" = "#FF7F00",        # Orange
  "Unknown" = "#A6A6A6",   # Gray
  "Control" = "#F781BF",   # Pink
  "Repeat pt" = "#999999"  # Dark gray
)




# Function to create a single scatterplot
create_biomarker_plot <- function(data, csf_var, plasma_var, biomarker_name) {
  # Check if both variables exist in the data
  if (!all(c(csf_var, plasma_var) %in% names(data))) {
    return(NULL)
  }

  # Prepare plot data
  plot_data <- data %>%
    select(all_of(c(csf_var, plasma_var, "dx_current"))) %>%
    filter(!is.na(.data[[csf_var]]) & !is.na(.data[[plasma_var]]))

  # Calculate overall correlation
  overall_cor <- cor(plot_data[[csf_var]], plot_data[[plasma_var]],
                     use = "complete.obs", method = "pearson")

  # Calculate correlation by diagnosis group
  group_cors <- plot_data %>%
    group_by(dx_current) %>%
    summarize(
      n = n(),
      r = cor(.data[[csf_var]], .data[[plasma_var]], use = "complete.obs"),
      .groups = 'drop'
    ) %>%
    filter(n >= 3)  # Only show correlation for groups with at least 3 samples

  # Create plot
  p <- ggplot(plot_data, aes(x = .data[[plasma_var]], y = .data[[csf_var]])) +
    # Add points colored by diagnosis
    geom_point(aes(color = dx_current), alpha = 0.6, size = 2) +
    # Add overall regression line
    geom_smooth(method = "lm", se = TRUE, color = "black",
                linetype = "solid", linewidth = 0.8, alpha = 0.3) +
    # Add regression lines for each diagnosis group
    scale_color_manual(values = dx_colors, name = "Diagnosis") +
    labs(
      title = biomarker_name,
      subtitle = sprintf("Overall r = %.3f, n = %d", overall_cor, nrow(plot_data)),
      x = "Plasma (log2)",
      y = "CSF (log2)"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      plot.title = element_text(size = 11, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 8, hjust = 0.5, color = "gray40"),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(fill = NA, color = "gray70"),
      legend.position = "right",
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 7),
      legend.key.size = unit(0.8, "lines"),
      axis.text = element_text(size = 7),
      axis.title = element_text(size = 8)
    )

  # Add correlation text for each group
  cor_text <- group_cors %>%
    mutate(label = sprintf("%s: r=%.2f (n=%d)", dx_current, r, n)) %>%
    summarize(label = paste(label, collapse = "\n"))

  if (nrow(group_cors) > 0) {
    p <- p +
      annotate("text", x = Inf, y = Inf, label = cor_text$label,
               hjust = 1.05, vjust = 1.05, size = 2.5, color = "gray30",
               fontface = "italic")
  }

  return(p)
}

# Create all biomarker pairs
biomarker_pairs <- create_biomarker_pairs()

# Generate plots organized by group
plots_by_group <- list()
group_order <- c("Matrix Metalloproteinases", "Angiogenic Factors",
                 "Cytokines", "Neurodegeneration Markers", "AD Biomarkers",
                 "Other Biomarkers")

for (group in group_order) {
  group_pairs <- biomarker_pairs %>% filter(Group == group)

  if (nrow(group_pairs) > 0) {
    group_plots <- list()

    for (i in seq_len(nrow(group_pairs))) {
      p <- create_biomarker_plot(
        data = dat_sub,
        csf_var = group_pairs$CSF[i],
        plasma_var = group_pairs$Plasma[i],
        biomarker_name = group_pairs$Biomarker[i]
      )

      if (!is.null(p)) {
        group_plots[[length(group_plots) + 1]] <- p
      }
    }

    if (length(group_plots) > 0) {
      plots_by_group[[group]] <- group_plots
    }
  }
}

# Create PDF with multiple plots per page
pdf("csf_plasma_biomarker_comparisons.pdf", width = 11, height = 8.5)

# Title page
plot.new()
text(0.5, 0.6, "CSF vs Plasma Biomarker Correlations",
     cex = 2.5, font = 2, col = "darkblue")
text(0.5, 0.5, "Comparison by Diagnosis Groups",
     cex = 1.5, font = 1, col = "darkblue")
text(0.5, 0.3, format(Sys.Date(), "%B %d, %Y"),
     cex = 1.2, font = 3, col = "gray40")

# Summary page
plot.new()
text(0.5, 0.9, "Summary of Biomarker Groups", cex = 1.8, font = 2)
y_pos <- 0.8
for (group in names(plots_by_group)) {
  text(0.1, y_pos, paste0("• ", group, ": ", length(plots_by_group[[group]]), " biomarkers"),
       cex = 1.2, adj = 0)
  y_pos <- y_pos - 0.08
}

# Plot pages - 4 plots per page
for (group in names(plots_by_group)) {
  group_plots <- plots_by_group[[group]]
  n_plots <- length(group_plots)
  n_pages <- ceiling(n_plots / 4)

  for (page in 1:n_pages) {
    start_idx <- (page - 1) * 4 + 1
    end_idx <- min(page * 4, n_plots)

    page_plots <- group_plots[start_idx:end_idx]

    # Add empty plots if needed to fill the grid
    while (length(page_plots) < 4) {
      page_plots[[length(page_plots) + 1]] <- plot_spacer()
    }

    # Create page layout
    page_layout <- wrap_plots(page_plots, ncol = 2, nrow = 2) +
      plot_annotation(
        title = group,
        subtitle = paste0("Page ", page, " of ", n_pages),
        theme = theme(
          plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
          plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray40")
        )
      )

    print(page_layout)
  }
}

# Close PDF
dev.off()

# 

# ---- ALSO SHOW PAGES IN R PLOT PANE ----

# If you used patchwork functions above, make sure it's loaded here too
if (!"package:patchwork" %in% search()) library(patchwork)

# Only show groups that actually exist (and keep your preferred order)
present_groups <- group_order[group_order %in% names(plots_by_group)]

for (group in present_groups) {
  group_plots <- plots_by_group[[group]]
  n_plots <- length(group_plots)
  n_pages <- ceiling(n_plots / 4)

  for (page in 1:n_pages) {
    start_idx <- (page - 1) * 4 + 1
    end_idx <- min(page * 4, n_plots)
    page_plots <- group_plots[start_idx:end_idx]

    # fill to 4 plots if needed
    while (length(page_plots) < 4) page_plots[[length(page_plots) + 1]] <- patchwork::plot_spacer()

    # same 2x2 layout + title/subtitle
    page_layout <- patchwork::wrap_plots(page_plots, ncol = 2, nrow = 2) +
      patchwork::plot_annotation(
        title = group,
        subtitle = paste0("Page ", page, " of ", n_pages),
        theme = theme(
          plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
          plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray40")
        )
      )

    # print to the RStudio/graphics device
    print(page_layout)
  }
}

# ---- AFTER dev.off(), preview each plot individually in the Plot pane ----

# (Optional) keep your preferred order but only for groups that exist
present_groups <- group_order[group_order %in% names(plots_by_group)]

for (grp in present_groups) {
  message("Previewing group: ", grp)
  for (p in plots_by_group[[grp]]) {
    print(p)  # <- this is the line you asked about
  }
}

# ---- Overall CSF–Plasma correlation per biomarker pair + one distribution plot ----
if (!requireNamespace("dplyr", quietly = TRUE)) library(dplyr)
if (!requireNamespace("purrr", quietly = TRUE)) library(purrr)
if (!requireNamespace("ggplot2", quietly = TRUE)) library(ggplot2)

safe_cor <- function(x, y) {
  if (is.null(x) || is.null(y)) return(NA_real_)
  suppressWarnings(stats::cor(x, y, use = "complete.obs", method = "pearson"))
}
safe_n <- function(x, y) sum(stats::complete.cases(x, y))

# Compute overall r and n for each CSF–plasma pair
corr_overall <- biomarker_pairs %>%
  mutate(
    r_overall = purrr::map2_dbl(CSF, Plasma, ~ safe_cor(dat_sub[[.x]], dat_sub[[.y]])),
    n_overall = purrr::map2_int(CSF, Plasma, ~ safe_n(dat_sub[[.x]], dat_sub[[.y]]))
  ) %>%
  filter(!is.na(r_overall) & n_overall >= 3)



# One plot: distribution of overall correlations
corr_overall <- corr_overall %>%
  dplyr::mutate(sign_r = ifelse(r_overall >= 0, "Positive", "Negative"))

p_corr_dist <- ggplot(corr_overall, aes(x = r_overall, fill = sign_r)) +
  geom_histogram(
    bins  = 24,
    color = "white",
    alpha = 0.9,
    position = "identity"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  scale_fill_manual(
    values = c("Negative" = "#377EB8", "Positive" = "#E41A1C"),
    name   = "Sign of r"
  ) +
  labs(
    title    = "Distribution of Overall CSF–Plasma Biomarker Correlations",
    subtitle = sprintf("One value per biomarker pair (n pairs = %d)", nrow(corr_overall)),
    x        = "Pearson r (overall)",
    y        = "Count"
  ) +
  theme_minimal(base_size = 9) +
  theme(
    plot.title    = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(color = "gray40", hjust = 0.5),
    axis.title    = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

print(p_corr_dist)




# Save the figure
ggplot2::ggsave(
  filename = "csf_plasma_correlation_distribution.pdf",
  plot = p_corr_dist, width = 7, height = 5, units = "in"
)


if (!requireNamespace("stringr", quietly = TRUE)) {
  install.packages("stringr")
}
library(stringr)

library(dplyr)
library(stringr)
# to extract variables from the dat_var_label_matching
dat_var_imp <-
  dat_var_label_matching %>%
  filter(Group %in% c("CSF", "Plasma")) %>%
  mutate(
    marker_id = case_when(
      # pTau181: CSF + Plasma
      Var %in% c("fl_csf__ALL__ptau_log2",
                 "fl_plasma__ALL__ptau_181_log2") ~ "pTau181",

      # The rest
      str_detect(Var, "mmp_9")   ~ "MMP-9",
      str_detect(Var, "vegf_d")  ~ "VEGF-D",
      str_detect(Var, "il_13")   ~ "IL-13",
      str_detect(Var, "nfl")     ~ "NfL",
      str_detect(Var, "gfap")    ~ "GFAP",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(marker_id))

dat_var_imp


corr_overall_marked <- corr_overall %>%
  # Attach marker info for CSF vars
  dplyr::left_join(
    dat_var_imp %>%
      dplyr::filter(Group == "CSF") %>%
      dplyr::select(Var, marker_id),
    by = c("CSF" = "Var")
  ) %>%
  dplyr::rename(CSF_marker = marker_id) %>%
  # Attach marker info for Plasma vars
  dplyr::left_join(
    dat_var_imp %>%
      dplyr::filter(Group == "Plasma") %>%
      dplyr::select(Var, marker_id),
    by = c("Plasma" = "Var")
  ) %>%
  dplyr::rename(Plasma_marker = marker_id)


corr_key <- corr_overall_marked %>%
  dplyr::filter(
    !is.na(CSF_marker),
    !is.na(Plasma_marker),
    CSF_marker == Plasma_marker
  ) %>%
  dplyr::mutate(
    pretty_label = paste0(CSF_marker, " (CSF vs plasma)")
  )

attr(dat_sub[["fl_cyto_csf_il_13_log2"]],        "label") <- "CSF IL-13"
attr(dat_sub[["fl_cyto_heparin_plasma_il_13_log2"]], "label") <- "Plasma IL-13"
make_pair_plot <- function(csf_var, plasma_var, label, r, n, dat = dat_sub) {
  e_plot_scatterplot(
    dat_plot  = dat,
    var_x     = csf_var,
    var_y     = plasma_var,
    var_color = "dx_current",    # AD / LA / Mixed / ...
    var_facet = NULL,
    text_title = sprintf("%s\nr = %.2f, n = %d", label, r, n),
    sw_print  = FALSE,
    smooth_all = "lm",              # linear regression line over all points
    sw_smooth_all_se = FALSE,       # no SE band
    smooth_by_var_color = "none",   # no separate lines by color
    sw_smooth_by_var_color_se = FALSE,
    sw_corr_in_subtitle = FALSE,    # we already show r in the title
    sw_var_x_name_str_wrap_width = 1e5,
    sw_var_y_name_str_wrap_width = 1e5
  )

}




highlight_plots <- purrr::pmap(
  list(
    csf_var = corr_key$CSF,
    plasma_var = corr_key$Plasma,
    label   = corr_key$pretty_label,
    r       = corr_key$r_overall,
    n       = corr_key$n_overall
  ),
  make_pair_plot
)

# After creating highlight_plots, remove legends from each plot
highlight_plots_no_legend <- lapply(highlight_plots, function(p) {
  p + theme(legend.position = "none")
})


# Extract legend 
shared_legend <- cowplot::get_legend(
  highlight_plots[[1]] + 
    guides(color = guide_legend(nrow = 1)) +  # horizontal legend
    theme(
      legend.position = "right",
      legend.title = element_text(face = "bold"),
      legend.text = element_text(size = 10)
    )
)

p_pairs_grid <- cowplot::plot_grid(
  plotlist = highlight_plots_no_legend,
  ncol     = 2,    # adjust rows/cols as you like
  labels   = NULL  # or "AUTO"
)

# Combine histogram and pairs grid (both without legends)
p_main <- cowplot::plot_grid(
  p_corr_dist,  # keep histogram legend if you want it, or use p_corr_dist_no_legend
  p_pairs_grid,
  ncol = 2,
  rel_widths = c(1, 2)
)

# Add the shared legend at the bottom
p_corr_with_examples <- cowplot::plot_grid(
  p_main,
  shared_legend,
  ncol = 1,
  rel_heights = c(1, 0.1)  # adjust ratio as needed
)

print(p_corr_with_examples)

ggsave(
  filename = "All_figures_for_paper/fig_csf_plasma_corr_and_key_pairs.png",
  plot     = p_corr_with_examples,
  width    = 14,
  height   = 6,
  dpi      = 300,
  bg       = "white"
)



# ---- (Optional) If you want the distribution by biomarker group (comment out if not needed) ----
# p_corr_by_group <- ggplot(corr_overall, aes(x = r_overall)) +
#   geom_histogram(bins = 18, alpha = 0.9) +
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   facet_wrap(~ Group, scales = "free_y") +
#   labs(
#     title = "CSF–Plasma Correlations by Biomarker Group",
#     x = "Pearson r (overall)", y = "Count"
#   ) +
#   theme_minimal(base_size = 11)
# print(p_corr_by_group)
# ggsave("csf_plasma_correlation_distribution_by_group.pdf", p_corr_by_group, width = 10, height = 7, units = "in")


# 
# library(tidyverse)
# library(ggplot2)
# library(patchwork)
# library(scales)
# library(cowplot)
# 
# # Extract all CSF and Plasma biomarkers from the data
# get_all_biomarkers <- function(data) {
#   # data = dat_sub
#   all_vars <- names(data)
#   
#   # Get CSF biomarkers (containing "csf" in the name)
#   csf_vars <- all_vars[str_detect(all_vars, "csf") & str_detect(all_vars, "fl_")]
#   
#   # Get Plasma biomarkers (containing "plasma" or "serum" in the name)
#   plasma_vars <- all_vars[(str_detect(all_vars, "plasma") | str_detect(all_vars, "serum")) & 
#                           str_detect(all_vars, "fl_")]
#   
#   # Add IgG variables that might be missed
#   if ("fl_serum_igg_log2" %in% all_vars) {
#     plasma_vars <- unique(c(plasma_vars, "fl_serum_igg_log2"))
# 
#   }
#   
#   list(csf = csf_vars, plasma = plasma_vars)
# }
# 
# # Define color palette for diagnosis groups
# dx_colors <- c(
#   "AD" = "#E41A1C",        # Red
#   "SIVD" = "#377EB8",      # Blue
#   "Mixed" = "#4DAF4A",     # Green
#   "LA" = "#984EA3",        # Purple
#   "MI" = "#FF7F00",        # Orange
#   "Unknown" = "#A6A6A6",   # Gray
#   "Control" = "#F781BF",   # Pink
#   "Repeat pt" = "#999999"  # Dark gray
# )
# 
# # Function to extract biomarker name from variable
# extract_biomarker_name <- function(var_name) {
#   # Remove prefixes and suffixes
#   name <- var_name %>%
#     str_remove("^fl_") %>%
#     str_remove("_log2$") %>%
#     str_remove("csf_") %>%
#     str_remove("plasma_") %>%
#     str_remove("serum_") %>%
#     str_remove("__ALL__") %>%
#     str_remove("msd_") %>%
#     str_remove("angio_") %>%
#     str_remove("cyto_") %>%
#     str_remove("heparin_") %>%
#     str_replace_all("_", "-") %>%
#     toupper()
#   
#   return(name)
# }
# 
# # Function to create a single scatterplot
# create_biomarker_plot <- function(data, csf_var, plasma_var, simplified = TRUE) {
#   # Check if both variables exist in the data
#   if (!all(c(csf_var, plasma_var) %in% names(data))) {
#     return(NULL)
#   }
#   
#   # Prepare plot data
#   plot_data <- data %>%
#     select(all_of(c(csf_var, plasma_var, "dx_current"))) %>%
#     filter(!is.na(.data[[csf_var]]) & !is.na(.data[[plasma_var]]))
#   
#   # Skip if too few observations
#   if (nrow(plot_data) < 10) {
#     return(NULL)
#   }
#   
#   # Calculate overall correlation
#   overall_cor <- cor(plot_data[[csf_var]], plot_data[[plasma_var]], 
#                      use = "complete.obs", method = "pearson")
#   
#   # Extract biomarker names
#   csf_name <- extract_biomarker_name(csf_var)
#   plasma_name <- extract_biomarker_name(plasma_var)
#   
#   # Create plot
#   if (simplified) {
#     # Simplified version for many plots
#     p <- ggplot(plot_data, aes(x = .data[[plasma_var]], y = .data[[csf_var]])) +
#       geom_point(aes(color = dx_current), alpha = 0.5, size = 1) +
#       geom_smooth(method = "lm", se = FALSE, color = "black", 
#                   linetype = "solid", linewidth = 0.5) +
#       scale_color_manual(values = dx_colors, name = "Diagnosis") +
#       labs(
#         title = paste0("CSF: ", csf_name, " vs Plasma: ", plasma_name),
#         subtitle = sprintf("r = %.3f, n = %d", overall_cor, nrow(plot_data)),
#         x = "Plasma (log2)",
#         y = "CSF (log2)"
#       ) +
#       theme_minimal(base_size = 7) +
#       theme(
#         plot.title = element_text(size = 7, face = "bold", hjust = 0.5),
#         plot.subtitle = element_text(size = 6, hjust = 0.5, color = "gray40"),
#         panel.grid.minor = element_blank(),
#         panel.border = element_rect(fill = NA, color = "gray80", linewidth = 0.5),
#         legend.position = "right",
#         axis.text = element_text(size = 5),
#         axis.title = element_text(size = 6),
#                  plot.margin = ggplot2::margin(t = 2, r = 2, b = 2, l = 2, unit = "pt")
#       )
#   } else {
#     # Detailed version
#     p <- ggplot(plot_data, aes(x = .data[[plasma_var]], y = .data[[csf_var]])) +
#       geom_point(aes(color = dx_current), alpha = 0.6, size = 2) +
#       geom_smooth(method = "lm", se = TRUE, color = "black", 
#                   linetype = "solid", linewidth = 0.8, alpha = 0.3) +
#       geom_smooth(aes(color = dx_current), method = "lm", se = FALSE, 
#                   linewidth = 0.6, alpha = 0.8) +
#       scale_color_manual(values = dx_colors, name = "Diagnosis") +
#       labs(
#         title = paste0("CSF: ", csf_name, " vs Plasma: ", plasma_name),
#         subtitle = sprintf("Overall r = %.3f, n = %d", overall_cor, nrow(plot_data)),
#         x = paste0("Plasma: ", plasma_name, " (log2)"),
#         y = paste0("CSF: ", csf_name, " (log2)")
#       ) +
#       theme_minimal(base_size = 9) +
#       theme(
#         plot.title = element_text(size = 10, face = "bold"),
#         plot.subtitle = element_text(size = 8, color = "gray40"),
#         panel.grid.minor = element_blank(),
#         panel.border = element_rect(fill = NA, color = "gray70"),
#         legend.position = "right",
#         legend.title = element_text(size = 8),
#         legend.text = element_text(size = 7),
#         legend.key.size = unit(0.8, "lines")
#       )
#   }
#   
#   return(p)
# }
# 
# # Get all biomarkers
# biomarkers <- get_all_biomarkers(dat_sub)
# n_csf <- length(biomarkers$csf)
# n_plasma <- length(biomarkers$plasma)
# total_combinations <- n_csf * n_plasma
# 
# cat(sprintf("Found %d CSF biomarkers and %d plasma biomarkers\n", n_csf, n_plasma))
# cat(sprintf("This will create %d total combinations\n", total_combinations))
# 
# # Create directory for individual plots
# dir.create("csf_plasma_individual_plots", showWarnings = FALSE)
# 
# # Group CSF biomarkers by type using dat_var_label_matching
# csf_grouped <- dat_var_label_matching %>%
#   filter(Var %in% biomarkers$csf) %>%
#   mutate(
#     CSF_Group = case_when(
#       str_detect(Type, "MMP") ~ "CSF MMP",
#       str_detect(Type, "Angiogenic") ~ "CSF Angiogenic Factors",
#       str_detect(Type, "Cytokines") ~ "CSF Cytokines",
#       str_detect(Type, "AD") ~ "CSF AD Markers",
#       str_detect(Type, "CSF Permeability") ~ "CSF Permeability",
#       TRUE ~ "CSF Other"
#     )
#   ) %>%
#   group_by(CSF_Group) %>%
#   summarize(vars = list(Var), .groups = 'drop')
# 
# # Create PDF with all combinations
# pdf("csf_plasma_all_combinations.pdf", width = 14, height = 10)
# 
# # Title page
# plot.new()
# text(0.5, 0.6, "CSF vs Plasma Biomarker Analysis", 
#      cex = 2.5, font = 2, col = "darkblue")
# text(0.5, 0.5, "All Possible Combinations", 
#      cex = 1.8, font = 1, col = "darkblue")
# text(0.5, 0.4, sprintf("%d CSF × %d Plasma = %d Total Plots", 
#                        n_csf, n_plasma, total_combinations), 
#      cex = 1.2, font = 1, col = "gray40")
# text(0.5, 0.3, format(Sys.Date(), "%B %d, %Y"), 
#      cex = 1.2, font = 3, col = "gray40")
# 
# # Summary page
# plot.new()
# text(0.5, 0.9, "Summary of Biomarker Groups", cex = 1.8, font = 2)
# y_pos <- 0.8
# for (i in seq_len(nrow(csf_grouped))) {
#   group_name <- csf_grouped$CSF_Group[i]
#   n_vars <- length(csf_grouped$vars[[i]])
#   text(0.1, y_pos, paste0("• ", group_name, ": ", n_vars, " biomarkers × ", 
#                           n_plasma, " plasma = ", n_vars * n_plasma, " plots"), 
#        cex = 1, adj = 0)
#   y_pos <- y_pos - 0.06
# }
# 
# # Legend page
# plot.new()
# text(0.5, 0.9, "Diagnosis Group Legend", cex = 1.8, font = 2)
# legend("center", 
#        legend = names(dx_colors), 
#        col = dx_colors, 
#        pch = 19, 
#        cex = 1.5, 
#        title = "Diagnosis Groups",
#        bty = "n")
# 
# # Create plots for each CSF group
# for (i in seq_len(nrow(csf_grouped))) {
#   group_name <- csf_grouped$CSF_Group[i]
#   csf_vars <- csf_grouped$vars[[i]]
# 
#   cat(sprintf("\nProcessing %s (%d CSF biomarkers)\n", group_name, length(csf_vars)))
# 
#   # For each CSF biomarker in this group
#   for (csf_var in csf_vars) {
#     csf_name <- extract_biomarker_name(csf_var)
# 
#          # Create plots for all plasma biomarkers
#     plots <- list()
#     plot_count <- 0
# 
#     for (plasma_var in biomarkers$plasma) {
#       p <- create_biomarker_plot(dat_sub, csf_var, plasma_var, simplified = TRUE)
#       if (!is.null(p)) {
#         plots[[length(plots) + 1]] <- p
#         plot_count <- plot_count + 1
# 
#         # Save individual plot as PNG
#         plasma_name <- extract_biomarker_name(plasma_var)
#         filename <- sprintf("csf_plasma_individual_plots/CSF_%s_vs_Plasma_%s.png",
#                           csf_name, plasma_name)
# 
#         # Create a larger version for individual saving
#         p_large <- create_biomarker_plot(dat_sub, csf_var, plasma_var, simplified = FALSE)
#         ggsave(filename, p_large, width = 6, height = 5, dpi = 300, bg = "white")
# 
#         if (plot_count %% 10 == 0) {
#           cat(sprintf("  Saved %d plots for CSF %s...\n", plot_count, csf_name))
#         }
#       }
#     }
# 
#     if (length(plots) > 0) {
#              # Create pages with 9 plots each (3x3 grid)
#       n_plots <- length(plots)
#       plots_per_page <- 9
#       n_pages <- ceiling(n_plots / plots_per_page)
# 
#       for (page in 1:n_pages) {
#         start_idx <- (page - 1) * plots_per_page + 1
#         end_idx <- min(page * plots_per_page, n_plots)
# 
#         page_plots <- plots[start_idx:end_idx]
# 
#         # Add empty plots if needed
#         while (length(page_plots) < plots_per_page) {
#           page_plots[[length(page_plots) + 1]] <- plot_spacer()
#         }
# 
#                  # Create page layout
#         page_layout <- wrap_plots(page_plots, ncol = 3, nrow = 3) +
#           plot_annotation(
#             title = sprintf("%s - CSF: %s", group_name, csf_name),
#             subtitle = sprintf("Page %d of %d (Plots %d-%d of %d)",
#                              page, n_pages, start_idx, end_idx, n_plots),
#             theme = theme(
#               plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
#               plot.subtitle = element_text(size = 10, hjust = 0.5, color = "gray40")
#             )
#           )
# 
#         print(page_layout)
#       }
#     }
#   }
# }
# 
# # Close PDF
# dev.off()
# 
# # Create a correlation matrix for all combinations
# cat("\nCalculating correlation matrix for all combinations...\n")
# 
# cor_matrix <- matrix(NA, nrow = n_csf, ncol = n_plasma,
#                      dimnames = list(biomarkers$csf, biomarkers$plasma))
# 
# for (i in seq_along(biomarkers$csf)) {
#   for (j in seq_along(biomarkers$plasma)) {
#     csf_var <- biomarkers$csf[i]
#     plasma_var <- biomarkers$plasma[j]
# 
#     # Calculate correlation if both variables have sufficient data
#     data_subset <- dat_sub %>%
#       select(all_of(c(csf_var, plasma_var))) %>%
#       na.omit()
# 
#     if (nrow(data_subset) >= 10) {
#       cor_matrix[i, j] <- cor(data_subset[[csf_var]], data_subset[[plasma_var]])
#     }
#   }
# }
# 
# # # Save correlation matrix as CSV
# cor_df <- as.data.frame(cor_matrix)
# cor_df$CSF_Biomarker <- rownames(cor_df)
# cor_df <- cor_df[, c("CSF_Biomarker", names(cor_df)[1:(ncol(cor_df)-1)])]
# 
# write.csv(cor_df, "csf_plasma_correlation_matrix.csv", row.names = FALSE)
# 
# # Create a heatmap of correlations
# library(pheatmap)
# library(RColorBrewer)
# 
# # # Clean names for plotting
# rownames(cor_matrix) <- sapply(rownames(cor_matrix), extract_biomarker_name)
# colnames(cor_matrix) <- sapply(colnames(cor_matrix), extract_biomarker_name)
# 
# # # Create heatmap
# png("csf_plasma_correlation_heatmap.png", width = 16, height = 12, units = "in", res = 300)
# pheatmap(cor_matrix,
#          color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
#          breaks = seq(-1, 1, length.out = 101),
#          cluster_rows = TRUE,
#          cluster_cols = TRUE,
#          show_rownames = TRUE,
#          show_colnames = TRUE,
#          fontsize_row = 8,
#          fontsize_col = 8,
#          main = "CSF vs Plasma Biomarker Correlations (All Combinations)",
#          na_col = "gray90",
#          angle_col = 45)
# dev.off()
# 
# # Find strongest correlations
# cat("\nTop 20 strongest positive correlations:\n")
# cor_long <- cor_matrix %>%
#   as.data.frame() %>%
#   rownames_to_column("CSF") %>%
#   pivot_longer(-CSF, names_to = "Plasma", values_to = "correlation") %>%
#   filter(!is.na(correlation)) %>%
#   arrange(desc(correlation))
# 
# print(head(cor_long, 20))
# 
# cat("\n\nTop 20 strongest negative correlations:\n")
# print(tail(cor_long, 20))
# 
# # Summary statistics
# cat("\n\nSummary of all correlations:\n")
# all_cors <- as.vector(cor_matrix)
# all_cors <- all_cors[!is.na(all_cors)]
# summary(all_cors)
# 
# cat(sprintf("\nPDF created: csf_plasma_all_combinations.pdf\n"))
# cat(sprintf("Individual PNG plots saved in: csf_plasma_individual_plots/\n"))
# cat(sprintf("Correlation matrix saved: csf_plasma_correlation_matrix.csv\n"))
# cat(sprintf("Heatmap created: csf_plasma_correlation_heatmap.png\n"))
# 
# # # Count PNG files created
# png_files <- list.files("csf_plasma_individual_plots", pattern = "\\.png$", full.names = FALSE)
# cat(sprintf("\nTotal individual PNG files created: %d\n", length(png_files)))


```

```{=html}
<!---
# Erik's compile commands in R:
  fn_qmd <- "Erhardt_Rosenberg_ADRC_Fluid_MRI_NPsy_Analysis_20231030.qmd"
  setwd("D:/Dropbox/StatAcumen/consult/Authorship/2014_GaryRosenberg_UNM_NIH_VascularDementia_MRI/2024_ADRC/Analyses/202311_ADRC_dx_Fluid_MRI_NPsy")
  quarto::quarto_render(input = fn_qmd)
-->
```

```{r chunk-01, echo=FALSE}
options(width = 80)
#options(warn  = -1)
options(str   = strOptions(list.len = 1e3))
options(knitr.kable.NA = '') # Display NAs as blanks
my_seed <- 34567
set.seed(my_seed)

sw_indy_figures = FALSE
size_plot_units        = "mm"
size_plot_1col_width   = 85     # * (1 / 25.4) # inches  Full page (2 columns)
size_plot_3_4col_width = NA     # # inches  3/4 page
size_plot_2col_width   = 180    # * (1 / 25.4) # inches  1 column (1/2 page)
size_plot_height       = 230    # * (1 / 25.4)  # * 607.582 / 486.612 # maximum height for space with caption
```

------------------------------------------------------------------------

# Next actions

------------------------------------------------------------------------

# Data preliminaries

```{r chunk-003}
library(erikmisc)
library(tidyverse)
library(labelled)

library(randomForestSRC)

# Parallel processing
library(parallel)
options(rf.cores = parallel::detectCores() - 2) # OpenMP Parallel Processing
options(mc.cores = parallel::detectCores() - 2) # R-side Parallel Processing
```

## Subset data

```{r chunk-004}
# load("../../Data/Sweden.RData")
# dat_sweden <-
#   dat_all |>
#   dplyr::select(
#     demo_study_id
#   , tidyselect::starts_with("fl_")
#   , -tidyselect::ends_with("_log2")
#   ) |>
#   dplyr::mutate(
#     I_Sweden    = TRUE
#   , ID          = 20000 + 1:dplyr::n()
#   , dx_current  = "Sweden"
#   ) |>
#   dplyr::rowwise() |>
#   dplyr::mutate(
#     prop_complete = mean(!is.na(dplyr::across(tidyselect::everything())))
#   ) |>
#   dplyr::ungroup()
# rm(dat_all)
# # e_plot_missing(dat_sweden)

load("../../Data/VCI.RData")
dat_vci <-
  dat_vci |>
  dplyr::select(
    demo_study_id
  , dx_biological
  #, tidyselect::any_of(names(dat_sweden))
  , tidyselect::starts_with("fl_")
  , tidyselect::starts_with("mri_")
  , tidyselect::starts_with("npsy_")
  , -dx_current
  ) |>
  dplyr::rename(
    dx_current = dx_biological
  ) |>
  dplyr::mutate(
    I_VCI       = TRUE
  , ID          = 10000 + 1:dplyr::n()
  , dx_current  = dx_current |> fct_recode(SIVD = "SVD")
  , dx_current  = dx_current |> as.character()
  ) |>
  dplyr::filter(
    dx_current %in% c("Control", "LA", "AD", "SIVD", "Mixed")
  ) |>
  dplyr::rowwise() |>
  dplyr::mutate(
    prop_complete = mean(!is.na(dplyr::across(tidyselect::everything())))
  ) |>
  dplyr::ungroup()

#hist(dat_vci$prop_complete, 100)

# keeping records with at least a proportion not NA, will impute NAs
# dat_vci <-
#   dat_vci |>
#   dplyr::filter(
#     VCI_prop_complete >= 0.6
#   )

# CSF labels
#labelled::var_label(dat_vci$fl_abeta_4240_ratio) <-
#  paste0(
#    "CSF "
#  , labelled::var_label(dat_vci$fl_abeta_4240_ratio)
#  )
#labelled::var_label(dat_vci$fl_ptau) <-
#  paste0(
#    "CSF "
#  , labelled::var_label(dat_vci$fl_ptau)
#  )

# e_plot_missing(
#     dat_plot       = dat_vci
#   , var_group      = "dx_current"
#   , sw_group_sort  = TRUE
#   , var2_sort      = "prop_complete"
#   )

# # log2 versions
# for (n_var in str_subset(names(dat_vci), pattern = "^fl_")) {
#   dat_vci[, paste0(n_var, "_log2") ] <-
#     e_log_shift(dat_vci[, n_var])
#
#   labelled::var_label(dat_vci[, paste0(n_var, "_log2") ]) <-
#     paste("log2", labelled::var_label( dat_vci[, n_var] ))
# }



# # bind data together
# dat_all <-
#   dat_vci |>
#   dplyr::bind_rows(
#     dat_sweden
#   ) |>
#   dplyr::mutate(
#     dx_current = dx_current |> factor(levels = c("Control", "LA", "AD", "SIVD", "Mixed", "Sweden"))
#   )

dat_all <-
  dat_vci

dat_all <-
  dat_all |>
  dplyr::select(
    ID
  , demo_study_id
  , dx_current
  , prop_complete
  , tidyselect::starts_with("fl_")
  , tidyselect::starts_with("mri_")
  , tidyselect::starts_with("npsy_")
  #, tidyselect::starts_with("I_")
  #, tidyselect::all_of(dat_all |> names() |> stringr::str_subset(pattern = stringr::fixed("_csf_"   )) |> sort())
  #, tidyselect::all_of(dat_all |> names() |> stringr::str_subset(pattern = stringr::fixed("_plasma_")) |> sort())
  , tidyselect::everything()
  )


dat_sub <-
  dat_all

p <-
  e_plot_missing(
    dat_plot       = dat_sub
  , var_group      = "dx_current"
  , sw_group_sort  = TRUE
  , var2_sort      = "prop_complete"
  )
p |> print()
ggsave(
      paste0("out/plot_missing", ".png")
    , plot   = p
    , width  = 12
    , height = 10
    ## png, jpeg
    , dpi    = 300
    , bg     = "white"
    ## pdf
    #, units  = "in"
    #, useDingbats = FALSE
    )
```

Variable lists and subset.

```{r chunk-009}
## Variable lists
Var <- list()

Var[["list"]][["var_fl_all"          ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("fl_"           )) |> sort()
Var[["list"]][["var_fl_csf"          ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("_csf_"         )) |> sort()
Var[["list"]][["var_fl_plasma"       ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("_plasma_"      )) |> sort()
Var[["list"]][["var_fl_angio_csf"    ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("angio_csf_"    )) |> sort()
Var[["list"]][["var_fl_angio_plasma" ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("angio_plasma_" )) |> sort()
Var[["list"]][["var_fl_cyto_csf"     ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("cyto_csf_"     )) |> sort()
Var[["list"]][["var_fl_cyto_plasma"  ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("cyto_plasma_"  )) |> sort()
Var[["list"]][["var_fl_msd_csf"      ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("msd_csf_"      )) |> sort()
Var[["list"]][["var_fl_msd_plasma"   ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("msd_plasma_"   )) |> sort()
Var[["list"]][["var_fl_simoa_csf"    ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("simoa_csf_"    )) |> sort()
Var[["list"]][["var_fl_simoa_plasma" ]] <- dat_sub |> names() |> stringr::str_subset(pattern = stringr::fixed("simoa_plasma_" )) |> sort()

### Exclude:
# fl_blood_date
# fl_lp_date

#Var[["list"]][["var_mri_all"         ]] <- dat_sub |> names() |> stringr::str_subset(pattern = "^mri_"                         ) |> sort()

Var[["list"]][["var_mri_dti"         ]] <-
  c(
    "mri_dti_fabnormal_z"
  , "mri_dti_freewater_mean"
  , "mri_dti_freewater_z"
  , "mri_dti_histpk_z"
  , "mri_dti_psmd"
  , "mri_dti_psmd_z"
  )
Var[["list"]][["var_mri_perm"         ]] <-
  c(
    "mri_perm_15t_mean_brain_ki"
  , "mri_perm_15t_tot_ki"
  , "mri_perm_3t_tot"
  #, "mri_permeability"
  , "mri_microbleeds"
  )
Var[["list"]][["var_mri_spect"         ]] <-
  c(
  #  "mri_spectroscopy"
    "mri_spectroscopy_cho_gpc_pch"
  , "mri_spectroscopy_cr_mrs"
  , "mri_spectroscopy_cr_pcr"
  , "mri_spectroscopy_gln"
  , "mri_spectroscopy_glu"
  , "mri_spectroscopy_glu_gln"
  , "mri_spectroscopy_ins"
  , "mri_spectroscopy_naa"
  , "mri_spectroscopy_naa_naag"
  )
Var[["list"]][["var_mri_vol"         ]] <-
  c(
    "mri_vol_atrophy"
  , "mri_vol_gm"
  , "mri_vol_hippocampal"
  , "mri_vol_intra_cranial_mm3"
  , "mri_vol_lesion"
  , "mri_vol_total_brain"
  , "mri_vol_wm"
  , "mri_vol_wm_jim"
  , "mri_vol_wm_ucd"
  , "mri_percent_hippo_vol"
  , "mri_percent_wmh_vol"
  , "mri_mean_cort_thick_mm"
  )

Var[["list"]][["var_mri_all"         ]] <-
  c(
    Var[["list"]][["var_mri_dti"         ]]
  , Var[["list"]][["var_mri_perm"        ]]
  , Var[["list"]][["var_mri_spect"       ]]
  , Var[["list"]][["var_mri_vol"         ]]
  )



#Var[["list"]][["var_npsy_all"        ]] <- dat_sub |> names() |> stringr::str_subset(pattern = "^npsy_"                        ) |> stringr::str_subset(pattern = "_t$" ) |> sort()
Var[["list"]][["var_npsy_sum_t"      ]] <-
  c(
    "npsy_overall_t"
  , "npsy_attention_t"
  , "npsy_executive_t"
  , "npsy_language_t"
  , "npsy_memory_t"
  , "npsy_processing_t"
  )
Var[["list"]][["var_npsy_ft_t"     ]] <-
  c(
    "npsy_ft_dom_ave"
  , "npsy_ft_dom_ave_t"
  , "npsy_ft_nondom_ave"
  , "npsy_ft_nondom_ave_t"
  )
Var[["list"]][["var_npsy_hvlt_t"     ]] <-
  c(
    "npsy_hvlt_delay_raw_tot"
  , "npsy_hvlt_delay_t"
  , "npsy_hvlt_immed_raw_tot"
  , "npsy_hvlt_immed_t"
  , "npsy_hvlt_rec_disc_index"
  , "npsy_hvlt_rec_disc_t"
  )
Var[["list"]][["var_npsy_peg_t"     ]] <-
  c(
    "npsy_peg_dom_t"
  , "npsy_peg_dom_time"
  , "npsy_peg_nondom_t"
  , "npsy_peg_nondom_time"
  )
Var[["list"]][["var_npsy_rcft_t"     ]] <-
  c(
    "npsy_rcft_copy"
  , "npsy_rcft_delay_recall"
  , "npsy_rcft_delay_recall_t"
  , "npsy_rcft_imm_recall"
  , "npsy_rcft_imm_recall_t"
  , "npsy_rcft_recog"
  , "npsy_rcft_recog_t"
  )
Var[["list"]][["var_npsy_strp_t"     ]] <-
  c(
    "npsy_strp_color_score"
  , "npsy_strp_color_score_t"
  , "npsy_strp_colorword_score"
  , "npsy_strp_colorword_score_t"
  , "npsy_strp_cw_interference"
  , "npsy_strp_cw_interference_t"
  , "npsy_strp_word_score"
  , "npsy_strp_word_score_t"
  )
Var[["list"]][["var_npsy_wcst_t"     ]] <-
  c(
    "npsy_wcst_cat"
  #, "npsy_wcst_cat_percent"           # ranges, need to be converted to numbers
  , "npsy_wcst_con_resp"
  , "npsy_wcst_con_resp_t"
  , "npsy_wcst_pers_err"
  , "npsy_wcst_pers_err_t"
  , "npsy_wcst_pers_resp"
  , "npsy_wcst_pers_resp_t"
  , "npsy_wcst_tot_err"
  , "npsy_wcst_tot_err_t"
  , "npsy_wrat_t"
  , "npsy_wrat_tot"
  )
Var[["list"]][["var_npsy_others1_t"     ]] <-
  c(
    "npsy_boston_t"
  , "npsy_boston_tot_correct"
  #, "npsy_clock_cutoff"
  , "npsy_clock_tot"
  , "npsy_dig_sym_t"
  , "npsy_dig_sym_tot"
  , "npsy_ds_tot_for_back"
  , "npsy_ds_tot_t"
  , "npsy_education"
  )
Var[["list"]][["var_npsy_others2_t"     ]] <-
  c(
    "npsy_faq_score"
  , "npsy_fas_t"
  , "npsy_fas_tot"
  , "npsy_gdi"
  #, "npsy_gdi_cutoff"
  , "npsy_mmse"
  , "npsy_mmse_t"
  , "npsy_sym_srch_t"
  , "npsy_sym_srch_tot"
  )

Var[["list"]][["var_npsy_all"        ]] <-
  c(
    Var[["list"]][["var_npsy_sum_t"      ]]
  , Var[["list"]][["var_npsy_ft_t"       ]]
  , Var[["list"]][["var_npsy_hvlt_t"     ]]
  , Var[["list"]][["var_npsy_peg_t"      ]]
  , Var[["list"]][["var_npsy_rcft_t"     ]]
  , Var[["list"]][["var_npsy_strp_t"     ]]
  , Var[["list"]][["var_npsy_wcst_t"     ]]
  , Var[["list"]][["var_npsy_others1_t"   ]]
  , Var[["list"]][["var_npsy_others2_t"   ]]
  )


Var[["list"]][["var_all_all"         ]] <-
  c(
    Var[["list"]][["var_fl_all"          ]]
  , Var[["list"]][["var_mri_all"         ]]
  , Var[["list"]][["var_npsy_all"        ]]
  )

var_groups <-
  c(
    "var_fl_angio_csf"
  , "var_fl_angio_plasma"
  , "var_fl_cyto_csf"
  , "var_fl_cyto_plasma"
  , "var_fl_msd_csf"
  , "var_fl_msd_plasma"
  , "var_fl_simoa_csf"
  , "var_fl_simoa_plasma"
  , "var_mri_dti"
  , "var_mri_perm"
  , "var_mri_spect"
  , "var_mri_vol"
  , "var_npsy_sum_t"
  , "var_npsy_ft_t"
  , "var_npsy_hvlt_t"
  , "var_npsy_peg_t"
  , "var_npsy_rcft_t"
  , "var_npsy_strp_t"
  , "var_npsy_wcst_t"
  , "var_npsy_others1_t"
  , "var_npsy_others2_t"
  )



var_groups <-
  c(
    "var_all_all"
  , "var_fl_all"
  , "var_fl_csf"
  , "var_fl_plasma"
  , var_groups
  ) |>
  unique()


```

Reorder dx categories, normalize volume variables to total brain volume.

```{r chunk-011}
# reorder dx categories
dat_sub <-
  dat_sub |>
  mutate(
    dx_current =
      dx_current |>
      forcats::fct_relevel(c("SIVD", "Mixed", "AD", "LA", "Control")) |>
      forcats::fct_drop()
  , dx_current_ADSIVDMX_Control =
      dx_current |>
      forcats::fct_collapse(
        AD_SIVD_MX = c("SIVD", "Mixed", "AD") #, "LA")
      , Control   = c("Control")
      ) |>
      factor(levels = c("AD_SIVD_MX", "Control")) |>
      forcats::fct_drop()
  , dx_current_ADMX_SIVD =
      dx_current |>
      forcats::fct_collapse(
        AD_MX   = c("AD", "Mixed")
      , SIVD = c("SIVD") #"LA",
      ) |>
      factor(levels = c("AD_MX", "SIVD")) |>
      forcats::fct_drop()
  , dx_current_AD_SIVDMX =
      dx_current |>
      forcats::fct_collapse(
        AD         = c("AD")
      , SIVD_MX = c("SIVD", "Mixed") # "LA",
      ) |>
      factor(levels = c("AD", "SIVD_MX")) |>
      forcats::fct_drop()
  , dx_current_AD_SIVD_MX =
      dx_current |>
      factor(levels = c("AD", "SIVD", "Mixed")) |>
      forcats::fct_drop()
  )

dx_groups <-
  c(
    "dx_current"
  , "dx_current_ADSIVDMX_Control"
  , "dx_current_ADMX_SIVD"
  , "dx_current_AD_SIVDMX"
  , "dx_current_AD_SIVD_MX"
  )

```

## Drop extreme outliers

```{r chunk-012, fig.height = 4, fig.width = 10}
#var_list_plot <-
#  c(Var$list$var_fl_csf, Var$list$var_fl_plasma)
var_list_plot <-
  Var |> unlist() |> unique() |> sort()

par(mfrow=c(1,3))
for (i_plot in 1:length(var_list_plot)) {
  ## i_plot = 1
  #if (i_plot > 3) {next}

  if( all(is.na(dat_sub[[ var_list_plot[i_plot] ]])) ) {
    next
  }

  plot(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , type  = "p"
  , main  = paste0("Before: ", var_list_plot[i_plot])
  , xlab  = ""
  , ylab  = ""
  , col   = dat_sub$dx_current
  , pch   = 20  #3
  , xlim  = c(1, nrow(dat_sub) * 1.2)
  )
  text(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , labels = 1:nrow(dat_sub)
  , pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
  )
}
par(mfrow=c(1,1))
```

Drop extreme outlying values.

```{r chunk-013}
# 10/30/2023 Erik
# Drop extreme outlying values
# Indicate range to remove (replace with NAs)

#dat_sub$fl_abeta_38                            [dat_sub$fl_abeta_38                                ] <- NA
#dat_sub$fl_abeta_40                            [dat_sub$fl_abeta_40                                ] <- NA
#dat_sub$fl_abeta_42                            [dat_sub$fl_abeta_42                                ] <- NA
#dat_sub$fl_abeta_4240_ratio                    [dat_sub$fl_abeta_4240_ratio                        ] <- NA
#dat_sub$fl_albumin_index                       [dat_sub$fl_albumin_index                           ] <- NA
dat_sub$fl_angio_csf_bfgf                      [dat_sub$fl_angio_csf_bfgf                       > 6   ] <- NA
dat_sub$fl_angio_csf_flt_1                     [dat_sub$fl_angio_csf_flt_1                      < 10   ] <- NA
#dat_sub$fl_angio_csf_plgf                      [dat_sub$fl_angio_csf_plgf                          ] <- NA
#dat_sub$fl_angio_csf_tie2                      [dat_sub$fl_angio_csf_tie2                          ] <- NA
dat_sub$fl_angio_csf_vegf_a                    [dat_sub$fl_angio_csf_vegf_a                     > 19   ] <- NA
#dat_sub$fl_angio_csf_vegf_c                    [dat_sub$fl_angio_csf_vegf_c                        ] <- NA
#dat_sub$fl_angio_csf_vegf_d                    [dat_sub$fl_angio_csf_vegf_d                        ] <- NA
dat_sub$fl_angio_plasma_bfgf                   [dat_sub$fl_angio_plasma_bfgf                    > 300   ] <- NA
dat_sub$fl_angio_plasma_flt_1                  [dat_sub$fl_angio_plasma_flt_1                   > 250   ] <- NA
#dat_sub$fl_angio_plasma_plgf                   [dat_sub$fl_angio_plasma_plgf                       ] <- NA
#dat_sub$fl_angio_plasma_tie2                   [dat_sub$fl_angio_plasma_tie2                       ] <- NA
#dat_sub$fl_angio_plasma_vegf_a                 [dat_sub$fl_angio_plasma_vegf_a                     ] <- NA
dat_sub$fl_angio_plasma_vegf_c                 [dat_sub$fl_angio_plasma_vegf_c                  > 1000   ] <- NA
#dat_sub$fl_angio_plasma_vegf_d                 [dat_sub$fl_angio_plasma_vegf_d                     ] <- NA
#dat_sub$fl_blood_date                          [dat_sub$fl_blood_date                              ] <- NA
#dat_sub$fl_csf_albumin                         [dat_sub$fl_csf_albumin                             ] <- NA
dat_sub$fl_csf_glucose                         [dat_sub$fl_csf_glucose                          < 20   ] <- NA
#dat_sub$fl_csf_igg                             [dat_sub$fl_csf_igg                                 ] <- NA
dat_sub$fl_csf_igg_synthesis                   [dat_sub$fl_csf_igg_synthesis                    > 0.5   ] <- NA
dat_sub$fl_csf_rbc                             [dat_sub$fl_csf_rbc                              > 500   ] <- NA
dat_sub$fl_csf_tnc                             [dat_sub$fl_csf_tnc                              > 15   ] <- NA
dat_sub$fl_csf_total_protein                   [dat_sub$fl_csf_total_protein                    > 200   ] <- NA
dat_sub$fl_cyto_csf_ifn_g                      [dat_sub$fl_cyto_csf_ifn_g                       > 6   ] <- NA
dat_sub$fl_cyto_csf_il_10                      [dat_sub$fl_cyto_csf_il_10                       > 0.45   ] <- NA
#dat_sub$fl_cyto_csf_il_12p70                   [dat_sub$fl_cyto_csf_il_12p70                       ] <- NA
#dat_sub$fl_cyto_csf_il_13                      [dat_sub$fl_cyto_csf_il_13                          ] <- NA
dat_sub$fl_cyto_csf_il_1b                      [dat_sub$fl_cyto_csf_il_1b                       > 1   ] <- NA
dat_sub$fl_cyto_csf_il_2                       [dat_sub$fl_cyto_csf_il_2                        > 0.5   ] <- NA
#dat_sub$fl_cyto_csf_il_4                       [dat_sub$fl_cyto_csf_il_4                           ] <- NA
dat_sub$fl_cyto_csf_il_6                       [dat_sub$fl_cyto_csf_il_6                        > 6   ] <- NA
dat_sub$fl_cyto_csf_il_8                       [dat_sub$fl_cyto_csf_il_8                        > 200   ] <- NA
dat_sub$fl_cyto_csf_tnf_a                      [dat_sub$fl_cyto_csf_tnf_a                       > 0.8   ] <- NA
#dat_sub$fl_cyto_plasma_ifn_g                   [dat_sub$fl_cyto_plasma_ifn_g                       ] <- NA
dat_sub$fl_cyto_plasma_il_10                   [dat_sub$fl_cyto_plasma_il_10                    > 3   ] <- NA
dat_sub$fl_cyto_plasma_il_12p70                [dat_sub$fl_cyto_plasma_il_12p70                 > 1.5   ] <- NA
#dat_sub$fl_cyto_plasma_il_13                   [dat_sub$fl_cyto_plasma_il_13                       ] <- NA
dat_sub$fl_cyto_plasma_il_1b                   [dat_sub$fl_cyto_plasma_il_1b                    > 10   ] <- NA
#dat_sub$fl_cyto_plasma_il_2                    [dat_sub$fl_cyto_plasma_il_2                        ] <- NA
dat_sub$fl_cyto_plasma_il_4                    [dat_sub$fl_cyto_plasma_il_4                     > 1   ] <- NA
dat_sub$fl_cyto_plasma_il_6                    [dat_sub$fl_cyto_plasma_il_6                     > 300   ] <- NA
dat_sub$fl_cyto_plasma_il_8                    [dat_sub$fl_cyto_plasma_il_8                     > 600   ] <- NA
#dat_sub$fl_cyto_plasma_tnf_a                   [dat_sub$fl_cyto_plasma_tnf_a                       ] <- NA
dat_sub$fl_igg_index                           [dat_sub$fl_igg_index                            < 0.1   ] <- NA
#dat_sub$fl_lp_date                             [dat_sub$fl_lp_date                                 ] <- NA
#dat_sub$fl_msd_csf_mmp_1                       [dat_sub$fl_msd_csf_mmp_1                           ] <- NA
dat_sub$fl_msd_csf_mmp_10                      [dat_sub$fl_msd_csf_mmp_10                       > 350   ] <- NA
dat_sub$fl_msd_csf_mmp_2                       [dat_sub$fl_msd_csf_mmp_2                        < 4000   ] <- NA
#dat_sub$fl_msd_csf_mmp_3                       [dat_sub$fl_msd_csf_mmp_3                           ] <- NA
#dat_sub$fl_msd_csf_mmp_9                       [dat_sub$fl_msd_csf_mmp_9                           ] <- NA
dat_sub$fl_msd_mmp_2_index                     [dat_sub$fl_msd_mmp_2_index                      > 1   ] <- NA
dat_sub$fl_msd_mmp_9_index                     [dat_sub$fl_msd_mmp_9_index                      > 0.01   ] <- NA
#dat_sub$fl_msd_plasma_mmp_1                    [dat_sub$fl_msd_plasma_mmp_1                        ] <- NA
#dat_sub$fl_msd_plasma_mmp_10                   [dat_sub$fl_msd_plasma_mmp_10                       ] <- NA
#dat_sub$fl_msd_plasma_mmp_2                    [dat_sub$fl_msd_plasma_mmp_2                        ] <- NA
dat_sub$fl_msd_plasma_mmp_3                    [dat_sub$fl_msd_plasma_mmp_3                     > 100000   ] <- NA
dat_sub$fl_msd_plasma_mmp_9                    [dat_sub$fl_msd_plasma_mmp_9                     > 500000   ] <- NA
dat_sub$fl_myelin_basic_protein                [dat_sub$fl_myelin_basic_protein                 > 8   ] <- NA
#dat_sub$fl_oligoclonal_bands                   [dat_sub$fl_oligoclonal_bands                       ] <- NA
#dat_sub$fl_ptau                                [dat_sub$fl_ptau                                    ] <- NA
dat_sub$fl_serum_albumin                       [dat_sub$fl_serum_albumin                        < 1500   ] <- NA
#dat_sub$fl_serum_igg                           [dat_sub$fl_serum_igg                               ] <- NA
dat_sub$fl_simoa_csf_il_1b                     [dat_sub$fl_simoa_csf_il_1b                      > 0.05   ] <- NA
#dat_sub$fl_simoa_csf_n2pb_gfap                 [dat_sub$fl_simoa_csf_n2pb_gfap                     ] <- NA
#dat_sub$fl_simoa_csf_n2pb_nfl                  [dat_sub$fl_simoa_csf_n2pb_nfl                      ] <- NA
#dat_sub$fl_simoa_csf_n4pe_abeta_40             [dat_sub$fl_simoa_csf_n4pe_abeta_40                 ] <- NA
#dat_sub$fl_simoa_csf_n4pe_abeta_42             [dat_sub$fl_simoa_csf_n4pe_abeta_42                 ] <- NA
#dat_sub$fl_simoa_csf_n4pe_gfap                 [dat_sub$fl_simoa_csf_n4pe_gfap                     ] <- NA
#dat_sub$fl_simoa_csf_n4pe_nfl                  [dat_sub$fl_simoa_csf_n4pe_nfl                      ] <- NA
#dat_sub$fl_simoa_csf_nfl                       [dat_sub$fl_simoa_csf_nfl                           ] <- NA
#dat_sub$fl_simoa_csf_plgf                      [dat_sub$fl_simoa_csf_plgf                          ] <- NA
#dat_sub$fl_simoa_csf_ptau_v2                   [dat_sub$fl_simoa_csf_ptau_v2                       ] <- NA
#dat_sub$fl_simoa_edta_plasma_il_1b             [dat_sub$fl_simoa_edta_plasma_il_1b                 ] <- NA
#dat_sub$fl_simoa_edta_plasma_n4pe_abeta_40     [dat_sub$fl_simoa_edta_plasma_n4pe_abeta_40         ] <- NA
#dat_sub$fl_simoa_edta_plasma_n4pe_abeta_42     [dat_sub$fl_simoa_edta_plasma_n4pe_abeta_42         ] <- NA
#dat_sub$fl_simoa_edta_plasma_n4pe_gfap         [dat_sub$fl_simoa_edta_plasma_n4pe_gfap             ] <- NA
#dat_sub$fl_simoa_edta_plasma_n4pe_nfl          [dat_sub$fl_simoa_edta_plasma_n4pe_nfl              ] <- NA
#dat_sub$fl_simoa_edta_plasma_nfl               [dat_sub$fl_simoa_edta_plasma_nfl                   ] <- NA
#dat_sub$fl_simoa_edta_plasma_plgf              [dat_sub$fl_simoa_edta_plasma_plgf                  ] <- NA
#dat_sub$fl_simoa_edta_plasma_ptau_v2           [dat_sub$fl_simoa_edta_plasma_ptau_v2               ] <- NA
#dat_sub$fl_simoa_heparin_plasma_il_1b          [dat_sub$fl_simoa_heparin_plasma_il_1b              ] <- NA
#dat_sub$fl_simoa_heparin_plasma_n4pe_abeta_40  [dat_sub$fl_simoa_heparin_plasma_n4pe_abeta_40      ] <- NA
#dat_sub$fl_simoa_heparin_plasma_n4pe_abeta_42  [dat_sub$fl_simoa_heparin_plasma_n4pe_abeta_42      ] <- NA
#dat_sub$fl_simoa_heparin_plasma_n4pe_gfap      [dat_sub$fl_simoa_heparin_plasma_n4pe_gfap          ] <- NA
#dat_sub$fl_simoa_heparin_plasma_n4pe_nfl       [dat_sub$fl_simoa_heparin_plasma_n4pe_nfl           ] <- NA
#dat_sub$fl_simoa_heparin_plasma_nfl            [dat_sub$fl_simoa_heparin_plasma_nfl                ] <- NA
#dat_sub$fl_simoa_heparin_plasma_plgf           [dat_sub$fl_simoa_heparin_plasma_plgf               ] <- NA
#dat_sub$fl_simoa_heparin_plasma_ptau_v2        [dat_sub$fl_simoa_heparin_plasma_ptau_v2            ] <- NA
#dat_sub$fl_simoa_serum_ptau_v2                 [dat_sub$fl_simoa_serum_ptau_v2                     ] <- NA
#dat_sub$fl_zym_csf_mmp_2_62_kda                [dat_sub$fl_zym_csf_mmp_2_62_kda                    ] <- NA
#dat_sub$fl_zym_csf_mmp_2_72_kda                [dat_sub$fl_zym_csf_mmp_2_72_kda                    ] <- NA
dat_sub$fl_zym_csf_mmp_2_84_kda                [dat_sub$fl_zym_csf_mmp_2_84_kda                 > 6   ] <- NA
#dat_sub$fl_zym_csf_mmp_2_92_kda                [dat_sub$fl_zym_csf_mmp_2_92_kda                    ] <- NA
#dat_sub$fl_zym_mmp_2_index                     [dat_sub$fl_zym_mmp_2_index                         ] <- NA
dat_sub$fl_zym_mmp_9_index                     [dat_sub$fl_zym_mmp_9_index                      > 0.30   ] <- NA
#dat_sub$fl_zym_plasma_mmp_2_62_kda             [dat_sub$fl_zym_plasma_mmp_2_62_kda                 ] <- NA
#dat_sub$fl_zym_plasma_mmp_2_72_kda             [dat_sub$fl_zym_plasma_mmp_2_72_kda                 ] <- NA
#dat_sub$fl_zym_plasma_mmp_2_84_kda             [dat_sub$fl_zym_plasma_mmp_2_84_kda                 ] <- NA
dat_sub$fl_zym_plasma_mmp_2_92_kda             [dat_sub$fl_zym_plasma_mmp_2_92_kda              > 300   ] <- NA
#dat_sub$mri_dti_fabnormal_z                    [dat_sub$mri_dti_fabnormal_z                        ] <- NA
#dat_sub$mri_dti_freewater_mean                 [dat_sub$mri_dti_freewater_mean                     ] <- NA
#dat_sub$mri_dti_freewater_z                    [dat_sub$mri_dti_freewater_z                        ] <- NA
#dat_sub$mri_dti_histpk_z                       [dat_sub$mri_dti_histpk_z                           ] <- NA
#dat_sub$mri_dti_psmd                           [dat_sub$mri_dti_psmd                               ] <- NA
#dat_sub$mri_dti_psmd_z                         [dat_sub$mri_dti_psmd_z                             ] <- NA
#dat_sub$mri_mean_cort_thick_mm                 [dat_sub$mri_mean_cort_thick_mm                     ] <- NA
#dat_sub$mri_microbleeds                        [dat_sub$mri_microbleeds                            ] <- NA
#dat_sub$mri_percent_hippo_vol                  [dat_sub$mri_percent_hippo_vol                      ] <- NA
#dat_sub$mri_percent_wmh_vol                    [dat_sub$mri_percent_wmh_vol                        ] <- NA
#dat_sub$mri_perm_15t_mean_brain_ki             [dat_sub$mri_perm_15t_mean_brain_ki                 ] <- NA
dat_sub$mri_perm_15t_tot_ki                    [dat_sub$mri_perm_15t_tot_ki                     > 15   ] <- NA
#dat_sub$mri_perm_3t_tot                        [dat_sub$mri_perm_3t_tot                            ] <- NA
#dat_sub$mri_permeability                       [dat_sub$mri_permeability                           ] <- NA
#dat_sub$mri_spectroscopy                       [dat_sub$mri_spectroscopy                           ] <- NA
#dat_sub$mri_spectroscopy_cho_gpc_pch           [dat_sub$mri_spectroscopy_cho_gpc_pch               ] <- NA
#dat_sub$mri_spectroscopy_cr_mrs                [dat_sub$mri_spectroscopy_cr_mrs                    ] <- NA
#dat_sub$mri_spectroscopy_cr_pcr                [dat_sub$mri_spectroscopy_cr_pcr                    ] <- NA
#dat_sub$mri_spectroscopy_gln                   [dat_sub$mri_spectroscopy_gln                       ] <- NA
#dat_sub$mri_spectroscopy_glu                   [dat_sub$mri_spectroscopy_glu                       ] <- NA
#dat_sub$mri_spectroscopy_glu_gln               [dat_sub$mri_spectroscopy_glu_gln                   ] <- NA
#dat_sub$mri_spectroscopy_ins                   [dat_sub$mri_spectroscopy_ins                       ] <- NA
#dat_sub$mri_spectroscopy_naa                   [dat_sub$mri_spectroscopy_naa                       ] <- NA
#dat_sub$mri_spectroscopy_naa_naag              [dat_sub$mri_spectroscopy_naa_naag                  ] <- NA
#dat_sub$mri_vol_atrophy                        [dat_sub$mri_vol_atrophy                            ] <- NA
#dat_sub$mri_vol_gm                             [dat_sub$mri_vol_gm                                 ] <- NA
#dat_sub$mri_vol_hippocampal                    [dat_sub$mri_vol_hippocampal                        ] <- NA
dat_sub$mri_vol_intra_cranial_mm3              [dat_sub$mri_vol_intra_cranial_mm3               > 3000000   ] <- NA
#dat_sub$mri_vol_lesion                         [dat_sub$mri_vol_lesion                             ] <- NA
#dat_sub$mri_vol_total_brain                    [dat_sub$mri_vol_total_brain                        ] <- NA
#dat_sub$mri_vol_wm                             [dat_sub$mri_vol_wm                                 ] <- NA
#dat_sub$mri_vol_wm_jim                         [dat_sub$mri_vol_wm_jim                             ] <- NA
dat_sub$mri_vol_wm_ucd                         [dat_sub$mri_vol_wm_ucd                          > 130000   ] <- NA
#dat_sub$npsy_attention_t                       [dat_sub$npsy_attention_t                           ] <- NA
#dat_sub$npsy_boston_t                          [dat_sub$npsy_boston_t                              ] <- NA
#dat_sub$npsy_boston_tot_correct                [dat_sub$npsy_boston_tot_correct                    ] <- NA
#dat_sub$npsy_clock_cutoff                      [dat_sub$npsy_clock_cutoff                          ] <- NA
#dat_sub$npsy_clock_tot                         [dat_sub$npsy_clock_tot                             ] <- NA
#dat_sub$npsy_dig_sym_t                         [dat_sub$npsy_dig_sym_t                             ] <- NA
#dat_sub$npsy_dig_sym_tot                       [dat_sub$npsy_dig_sym_tot                           ] <- NA
#dat_sub$npsy_ds_tot_for_back                   [dat_sub$npsy_ds_tot_for_back                       ] <- NA
#dat_sub$npsy_ds_tot_t                          [dat_sub$npsy_ds_tot_t                              ] <- NA
#dat_sub$npsy_education                         [dat_sub$npsy_education                             ] <- NA
#dat_sub$npsy_executive_t                       [dat_sub$npsy_executive_t                           ] <- NA
#dat_sub$npsy_faq_score                         [dat_sub$npsy_faq_score                             ] <- NA
#dat_sub$npsy_fas_t                             [dat_sub$npsy_fas_t                                 ] <- NA
#dat_sub$npsy_fas_tot                           [dat_sub$npsy_fas_tot                               ] <- NA
#dat_sub$npsy_ft_dom_ave                        [dat_sub$npsy_ft_dom_ave                            ] <- NA
#dat_sub$npsy_ft_dom_ave_t                      [dat_sub$npsy_ft_dom_ave_t                          ] <- NA
#dat_sub$npsy_ft_nondom_ave                     [dat_sub$npsy_ft_nondom_ave                         ] <- NA
#dat_sub$npsy_ft_nondom_ave_t                   [dat_sub$npsy_ft_nondom_ave_t                       ] <- NA
#dat_sub$npsy_gdi                               [dat_sub$npsy_gdi                                   ] <- NA
#dat_sub$npsy_hvlt_delay_raw_tot                [dat_sub$npsy_hvlt_delay_raw_tot                    ] <- NA
#dat_sub$npsy_hvlt_delay_t                      [dat_sub$npsy_hvlt_delay_t                          ] <- NA
#dat_sub$npsy_hvlt_immed_raw_tot                [dat_sub$npsy_hvlt_immed_raw_tot                    ] <- NA
#dat_sub$npsy_hvlt_immed_t                      [dat_sub$npsy_hvlt_immed_t                          ] <- NA
#dat_sub$npsy_hvlt_rec_disc_index               [dat_sub$npsy_hvlt_rec_disc_index                   ] <- NA
#dat_sub$npsy_hvlt_rec_disc_t                   [dat_sub$npsy_hvlt_rec_disc_t                       ] <- NA
#dat_sub$npsy_language_t                        [dat_sub$npsy_language_t                            ] <- NA
#dat_sub$npsy_memory_t                          [dat_sub$npsy_memory_t                              ] <- NA
#dat_sub$npsy_mmse                              [dat_sub$npsy_mmse                                  ] <- NA
#dat_sub$npsy_mmse_t                            [dat_sub$npsy_mmse_t                                ] <- NA
#dat_sub$npsy_overall_t                         [dat_sub$npsy_overall_t                             ] <- NA
dat_sub$npsy_peg_dom_t                         [dat_sub$npsy_peg_dom_t                         > 90    ] <- NA
#dat_sub$npsy_peg_dom_time                      [dat_sub$npsy_peg_dom_time                          ] <- NA
dat_sub$npsy_peg_nondom_t                      [dat_sub$npsy_peg_nondom_t                      > 80    ] <- NA
dat_sub$npsy_peg_nondom_time                   [dat_sub$npsy_peg_nondom_time                   > 250    ] <- NA
#dat_sub$npsy_processing_t                      [dat_sub$npsy_processing_t                          ] <- NA
#dat_sub$npsy_rcft_copy                         [dat_sub$npsy_rcft_copy                             ] <- NA
#dat_sub$npsy_rcft_delay_recall                 [dat_sub$npsy_rcft_delay_recall                     ] <- NA
#dat_sub$npsy_rcft_delay_recall_t               [dat_sub$npsy_rcft_delay_recall_t                   ] <- NA
#dat_sub$npsy_rcft_imm_recall                   [dat_sub$npsy_rcft_imm_recall                       ] <- NA
#dat_sub$npsy_rcft_imm_recall_t                 [dat_sub$npsy_rcft_imm_recall_t                     ] <- NA
#dat_sub$npsy_rcft_recog                        [dat_sub$npsy_rcft_recog                            ] <- NA
#dat_sub$npsy_rcft_recog_t                      [dat_sub$npsy_rcft_recog_t                          ] <- NA
#dat_sub$npsy_strp_color_score                  [dat_sub$npsy_strp_color_score                      ] <- NA
#dat_sub$npsy_strp_color_score_t                [dat_sub$npsy_strp_color_score_t                    ] <- NA
#dat_sub$npsy_strp_colorword_score              [dat_sub$npsy_strp_colorword_score                  ] <- NA
#dat_sub$npsy_strp_colorword_score_t            [dat_sub$npsy_strp_colorword_score_t                ] <- NA
#dat_sub$npsy_strp_cw_interference              [dat_sub$npsy_strp_cw_interference                  ] <- NA
#dat_sub$npsy_strp_cw_interference_t            [dat_sub$npsy_strp_cw_interference_t                ] <- NA
#dat_sub$npsy_strp_word_score                   [dat_sub$npsy_strp_word_score                       ] <- NA
#dat_sub$npsy_strp_word_score_t                 [dat_sub$npsy_strp_word_score_t                     ] <- NA
#dat_sub$npsy_sym_srch_t                        [dat_sub$npsy_sym_srch_t                            ] <- NA
#dat_sub$npsy_sym_srch_tot                      [dat_sub$npsy_sym_srch_tot                          ] <- NA
#dat_sub$npsy_wcst_cat                          [dat_sub$npsy_wcst_cat                              ] <- NA
#dat_sub$npsy_wcst_cat_percent                  [dat_sub$npsy_wcst_cat_percent                      ] <- NA
#dat_sub$npsy_wcst_con_resp                     [dat_sub$npsy_wcst_con_resp                         ] <- NA
#dat_sub$npsy_wcst_con_resp_t                   [dat_sub$npsy_wcst_con_resp_t                       ] <- NA
#dat_sub$npsy_wcst_pers_err                     [dat_sub$npsy_wcst_pers_err                         ] <- NA
#dat_sub$npsy_wcst_pers_err_t                   [dat_sub$npsy_wcst_pers_err_t                       ] <- NA
#dat_sub$npsy_wcst_pers_resp                    [dat_sub$npsy_wcst_pers_resp                        ] <- NA
#dat_sub$npsy_wcst_pers_resp_t                  [dat_sub$npsy_wcst_pers_resp_t                      ] <- NA
#dat_sub$npsy_wcst_tot_err                      [dat_sub$npsy_wcst_tot_err                          ] <- NA
#dat_sub$npsy_wcst_tot_err_t                    [dat_sub$npsy_wcst_tot_err_t                        ] <- NA
dat_sub$npsy_wrat_t                            [dat_sub$npsy_wrat_t                             > 80   ] <- NA
dat_sub$npsy_wrat_tot                          [dat_sub$npsy_wrat_tot                           > 60   ] <- NA

```

Plot again

```{r chunk-014, fig.height = 4, fig.width = 10}
# var_list_plot <-
#   c(Var$list$var_npsy, Var$list$var_fl_csf, Var$list$var_fl_plasma)

par(mfrow=c(1,3))
for (i_plot in 1:length(var_list_plot)) {
  ## i_plot = 1
  #if (i_plot > 3) {next}

  if( all(is.na(dat_sub[[ var_list_plot[i_plot] ]])) ) {
    next
  }

  plot(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , type  = "p"
  , main  = paste0("After: ", var_list_plot[i_plot])
  , xlab  = ""
  , ylab  = ""
  , col   = dat_sub$dx_current
  , pch   = 20  #3
  , xlim  = c(1, nrow(dat_sub) * 1.2)
  )
  text(
    x     = 1:nrow(dat_sub)
  , y     = dat_sub[[ var_list_plot[i_plot] ]]
  , labels = 1:nrow(dat_sub)
  , pos   = 4 # 1 = below, 2 = left, 3 = above, 4 = right
  )
}
```

## Plot data distributions

See `out/plot_medians_*.png` plots.

```{r}

# log2 versions
for (n_var in str_subset(names(dat_sub), pattern = "^fl_")) {
  dat_sub[, n_var ] <-  # paste0(n_var, "_log2")
    e_log_shift(dat_sub[, n_var])

  #labelled::var_label(dat_sub[, paste0(n_var, "_log2") ]) <-
  #  paste("log2", labelled::var_label( dat_sub[, n_var] ))
}

```

```{r}
#| fig-width:  12
#| fig-height: 14

for (i_var_group in seq_along(var_groups)) {
  ## i_var_group = 1

  print(paste0(var_groups[i_var_group], "                ===================="))

  p_list <- list()

  for (i_var in seq_along(Var[["list"]][[ var_groups[i_var_group] ]])) {
    ## i_var = 1

    var_this <- Var[["list"]][[ var_groups[i_var_group] ]][i_var]

    print(paste0(var_this, "                ----------"))

    if (
      (length(var_this) == 0)           |
      all(is.na(dat_sub[[ var_this ]]))
    ) {
      next
    }

    p_list[[ i_var ]] <-
      e_plot_compare_medians(
        dat_plot            = dat_sub |> tidyr::drop_na(var_this) |> dplyr::mutate(dx_current = dx_current |> fct_drop())
      , var_response        = var_this
      , var_groups          = "dx_current"
      #, label_response      = labelled::var_label(dat_mtcars_e$mpg)
      #, label_groups        = labelled::var_label(dat_mtcars_e$cyl)
      , plot_title          = var_this
      , plot_subtitle       = "Comparison of medians"
      , comparisons         = NULL #c("Control", "Sweden")) #"all"
      , cm_method           = "wilcox.test"
      , cm_p.adjust.method  = "holm"
      , cm_dot              = "median"
      , cm_fun              = "median_iqr"
      , cm_error.plot       = "errorbar"
      , sw_add              = "dotplot"
      , sw_caption_desc     = TRUE
      )


  } # i_var

  if (length(p_list) == 0) { next }

  p_arranged <-
    cowplot::plot_grid(
      plotlist  = p_list
    , nrow      = 3
    , ncol      = 4
    , labels    = "AUTO"
    )

  #p_arranged %>% print()

  ggsave(
      paste0("out/plot_medians_VCI_", var_groups[i_var_group], ".png")
    , plot   = p_arranged
    , width  = 15
    , height = 12    #4*ceiling(length(Var[["list"]][[ var_groups[i_var_group] ]])/2)
    ## png, jpeg
    , dpi    = 300
    , bg     = "white"
    ## pdf
    #, units  = "in"
    #, useDingbats = FALSE
    )

} # i_var_group
```

# Tables and plots

## Demographic summaries

```{r chunk-029, results='asis', echo=TRUE}
dat_sub |>
  mutate(
    dx_current = dx_current |> fct_drop()
  ) |>
  pull(dx_current) |>
  table() |>
  as.data.frame() |>
  t() |>
  as.data.frame()
  # |>
  # readr::write_csv(
  #   file = "out/tab_dx_current_freq.csv"
  # , col_names = FALSE
  # )

table1_out <-
  dat_sub |>
  gtsummary::tbl_summary(
    by           = "dx_current"        # use gtsummary::tbl_strata for more than 1
  # , label        = NULL                          # automatically uses the label attribute
  # , statistic    = NULL
  # , digits       = NULL
  # , type         = NULL
  # , value        = NULL
  # , missing      = NULL
  # , missing_text = NULL
  # , sort         = NULL
  # , percent      = NULL
  , include      = Var |> unlist() |> unique() |> sort()
  ) |>
  gtsummary::add_p()          |>  # add p-values to the output comparing values across groups
  #gtsummary::add_q()          |>  # add a column of q values to control for multiple comparisons
  gtsummary::bold_p(t = 0.05, q = FALSE) |> # Bold significant p-values or q-values
  #gtsummary::add_difference() |>  # add column for difference between two group, confidence interval, and p-value
  #gtsummary::add_ci()         |>  # add confidence intervals
  gtsummary::add_overall()    |>  # add a column with overall summary statistics
  gtsummary::add_stat_label() |>  # add label for the summary statistics shown in each row
  gtsummary::add_n()          |>  # add a column with N (or N missing) for each variable
  gtsummary::bold_labels()    |>  # bold variable name labels
  gtsummary::italicize_levels() |> # italicize levels
  gtsummary::modify_caption("Table title")
# print to Viewer
table1_out
# export to Excel
table1_out |>
  gtsummary::as_hux_xlsx(file = paste0("out/tab1_vars", "_", "all", ".xlsx"))
```

## Remove variables that greatly differ between VCI and Sweden

```{r}
# var_VCI_Sweden_different_list <-
#   c(
#     "fl_angio_csf_bfgf"
#   , "fl_angio_csf_flt_1"
#   , "fl_angio_csf_plgf"
#     "fl_angio_csf_tie2"
#   , "fl_angio_csf_vegf_a"
#   , "fl_angio_csf_vegf_c"
#   , "fl_angio_csf_vegf_d"
#   , "fl_angio_plasma_bfgf"
#   , "fl_angio_plasma_flt_1"
#   , "fl_angio_plasma_plgf"
#   , "fl_angio_plasma_tie2"
#   , "fl_angio_plasma_vegf_a"
#   , "fl_angio_plasma_vegf_c"
#   , "fl_angio_plasma_vegf_d"
#   , "fl_cyto_csf_ifn_g"
#   , "fl_cyto_csf_il_10"
#   , "fl_cyto_csf_il_12p70"
#   , "fl_cyto_csf_il_13"
#   , "fl_cyto_csf_il_1b"
#   , "fl_cyto_csf_il_2"
#   , "fl_cyto_csf_il_4"
#   , "fl_cyto_csf_il_6"
#   , "fl_cyto_csf_il_8"
#   , "fl_cyto_csf_tnf_a"
#   , "fl_cyto_plasma_ifn_g"
#   , "fl_cyto_plasma_il_10"
#   , "fl_cyto_plasma_il_12p70"
#   , "fl_cyto_plasma_il_13"
#   , "fl_cyto_plasma_il_1b"
#   , "fl_cyto_plasma_il_2"
#   , "fl_cyto_plasma_il_4"
#   , "fl_cyto_plasma_il_6"
#   , "fl_cyto_plasma_il_8"
#   , "fl_cyto_plasma_tnf_a"
#   , "fl_msd_csf_mmp_1"
#   , "fl_msd_csf_mmp_10"
#   , "fl_msd_csf_mmp_2"
#   , "fl_msd_csf_mmp_3"
#   , "fl_msd_csf_mmp_9"
#   , "fl_msd_plasma_mmp_1"
#   , "fl_msd_plasma_mmp_10"
#   , "fl_msd_plasma_mmp_2"
#   , "fl_msd_plasma_mmp_3"
#   , "fl_msd_plasma_mmp_9"
#   , "fl_simoa_csf_n2pb_gfap"
#   , "fl_simoa_csf_n2pb_nfl"
#   , "fl_simoa_csf_n4pe_nfl"
#   , "fl_simoa_csf_nfl"
#   , "fl_simoa_csf_plgf"
#   , "fl_simoa_edta_plasma_n4pe_nfl"
#   , "fl_simoa_edta_plasma_nfl"
#   , "fl_simoa_edta_plasma_plgf"
#   , "fl_simoa_heparin_plasma_n4pe_nfl"
#   , "fl_simoa_heparin_plasma_nfl"
#   , "fl_simoa_heparin_plasma_plgf"
#   , "fl_simoa_plasma_n2pb_gfap"
#   )
#
# var_VCI_Sweden_different_list
```

```{r}

# # remove different variables from lists
# for (n_var_list in var_groups) {
#   Var[["list"]][[ n_var_list ]] <-
#     Var[["list"]][[ n_var_list ]][ Var[["list"]][[ n_var_list ]] %notin% var_VCI_Sweden_different_list ]
# }
```

## --------------------------------------------------------------------------------

# Random forest, which fluid variable help predict DX

-   Keep observations with at least 50% of the variables non-NA
-   Impute the missing values

```{r chunk-044}
library(doParallel)  # for %dopar% operator
num_cores <- 6 # length(var_groups)
#num_cores <- parallel::detectCores() - 2
doParallel::registerDoParallel(num_cores)

tictoc::tic(msg = "Timer") # start timer

## don't prespecify outer "results" list

results <-
  foreach::foreach(
    n_var_group = var_groups
  #, .combine =                             # default is a list
  , .inorder = TRUE                         # FALSE is faster
  , .packages = c("erikmisc", "tidyverse", "labelled", "patchwork", "randomForestSRC")  # character vector of packages that the tasks depend on
  , .export = NULL                          # character vector of variables to export
  #) %do% # sequential
  ) %dopar%  # parallel
  {

#out_rf_results <- list()
#for (n_var_group in var_groups) {
  ## n_var_group = var_groups[1]

  this_result <- list()
  #out_rf_results[[ n_var_group ]] <- list()
  for (n_dx_group in dx_groups) {
    ## n_dx_group = dx_groups[5]
    this_result[[ n_dx_group ]] <- list()
    #out_rf_results[[ n_var_group ]][[ n_dx_group ]] <- list()

    text_var_dx <- paste0(n_var_group, ",  ", n_dx_group)

    print("")
    print("")
    print("")
    print(paste0(
      "RF Training:  "
    , n_var_group
    , ",  "
    , n_dx_group
    , "  ########################################"
    ))


    dat_rf_class <-
      dat_sub |>
      dplyr::select(
        tidyselect::all_of(n_dx_group)
      , tidyselect::all_of(Var[["list"]][[ n_var_group ]])
      ) |>
      dplyr::rename(
        "dx_current" = n_dx_group
      ) |>
      dplyr::rowwise() |>
      dplyr::mutate(
        prop_complete = mean(!is.na(dplyr::across(tidyselect::everything())))
      ) |>
      dplyr::ungroup() |>
      dplyr::filter(
        # keep observations where at least 50% have data across columns
        prop_complete >= 0.50
      #, dx_current %notin% "Sweden"
      , !is.na(dx_current)
      ) |>
      dplyr::select(
        -prop_complete
      ) |>
      dplyr::mutate(
        dx_current = dx_current |> forcats::fct_drop()
      ) |>
      as.data.frame()

    col_prop_complete <- colMeans(!is.na(dat_rf_class))

    dat_rf_class <-
      dat_rf_class |>
      dplyr::select(
        dx_current
        # keep variables where at least 50% have data across observations (already filtered)
      ,  tidyselect::all_of( names(col_prop_complete)[col_prop_complete >= 0.50] )
      )

    dim(dat_rf_class)
    table(dat_rf_class[, 1])

    # missing either diagnosis or predictor variables, skip
    if (ncol(dat_rf_class) < 2) {
      out_rf_results[[ n_var_group ]][[ n_dx_group ]] <- NA
      next
    }


    # dat_miss_impute_fast <-
    #   randomForestSRC::impute(
    #     data      = dat_rf_class
    #   )
    ## Random Forest Slow imputation mode is more accurate
    #    uses randomForestSRC::rfsrc(na.action = "na.impute")
    o_dat_rf_class_impute <-
      randomForestSRC::rfsrc(
        data      = dat_rf_class
      , na.action = "na.impute"
      )

    ind_row_impute <-
      which(!complete.cases(o_dat_rf_class_impute$xvar))

    dat_rf_class_impute <-
      o_dat_rf_class_impute$xvar
    dat_rf_class_impute[ ind_row_impute, ] <-
      o_dat_rf_class_impute$imputed.data


    #dat_rf_class

    p1 <-
      e_plot_missing(
        dat_plot       = dat_rf_class
      , var_group      = "dx_current"
      , sw_group_sort  = TRUE
      #, var2_sort      = "prop_complete"
      )
    p2 <-
      e_plot_missing(
        dat_plot       = dat_rf_class_impute
      , var_group      = "dx_current"
      , sw_group_sort  = TRUE
      #, var2_sort      = "prop_complete"
      )

    p_arranged <-
      cowplot::plot_grid(
        plotlist  = list(p1, p2)
      , nrow      = 1
      #, ncol      = 2
      , labels    = "AUTO"
      )

    #p_arranged %>% print()

    this_result[[ n_dx_group ]][[ "plot_missing" ]] <-
    #out_rf_results[[ n_var_group ]][[ n_dx_group ]][[ "plot_missing" ]] <-
      p_arranged

    ggsave(
        paste0("out/plot_rf_train__", n_dx_group, "_-_", n_var_group, "__missing",  ".png")
      , plot   = p_arranged
      , width  = 12
      , height = 8
      ## png, jpeg
      , dpi    = 300
      , bg     = "white"
      ## pdf
      #, units  = "in"
      #, useDingbats = FALSE
      )


    # Perform RF
    out_e_rf <-
      e_rfsrc_classification(
        dat_rf_class    = dat_rf_class_impute
      , rf_y_var        = NULL  # if not specified, assumed to be first column
      , rf_x_var        = NULL  # if not specified, assumed to be second through last columns
      , sw_rfsrc_ntree  = 1000
      , sw_alpha        = 0.20  # for subsample bootsample model selection and VIMP plot
      )

    this_result[[ n_dx_group ]][[ "out_e_rf" ]] <-
    #out_rf_results[[ n_var_group ]][[ n_dx_group ]][[ "out_e_rf" ]] <-
      out_e_rf

    # # Compile summary plot
    # text_formula <-
    #   paste(as.character(out_e_rf$rf_formula)[c(2, 1, 3)], collapse = " ")
    #   #|>
    #   #stringr::str_wrap(width = 60)
    #
    # library(patchwork)
    # plot_design <-
    #   "AAAABB
    #    AAAABB
    #    CCCCDD
    #    CCCCDD
    #    EEEEEE
    #    EEEEEE"
    #
    # # If selected model has x variables
    # if (length(out_e_rf$rf_x_var_sel)) {
    #   #this_result[[ n_dx_group ]][[ "plot_rf_train_all_summary" ]] <-
    #   out_rf_results[[ n_var_group ]][[ n_dx_group ]][[ "plot_rf_train_all_summary" ]] <-
    #     cowplot::plot_grid(out_e_rf$plot_o_class              ) +
    #     cowplot::plot_grid(out_e_rf$plot_o_class_subsample    ) +
    #     cowplot::plot_grid(out_e_rf$plot_o_class_sel          ) +
    #     cowplot::plot_grid(out_e_rf$plot_o_class_sel_subsample) +
    #     cowplot::plot_grid(plotlist = out_e_rf$plot_o_class_sel_ROC$plot_roc, nrow = 1) +
    #     plot_layout(design = plot_design) +
    #     plot_annotation(
    #       title       = text_var_dx
    #     , subtitle    = text_formula
    #     , caption     = paste0("Overall AUC = ", round(out_e_rf$o_class_sel_AUC, 3))
    #     , tag_levels  = "A"
    #     )
    # } else {
    #   #this_result[[ n_dx_group ]][[ "plot_rf_train_all_summary" ]] <-
    #   out_rf_results[[ n_var_group ]][[ n_dx_group ]][[ "plot_rf_train_all_summary" ]] <-
    #     cowplot::plot_grid(out_e_rf$plot_o_class              ) +
    #     cowplot::plot_grid(out_e_rf$plot_o_class_subsample    ) +
    #     #cowplot::plot_grid(out_e_rf$plot_o_class_sel          ) +
    #     #cowplot::plot_grid(out_e_rf$plot_o_class_sel_subsample) +
    #     #cowplot::plot_grid(plotlist = out_e_rf$plot_o_class_sel_ROC$plot_roc, nrow = 1) +
    #     plot_spacer() +
    #     plot_spacer() +
    #     plot_spacer() +
    #     plot_layout(design = plot_design) +
    #     plot_annotation(
    #       title       = text_var_dx
    #     , subtitle    = text_formula
    #     , caption     = "No x variables selected"
    #     , tag_levels  = "A"
    #     )
    # }

    ggsave(
        paste0("out/plot_rf_train__", n_dx_group, "_-_", n_var_group, "__all_summary",  ".png")
      , plot   = this_result[[ n_dx_group ]][[ "out_e_rf" ]][[ "plot_rf_train_all_summary" ]]
      #, plot   = out_rf_results[[ n_var_group ]][[ n_dx_group ]][[ "out_e_rf" ]][[ "plot_rf_train_all_summary" ]]
      , width  = 16
      , height = 16
      ## png, jpeg
      , dpi    = 300
      , bg     = "white"
      ## pdf
      #, units  = "in"
      #, useDingbats = FALSE
      )

    out_save <-
      this_result[[ n_dx_group ]]
      #out_rf_results[[ n_var_group ]][[ n_dx_group ]]

    save(
      out_save
    , file = paste0("out/RData__", n_dx_group, "_-_", n_var_group, "__results", ".RData")
    )

    ## Do not save result
    this_result[[ n_dx_group ]] <-
    #out_rf_results[[ n_var_group ]][[ n_dx_group ]] <-
      NULL

  } # n_dx_group
#} # n_var_group

  return(this_result)
} # foreach

tictoc::toc() # end timer

# explicitly close the implicitly created cluster
doParallel::stopImplicitCluster()


#save.image(
#    file = "out/RF_Trained.RData"
#  )

```

# Predictions

```{r}

## load(
##   "out/RData__var_fl_all_-_dx_current_ADSIVDMX_Control.RData"
## )
##
## dat_to_pred <-
##   dat_sub |>
##   dplyr::filter(
##     I_Sweden
##   ) |>
##   dplyr::select(
##     ID
##   , tidyselect::all_of(out_save$out_e_rf$o_class_sel$xvar.names)
##   ) |>
##   tidyr::drop_na()
##
##
## out_save$out_e_rf$o_class_sel
##
## out_e_rf_pred <-
##   predict(
##     object        = out_save$out_e_rf$o_class_sel
##   , newdata       = dat_sub |> dplyr::filter(I_Sweden)
##   )
##
## out_e_rf_pred$predicted
## out_e_rf_pred$class
##
## dat_to_pred <-
##   dat_to_pred |>
##   dplyr::bind_cols(
##     out_e_rf_pred$predicted
##   , tibble::tibble(class = out_e_rf_pred$class)
##   ) |>
##   dplyr::select(
##     ID
##   , class
##   , tidyselect::all_of(colnames(out_e_rf_pred$predicted))
##   , tidyselect::everything()
##   )
##
## readr::write_csv(
##     dat_to_pred
##   , file = "out/dat_Sweden_pred__var_fl_all_-_dx_current_ADSIVDMX_Control.csv"
##   )
##
## fl_angio_csf_flt_1
## fl_angio_csf_plgf
## fl_angio_csf_vegf_a
## fl_cyto_csf_il_8
## fl_cyto_plasma_il_6
## fl_cyto_plasma_tnf_a
## fl_msd_csf_mmp_1
## fl_msd_csf_mmp_10
## fl_msd_csf_mmp_9
## fl_msd_plasma_mmp_1
## fl_msd_plasma_mmp_3


```
