---
title:        "Composite Inflammatory Score: BBB Disruption Factor"
subtitle:     "Based on Hilmer, et al., (2022). *Blood-brain barrier disruption measured by albumin index correlates with inflammatory fluid biomarkers*"
author:       Azadeh Golduzian
date:         last-modified # today, now, last-modified
date-format:  long          # full, long, medium, short, iso,   https://quarto.org/docs/reference/dates.html
format:
  html:
    theme:                  litera
    highlight-style:        atom-one
    page-layout:            full      # article, full   # https://quarto.org/docs/output-formats/page-layout.html
    toc:                    true
    toc-depth:              4
    toc-location:           body      # left, body, right
    number-sections:        true      # true, false
    number-depth:           3
    code-fold:              show      # true (initially hidden), false, show (initially shown)
    code-tools:                       # menu top-right to show/hide all code
      toggle:               true
      caption:              "Code"    # none
      source:               false
    code-overflow:          scroll    # scroll, wrap
    code-block-bg:          true
    code-block-border-left: "#30B0E0"
    df-print:               paged     # default, kable, tibble, paged   # https://quarto.org/docs/computations/r.html
fig-width:                  6
fig-height:                 4
execute: # https://quarto.org/docs/computations/execution-options.html, https://quarto.org/docs/computations/r.html
  cache:    false   # false, true
  eval:     true    # true, false  Evaluate the code chunk (if false, just echos the code into the output)
  echo:     true    # true, false  Include the source code in output
---

<!---
# Erik's compile commands in R:
  setwd("D:/Dropbox/StatAcumen/consult/Authorship/2014_GaryRosenberg_UNM_NIH_VascularDementia_MRI/2024_ADRC/Analyses/202501_ADRC_Permeability")
  fn_qmd <- "Golduzian_Rosenberg_ADRC_Inflammation_Analysis_4-Composite_Inflammation_Score_20260326.qmd"
  quarto::quarto_render(input = fn_qmd)
-->

```{r init, echo=FALSE}
#options(width = 80)
#options(warn  = -1)
options(str   = strOptions(list.len = 1e3))
options(knitr.kable.NA = '') # Display NAs as blanks
my_seed <- 34567
set.seed(my_seed)
```

--------------------------------------------------------------------------------

# Next actions

- Identify best-correlated biomarkers with albumin index (BBB disruption marker)
- Build composite inflammatory score using backward BIC regression
  - Model 1: `fl_albumin_index_log2 ~ CSF + Plasma + MRI` predictors
  - Model 2: `mri_perm_3t_tot_log2 ~ CSF + Plasma + MRI` predictors (Ktrans)
- Target: Beat R² = 0.53
- Save composite score `score_inflammation` for downstream use

--------------------------------------------------------------------------------

# Data preliminaries

```{r chunk-003}
library(erikmisc)
library(tidyverse)
library(labelled)
library(corrplot)
library(broom)
library(dunn.test)
library(ggpubr)
library(modelsummary)
library(patchwork)

dir.create("All_figures_for_paper",      showWarnings = FALSE)
dir.create("All_figures_for_paper/composite", showWarnings = FALSE)
```

## Load saved data

```{r chunk-004}
load("dat_sub_Data_inflammation_analysis_paper_06052025.RData")
load("dat_sub_and_Var_for_inflammation_analysis.RData")
```

## Variable list for this analysis

Candidate predictors are all CSF, Plasma, and MRI biomarkers
(excluding the two response variables themselves).

```{r chunk-005}
var_resp_albumin <- "fl_albumin_index_log2"
var_resp_ktrans  <- "mri_perm_3t_tot_log2"

# All candidate predictors: CSF + Plasma + MRI, excluding response vars
var_list_all <-
  c(Var$csf, Var$plasma, Var$mri)

vars_exclude_from_predictors <-
  c(
    "ptid", "age", "sex", "demo_edu_yrs", "dx_current"
  , var_resp_albumin
  , var_resp_ktrans
  , "mri_perm_3t_tot"   # raw (non-log2) version of Ktrans
  )

list_var_covar <-
  var_list_all[
    var_list_all %notin% vars_exclude_from_predictors
  ]

cat("Number of candidate predictors:", length(list_var_covar), "\n")
print(list_var_covar)
```

## Z-score scale predictors (keep responses on log2 scale)

```{r chunk-006}
# data for demographics, before z-score scaling
dat_sheet_demo <- dat_sub

dat_sheet <-
  dat_sub |>
  dplyr::mutate(
    across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), scale)
  , across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), as.numeric)
  )

# restore variable labels
for (n_label in as.character(unlist(Var))) {
  labelled::var_label(dat_sheet     [[ n_label ]]) <-
    labelled::var_label(dat_sheet_demo[[ n_label ]])
}
```

--------------------------------------------------------------------------------

# Step 1: Visualize Missingness

Using `e_plot_missing` (from erikmisc) sorted by albumin index to identify which
biomarkers have the most complete data in subjects with BBB measures.

```{r chunk-007, fig.height=10, fig.width=12}
p_missing <-
  e_plot_missing(
    dat_plot       = dat_sub
  , var_group      = "dx_current"
  , sw_group_sort  = TRUE
  , var2_sort      = "fl_albumin_index_log2"
  )
p_missing |> print()

ggsave(
    "All_figures_for_paper/composite/plot_missing_for_composite.png"
  , plot   = p_missing
  , width  = 14
  , height = 10
  , dpi    = 300
  , bg     = "white"
  )
```

--------------------------------------------------------------------------------

# Step 2: Best Subset — Observations vs. Variables Trade-off

`e_plot_complete_by_variable_subset` shows how sample size changes as we add
more biomarkers.  We search over **all CSF + Plasma + MRI** candidates
(the full variable universe Erik approved) and pick the row that gives the
best balance between number of predictors and complete observations.

## For Albumin Index (BBB)

```{r chunk-008, fig.height=6, fig.width=10}
vars_for_subset_albumin <-
  c(var_resp_albumin, list_var_covar) |> unique()

result_subset_albumin <-
  e_plot_complete_by_variable_subset(
    dat      = dat_sheet
  , var_list = vars_for_subset_albumin
  , var_resp = var_resp_albumin
  , sw_plot_complete = 20   # show top 20 subsets
  )

# Inspect the table to choose the best row
result_subset_albumin$tab |> print(n = 30)
```

Choose the subset row that maximises predictors while keeping n ≥ 60.
Adjust `i_best_albumin` after inspecting the table above.

```{r chunk-009}
# ---- USER: adjust index after inspecting table above ----
i_best_albumin <- 5   # row index in result_subset_albumin$tab

best_vars_albumin <- result_subset_albumin$var_names[[ i_best_albumin ]]
n_best_albumin    <- result_subset_albumin$tab$n_complete[ i_best_albumin ]

cat("Selected subset for Albumin model:\n")
cat("  n complete:", n_best_albumin, "\n")
cat("  Variables :", paste(best_vars_albumin, collapse = ", "), "\n")
```

## For Ktrans / MRI Permeability

```{r chunk-010, fig.height=6, fig.width=10}
vars_for_subset_ktrans <-
  c(var_resp_ktrans, list_var_covar) |> unique()

result_subset_ktrans <-
  e_plot_complete_by_variable_subset(
    dat      = dat_sheet
  , var_list = vars_for_subset_ktrans
  , var_resp = var_resp_ktrans
  , sw_plot_complete = 20
  )

result_subset_ktrans$tab |> print(n = 30)
```

```{r chunk-011}
# ---- USER: adjust index after inspecting table above ----
i_best_ktrans <- 5   # row index in result_subset_ktrans$tab

best_vars_ktrans <- result_subset_ktrans$var_names[[ i_best_ktrans ]]
n_best_ktrans    <- result_subset_ktrans$tab$n_complete[ i_best_ktrans ]

cat("Selected subset for Ktrans model:\n")
cat("  n complete:", n_best_ktrans, "\n")
cat("  Variables :", paste(best_vars_ktrans, collapse = ", "), "\n")
```

--------------------------------------------------------------------------------

# Step 3: Backward BIC Regression — Albumin Index

## Define helper: BIC backward selection

```{r chunk-012}
# Reusable backward BIC selector (same pattern as Analysis_1)
f_model_sel <-
  function(
    dat_this      = dat_lm
  , lm_criterion  = c("AIC", "BIC")[2]
  ) {

  if (lm_criterion == "AIC") { AIC_k = 2                     }
  if (lm_criterion == "BIC") { AIC_k = log(nrow( dat_this )) }

  # Full model with all candidate predictors
  formula_full <-
    as.formula(
      paste0(
        "Y ~ age + sex + "
      , paste(setdiff(names(dat_this), c("Y", "age", "sex")), collapse = " + ")
      )
    )

  fit_full <- lm(formula = formula_full, data = dat_this)

  # Step-wise backward with BIC / AIC penalty
  fit_red <-
    stats::step(
      object    = fit_full
    , direction = "backward"
    , trace     = 0
    , k         = AIC_k
    )

  return(fit_red)
} # f_model_sel

lm_criterion <- c("AIC", "BIC")[2]
```

## Fit full → reduced model (Albumin Index)

```{r chunk-013}
# Build complete-case dataset for the chosen subset
predictor_vars_albumin <-
  setdiff(best_vars_albumin, var_resp_albumin)

dat_lm_albumin <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_albumin)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_albumin)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model — complete observations:", nrow(dat_lm_albumin), "\n")

# Backward BIC selection
fit_albumin_full <-
  lm(
    formula =
      as.formula(
        paste0(
          "Y ~ age + sex + "
        , paste(predictor_vars_albumin, collapse = " + ")
        )
      )
  , data = dat_lm_albumin
  )

fit_albumin_red <-
  stats::step(
    object    = fit_albumin_full
  , direction = "backward"
  , trace     = 0
  , k         = log(nrow(dat_lm_albumin))  # BIC
  )

cat("\n=== Reduced Albumin Model (BIC) ===\n")
summary(fit_albumin_red)

cat("\nR² =", summary(fit_albumin_red)$r.squared |> round(3), "\n")
cat("Adj R² =", summary(fit_albumin_red)$adj.r.squared |> round(3), "\n")
```

## Residual diagnostics (Albumin model)

```{r chunk-014, fig.height=4, fig.width=10}
par(mfrow = c(1, 3))
plot(fit_albumin_red, which = c(1, 2, 3))
par(mfrow = c(1, 1))
```

## Coefficient plot (Albumin model)

```{r chunk-015, fig.height=5, fig.width=7}
p_coef_albumin <-
  fit_albumin_red |>
  broom::tidy(conf.int = TRUE) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::mutate(
    term = term |>
      stringr::str_replace_all(stringr::fixed("sexMale"), "Sex (Male)") |>
      stringr::str_replace_all("^age$", "Age")
  , sig = dplyr::if_else(p.value < 0.05, "p < 0.05", "n.s.")
  ) |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = estimate, y = reorder(term, estimate)
    , colour = sig
    )
  ) +
  ggplot2::theme_bw() +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed", colour = "gray60") +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(xmin = conf.low, xmax = conf.high)
  , height = 0.3
  ) +
  ggplot2::scale_colour_manual(
    values = c("p < 0.05" = "#E31A1C", "n.s." = "gray50")
  ) +
  ggplot2::labs(
    title    = "Reduced Model — Albumin Index (BBB)"
  , subtitle = paste0("BIC backward selection  |  R² = ",
                      round(summary(fit_albumin_red)$r.squared, 3),
                      "  |  n = ", nrow(dat_lm_albumin))
  , x        = "Standardized Coefficient (95% CI)"
  , y        = NULL
  , colour   = NULL
  ) +
  ggplot2::theme(legend.position = "bottom")

p_coef_albumin |> print()

ggsave(
    "All_figures_for_paper/composite/p_coef_albumin_composite.png"
  , plot   = p_coef_albumin
  , width  = 7
  , height = 5
  , dpi    = 300
  , bg     = "white"
  )
```

--------------------------------------------------------------------------------

# Step 4: Backward BIC Regression — Ktrans (MRI Permeability)

```{r chunk-016}
predictor_vars_ktrans <-
  setdiff(best_vars_ktrans, var_resp_ktrans)

dat_lm_ktrans <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_ktrans)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_ktrans)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model — complete observations:", nrow(dat_lm_ktrans), "\n")

fit_ktrans_full <-
  lm(
    formula =
      as.formula(
        paste0(
          "Y ~ age + sex + "
        , paste(predictor_vars_ktrans, collapse = " + ")
        )
      )
  , data = dat_lm_ktrans
  )

fit_ktrans_red <-
  stats::step(
    object    = fit_ktrans_full
  , direction = "backward"
  , trace     = 0
  , k         = log(nrow(dat_lm_ktrans))  # BIC
  )

cat("\n=== Reduced Ktrans Model (BIC) ===\n")
summary(fit_ktrans_red)

cat("\nR² =", summary(fit_ktrans_red)$r.squared |> round(3), "\n")
cat("Adj R² =", summary(fit_ktrans_red)$adj.r.squared |> round(3), "\n")
```

## Residual diagnostics (Ktrans model)

```{r chunk-017, fig.height=4, fig.width=10}
par(mfrow = c(1, 3))
plot(fit_ktrans_red, which = c(1, 2, 3))
par(mfrow = c(1, 1))
```

## Coefficient plot (Ktrans model)

```{r chunk-018, fig.height=5, fig.width=7}
p_coef_ktrans <-
  fit_ktrans_red |>
  broom::tidy(conf.int = TRUE) |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::mutate(
    term = term |>
      stringr::str_replace_all(stringr::fixed("sexMale"), "Sex (Male)") |>
      stringr::str_replace_all("^age$", "Age")
  , sig = dplyr::if_else(p.value < 0.05, "p < 0.05", "n.s.")
  ) |>
  ggplot2::ggplot(
    ggplot2::aes(
      x = estimate, y = reorder(term, estimate)
    , colour = sig
    )
  ) +
  ggplot2::theme_bw() +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed", colour = "gray60") +
  ggplot2::geom_point(size = 3) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(xmin = conf.low, xmax = conf.high)
  , height = 0.3
  ) +
  ggplot2::scale_colour_manual(
    values = c("p < 0.05" = "#E31A1C", "n.s." = "gray50")
  ) +
  ggplot2::labs(
    title    = "Reduced Model — Ktrans (MRI WM Permeability)"
  , subtitle = paste0("BIC backward selection  |  R² = ",
                      round(summary(fit_ktrans_red)$r.squared, 3),
                      "  |  n = ", nrow(dat_lm_ktrans))
  , x        = "Standardized Coefficient (95% CI)"
  , y        = NULL
  , colour   = NULL
  ) +
  ggplot2::theme(legend.position = "bottom")

p_coef_ktrans |> print()

ggsave(
    "All_figures_for_paper/composite/p_coef_ktrans_composite.png"
  , plot   = p_coef_ktrans
  , width  = 7
  , height = 5
  , dpi    = 300
  , bg     = "white"
  )
```

--------------------------------------------------------------------------------

# Step 5: Build Composite Inflammatory Score

The composite score is the **fitted value** from the reduced Albumin model,
restricted to the biomarker terms only (i.e., removing the intercept,
age, and sex contributions so the score reflects **inflammation** alone,
not demographics).

This mirrors how the VD and AD composite scores are computed: a
weighted linear combination of the most informative biomarkers,
with weights = regression coefficients.

```{r chunk-019}
# Extract biomarker-only terms from the reduced model
coef_albumin_red <- coef(fit_albumin_red)

# Identify terms that are NOT intercept, age, sex
terms_demo <- c("(Intercept)", "age", "sexMale", "sexFemale")
terms_biomarker <-
  names(coef_albumin_red)[
    names(coef_albumin_red) %notin% terms_demo
  ]

cat("Biomarker terms retained in Albumin composite:\n")
print(terms_biomarker)
cat("\nCorresponding coefficients:\n")
print(coef_albumin_red[ terms_biomarker ])
```

## Compute score for all subjects (with NA propagation)

```{r chunk-020}
# score = sum( beta_j * X_j ) for each retained biomarker j
# Uses z-scored dat_sheet (consistent with model fitting)

dat_sub <-
  dat_sub |>
  dplyr::mutate(
    score_inflammation = NA_real_
  )

# Build a z-scored version of dat_sub for scoring
dat_score_input <-
  dat_sub |>
  dplyr::mutate(
    across(
      tidyselect::all_of(terms_biomarker)
    , ~ as.numeric(scale(.x))
    )
  )

# Compute weighted sum for each row (NA if any biomarker is missing)
score_matrix <-
  dat_score_input |>
  dplyr::select(tidyselect::all_of(terms_biomarker)) |>
  as.matrix()

score_weights <- coef_albumin_red[ terms_biomarker ]

dat_sub$score_inflammation <-
  as.numeric(score_matrix %*% score_weights)

cat("Score computed for", sum(!is.na(dat_sub$score_inflammation)),
    "of", nrow(dat_sub), "subjects\n")

# Label the new variable
labelled::var_label(dat_sub$score_inflammation) <-
  "Composite Inflammatory Score (BBB disruption)"

summary(dat_sub$score_inflammation)
```

## Distribution of the composite score

```{r chunk-021, fig.height=4, fig.width=8}
p_score_dist <-
  dat_sub |>
  dplyr::filter(!is.na(score_inflammation)) |>
  ggplot2::ggplot(
    ggplot2::aes(x = score_inflammation, fill = dx_current, colour = dx_current)
  ) +
  ggplot2::theme_bw() +
  ggplot2::geom_density(alpha = 0.25) +
  ggplot2::geom_rug(alpha = 0.4) +
  ggplot2::labs(
    title    = "Composite Inflammatory Score by Diagnosis"
  , x        = "Composite Inflammatory Score"
  , y        = "Density"
  , fill     = "Diagnosis"
  , colour   = "Diagnosis"
  ) +
  ggplot2::theme(legend.position = "bottom")

p_score_dist |> print()

ggsave(
    "All_figures_for_paper/composite/p_score_distribution.png"
  , plot   = p_score_dist
  , width  = 8
  , height = 4
  , dpi    = 300
  , bg     = "white"
  )
```

## Composite score vs. Albumin Index (validation)

```{r chunk-022, fig.height=4, fig.width=6}
p_score_vs_albumin <-
  e_plot_scatterplot(
    dat_plot            = dat_sub
  , var_x               = "score_inflammation"
  , var_y               = "fl_albumin_index_log2"
  , text_title          = "Composite Inflammatory Score vs. Albumin Index"
  , smooth_all          = c("none", "loess", "lm", "glm", "gam")[3]
  , sw_smooth_all_se    = TRUE
  , sw_corr_in_subtitle = TRUE
  )

p_score_vs_albumin |> print()

ggsave(
    "All_figures_for_paper/composite/p_score_vs_albumin.png"
  , plot   = p_score_vs_albumin
  , width  = 6
  , height = 4
  , dpi    = 300
  , bg     = "white"
  )
```

## Composite score vs. Ktrans (cross-validation)

```{r chunk-023, fig.height=4, fig.width=6}
p_score_vs_ktrans <-
  e_plot_scatterplot(
    dat_plot            = dat_sub
  , var_x               = "score_inflammation"
  , var_y               = "mri_perm_3t_tot_log2"
  , text_title          = "Composite Inflammatory Score vs. Ktrans"
  , smooth_all          = c("none", "loess", "lm", "glm", "gam")[3]
  , sw_smooth_all_se    = TRUE
  , sw_corr_in_subtitle = TRUE
  )

p_score_vs_ktrans |> print()

ggsave(
    "All_figures_for_paper/composite/p_score_vs_ktrans.png"
  , plot   = p_score_vs_ktrans
  , width  = 6
  , height = 4
  , dpi    = 300
  , bg     = "white"
  )
```

--------------------------------------------------------------------------------

# Step 6: Combined Figure — Reduced Model Summary

Side-by-side coefficient plots for both response variables.

```{r chunk-024, fig.height=6, fig.width=14}
p_combined_coef <-
  p_coef_albumin + p_coef_ktrans +
  patchwork::plot_annotation(
    title      = "Backward BIC Regression — Composite Inflammatory Biomarkers"
  , subtitle   = "Left: BBB Albumin Index  |  Right: Ktrans MRI Permeability"
  , tag_levels = "A"
  , theme      = ggplot2::theme(plot.caption = ggplot2::element_text(hjust = 0))
  )

p_combined_coef |> print()

ggsave(
    "All_figures_for_paper/composite/p_combined_coef_both_responses.png"
  , plot   = p_combined_coef
  , width  = 14
  , height = 6
  , dpi    = 300
  , bg     = "white"
  )

ggsave(
    "All_figures_for_paper/composite/p_combined_coef_both_responses.pdf"
  , plot   = p_combined_coef
  , width  = 14
  , height = 6
  , units  = "in"
  , useDingbats = FALSE
  )
```

--------------------------------------------------------------------------------

# Step 7: Model Performance Summary

```{r chunk-025}
tibble::tribble(
  ~Response                      , ~Model            , ~`R²`                                          , ~`Adj R²`                                         , ~n
, "Albumin Index (fl_albumin_index_log2)"
  , "Full (all candidate predictors)"
  , round(summary(fit_albumin_full)$r.squared,    3)
  , round(summary(fit_albumin_full)$adj.r.squared, 3)
  , nrow(dat_lm_albumin)
, "Albumin Index (fl_albumin_index_log2)"
  , "Reduced (BIC backward)"
  , round(summary(fit_albumin_red)$r.squared,     3)
  , round(summary(fit_albumin_red)$adj.r.squared,  3)
  , nrow(dat_lm_albumin)
, "Ktrans (mri_perm_3t_tot_log2)"
  , "Full (all candidate predictors)"
  , round(summary(fit_ktrans_full)$r.squared,     3)
  , round(summary(fit_ktrans_full)$adj.r.squared,  3)
  , nrow(dat_lm_ktrans)
, "Ktrans (mri_perm_3t_tot_log2)"
  , "Reduced (BIC backward)"
  , round(summary(fit_ktrans_red)$r.squared,      3)
  , round(summary(fit_ktrans_red)$adj.r.squared,   3)
  , nrow(dat_lm_ktrans)
) |>
  knitr::kable(caption = "Model R² Summary — Composite Inflammatory Score Analysis")
```

```{r chunk-026}
cat("=== Biomarkers retained in Albumin reduced model ===\n")
broom::tidy(fit_albumin_red, conf.int = TRUE) |>
  dplyr::mutate(
    across(where(is.numeric), ~ round(.x, 3))
  ) |>
  dplyr::filter(term != "(Intercept)") |>
  print(n = Inf)

cat("\n=== Biomarkers retained in Ktrans reduced model ===\n")
broom::tidy(fit_ktrans_red, conf.int = TRUE) |>
  dplyr::mutate(
    across(where(is.numeric), ~ round(.x, 3))
  ) |>
  dplyr::filter(term != "(Intercept)") |>
  print(n = Inf)
```

--------------------------------------------------------------------------------

# Step 8: Save updated dataset with composite score

```{r chunk-027}
save(
    dat_sub
  , Var
  , fit_albumin_red
  , fit_ktrans_red
  , terms_biomarker
  , file = "dat_sub_with_composite_inflammation_score.RData"
  )

cat("Saved: dat_sub_with_composite_inflammation_score.RData\n")
cat("New variable added to dat_sub: score_inflammation\n")
```

--------------------------------------------------------------------------------

# Session Info

```{r chunk-028}
sessionInfo()
```
