---
title:        "Composite Inflammatory Score: Correlation-First Approach"
subtitle:     "Based on Hilmer, et al., (2022). *Blood-brain barrier disruption measured by albumin index correlates with inflammatory fluid biomarkers*"
author:       Azadeh Golduzian
date:         last-modified
date-format:  long
format:
  html:
    theme:                  litera
    highlight-style:        atom-one
    page-layout:            full
    toc:                    true
    toc-depth:              4
    toc-location:           body
    number-sections:        true
    number-depth:           3
    code-fold:              show     
    code-tools:
      toggle:               true
      caption:              "Code"
      source:               false
    code-overflow:          scroll
    code-block-bg:          true
    code-block-border-left: "#30B0E0"
    df-print:               paged
fig-width:                  8
fig-height:                 6
execute:
  cache:    false
  eval:     true
  echo:     true
---

<!---
# Erik's compile commands in R:
  setwd("D:/Dropbox/StatAcumen/consult/Authorship/2014_GaryRosenberg_UNM_NIH_VascularDementia_MRI/2024_ADRC/Analyses/202501_ADRC_Permeability")
  fn_qmd <- "Golduzian_Rosenberg_ADRC_Inflammation_Analysis_4b-Composite_Inflammation_Score_CorrFirst_20260326.qmd"
  quarto::quarto_render(input = fn_qmd)
-->

```{r init, echo=FALSE}
options(str   = strOptions(list.len = 1e3))
options(knitr.kable.NA = '')
my_seed <- 34567
set.seed(my_seed)

# The development version from GitHub
# install.packages("remotes")
remotes::install_github("erikerhardt/erikmisc", force= TRUE)
```

--------------------------------------------------------------------------------

# Approach


**New strategy — 3 panels:**

1. **Panel A:** Rank all biomarkers by |Spearman ρ| with Albumin Index AND Ktrans.
   This shows which markers are biologically relevant to BBB disruption.

2. **Panel B:** For each "top-K" set (take the top 1, top 2, … top K biomarkers
   by correlation), compute how many subjects have complete data on all K.
   This is the N-vs-quality trade-off curve.

3. **Decision:** Pick a cutoff from Panel B that gives n ≥ ~60 with the
   highest-correlated markers, then run backward BIC regression on that set.

--------------------------------------------------------------------------------

# Data Preliminaries
    
```{r chunk-003}
library(erikmisc)
library(tidyverse)
library(labelled)
library(broom)
library(patchwork)

dir.create("All_figures_for_paper",            showWarnings = FALSE)
dir.create("All_figures_for_paper/composite",  showWarnings = FALSE)
```



```{r chunk-004}

load("dat_sub_Data_inflammation_analysis_paper_06052025.RData")
load("dat_sub_and_Var_for_inflammation_analysis.RData")

```



```{r chunk-005}

var_resp_albumin <- "fl_albumin_index_log2"
var_resp_ktrans  <- "mri_perm_3t_tot_log2"

# Exclude response variables from candidate pool
vars_exclude <-
  c( var_resp_albumin
  , var_resp_ktrans
  , "mri_perm_3t_tot"       # raw Ktrans
  , "fl_igg_index_log2"     # excluded per earlier decision
  , "fl_csf_igg_log2"       # excluded per earlier decision
  )

list_var_covar <-
  c(Var$csf, Var$plasma, Var$mri) |>
  unique()

list_var_covar <- list_var_covar[ list_var_covar %notin% vars_exclude ]

cat("Candidate predictors:", length(list_var_covar), "\n")
```


```{r chunk-006}
# Z-score predictors; keep responses on original log2 scale
dat_sheet_demo <- dat_sub

dat_sheet <-
  dat_sub |>
  dplyr::mutate(
    across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), scale)
  , across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), as.numeric)
  )

for (n_label in as.character(unlist(Var))) {
  labelled::var_label(dat_sheet[[ n_label ]]) <-
    labelled::var_label(dat_sheet_demo[[ n_label ]])
}
```

--------------------------------------------------------------------------------

# Panel A — Biomarker Correlation Ranking

```{r}

compute_corr_rank <- function(dat, candidates, resp_var) {
  results <- list()
  for (v in candidates) {
    tmp <- dat |> select(all_of(c(resp_var, v))) |> drop_na()
    if (nrow(tmp) < 5) next
    ct <- cor.test(tmp[[resp_var]], tmp[[v]], method = "spearman", exact = FALSE)
    results[[v]] <- tibble(variable = v, rho = ct$estimate, n_pairwise = nrow(tmp))
  }
  bind_rows(results) |>
  mutate(
    abs_rho  = abs(rho)
  , response = resp_var
  ) |>
  arrange(desc(abs_rho))
}

corr_albumin <- compute_corr_rank(dat_sheet, list_var_covar, var_resp_albumin)
corr_ktrans  <- compute_corr_rank(dat_sheet, list_var_covar, var_resp_ktrans)

```


```{r chunk-008, fig.height=10, fig.width=14}

#keep only biomarkers with individual n >= 60
n_min_individual <- 60
corr_albumin <- corr_albumin |> filter(n_pairwise >= n_min_individual)
corr_ktrans  <- corr_ktrans  |> filter(n_pairwise >= n_min_individual)

print(corr_albumin, n = 20)
print(corr_ktrans,  n = 20)

# Plot
top_n_show <- 20

make_corr_plot <- function(corr_df, title_str) {
  corr_df |>
  slice_head(n = top_n_show) |>
  mutate(
    variable  = variable |> str_replace_all("_log2$", "") |> str_replace_all("_", " ")
  , direction = if_else(rho >= 0, "Positive", "Negative")
  ) |>
  ggplot(aes(x = abs_rho, y = reorder(variable, abs_rho), fill = direction)) +
  theme_bw(base_size = 11) +
  geom_col(alpha = 0.85) +
  geom_text(aes(label = paste0("  n=", n_pairwise)), hjust = -0.1, size = 3, colour = "gray30") +
  scale_x_continuous(limits = c(0, 0.75), expand = expansion(mult = c(0, 0.15))) +
  scale_fill_manual(values = c("Positive" = "#2166AC", "Negative" = "#D6604D")) +
  labs(title = title_str, x = "|Spearman ρ|", y = NULL, fill = "Direction",
       caption = "Stars: * p<0.05  ** p<0.01  *** p<0.001  |  n = pairwise complete obs") +
  theme(legend.position = "bottom", plot.caption = element_text(hjust = 0),
        axis.text.y = element_text(size = 9))
}

p_corr_albumin <- make_corr_plot(corr_albumin, "Spearman ρ with Albumin Index (BBB)")
p_corr_ktrans  <- make_corr_plot(corr_ktrans,  "Spearman ρ with Ktrans (MRI Permeability)")

p_corr_combined <-
  p_corr_albumin + p_corr_ktrans +
  plot_annotation(title = "Panel A — Biomarker Correlation Ranking", tag_levels = "A")

p_corr_combined |> print()

ggsave("All_figures_for_paper/composite/PanelA_corr_ranking.png",
       plot = p_corr_combined, width = 14, height = 10, dpi = 300, bg = "white")
```

-----------------------------------------------------------------
--------------------------------------------------------------------------------

# Decision Point

Look at Panel A  then set `k_albumin` and `k_ktrans` below.


```{r chunk-011}

# Run Erik's complete-by-variable plot on the correlation-ranked candidates
result_albumin <- e_plot_complete_by_variable_subset(
    dat              = dat_sheet
  , var_list         = c(var_resp_albumin, corr_albumin$variable)
  , var_resp         = var_resp_albumin
  , sw_plot_complete = 20
)


#  set k after looking at the table above ----
k_albumin <- 19 # adjust this

predictor_vars_albumin <- result_albumin$var_names[[ k_albumin ]]
predictor_vars_albumin <- predictor_vars_albumin[ predictor_vars_albumin != var_resp_albumin ]
cat("Chosen variables:", paste(chosen_albumin, collapse = ", "), "\n")
cat("Complete-case N:", result_albumin$n_complete[ k_albumin ], "\n")

# For Ktrans

result_ktrans <- e_plot_complete_by_variable_subset(
    dat              = dat_sheet
  , var_list         = c(var_resp_ktrans, corr_ktrans$variable)
  , var_resp         = var_resp_ktrans
  , sw_plot_complete = 20
)


# set k after looking at the table above ----
k_ktrans <- 20   # adjust this

predictor_vars_ktrans <- result_ktrans$var_names[[ k_ktrans ]]
predictor_vars_ktrans <- predictor_vars_ktrans[ predictor_vars_ktrans != var_resp_ktrans ]
cat("Chosen variables:", paste(chosen_ktrans, collapse = ", "), "\n")
cat("Complete-case N:", result_ktrans$n_complete[ k_ktrans ], "\n")

```

--------------------------------------------------------------------------------

# Backward AIC Regression

## Albumin Index model

```{r chunk-013}


dat_lm_albumin <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_albumin)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_albumin)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model — complete observations:", nrow(dat_lm_albumin), "\n")

fit_albumin_full <-
  lm(
    formula = as.formula(
      paste0("Y ~  ", paste(predictor_vars_albumin, collapse = " + "))
    )
  , data = dat_lm_albumin
  )

fit_albumin_red <-
  stats::step(
    object    = fit_albumin_full
  , direction = "backward"
  , trace     = 0
  , k         =2   # AIC
  )


cat("\n=== Reduced Albumin Model (BIC backward) ===\n")
summary(fit_albumin_red)




```

## Ktrans model

```{r chunk-014}

dat_lm_ktrans <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_ktrans)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_ktrans)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model — complete observations:", nrow(dat_lm_ktrans), "\n")

fit_ktrans_full <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", paste(predictor_vars_ktrans, collapse = " + "))
    )
  , data = dat_lm_ktrans
  )

fit_ktrans_red <-
  stats::step(
    object    = fit_ktrans_full
  , direction = "backward"
  , trace     = 0
  , k         = 2    # AIC
  )

cat("\n=== Reduced Ktrans Model (BIC backward) ===\n")
summary(fit_ktrans_red)

```

## Residual diagnostics

```{r chunk-015, fig.height=4, fig.width=12}

e_plot_lm_diagnostics(fit_albumin_red, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_model_contrasts(
   fit = fit_albumin_red,
  dat_cont = dat_lm_albumin,
)


e_plot_lm_diagnostics(fit_ktrans_red, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_model_contrasts(
   fit = fit_ktrans_red,
  dat_cont = dat_lm_ktrans,
)
```



# Build Composite Inflammatory Score

Biomarker-only weighted sum (age and sex terms removed so the score reflects
**inflammation / BBB biology** rather than demographics).

```{r chunk-017}
coef_albumin_red <- coef(fit_albumin_red)

terms_demo      <- c("(Intercept)", "age", "sexMale", "sexFemale")
terms_biomarker <- names(coef_albumin_red)[ names(coef_albumin_red) %notin% terms_demo ]

cat("Biomarker terms in composite:\n")
print(coef_albumin_red[ terms_biomarker ])

# Score on z-scaled predictors (consistent with model fitting) they were not same as the datsheet
# dat_score_input <-
#   dat_sub |>
#   dplyr::mutate(
#     across(tidyselect::all_of(terms_biomarker), ~ as.numeric(scale(.x)))
#   )

score_matrix  <- dat_sheet |> dplyr::select(tidyselect::all_of(terms_biomarker)) |> as.matrix()
score_weights <- coef_albumin_red[ terms_biomarker ]

#score = (MMP-2 × 0.33) + (IL-10 × 0.79) + (MMP-3 × 0.32) + (GFAP × 0.53) + ...
dat_sheet$score_inflammation <- as.numeric(score_matrix %*% score_weights)
labelled::var_label(dat_sheet$score_inflammation) <- "Composite Inflammatory Score (BBB disruption)"

cat("\nScore computed for", sum(!is.na(dat_sheet$score_inflammation)),
    "of", nrow(dat_sheet), "subjects\n")
summary(dat_sheet$score_inflammation)
```


###  composite scores for ktrans 

```{r}
coef_ktrans_red <- coef(fit_ktrans_red)

terms_demo      <- c("(Intercept)", "age", "sexMale", "sexFemale")
terms_biomarker <- names(coef_ktrans_red)[ names(coef_ktrans_red) %notin% terms_demo ]

cat("Biomarker terms in composite:\n")
print(coef_albumin_red[ terms_biomarker ])

# Score on z-scaled predictors (consistent with model fitting) they were not same as the datsheet
# dat_score_input <-
#   dat_sub |>
#   dplyr::mutate(
#     across(tidyselect::all_of(terms_biomarker), ~ as.numeric(scale(.x)))
#   )

score_matrix_ktrans  <- dat_sheet |> dplyr::select(tidyselect::all_of(terms_biomarker)) |> as.matrix()
score_weights_ktrans <- coef_ktrans_red[ terms_biomarker ]

#score = (MMP-2 × 0.33) + (IL-10 × 0.79) + (MMP-3 × 0.32) + (GFAP × 0.53) + ...
dat_sheet$score_inflammation <- as.numeric(score_matrix_ktrans %*% score_weights_ktrans)
labelled::var_label(dat_sheet$score_inflammation) <- "Composite Inflammatory Score (BBB disruption)"

cat("\nScore computed for", sum(!is.na(dat_sheet$score_inflammation)),
    "of", nrow(dat_sheet), "subjects\n")
summary(dat_sheet$score_inflammation)
```


## Validation plots

```{r chunk-018, fig.height=4, fig.width=13}

p_vs_alb <-
  e_plot_scatterplot(
    dat_plot            = dat_sheet
  , var_x               = "score_inflammation"
  , var_y               = "fl_albumin_index_log2"
  , text_title          = "Score vs. Albumin Index"
  , smooth_all          = "lm"
  , sw_smooth_all_se    = TRUE
  , sw_corr_in_subtitle = TRUE
  )

p_vs_ktr <-
  e_plot_scatterplot(
    dat_plot            = dat_sheet
  , var_x               = "score_inflammation"
  , var_y               = "mri_perm_3t_tot_log2"
  , text_title          = "Score vs. Ktrans"
  , smooth_all          = "lm"
  , sw_smooth_all_se    = TRUE
  , sw_corr_in_subtitle = TRUE
  )

p_valid <-
  p_vs_alb + p_vs_ktr +
  patchwork::plot_annotation(title = "Composite Inflammatory Score — Validation")

p_valid |> print()

ggsave(
    "All_figures_for_paper/composite/p_score_validation.png"
  , plot = p_valid, width = 13, height = 4, dpi = 300, bg = "white"
  )
```


# Save

```{r chunk-020}
save(
    dat_sub
  , Var
  , fit_albumin_red
  , fit_ktrans_red
  , terms_biomarker
  , corr_albumin
  , corr_ktrans
  , file = "dat_sub_with_composite_inflammation_score.RData"
  )

cat("Saved: dat_sub_with_composite_inflammation_score.RData\n")
```

# Plasma MMPs 2, 3 and 10 with albumin as response

```{r}

predictor_vars_albumin_mmp <- c("fl_plasma__ALL__mmp_2_log2", 
                                "fl_plasma__ALL__mmp_10_log2", 
                                "fl_plasma__ALL__mmp_3_log2" )

dat_lm_albumin_mmps <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_albumin)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_albumin_mmp)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model — complete observations:", nrow(dat_lm_albumin_mmps), "\n")

fit_albumin_full_mmp <-
  lm(
    formula = as.formula(
      paste0("Y ~  ", paste(predictor_vars_albumin_mmp, collapse = " + "))
    )
  , data = dat_lm_albumin_mmps
  )

fit_albumin_full_mmp_int <-
  lm(
    formula = as.formula(
      paste0("Y ~ (", paste(predictor_vars_albumin_mmp, collapse = " + "), ")^2")
    ),
    data = dat_lm_albumin_mmps
  )

summary(fit_albumin_full_mmp_int)

fit_albumin_red_mmps <-
  stats::step(
    object    = fit_albumin_full_mmp_int
  , direction = "backward"
  , trace     = 0
  , k         =2   # AIC
  )


summary(fit_albumin_full_mmp)

summary(fit_albumin_red_mmps)

# Now individual



```


# Plasma MMPs 2, 3 and 10 with Ktrans as response
```{r}
predictor_vars_ktrans_mmp <- c("fl_plasma__ALL__mmp_2_log2", 
                                "fl_plasma__ALL__mmp_10_log2", 
                                "fl_plasma__ALL__mmp_3_log2" )
dat_lm_ktrans_mmp <-
  dat_sheet |>
  dplyr::select(
    Y   = tidyselect::all_of(var_resp_ktrans)
  , age
  , sex
  , tidyselect::all_of(predictor_vars_ktrans_mmp)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model — complete observations:", nrow(dat_lm_ktrans_mmp), "\n")

fit_ktrans_full_mmp <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", paste(predictor_vars_ktrans_mmp, collapse = " + "))
    )
  , data = dat_lm_ktrans_mmp
  )


fit_ktrans_full_mmp_int <-
  lm(
    formula = as.formula(
      paste0("Y ~ (", paste(predictor_vars_ktrans_mmp, collapse = " + "), ")^2")
    ),
    data = dat_lm_ktrans_mmp
  )

fit_ktrans_red_mmp <-
  stats::step(
    object    = fit_ktrans_full_mmp_int
  , direction = "backward"
  , trace     = 0
  , k         = 2    # AIC
  )
summary(fit_ktrans_full_mmp)


summary(fit_ktrans_red_mmp)

```

# Plasma individual

```{r}
# Individual MMP models for Albumin Index (no age, no sex)

predictor_mmp2  <- "fl_plasma__ALL__mmp_2_log2"
predictor_mmp10 <- "fl_plasma__ALL__mmp_10_log2"
predictor_mmp3  <- "fl_plasma__ALL__mmp_3_log2"


# ----------------------------
# MMP-2
# ----------------------------
dat_lm_albumin_mmp2 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_albumin),
    tidyselect::all_of(predictor_mmp2)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model with MMP-2 — complete observations:", nrow(dat_lm_albumin_mmp2), "\n")

fit_albumin_mmp2 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp2)
    ),
    data = dat_lm_albumin_mmp2
  )

summary(fit_albumin_mmp2)



# ----------------------------
# MMP-10
# ----------------------------
dat_lm_albumin_mmp10 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_albumin),
    tidyselect::all_of(predictor_mmp10)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model with MMP-10 — complete observations:", nrow(dat_lm_albumin_mmp10), "\n")

fit_albumin_mmp10 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp10)
    ),
    data = dat_lm_albumin_mmp10
  )

summary(fit_albumin_mmp10)



# ----------------------------
# MMP-3
# ----------------------------
dat_lm_albumin_mmp3 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_albumin),
    tidyselect::all_of(predictor_mmp3)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Albumin model with MMP-3 — complete observations:", nrow(dat_lm_albumin_mmp3), "\n")

fit_albumin_mmp3 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp3)
    ),
    data = dat_lm_albumin_mmp3
  )

summary(fit_albumin_mmp3)
```

# Ktrans 

```{r}
# Individual MMP models for Ktrans (no age, no sex)

predictor_mmp2  <- "fl_plasma__ALL__mmp_2_log2"
predictor_mmp10 <- "fl_plasma__ALL__mmp_10_log2"
predictor_mmp3  <- "fl_plasma__ALL__mmp_3_log2"


# ----------------------------
# MMP-2
# ----------------------------
dat_lm_ktrans_mmp2 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_ktrans),
    tidyselect::all_of(predictor_mmp2)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model with MMP-2 — complete observations:", nrow(dat_lm_ktrans_mmp2), "\n")

fit_ktrans_mmp2 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp2)
    ),
    data = dat_lm_ktrans_mmp2
  )

summary(fit_ktrans_mmp2)



# ----------------------------
# MMP-10
# ----------------------------
dat_lm_ktrans_mmp10 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_ktrans),
    tidyselect::all_of(predictor_mmp10)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model with MMP-10 — complete observations:", nrow(dat_lm_ktrans_mmp10), "\n")

fit_ktrans_mmp10 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp10)
    ),
    data = dat_lm_ktrans_mmp10
  )

summary(fit_ktrans_mmp10)



# ----------------------------
# MMP-3
# ----------------------------
dat_lm_ktrans_mmp3 <-
  dat_sheet |>
  dplyr::select(
    Y = tidyselect::all_of(var_resp_ktrans),
    tidyselect::all_of(predictor_mmp3)
  ) |>
  tidyr::drop_na() |>
  tibble::as_tibble()

cat("Ktrans model with MMP-3 — complete observations:", nrow(dat_lm_ktrans_mmp3), "\n")

fit_ktrans_mmp3 <-
  lm(
    formula = as.formula(
      paste0("Y ~ ", predictor_mmp3)
    ),
    data = dat_lm_ktrans_mmp3
  )

summary(fit_ktrans_mmp3)
```



# Table
```{r}


dat_sub |>
  tidyr::separate_wider_delim(
    cols        = ptid
  , delim       = "-"
  , names       = c("ptid_Prefix", "ptid_Number")
  , cols_remove = FALSE
  ) |>
  dplyr::mutate(
    Cohort = dplyr::case_when(
      ptid_Prefix == "UP"  ~ "MarkVCID"
    , ptid_Prefix == "VCI" ~ "VCI"
    , ptid_Prefix == "VDP" ~ "VDP"
    , TRUE                 ~ ptid_Prefix
    )
  ) |>
  dplyr::group_by(Cohort, ptid_Prefix) |>
  dplyr::summarise(
    N                  = dplyr::n()
  , `Albumin Index`    = sum(!is.na(fl_albumin_index_log2))
  , `MRI Permeability` = sum(!is.na(mri_perm_3t_tot_log2))
  , .groups = "drop"
  ) |>
  dplyr::rename(`Grant ID` = ptid_Prefix) |>
  knitr::kable(caption = "Biomarker Availability by Cohort and Grant ID")


dat_sub |>
  tidyr::separate_wider_delim(
    cols        = ptid
  , delim       = "-"
  , names       = c("ptid_Prefix", "ptid_Number")
  , cols_remove = FALSE
  ) |>
  dplyr::mutate(
    Cohort = dplyr::case_when(
      ptid_Prefix == "UP"  ~ "MarkVCID"
    , ptid_Prefix == "VCI" ~ "VCI"
    , ptid_Prefix == "VDP" ~ "VDP"
    , TRUE                 ~ ptid_Prefix
    )
  ) |>
  dplyr::group_by(Cohort, ptid_Prefix) |>
  dplyr::summarise(
    N                  = dplyr::n()
  , `Albumin Index`    = sum(!is.na(fl_albumin_index_log2))
  , `MRI Permeability` = sum(!is.na(mri_perm_3t_tot_log2))
  , .groups = "drop"
  ) |>
  dplyr::rename(`Grant ID` = ptid_Prefix) |>
  gt() |>
  tab_header(title = "Biomarker Availability by Cohort and Grant ID") |>
  cols_align(align = "center", columns = c(N, `Albumin Index`, `MRI Permeability`)) |>
  tab_style(
    style     = cell_text(weight = "bold")
  , locations = cells_column_labels()
  )

```

