---
title: "Multivariate_analysis_groups"
author: "Azadeh"
format: html
editor: visual
---

![]()

```{r}
# remotes::install_github("erikerhardt/erikmisc")

library(ggplot2)
library(dplyr)
library(tidyr)
library(car)
library(broom)
library(emmeans)
library(pROC)
library(rlang)
library(patchwork)
library(erikmisc)
```

#load data

```{r}

load("dat_sub_Data_inflammation_analysis_paper_06052025.RData")
load("dat_sub_and_Var_for_inflammation_analysis.RData")
#hist(dat_all_combine$mri_perm_3t_tot |> e_log_shift(min_add = 0.1))
#hist(dat_all_combine$mri_perm_3t_tot_log2)

# update log2 variable using different offset from 0
dat_sub$mri_perm_3t_tot_log2 <- 
  dat_sub$mri_perm_3t_tot |> 
  e_log_shift(min_add = 0.1)

```

# create the dat_sheet with scaling fixed
```{r}
# This code does provide zscale Y or responses
dat_sheet <-
  #dat_sheet %>%
  dat_sub %>%
  dplyr::mutate(
    # to allow log transformation when wmhv==0
  #  percent_wmhv  = percent_wmhv + 0.00001
  #  fl_cyto_csf_il_4 = ifelse(fl_cyto_csf_il_4 == 0.001, NA, fl_cyto_csf_il_4)
  #  across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), e_log_shift, min_add = 1, base_log = 2)
    across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), scale)
  , across(tidyselect::all_of(c(Var$csf, Var$plasma, Var$mri)), as.numeric)
  )
```

# For this code we use the plot that we created for univariate model 
# for Panel A ; response MrI 3T ~ AD Group 
```{r}

# Define variables
CSF_AD <- c("fl_abeta_4240_ratio_log2", "fl_csf__ALL__ptau_log2")
vars_to_analyze <- c("mri_perm_3t_tot", CSF_AD) |> unique()



# Filter complete cases
dat_model <- dat_sheet[, vars_to_analyze] |> na.omit()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
# result_csf_AD <- e_plot_complete_by_variable_subset(
#   dat = dat_model,
#   var_list = vars_to_analyze,
#   var_resp = "mri_perm_3t_tot",
#   sw_plot_complete = 10 
# )

# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(CSF_AD, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_model)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

```

# for Panel A ; response MrI 3T ~ Angio

```{r}

  Angio_csf <-  c("fl_angio_csf_bfgf_log2"               
  , "fl_angio_csf_flt_1_log2"               
  , "fl_angio_csf_tie2_log2"                
  , "fl_angio_csf_vegf_a_log2"              
  , "fl_angio_csf_vegf_c_log2"              
  , "fl_angio_csf_vegf_d_log2" )


vars_to_analyze <- c("mri_perm_3t_tot", Angio_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_AD <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "mri_perm_3t_tot",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(Angio_csf, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)


```

#for Panel A ; response MrI 3T ~ Cytokins

```{r}

  cyto_csf <-  c( 
    "fl_cyto_csf_ifn_g_log2"                
  , "fl_csf__ALL__il_1b_log2"              
  , "fl_cyto_csf_il_2_log2"                
  , "fl_cyto_csf_il_4_log2"        
  , "fl_cyto_csf_il_6_log2"                 
  , "fl_cyto_csf_il_8_log2"     
  , "fl_cyto_csf_il_10_log2"                
  , "fl_cyto_csf_il_12p70_log2"           
  , "fl_cyto_csf_il_13_log2"                
  , "fl_cyto_csf_tnf_a_log2" )


vars_to_analyze <- c("mri_perm_3t_tot", cyto_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_cyto <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "mri_perm_3t_tot",
  sw_plot_complete = 10
)



# Extract the best subset (row 2 with 113 observations and 4 plasma variables)
best_subset_vars <- result_csf_cyto$var_names[[9]]  # Row 5 from results
cat("Using subset with", result_plasma$n_complete[9], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n\n")

# Create dataset with complete cases for this subset
dat_complete <- dat_sheet |> 
  select(all_of(best_subset_vars)) |> 
  filter(complete.cases(pick(everything())))

# ============================================
# STEP 2: Variable importance with Random Forest
# ============================================

# Remove response from predictor list
predictors <- setdiff(best_subset_vars, "mri_perm_3t_tot")

# Run Random Forest
set.seed(123)
rf_model <- randomForest(
  mri_perm_3t_tot ~ .,
  data = dat_complete,
  ntree = 500,
  importance = TRUE
)

# Get variable importance
var_imp <- importance(rf_model)
var_imp_df <- data.frame(
  variable = rownames(var_imp),
  importance = var_imp[, "%IncMSE"]
) |> 
  arrange(desc(importance))

cat("Variable Importance from Random Forest:\n")
print(var_imp_df)

# ============================================
# STEP 3: Build linear model with top variables
# ============================================
top_vars <- var_imp_df$variable[1:5]



# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(top_vars, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_complete)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

```

#for Panel A ; response MrI 3T ~ MMP

```{r}

  MMP_csf <-  c(
    "fl_plasma__ALL__mmp_1_log2",
    "fl_plasma__ALL__mmp_2_log2",
    "fl_plasma__ALL__mmp_3_log2",
    "fl_plasma__ALL__mmp_9_log2",
    "fl_plasma__ALL__mmp_10_log2" 
    
  )


vars_to_analyze <- c("mri_perm_3t_tot", MMP_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_AD <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "mri_perm_3t_tot",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(MMP_csf, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)
```

#for Panel A ; response MrI 3T ~ OTHER

```{r}

other_csf <-  c(
 "fl_csf__ALL__gfap_log2" ,
 "fl_csf_igg_log2" ,
 "fl_igg_index_log2"    ,
 "fl_myelin_basic_protein_log2",
 "fl_csf__ALL__nfl_log2" ,
 "fl_csf__ALL__plgf_log2" 
    
  )


vars_to_analyze <- c("mri_perm_3t_tot", other_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_other <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "mri_perm_3t_tot",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(other_csf, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)


```

#for Panel A ; response MrI 3T ~ MRI

```{r}

MRI_csf <-  c(
  "mri_percent_vol_hippo" ,
  "mri_mean_cort_thick_mm" ,
  "mri_dti_psmd"    ,
  "mri_dti_freewater_mean" 
    
  )


vars_to_analyze <- c("mri_perm_3t_tot", MRI_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_mri <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "mri_perm_3t_tot",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(MRI_csf, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)


```

#for Panel A ; Albumin ~ AD 

```{r}
AD_csf <-  c(
  "fl_abeta_4240_ratio_log2",
  "fl_csf__ALL__ptau_log2"
    
  )


vars_to_analyze <- c("fl_albumin_index_log2", AD_csf) |> unique()

# # Plotting bug in function bug in the function,. It assumes some missingness will always be present, but throws an error if there’s none.
result_csf_mri <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(AD_csf, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)


```



Now this section focuses only on the variables that were significant in the univariate models, based on the results we observed in the plots.

# 3T VS all CSF Significant

```{r}

CSF_Vars_Sign <-  c( 
  
    "fl_csf__ALL__il_1b_log2"         
  , "mri_dti_freewater_mean" 
  ,  "mri_dti_psmd"
  )


vars_to_analyze <- c("mri_perm_3t_tot_log2", CSF_Vars_Sign) |> unique()



# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot_log2 ~ (", paste(CSF_Vars_Sign, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])


```


 3T VS all plasma Significant panel B, now only VEGF A only matter 
do I need to check that as well? 
```{r}

plasma_Vars_Sign <-  c( 
           

  )


vars_to_analyze <- c("mri_perm_3t_tot", plasma_Vars_Sign) |> unique()



# Modeling
formula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(plasma_Vars_Sign, collapse = " + "), ")^2"
))

# Modeling
mainformula_full <- as.formula(paste0(
  "mri_perm_3t_tot ~ (", paste(plasma_Vars_Sign, collapse = " + "), ")"
))
fit_main <- lm(mainformula_full, data = dat_sheet)


fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_main)
summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
```



# Albumin VS All CSF panel A

```{r}

CSF_Vars_Sign <-  c( 
           
   "fl_cyto_csf_il_8_log2" ,
   "fl_cyto_csf_il_6_log2",
   "fl_msd_csf_mmp_2_log2" ,
   "fl_msd_csf_mmp_3_log2" ,
   "fl_csf_igg_log2", 
   "fl_igg_index_log2" ,
   "fl_csf__ALL__plgf_log2",
   "mri_percent_vol_hippo"  
  , "fl_cyto_csf_il_10_log2"                

  )


vars_to_analyze <- c("fl_albumin_index_log2", CSF_Vars_Sign) |> unique()

result_csf <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2"
  # sw_plot_complete = 30
)
dat_model <- dat_sheet |> tidyr::drop_na(tidyselect::all_of(vars_to_analyze))

# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(CSF_Vars_Sign, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_model)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[3])


# WITHOUT IGG

CSF_Vars_Sign <-  c( 
           
   "fl_cyto_csf_il_8_log2" ,
   "fl_cyto_csf_il_6_log2",
   "fl_msd_csf_mmp_2_log2" ,
   "fl_msd_csf_mmp_3_log2" ,
   # "fl_csf_igg_log2", 
   # "fl_igg_index_log2" ,
   "fl_csf__ALL__plgf_log2",
   "mri_percent_vol_hippo"  
  , "fl_cyto_csf_il_10_log2"                

  )


vars_to_analyze <- c("fl_albumin_index_log2", CSF_Vars_Sign) |> unique()

result_csf <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2"
  # sw_plot_complete = 30
)
dat_model <- dat_sheet |> tidyr::drop_na(tidyselect::all_of(vars_to_analyze))

# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(CSF_Vars_Sign, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_model)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[3])

```

# Albumin VS All plasma panel B

```{r}

plasma_Vars_Sign <-  c( 
           
 "fl_cyto_heparin_plasma_il_10_log2" ,
 "fl_serum_igg_log2"

  )


vars_to_analyze <- c("fl_albumin_index_log2", plasma_Vars_Sign) |> unique()

result_csf <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2",
  sw_plot_complete = 10
)

# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(plasma_Vars_Sign, collapse = " + "), ")^2"
))

fit_full <- lm(formula_full, data = dat_sheet)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])


```


#AIC Check only sig one

#1. Start with albumin as response 

1.1 CSF AND MRI AS PREDICTORS 

```{r}
csf_mri_vars <-  c( 
           
   "fl_msd_csf_mmp_2_log2" ,
   "fl_msd_csf_mmp_3_log2" ,
   "fl_csf__ALL__plgf_log2",
   "mri_dti_psmd",
   "fl_cyto_csf_il_10_log2",
   "fl_angio_csf_tie2_log2" 
  )


vars_to_analyze <- c("fl_albumin_index_log2", csf_mri_vars) |> unique()

result_csf <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2"
  # sw_plot_complete = 30
)
dat_model <- dat_sheet |> tidyr::drop_na(tidyselect::all_of(vars_to_analyze))

#RF 
dat_model <- dat_sheet %>%
  select(all_of(c("fl_albumin_index_log2", csf_mri_vars))) %>%
  tidyr::drop_na()
library(randomForest)
?randomForest
# 3. Random Forest
set.seed(123)
rf_fit <- randomForest(
  fl_albumin_index_log2 ~ .,
  data = dat_model,
  ntree = 1000,
  importance = TRUE
)

# 4. Results
print(rf_fit)
varImpPlot(rf_fit)
importance(rf_fit)

#result: 
#Call:
#  randomForest(formula = fl_albumin_index_log2 ~ ., data = dat_model,      ntree = 1000, importance = TRUE) 
#                Type of random forest: regression
#                      Number of trees: 1000
# No. of variables tried at each split: 2


# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(csf_mri_vars, collapse = " + "), ")"
))

fit_full <- lm(formula_full, data = dat_model)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[3])


```

1.2 All Plasma now 
```{r}
plasma_vars <-  c( 
  "fl_plasma__ALL__vegf_d_log2",
  "fl_plasma__ALL__il_1b_log2",
  "fl_cyto_heparin_plasma_il_10_log2",
  "fl_serum_igg_log2" ,
  "fl_plasma__ALL__abeta_42_log2"

  )


vars_to_analyze <- c("fl_albumin_index_log2", plasma_vars) |> unique()

result_plasma <- e_plot_complete_by_variable_subset(
  dat = dat_sheet,
  var_list = vars_to_analyze,
  var_resp = "fl_albumin_index_log2"
  # sw_plot_complete = 30
)
dat_model <- dat_sheet |> tidyr::drop_na(tidyselect::all_of(vars_to_analyze))

#RF 
dat_model <- dat_sheet %>%
  select(all_of(c("fl_albumin_index_log2", plasma_vars))) %>%
  tidyr::drop_na()

library(randomForest)
# 3. Random Forest
set.seed(123)
rf_fit <- randomForest(
  fl_albumin_index_log2 ~ .,
  data = dat_model,
  ntree = 1000,
  importance = TRUE
)

# 4. Results
print(rf_fit)
varImpPlot(rf_fit)
importance(rf_fit)

# with imputation 


# 1. Check for problematic rows
dat_for_impute <- dat_sheet %>%
  select(all_of(c("fl_albumin_index_log2", plasma_vars)))

# Remove rows where ALL values are missing
dat_for_impute <- dat_for_impute %>%
  filter(rowSums(!is.na(.)) > 0)

# Check column types
sapply(dat_for_impute, class)

# Force numeric
dat_for_impute <- dat_for_impute %>%
  mutate(across(everything(), ~as.numeric(as.character(.))))

# 2. Now try imputation
set.seed(123)
imputed_result <- missForest(
  xmis = as.data.frame(dat_for_impute),  # Convert to data.frame
  ntree = 100
)

dat_model_imputed <- imputed_result$ximp

# 3. Random Forest
set.seed(123)
rf_fit <- randomForest(
  fl_albumin_index_log2 ~ .,
  data = dat_model_imputed,
  ntree = 1000,
  importance = TRUE
)

print(rf_fit)

importance(rf_fit)
varImpPlot(rf_fit)

# Get actual values
imp_df <- importance(rf_fit) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("variable") %>%
  arrange(desc(`%IncMSE`))

print(imp_df)
#result: 
#Call:
#  randomForest(formula = fl_albumin_index_log2 ~ ., data = dat_model,      ntree = 1000, importance = TRUE) 
#                Type of random forest: regression
#                      Number of trees: 1000
# No. of variables tried at each split: 2


# Modeling
formula_full <- as.formula(paste0(
  "fl_albumin_index_log2 ~ (", paste(csf_mri_vars, collapse = " + "), ")"
))

fit_full <- lm(formula_full, data = dat_model)

fit_selected <- step(
  fit_full,
  direction = "both",
  trace = 0,
  k = log(nrow(dat_model))
)

summary(fit_full)
summary(fit_selected)

e_plot_lm_diagnostics(fit_full, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_lm_diagnostics(fit_selected, sw_plot_set = c(NA, "simple", "simpleAV", "all")[3])

```


#Now base on AIC do the group analysis


#Shared helper 

```{r}

# ── Global panel size — same for every figure in the paper ───────────────────
PANEL_IN <- 4.5   # every scatter plot = 4.5 × 4.5 inches, always

save_contrast_grid <- function(p_list, filename, remove = integer(0), ncol = 2) {
  
  # Expand interaction terms
  p_expanded <- list()
  for (i_term in seq_along(p_list)) {
    n_list <- names(p_list)[i_term]
    if (stringr::str_detect(n_list, stringr::fixed(":"))) {
      p_expanded <- c(p_expanded, list(p_list[[i_term]][[1]], p_list[[i_term]][[2]]))
    } else {
      p_expanded <- c(p_expanded, list(p_list[[i_term]]))
    }
  }

  # Remove unwanted panels
  if (length(remove) > 0) p_expanded <- p_expanded[-remove]

  # Add top margin to every panel
  p_expanded <- lapply(p_expanded, function(p) {
    p + ggplot2::theme(plot.margin = ggplot2::margin(t = 22, r = 5, b = 5, l = 5, unit = "pt"))
  })

  # Extract shared legend if any panel has one, then strip from ALL panels
  shared_legend <- NULL
  for (p in p_expanded) {
    leg <- cowplot::get_legend(p)
    if (!is.null(leg)) { shared_legend <- leg; break }
  }
  p_expanded <- lapply(p_expanded, function(p) {
    p + ggplot2::theme(legend.position = "none")
  })

  # ── Size derived from GLOBAL constant — same across every analysis ──────────
  n_panels <- length(p_expanded)
  nrow_val <- ceiling(n_panels / ncol)
  w        <- ncol   * PANEL_IN   # e.g. 2 × 4.5 = 9 in
  h        <- nrow_val * PANEL_IN # e.g. 3 × 4.5 = 13.5 in
  # ─────────────────────────────────────────────────────────────────────────────

  p_grid <- cowplot::plot_grid(
    plotlist       = p_expanded,
    labels         = "AUTO",
    label_size     = 14,
    label_fontface = "bold",
    label_x        = 0.02,
    label_y        = 0.995,
    nrow           = nrow_val,
    ncol           = ncol,
    align          = "hv",
    axis           = "tblr"
  )

  # Attach legend to the right if it exists
  if (!is.null(shared_legend)) {
    p_grid <- cowplot::plot_grid(
      p_grid, shared_legend,
      ncol       = 2,
      rel_widths = c(1, 0.12)
    )
    w <- w + 1
  }

  ggsave(
    file.path("All_figures_for_paper", filename),
    plot   = p_grid,
    width  = w,
    height = h,
    dpi    = 300,
    bg     = "white"
  )
  message("Saved: ", filename, " | ", n_panels, " panels | ", w, "x", h, " in")
  invisible(p_grid)
}
```


# Albumin ~ CSF
```{r}

# --- ANALYSIS 1: Albumin Index ~ Significant CSF variables ---
sig_csf_alb <- c(
  
  "fl_abeta_4240_ratio_log2",        # Aβ42/40 Ratio:
  "fl_csf__ALL__ptau_log2",          # p-Tau 181:
  "fl_angio_csf_bfgf_log2" ,
  "fl_angio_csf_flt_1_log2",
  "fl_angio_csf_tie2_log2",          # Tie-2: * (X + age:X)
  # "fl_angio_csf_vegf_d_log2",        # VEGF-D: *
  "fl_cyto_csf_ifn_g_log2",
  "fl_csf__ALL__il_1b_log2",
  "fl_cyto_csf_il_6_log2",
  "fl_cyto_csf_il_8_log2",           # IL-8: * (X + Age:X)
  "fl_cyto_csf_il_10_log2",          # IL-10: * (X + Age:X)
  # "fl_cyto_csf_il_12p70_log2",       # IL-12p70: * (X + ***)
  # "fl_cyto_csf_tnf_a_log2",          # TNF-α: *
  "fl_msd_csf_mmp_1_log2",           # MMP-1: **
  "fl_msd_csf_mmp_2_log2",           # MMP-2: ***
  "fl_msd_csf_mmp_3_log2",           # MMP-3: *
  "fl_csf__ALL__mmp_10_log2",        # MMP-10: *
  "fl_csf__ALL__plgf_log2",          # PlGF: *
  "fl_csf__ALL__gfap_log2",
  # "fl_csf_igg_log2",                 # IgG: ***
  # "fl_igg_index_log2",               # IgG Index: ***
  # "fl_myelin_basic_protein_log2",    # Myelin Basic Protein: **
  "fl_csf__ALL__nfl_log2"
  

  )

# DO it later 

  # "mri_dti_psmd" ,                    # DTI PSMD: * (X + Age:X)
  # "mri_dti_freewater_mean" ,
  # "mri_percent_vol_hippo",
  # "mri_mean_cort_thick_mm"

# Analysis 1: Albumin ~ CSF
vars_1 <- c("fl_albumin_index_log2", sig_csf_alb) |> unique()
result_csf_alb <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_1,
  var_resp = "fl_albumin_index_log2"
)


plot_capture_csf <- grid::grid.grabExpr({
  result_csf <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_1,
    var_resp = "fl_albumin_index_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  
plot_capture_gg_csf <- cowplot::ggdraw(plot_capture_csf)  # <- ggplot now


ggsave(
  "All_figures_for_paper/result_csf_alb.png",
  plot   = plot_capture_gg_csf,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)




best_subset_vars <- result_csf_alb$var_names[[17]]  # Row 2 from your results
cat("Using subset with", result_csf_alb$n_complete[17], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")


# Extract response and predictors from best_subset_vars
var_resp_1 <- "fl_albumin_index_log2"

# # Create dataset with complete cases for this subset
dat_complete <- dat_sheet |>
  select(all_of(best_subset_vars)) |>
  filter(complete.cases(pick(everything())))



# Extract response and predictors from best_subset_vars
var_resp_1 <- "fl_albumin_index_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_1, "age", "sex")]

cat("Response:", var_resp_2, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)


# Make sure age and sex are in dat_complete
dat_lm_1 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_1),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()


cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_1), "\n")

# Full model: age + sex + each X + age:X + sex:X
form_main <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(best_subset_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))


form_upper <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(best_subset_vars, collapse = " + "), "+",
  paste0("age:", best_subset_vars, collapse = " + "), "+",
  paste0("sex:", best_subset_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_1, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_1),
  scope = list(lower = form_lower, upper = form_main),
  direction = "both",
  k = 2, #log(nrow(dat_lm_1)), #2,
  # k = log(nrow(dat_lm_1)),
  trace = 0
)



summary(fit_step)
# car::Anova(fit_step, type = 3)
# broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_1,
)




# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_1,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")


plot_albumin_csf <- save_contrast_grid(fit_contrasts$plots, "plot_albumin_csf.png")





```

#  Ktrans ~ CSF
```{r}
# --- ANALYSIS 2: 3T Permeability ~ Significant CSF variables ---
sig_csf_perm <- c(
  
  "fl_angio_csf_vegf_c_log2" ,
  "fl_cyto_csf_ifn_g_log2",
  "fl_cyto_csf_il_10_log2" ,
  "fl_csf__ALL__gfap_log2",   
  "fl_csf_igg_log2" ,
  "fl_csf__ALL__il_1b_log2",         # IL-1β: *
  "fl_msd_csf_mmp_2_log2",          # MMP-2: *
  "fl_csf__ALL__nfl_log2"        # NFL: * (Gender:X + Age:X)
  # "mri_dti_psmd",                   # DTI PSMD: * (X + Age:X + **)
  # "mri_dti_freewater_mean"          # Mean Free Water: * (Age:X + **)
)

# Analysis 2: perm ~ CSF
vars_2 <- c("mri_perm_3t_tot_log2", sig_csf_perm) |> unique()
result_csf_perm <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_2,
  var_resp = "mri_perm_3t_tot_log2"
)


plot_capture_csf <- grid::grid.grabExpr({
  result_csf <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_2,
    var_resp = "mri_perm_3t_tot_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  
plot_capture_ktrans_csf<- cowplot::ggdraw(plot_capture_csf)  # <- ggplot now


ggsave(
  "All_figures_for_paper/plot_capture_ktrans_csf.png",
  plot   = plot_capture_ktrans_csf,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)


best_subset_vars <- result_csf_perm$var_names[[7]]  # Row 2 from your results
cat("Using subset with", result_csf_perm$n_complete[7], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n\n")


# # Create dataset with complete cases for this subset
dat_complete <- dat_sheet |>
  select(all_of(best_subset_vars)) |>
  filter(complete.cases(pick(everything())))



# Extract response and predictors from best_subset_vars
var_resp_2 <- "mri_perm_3t_tot_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_2, "age", "sex")]

cat("Response:", var_resp_2, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

# Make sure age and sex are in dat_complete
dat_lm_2 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_2),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()

cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_2), "\n")

form_main <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(pred_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))


# Full model: age + sex + each X + age:X + sex:X
form_upper <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_2, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_2),
  scope = list(lower = form_lower, upper = form_upper),
  direction = "both",
  k = 2,
  trace = 0
)

summary(fit_step)
# car::Anova(fit_step, type = 3)
# broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_2,
)









# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_2,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")

plot_ktrans_csf <- save_contrast_grid(
  p_list   = fit_contrasts$plots,
  filename = "plot_ktrans_csf.png",
  remove   = c(1, 5)        # same positions_to_remove you had
)
```


# Albumin ~ Plsama

```{r}
# --- ANALYSIS 3: Albumin Index ~ Significant Plasma variables ---
sig_plasma_alb <- c(
  # "fl_plasma__ALL__ptau_181_log2",       # p-Tau 181: *
  "fl_plasma__ALL__abeta_40_log2",       # Aβ40: * (X + colored Age:X, Gender:X)
  "fl_plasma__ALL__abeta_42_log2",       # Aβ42: * (X + *)
  # "fl_plasma__ALL__bfgf_log2",           # bFGF: *
  # "fl_plasma__ALL__flt_1_log2",          # Flt-1: *
  # "fl_plasma__ALL__tie2_log2",           # Tie-2: *
  # "fl_plasma__ALL__vegf_a_log2",         # VEGF-A: *
  # "fl_plasma__ALL__vegf_c_log2",         # VEGF-C: *
  "fl_plasma__ALL__vegf_d_log2",         # VEGF-D: * (X + *)
  # "fl_cyto_heparin_plasma_ifn_g_log2",   # IFN-γ: *
  "fl_plasma__ALL__il_1b_log2",          # IL-1β: * (X + Gender:X)
  # "fl_cyto_heparin_plasma_il_2_log2",    # IL-2: *
  # "fl_cyto_heparin_plasma_il_6_log2",    # IL-6: **
  # "fl_cyto_heparin_plasma_il_8_log2",    # IL-8: *
  "fl_cyto_heparin_plasma_il_10_log2",   # IL-10: *
  # "fl_cyto_heparin_plasma_il_12p70_log2",# IL-12p70: *
  # "fl_cyto_heparin_plasma_il_13_log2",   # IL-13: *
  # "fl_plasma__ALL__mmp_1_log2",          # MMP-1: *
  "fl_plasma__ALL__mmp_2_log2",          # MMP-2: * (X + Age:X)
  "fl_plasma__ALL__mmp_3_log2",          # MMP-3: **
  # "fl_plasma__ALL__mmp_9_log2",          # MMP-9: *
  "fl_plasma__ALL__mmp_10_log2",         # MMP-10: * (X + *)
  # "fl_plasma__ALL__plgf_log2",           # PlGF: *
  # "fl_plasma__ALL__nfl_log2",            # NFL: *
  "fl_serum_igg_log2"                    # Serum IgG: *** (X + **)
)




# Analysis 1: Albumin ~ CSF
vars_1 <- c("fl_albumin_index_log2", sig_plasma_alb) |> unique()
result_plasma_alb <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_1,
  var_resp = "fl_albumin_index_log2"
)

plot_capture_plasma <- grid::grid.grabExpr({
  result_csf <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_1,
    var_resp = "fl_albumin_index_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  
plot_capture_gg_plasma <- cowplot::ggdraw(plot_capture_plasma)  # <- ggplot now


ggsave(
  "All_figures_for_paper/plot_capture_plasma.png",
  plot   = plot_capture_gg_plasma,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)



best_subset_vars <- result_plasma_alb$var_names[[15]]  # Row 2 from your results
cat("Using subset with", result_plasma_alb$n_complete[15], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n\n")


# Create dataset with complete cases for this subset
dat_complete <- dat_sheet |> 
  select(all_of(best_subset_vars)) |> 
  filter(complete.cases(pick(everything())))




# Extract response and predictors from best_subset_vars
var_resp_1 <- "fl_albumin_index_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_1, "age", "sex")]

cat("Response:", var_resp_1, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

# Make sure age and sex are in dat_complete
dat_lm_1 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_1),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()

cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_1), "\n")


form_main <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(pred_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))

# Full model: age + sex + each X + age:X + sex:X
form_upper <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_1, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_1),
  scope = list(lower = form_lower, upper = form_main),
  direction = "both",
  k = 2,
  # k= log(nrow(dat_lm_1))
  trace = 0
)

summary(fit_step)
car::Anova(fit_step, type = 3)
broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_1,
)



# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_1,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")


plot_albumin_plasma <- save_contrast_grid(
  p_list   = fit_contrasts$plots,
  filename = "plot_Albumin_plasma.png",
  remove   = c(1)
)



```


# Ktrans~ plasma

```{r}

# --- ANALYSIS 4: 3T Permeability ~ Significant Plasma variables ---
sig_plasma_perm <- c(
  "fl_plasma__ALL__ptau_181_log2",   # p-Tau 181: *
  "fl_plasma__ALL__abeta_40_log2",   # Aβ40: * (X + colored tiles)
  "fl_plasma__ALL__vegf_a_log2",     # VEGF-A: * (Gender:X)
  "fl_plasma__ALL__vegf_c_log2",     # VEGF-C: large positive (Age:X?)
  "fl_cyto_heparin_plasma_ifn_g_log2",
   "fl_cyto_heparin_plasma_il_13_log2" ,
  "fl_plasma__ALL__mmp_9_log2",      # MMP-9: * (X + Gender:X)
  "fl_plasma__ALL__plgf_log2"        # PlGF: * (X + red/blue tiles)
)


# Analysis4 : perm ~ plasma
vars_2 <- c("mri_perm_3t_tot_log2", sig_plasma_perm) |> unique()
result_plasma_perm <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_2,
  var_resp = "mri_perm_3t_tot_log2"
)


plot_capture_plasma <- grid::grid.grabExpr({
  result_csf <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_2,
    var_resp = "mri_perm_3t_tot_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  
plot_capture_ktrans_plasma<- cowplot::ggdraw(plot_capture_plasma)  # <- ggplot now


ggsave(
  "All_figures_for_paper/plot_capture_ktrans_plasma.png",
  plot   = plot_capture_ktrans_plasma,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)



best_subset_vars <- result_plasma_perm$var_names[[6]]  # Row 2 from your results
cat("Using subset with", result_plasma_perm$n_complete[6], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n\n")


# Create dataset with complete cases for this subset
dat_complete <- dat_sheet |> 
  select(all_of(best_subset_vars)) |> 
  filter(complete.cases(pick(everything())))




# Extract response and predictors from best_subset_vars
var_resp_2 <- "mri_perm_3t_tot_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_2, "age", "sex")]

cat("Response:", var_resp_2, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

# Make sure age and sex are in dat_complete
dat_lm_2 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_2),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()

cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_2), "\n")


form_main <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(pred_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))

# Full model: age + sex + each X + age:X + sex:X
form_upper <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_2, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_2),
  scope = list(lower = form_lower, upper = form_main),
  direction = "both",
  k = 2,
  trace = 0
)

summary(fit_step)
# car::Anova(fit_step, type = 3)
# broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])


e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_2,
)

# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_2,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")


plot_ktrans_plasma <- save_contrast_grid(
  p_list   = fit_contrasts$plots,
  filename = "plot_ktrans_plasma.png"
  # no removals
)


```

# mri realtionship to albumin 

```{r}
# --- ANALYSIS 1: Albumin Index ~ Significant CSF variables ---
sig_csf_alb <- c(
  
  "mri_dti_psmd" ,                    # DTI PSMD: * (X + Age:X)
  "mri_dti_freewater_mean" ,
  "mri_percent_vol_hippo",
  "mri_mean_cort_thick_mm"

  )


# Analysis 1: Albumin ~ CSF
vars_1 <- c("fl_albumin_index_log2", sig_csf_alb) |> unique()
result_csf_alb <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_1,
  var_resp = "fl_albumin_index_log2"
)


plot_capture_mri <- grid::grid.grabExpr({
  result_csf <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_1,
    var_resp = "fl_albumin_index_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  
plot_capture_gg_mri<- cowplot::ggdraw(plot_capture_mri)  # <- ggplot now


ggsave(
  "All_figures_for_paper/plot_capture_mri.png",
  plot   = plot_capture_gg_mri,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)


best_subset_vars <- result_csf_alb$var_names[[5]]  # Row 2 from your results
cat("Using subset with", result_csf_alb$n_complete[5], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")


# Extract response and predictors from best_subset_vars
var_resp_1 <- "fl_albumin_index_log2"

# # Create dataset with complete cases for this subset
dat_complete <- dat_sheet |>
  select(all_of(best_subset_vars)) |>
  filter(complete.cases(pick(everything())))



# Extract response and predictors from best_subset_vars
var_resp_1 <- "fl_albumin_index_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_1, "age", "sex")]

cat("Response:", var_resp_2, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)


# Make sure age and sex are in dat_complete
dat_lm_1 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_1),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()


cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_1), "\n")

# Full model: age + sex + each X + age:X + sex:X
form_main <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(best_subset_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))


form_upper <- as.formula(paste(
  var_resp_1, "~ age + sex +",
  paste(best_subset_vars, collapse = " + "), "+",
  paste0("age:", best_subset_vars, collapse = " + "), "+",
  paste0("sex:", best_subset_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_1, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_1),
  scope = list(lower = form_lower, upper = form_upper),
  direction = "both",
  k = 2, #log(nrow(dat_lm_1)), #2,
  # k = log(nrow(dat_lm_1)),
  trace = 0
)



summary(fit_step)
# car::Anova(fit_step, type = 3)
# broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_1,
)





# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_1,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")

plot_albumin_mri <- save_contrast_grid(
  p_list   = fit_contrasts$plots,
  filename = "plot_Albumin_mri.png",
  remove   = c(1, 4, 6)
)
```

#mri to permablity

```{r}
# --- ANALYSIS 1: Albumin Index ~ Significant CSF variables ---
sig_csf_alb <- c(
  
  "mri_dti_psmd" ,                    # DTI PSMD: * (X + Age:X)
  "mri_dti_freewater_mean" ,
  "mri_percent_vol_hippo",
  "mri_mean_cort_thick_mm"

  )


# Analysis 1: Albumin ~ CSF
vars_2 <- c("mri_perm_3t_tot_log2", sig_csf_alb) |> unique()
result_ktrans_mri <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_2,
  var_resp = "mri_perm_3t_tot_log2"
)

plot_capture_mri <- grid::grid.grabExpr({
  result_mru <- e_plot_complete_by_variable_subset(
    dat = dat_sheet,
    var_list = vars_2,
    var_resp = "mri_perm_3t_tot_log2",
    sw_plot_complete = TRUE  # Or TRUE for all
  )
})
  

plot_capture_ktrans_mri<- cowplot::ggdraw(plot_capture_mri)  # <- ggplot now


ggsave(
  "All_figures_for_paper/plot_capture_ktrans_mri.png",
  plot   = plot_capture_ktrans_mri,
  width  = 10,
  height = 10,
  dpi    = 300,
  bg     = "white"
)



best_subset_vars <- result_csf_alb$var_names[[3]]  # Row 2 from your results
cat("Using subset with", result_csf_alb$n_complete[3], "complete observations\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")


# Extract response and predictors from best_subset_vars
var_resp_2 <- "mri_perm_3t_tot_log2"

# # Create dataset with complete cases for this subset
dat_complete <- dat_sheet |>
  select(all_of(best_subset_vars)) |>
  filter(complete.cases(pick(everything())))



# Extract response and predictors from best_subset_vars
var_resp_2 <- "mri_perm_3t_tot_log2"

# Predictors = everything in best_subset_vars EXCEPT response, age, sex
# (age and sex should already be in dat_sheet but check if in best_subset_vars)
pred_vars <- best_subset_vars[!best_subset_vars %in% 
               c(var_resp_2, "age", "sex")]

cat("Response:", var_resp_2, "\n")
cat("N complete:", nrow(dat_complete), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)


# Make sure age and sex are in dat_complete
dat_lm_2 <- dat_sheet |>
  dplyr::select(
    all_of(var_resp_2),
    all_of(pred_vars),
    age, sex
  ) |>
  tidyr::drop_na()


cat("\nN after adding age/sex and dropping NA:", nrow(dat_lm_2), "\n")

# Full model: age + sex + each X + age:X + sex:X
form_main <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(best_subset_vars, collapse = " + ") #, "+",
  #paste0("age:", pred_vars, collapse = " + "), "+",
  #paste0("sex:", pred_vars, collapse = " + ")
))


form_upper <- as.formula(paste(
  var_resp_2, "~ age + sex +",
  paste(best_subset_vars, collapse = " + "), "+",
  paste0("age:", best_subset_vars, collapse = " + "), "+",
  paste0("sex:", best_subset_vars, collapse = " + ")
))

form_lower <- as.formula(paste(var_resp_2, "~ 1"))

# Stepwise AIC
fit_step <- step(
  lm(form_main, data = dat_lm_2),
  scope = list(lower = form_lower, upper = form_upper),
  direction = "both",
  k = 2, #log(nrow(dat_lm_1)), #2,
  # k = log(nrow(dat_lm_1)),
  trace = 0
)



summary(fit_step)
# car::Anova(fit_step, type = 3)
# broom::tidy(fit_step) |> dplyr::arrange(p.value) |> print(n = Inf)



e_plot_lm_diagnostics(fit_step, sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])
e_plot_model_contrasts(
   fit = fit_step,
  dat_cont = dat_lm_2,
)




# Apply Erik's contrast function to get all statistical outputs
fit_contrasts <- e_plot_model_contrasts(
  fit = fit_step,
  dat_cont = dat_lm_2,
  sw_print = FALSE,
  sw_table_in_plot = FALSE,
  adjust_method = "bonferroni",  # Adjust for multiple comparisons
  CI_level = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep = "singles",
  sw_produce_plots = TRUE
)

# Remove the specific unwanted plot by name
# fit_contrasts$plots[["sex"]] <- NULL
library(cowplot)

# ── 1. Recursive extractor: pulls every ggplot out of a nested list ──────────
extract_all_gg <- function(x) {
  if (inherits(x, c("gg", "ggplot"))) return(list(x))
  if (is.list(x)) return(unlist(lapply(x, extract_all_gg), recursive = FALSE))
  NULL
}

all_contrast_plots <- extract_all_gg(fit_contrasts$plots)
n_plots <- length(all_contrast_plots)
cat("Found", n_plots, "ggplot(s) in fit_contrasts$plots\n")


plot_ktrans_mri <- save_contrast_grid(
  p_list   = fit_contrasts$plots,
  filename = "plot_ktrans_mri.png"
  # no removals
)
```

# GFAP as response :GFAP CSF VS CSF

```{r}
# ── CSF GFAP ~ CSF + Plasma + MRI ───────────────────────────────────────────

sig_csf_gfap <- c(
  # CSF AD (colored)
  "fl_abeta_4240_ratio_log2",
  "fl_csf__ALL__ptau_log2",
  # CSF Angio (colored)
  "fl_angio_csf_tie2_log2",
  "fl_angio_csf_vegf_a_log2",
  "fl_angio_csf_flt_1_log2",
  # CSF Cytokines (colored)
  "fl_cyto_csf_ifn_g_log2",
  "fl_csf__ALL__il_1b_log2",
  "fl_cyto_csf_il_4_log2",
  "fl_cyto_csf_il_8_log2",
  "fl_cyto_csf_il_10_log2" ,
  "fl_cyto_csf_il_12p70_log2",
  # CSF MMP (colored)
   "fl_msd_csf_mmp_3_log2" ,
  "fl_msd_csf_mmp_9_log2",
  "fl_csf__ALL__mmp_10_log2",
  
  # CSF Other (colored)
  "fl_myelin_basic_protein_log2" ,
  "fl_csf__ALL__nfl_log2",
  "fl_albumin_index_log2"

)

# MRi 
  # MRI (colored)
  # "mri_mean_cort_thick_mm",
  # "mri_dti_freewater_mean",
  #  "mri_dti_psmd"   ,
  #  "mri_percent_vol_hippo" 

vars_gfap_csf <- c("fl_csf__ALL__gfap_log2", sig_csf_gfap) |> unique()

result_csf_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_csf,
  var_resp = "fl_csf__ALL__gfap_log2"
)

#later I can add that if need  to save it
# plot_capture_csf_gfap <- grid::grid.grabExpr({
#   e_plot_complete_by_variable_subset(
#     dat              = dat_sheet,
#     var_list         = vars_gfap_csf,
#     var_resp         = "fl_csf__ALL__gfap_log2",
#     sw_n_complete = 5
#     )
# })
# 
# 
# ggsave(
#   "All_figures_for_paper/plot_capture_csf_gfap_tradeoff.png",
#   plot   = cowplot::ggdraw(plot_capture_csf_gfap),
#   width  = 10, height = 10, dpi = 300, bg = "white"
# )

# Print to choose best row
print(result_csf_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_csf_gfap$table$var_names[[5]]
cat("N complete:", result_csf_gfap$table$n_complete[5], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_csf <- "fl_csf__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_csf, "age", "sex")]

dat_lm_csf_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_csf), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_csf_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))


form_lower_csf_gfap <- as.formula(paste(var_resp_gfap_csf, "~ 1"))

fit_step_csf_gfap <- step(
  lm(form_main_csf_gfap, data = dat_lm_csf_gfap),
  scope     = list(lower = form_lower_csf_gfap, upper = form_upper_csf_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)


summary(fit_step_csf_gfap)

e_plot_lm_diagnostics(fit_step_csf_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_csf_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_csf_gfap,
  dat_cont                       = dat_lm_csf_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_csf_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_csf_gfap$plots,
  filename = "plot_csf_gfap_vs_all.png"
)



```

# GFAP as response :GFAP PLASMA VS PLASMA

```{r}
# ── PLASMA GFAP ~ CSF + Plasma + MRI ────────────────────────────────────────

sig_plasma_gfap <- c(

  # "fl_plasma__ALL__abeta_40_log2",
  "fl_plasma__ALL__ptau_181_log2",
  "fl_plasma__ALL__abeta_42_log2",
  "fl_plasma__ALL__vegf_a_log2",
  "fl_cyto_heparin_plasma_ifn_g_log2",
  "fl_plasma__ALL__il_1b_log2",
  "fl_cyto_heparin_plasma_il_2_log2",
  "fl_cyto_heparin_plasma_il_4_log2",
  "fl_plasma__ALL__mmp_9_log2",
    "fl_albumin_index_log2"

  # MRI: very few participants overlap with plasma GFAP — skip unless n allows
)

#     "fl_csf__ALL__ptau_log2",
#   "fl_cyto_csf_ifn_g_log2",
#   "fl_msd_csf_mmp_9_log2",
#   "fl_csf__ALL__mmp_10_log2"
vars_gfap_plasma <- c("fl_plasma__ALL__gfap_log2", sig_plasma_gfap) |> unique()

result_plasma_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_plasma,
  var_resp = "fl_plasma__ALL__gfap_log2"
)

plot_capture_plasma_gfap <- grid::grid.grabExpr({
  e_plot_complete_by_variable_subset(
    dat              = dat_sheet,
    var_list         = vars_gfap_plasma,
    var_resp         = "fl_plasma__ALL__gfap_log2",
  )
})

ggsave(
  "All_figures_for_paper/plot_capture_plasma_gfap_tradeoff.png",
  plot   = cowplot::ggdraw(plot_capture_plasma_gfap),
  width  = 10, height = 10, dpi = 300, bg = "white"
)

# Print to choose best row
print(result_plasma_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_plasma_gfap$table$var_names[[1]]
cat("N complete:", result_plasma_gfap$table$n_complete[1], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_plasma <- "fl_plasma__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_plasma, "age", "sex")]

dat_lm_plasma_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_plasma), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_plasma_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower_plasma_gfap <- as.formula(paste(var_resp_gfap_plasma, "~ 1"))

fit_step_plasma_gfap <- step(
  lm(form_main_plasma_gfap, data = dat_lm_plasma_gfap),
  scope     = list(lower = form_lower_plasma_gfap, upper = form_upper_plasma_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_plasma_gfap)

e_plot_lm_diagnostics(fit_step_plasma_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_plasma_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_plasma_gfap,
  dat_cont                       = dat_lm_plasma_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_plasma_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_plasma_gfap$plots,
  filename = "plot_plasma_gfap_vs_all.png"
)
```

# GFAP as response :GFAP CSF VS PLASMA

```{r}
# ── CSF GFAP ~ CSF + Plasma + MRI ───────────────────────────────────────────

sig_plasma_gfap <- c(
  
"fl_plasma__ALL__ptau_181_log2",
"fl_plasma__ALL__ptau_217_log2",
"fl_plasma__ALL__tie2_log2",
"fl_plasma__ALL__vegf_a_log2",
"fl_plasma__ALL__vegf_d_log2",
"fl_plasma__ALL__vegf_c_log2",
"fl_cyto_heparin_plasma_ifn_g_log2",
"fl_cyto_heparin_plasma_il_2_log2",
"fl_cyto_heparin_plasma_il_10_log2",
"fl_plasma__ALL__mmp_9_log2",
"fl_plasma__ALL__nfl_log2"
)

# MRi 
  # MRI (colored)
  # "mri_mean_cort_thick_mm",
  # "mri_dti_freewater_mean",
  #  "mri_dti_psmd"   ,
  #  "mri_percent_vol_hippo" 

vars_gfap_plasma <- c("fl_csf__ALL__gfap_log2", sig_plasma_gfap) |> unique()

result_csf_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_plasma,
  var_resp = "fl_csf__ALL__gfap_log2"
)

#later I can add that if need  to save it
# plot_capture_csf_gfap <- grid::grid.grabExpr({
#   e_plot_complete_by_variable_subset(
#     dat              = dat_sheet,
#     var_list         = vars_gfap_csf,
#     var_resp         = "fl_csf__ALL__gfap_log2",
#     sw_n_complete = 5
#     )
# })
# 
# 
# ggsave(
#   "All_figures_for_paper/plot_capture_csf_gfap_tradeoff.png",
#   plot   = cowplot::ggdraw(plot_capture_csf_gfap),
#   width  = 10, height = 10, dpi = 300, bg = "white"
# )

# Print to choose best row
print(result_csf_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_csf_gfap$table$var_names[[3]]
cat("N complete:", result_csf_gfap$table$n_complete[3], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_csf <- "fl_csf__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_csf, "age", "sex")]

dat_lm_csf_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_csf), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_csf_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower_csf_gfap <- as.formula(paste(var_resp_gfap_csf, "~ 1"))

fit_step_csf_gfap <- step(
  lm(form_main_csf_gfap, data = dat_lm_csf_gfap),
  scope     = list(lower = form_lower_csf_gfap, upper = form_upper_csf_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_csf_gfap)

e_plot_lm_diagnostics(fit_step_csf_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_csf_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_csf_gfap,
  dat_cont                       = dat_lm_csf_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_csf_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_csf_gfap$plots,
  filename = "plot_csf_gfap_vs_all.png"
)



```

# GFAP as response :GFAP plasma VS csf

```{r}
# ── PLASMA GFAP ~ CSF + Plasma + MRI ────────────────────────────────────────

sig_csf_gfap <- c(

  "fl_abeta_4240_ratio_log2",
  "fl_csf__ALL__ptau_log2",
  "fl_angio_csf_vegf_a_log2",
  "fl_angio_csf_vegf_c_log2",
  "fl_cyto_csf_ifn_g_log2",
  "fl_csf__ALL__il_1b_log2",
  "fl_cyto_csf_il_2_log2",
  "fl_cyto_csf_il_4_log2",
  "fl_cyto_csf_il_8_log2",
  "fl_cyto_csf_il_13_log2",
  "fl_msd_csf_mmp_1_log2",
  "fl_msd_csf_mmp_9_log2",
  "fl_csf_igg_log2",
  "fl_albumin_index_log2"
  
)


vars_gfap_csf <- c("fl_plasma__ALL__gfap_log2", sig_csf_gfap) |> unique()

result_plasma_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_csf,
  var_resp = "fl_plasma__ALL__gfap_log2"
)




plot_capture_plasma_gfap <- grid::grid.grabExpr({
  e_plot_complete_by_variable_subset(
    dat              = dat_sheet,
    var_list         = vars_gfap_plasma,
    var_resp         = "fl_plasma__ALL__gfap_log2",
  )
})

ggsave(
  "All_figures_for_paper/plot_capture_plasma_gfap_tradeoff.png",
  plot   = cowplot::ggdraw(plot_capture_plasma_gfap),
  width  = 10, height = 10, dpi = 300, bg = "white"
)

# Print to choose best row
print(result_plasma_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_plasma_gfap$table$var_names[[2]]
cat("N complete:", result_plasma_gfap$table$n_complete[2], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_plasma <- "fl_plasma__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_plasma, "age", "sex")]

dat_lm_plasma_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_plasma), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_plasma_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower_plasma_gfap <- as.formula(paste(var_resp_gfap_plasma, "~ 1"))

fit_step_plasma_gfap <- step(
  lm(form_main_plasma_gfap, data = dat_lm_plasma_gfap),
  scope     = list(lower = form_lower_plasma_gfap, upper = form_upper_plasma_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_plasma_gfap)

e_plot_lm_diagnostics(fit_step_plasma_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_plasma_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_plasma_gfap,
  dat_cont                       = dat_lm_plasma_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_plasma_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_plasma_gfap$plots,
  filename = "plot_plasma_gfap_vs_all.png"
)
```

# Plasma GFAP vs MRI 

```{r}


# ── PLASMA GFAP ~ CSF + Plasma + MRI ────────────────────────────────────────

sig_mri_gfap <- c(

"mri_dti_psmd",
"mri_dti_freewater_mean",
"mri_mean_cort_thick_mm"
  
)


vars_gfap_mri <- c("fl_plasma__ALL__gfap_log2", sig_mri_gfap) |> unique()

result_plasma_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_mri,
  var_resp = "fl_plasma__ALL__gfap_log2"
)

plot_capture_plasma_gfap <- grid::grid.grabExpr({
  e_plot_complete_by_variable_subset(
    dat              = dat_sheet,
    var_list         = vars_gfap_plasma,
    var_resp         = "fl_plasma__ALL__gfap_log2",
  )
})

ggsave(
  "All_figures_for_paper/plot_capture_plasma_gfap_tradeoff.png",
  plot   = cowplot::ggdraw(plot_capture_plasma_gfap),
  width  = 10, height = 10, dpi = 300, bg = "white"
)

# Print to choose best row
print(result_plasma_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_plasma_gfap$table$var_names[[2]]
cat("N complete:", result_plasma_gfap$table$n_complete[2], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_plasma <- "fl_plasma__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_plasma, "age", "sex")]

dat_lm_plasma_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_plasma), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_plasma_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_plasma_gfap <- as.formula(paste(
  var_resp_gfap_plasma, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower_plasma_gfap <- as.formula(paste(var_resp_gfap_plasma, "~ 1"))

fit_step_plasma_gfap <- step(
  lm(form_main_plasma_gfap, data = dat_lm_plasma_gfap),
  scope     = list(lower = form_lower_plasma_gfap, upper = form_upper_plasma_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_plasma_gfap)

e_plot_lm_diagnostics(fit_step_plasma_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_plasma_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_plasma_gfap,
  dat_cont                       = dat_lm_plasma_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_plasma_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_plasma_gfap$plots,
  filename = "plot_plasma_gfap_vs_all.png"
)
```


#CSF GFAP MRI 

```{r}
# ── PLASMA GFAP ~ CSF + Plasma + MRI ────────────────────────────────────────

sig_mri_csfgfap <- c(

"mri_dti_psmd",
"mri_dti_freewater_mean",
"mri_mean_cort_thick_mm",
"mri_percent_vol_hippo" 

  
)


vars_gfap_mri <- c("fl_csf__ALL__gfap_log2", sig_mri_csfgfap) |> unique()

result_csf_gfap <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_gfap_mri,
  var_resp = "fl_csf__ALL__gfap_log2"
)

plot_capture_plasma_gfap <- grid::grid.grabExpr({
  e_plot_complete_by_variable_subset(
    dat              = dat_sheet,
    var_list         = vars_gfap_csf,
    var_resp         = "fl_csf__ALL__gfap_log2",
  )
})

ggsave(
  "All_figures_for_paper/plot_capture_plasma_gfap_tradeoff.png",
  plot   = cowplot::ggdraw(plot_capture_plasma_gfap),
  width  = 10, height = 10, dpi = 300, bg = "white"
)

# Print to choose best row
print(result_csf_gfap[, c("n_complete", "n_vars")])

# ── Adjust row index after inspecting the trade-off plot ────────────────────
best_subset_vars <- result_csf_gfap$table$var_names[[5]]
cat("N complete:", result_csf_gfap$table$n_complete[5], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_gfap_csf <- "fl_csf__ALL__gfap_log2"

pred_vars <- best_subset_vars[!best_subset_vars %in%
               c(var_resp_gfap_csf, "age", "sex")]

dat_lm_csf_gfap <- dat_sheet |>
  dplyr::select(all_of(var_resp_gfap_csf), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_csf_gfap), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_csf_gfap <- as.formula(paste(
  var_resp_gfap_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

form_lower_csf_gfap <- as.formula(paste(var_resp_gfap_csf, "~ 1"))

fit_step_csf_gfap <- step(
  lm(form_main_csf_gfap, data = dat_lm_csf_gfap),
  scope     = list(lower = form_lower_csf_gfap, upper = form_upper_csf_gfap),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_csf_gfap)

e_plot_lm_diagnostics(fit_step_plasma_gfap,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2])

fit_contrasts_plasma_gfap <- e_plot_model_contrasts(
  fit                            = fit_step_csf_gfap,
  dat_cont                       = dat_lm_csf_gfap,
  sw_print                       = FALSE,
  sw_table_in_plot               = FALSE,
  adjust_method                  = "bonferroni",
  CI_level                       = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep              = "singles",
  sw_produce_plots               = TRUE
)

plot_plasma_gfap <- save_contrast_grid(
  p_list   = fit_contrasts_plasma_gfap$plots,
  filename = "plot_plasma_gfap_vs_all.png"
)
```




#pc1 VS CSF

```{r}

sig_composite_csf <- c(
  "fl_abeta_4240_ratio_log2",
  # "fl_csf__ALL__ptau_log2",
  # "fl_angio_csf_bfgf_log2",
  # "fl_angio_csf_flt_1_log2",
  # "fl_angio_csf_tie2_log2",
  # "fl_angio_csf_vegf_a_log2",
  # "fl_angio_csf_vegf_c_log2",
  # "fl_angio_csf_vegf_d_log2",
  # "fl_cyto_csf_ifn_g_log2",
  "fl_csf__ALL__il_1b_log2",
  "fl_cyto_csf_il_4_log2",
  # "fl_cyto_csf_il_6_log2",
  # "fl_cyto_csf_il_8_log2",
  "fl_cyto_csf_il_10_log2",
  # "fl_cyto_csf_il_12p70_log2",
  # "fl_cyto_csf_il_13_log2",
  "fl_cyto_csf_tnf_a_log2",
  "fl_msd_csf_mmp_1_log2",
  "fl_msd_csf_mmp_2_log2"
  # "fl_msd_csf_mmp_3_log2",
  # "fl_msd_csf_mmp_9_log2",
  # "fl_csf__ALL__mmp_10_log2",
  # "fl_csf__ALL__plgf_log2",
  # "fl_csf_igg_log2",
  # "fl_igg_index_log2"
  # "fl_myelin_basic_protein_log2",
  # "fl_csf__ALL__nfl_log2"
)



vars_pc1_csf <- c("PC1", sig_composite_csf) |> unique()

result_csf_pc1 <- e_plot_complete_by_variable_subset(
  dat      = dat_sheet,
  var_list = vars_pc1_csf,
  var_resp = "PC1"
)

plot_capture_csf_pc1 <- grid::grid.grabExpr({
  e_plot_complete_by_variable_subset(
    dat      = dat_sheet,
    var_list = vars_pc1_csf,
    var_resp = "PC1"
  )
})

ggsave(
  "All_figures_for_paper/plot_capture_csf_pc1_tradeoff.png",
  plot   = cowplot::ggdraw(plot_capture_csf_pc1),
  width  = 10, height = 10, dpi = 300, bg = "white"
)

# Print table to choose best row
print(result_csf_pc1$table[, c("n_complete", "n_vars")])

# Adjust row index after inspecting the trade-off plot
best_subset_vars <- result_csf_pc1$table$var_names[[6]]
cat("N complete:", result_csf_pc1$table$n_complete[6], "\n")
cat("Variables:", paste(best_subset_vars, collapse = ", "), "\n")

var_resp_pc1_csf <- "PC1"

pred_vars <- best_subset_vars[!best_subset_vars %in%
  c(var_resp_pc1_csf, "age", "sex")]

dat_lm_csf_pc1 <- dat_sheet |>
  dplyr::select(all_of(var_resp_pc1_csf), all_of(pred_vars), age, sex) |>
  tidyr::drop_na()

cat("N after drop_na:", nrow(dat_lm_csf_pc1), "\n")
cat("Predictors (", length(pred_vars), "):\n")
print(pred_vars)

form_main_csf_pc1 <- as.formula(paste(
  var_resp_pc1_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + ")
))

form_upper_csf_pc1 <- as.formula(paste(
  var_resp_pc1_csf, "~ age + sex +",
  paste(pred_vars, collapse = " + "), "+",
  paste0("age:", pred_vars, collapse = " + "), "+",
  paste0("sex:", pred_vars, collapse = " + ")
))

# Keep age and sex in all models
form_lower_csf_pc1 <- as.formula(paste(var_resp_pc1_csf, "~ age + sex"))

fit_step_csf_pc1 <- step(
  lm(form_main_csf_pc1, data = dat_lm_csf_pc1),
  scope     = list(lower = form_lower_csf_pc1, upper = form_upper_csf_pc1),
  direction = "both",
  k         = 2,
  trace     = 0
)

summary(fit_step_csf_pc1)

e_plot_lm_diagnostics(
  fit_step_csf_pc1,
  sw_plot_set = c(NA, "simple", "simpleAV", "all")[2]
)

fit_contrasts_csf_pc1 <- e_plot_model_contrasts(
  fit                             = fit_step_csf_pc1,
  dat_cont                        = dat_lm_csf_pc1,
  sw_print                        = FALSE,
  sw_table_in_plot                = FALSE,
  adjust_method                   = "bonferroni",
  CI_level                        = 0.95,
  sw_marginal_even_if_interaction = TRUE,
  sw_TWI_plots_keep               = "singles",
  sw_produce_plots                = TRUE
)

plot_csf_pc1 <- save_contrast_grid(
  p_list   = fit_contrasts_csf_pc1$plots,
  filename = "plot_csf_pc1_vs_all.png"
)



```


